### 2020-01-15 ms/ms.tex
Batch commit of manuscript draft first sent to co-authors

\documentclass[12pt]{article}
\usepackage[top=0.85in,left=1.0in,right=1.0in,footskip=0.75in]{geometry}
%\usepackage[parfill]{parskip}
\usepackage{setspace}
\usepackage{lineno}
\usepackage[hidelinks]{hyperref}
%\onehalfspacing
\doublespacing

\usepackage[round,sectionbib]{natbib}
\setcitestyle{authoryear}
\bibpunct{(}{)}{;}{a}{}{;}
\bibliographystyle{mee}

%\usepackage{textcomp}
%\usepackage{libertine}
%\usepackage{inconsolata} % sans serif typewriter

\usepackage{mathtools} % for dcases
\usepackage{xcolor} % for textcolor
\usepackage{makecell} % for \makecell and within cell line breaks (\thead)

\usepackage[T1]{fontenc}

%\makeatletter\let\expandableinput\@@input\makeatother % expandable input for \input inside tables

% Linux Libertine:
\usepackage{textcomp}
\usepackage[sb]{libertine}
\usepackage[varqu,varl]{inconsolata}% sans serif typewriter
\usepackage[libertine,bigdelims,vvarbb]{newtxmath} % bb from STIX
\usepackage[vvarbb]{newtxmath} % bb from STIX, removed bigdelims for ScholarOne rendering
\usepackage[cal=boondoxo]{mathalfa} % mathcal
%\useosf % osf for text, not math
%\usepackage[supstfm=libertinesups,%
%  supscaled=1.2,%
%  raised=-.13em]{superiors}

\usepackage{xspace}
\usepackage{xfrac} % for diagonal inline fractions in text
%\usepackage{array} % for making whole row bold in table
\usepackage{colortbl} % for background colours in table rows
\usepackage{longtable}
\usepackage{amssymb} % for \checkmark 
\usepackage{amsmath} % for \checkmark 
\usepackage{rotating}
\usepackage[nolists,tablesfirst]{endfloat} % for putting figs and tables at end of document
\DeclareDelayedFloatFlavor{sidewaystable}{table}
\usepackage{makecell} % for \makecell in tables
\usepackage{doi}

%\usepackage{xr} % to obtain label references from supp materials file
%\externaldocument[S-]{suppmat}

\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{array} % for table wrapping of columns
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
%\usepackage[detect-all]{siunitx} % for SI units

% custom hyphenation:
\hyphenation{inverse}
\hyphenation{At-lantic}
\hyphenation{elasmo-branchs}

% Macros
\newcommand{\So}{$S_{1}$\xspace}
\newcommand{\St}{$S_{2}$\xspace}
\newcommand{\Pg}{$P_r$\xspace}
\newcommand{\prmu}{$\mu_r$\xspace}
\newcommand{\prsig}{$\sigma_r$\xspace}
\newcommand{\Linf}{$L_{\infty}$}
\newcommand{\DWinf}{$DW_{\infty}$}
\newcommand{\alphat}{$\tilde{\alpha}$}
\newcommand{\lamat}{$l_{\alpha_{mat}}$}
\newcommand{\lamatb}{$l_{\alpha_{mat}}b$}
\newcommand{\rmax}{$r_{max}$\xspace}
\newcommand{\ageratio}{$\alpha_{mat}/\alpha_{max}$}
\newcommand{\yr}{year\textsuperscript{-1}}
\newcommand{\rsq}{$R^2$\xspace}
\newcommand{\mytilde}{\raise.17ex\hbox{$\scriptstyle\mathtt{\sim}$}}
% Select what to do with command \comment:  
%\newcommand{\comment}[1]{}  % comment not shown
\newcommand{\comment}[1]{\par {\bfseries \color{blue} #1 \par}} % comment shown
%% END MACROS SECTION

\begin{document}
\linenumbers

\section*{Trends in marine survival of Altantic salmon in eastern Canada}

\textbf{Sebasti\'{a}n A. Pardo\textsuperscript{1*}, 
        Geir H. Bolstad\textsuperscript{2}, 
        J. Brian Dempson\textsuperscript{3}, 
        Julien April\textsuperscript{4}, 
        Ross A. Jones\textsuperscript{5}, 
        Martha J. Robertson\textsuperscript{3}, 
        Dustin Raab\textsuperscript{6}, 
Jeffrey A. Hutchings\textsuperscript{1}}

\noindent\small{\textsuperscript{1} Department of Biology, Dalhousie University, Halifax, NS, B3H 4R2, Canada\\}
\small{\textsuperscript{2} Norwegian Institute for Nature Research (NINA), Trondheim, Norway\\}
\small{\textsuperscript{3} Department of Fisheries and Oceans Canada, St. John's, NL, Canada\\}
\small{\textsuperscript{4} Minist\`{e}re des For\^{e}ts, de la Faune et des Parcs, Qu\`{e}bec, QC, Canada\\}
\small{\textsuperscript{5} Department of Fisheries and Oceans Canada, Moncton, NB, Canada\\}
\small{\textsuperscript{6} Department of Fisheries and Oceans Canada, Dartmouth, NS, Canada\\}
\small{\textsuperscript{*} Corresponding author: spardo@dal.ca}

\section*{Abstract}

The marine phase of anadromous Atlantic salmon (\emph{Salmo salar}) is the
least known yet one of the most crucial with regards to population
persistence.
Recently, declines in many Atlantic salmon populations in eastern Canada have
been attributed to changes in the conditions at sea, thus affecting their
survival.
However, marine survival estimates are difficult to obtain given that many
individuals spend multiple winters in the ocean before returning to freshwater to
spawn; therefore multiple parameters need to be estimated.
We used smolt and adult abundance time series for seven populations, ranging between 19 and 42 years,
to fit a hierarchical Bayesian maturity schedule model and estimate three components of
marine survival: survival in the first year at sea (\So), survival in the second year
at sea (\St), and proportion returning as one sea-winter (\Pg).
Trends in estimates of \So were not consistent among rivers, with some population showing increases in \So over
time, other showing decreases, and some being relatively stable.
Estimates of \St were highly uncertain and therefore trends could not be assessed, while 
estimates of \Pg were mostly stable across time, except for the LaHave and Nashwaak Rivers.
Our findings challenge the assumption that marine survival has decreased for
all salmon populations, and suggest that trends can be river-specific. 
Further work exploring potential correlates of these trends in marine survival
is warranted.

Keywords: salmonid, survival at sea, natural mortality, marine mortality

%Running Head:

\section*{Introduction} % (4-5 paragraphs)

%1. Declining salmon numbers
Across eastern Canada, many wild populations of Atlantic salmon (\emph{Salmo salar})
have declined to the point of being considered Threatened by the 
Committee on the Status of Endangered Wildlife in Canada \citep[COSEWIC, ][]{Cosewic2010}.
While it is not fully understood what is driving these declines, the potential
drivers of these are many \citep[see ][for a detailed discussion of possible
causes]{Cairns2001}: fishing mortality, damming of waterways
and changes in their freshwater habitats \citep{Dunfield1985}, acidification
of freshwater habitats \citep[particularly in the Southern Uplands region of
NS, see][]{Gibson2010}, predation by seals and birds \citep{Cairns2000}, negative
effects of interbreeding or interactions with escaped farmed salmon
\citep{Keyser2018}, and changes in their marine survival.

% 2. 
Theres a common perception that marine survival of Atlantic salmon has
declined throughout the North Atlantic \citep{OMaoileidigh2003, Hansen1998,
    Chaput2012a}. This perception is widespread both in the scientific
literature as well as federal reports. The latest status report on Atlantic
salmon by COSEWIC states that ``While the mechanism(s) of marine mortality is
uncertain, what is clear is that the recent period of poor sea survival is
occurring in parallel with many widespread changes in the North Atlantic
ecosystem.'' \citep{Cosewic2010}. Consequently, there have several lines of
inquiry as to what the causes behind these declines might be attributable to
\citep{Friedland1993, Friedland1998}.

Nonetheless, most of the populations declines assessed by COSEWIC have
occurred in populations in the southern extent of its distribution:
populations in the Bay of Fundy, Anticosti Island, and the Atlantic coast of
Nova Scotia being assessed as Endangered, populations in the south coast of
Newfoundland assessed as Threatened, And the populations in the New Brunswick
and Qu\`{e}bec coasts of the Gulf of St. Lawrence assessed as Special Concern
\citep{Cosewic2010}. On the other hand, northermost populations have shown
stable, or even increasing population trends, suggesting that, if marine
survival were to be a factor in many of these declines, these changes are not
uniformly distributed across all populations.

% 3. Nonetheless, this premise has never been examined in detailed.
While the purported decline of marine survival is mentioned widely in the 
scientific literature, only relatively few studies have quantified it in detail.
The basis for this premise is a limited number of time series that exhibit
temporal declines in a proxy of marine survival (i.e. return rates or
post-smolt survival).
\citet{Chaput2012a} looked at return rates of smolts to adult salmon to assess
declines in marine survival and found that most populations from Canadian
rivers had declining return rates. 
However, looking at trends in return rates
can mask changes in differential survival in different years at sea, as well
as changes in proportion of adults returning after one or two years at sea.
A more recent study by \citet{Olmos2019} suggested that trends in post-smolt
survival, when estimated at the stock unit level, are synchronously declining
for all Atlantic salmon in Eastern North America and Canada. Nonetheless,
their analysis relies on stock-recruitment relationships rather than smolt
abundance indices.

%4. 
In this study, we compile smolt and returns times series data from seven wild
populations of Atlantic salmon in eastern Canada to model trends in marine survival.
While previous studies have used maturity schedule models to estimate marine
survival for a couple of Atlantic salmon populations \citep{Chaput2003b}, none
have incorporated data from the last decades and they have not examined
trends among more than a couple of populations.

\section*{Methods}

\subsection*{Data}

We obtained time series data of outmigrating smolt and returning adult
abundances for seven Atlantic salmon populations in eastern Canada, covering a
wide range of the species' distribution (Fig.~\ref{fig:map}). These
populations included the LaHave River in the Southern Uplands region of Nova
Scotia (NS), The Nashwaak River in New Brunswick (NS) which connects to the
Bay of Fundy, The Trinit\'{e} and Saint-Jean Rivers in Qu\`{e}bec (QC), which
flow to the St Lawrence River and Bay of St Lawrence, respectively, and the
Western Arm Brook, Campbellton, and Conne in Newfoundland, which flow to the
Bay of St Lawrence, Labrador Sea, and Atlantic Ocean, respectively.
Data were collected in NS, NB, and Newfoundland
and Labrador (NL) by Fisheries and Oceans Canada (DFO) and in QC
by the Minist\`{e}re des For\^{e}ts, de la Faune et des Parcs, Qu\`{e}bec.

\begin{figure}[htbp] \centering
    \includegraphics[width=0.85\linewidth]{figures/rivers-map2.png}
    \caption{Locations of the seven rivers in eastern Canada with time series abundance data of outmigrating smolts and 
    returning adults.} \label{fig:map} 
\end{figure}

\subsubsection*{Smolt data}

\comment{Co-authors: please check data sources}

Smolt estimates originate from a variety of sources.
Estimates from the XXXX populations were obtained using a capture-mark-recapture approach, while
Estimates from the XXXX populations were obtained by direct counts using a XXXX.
Refer to ... for more details on the data collection details.

\subsubsection*{Returns data}

Yearly return data is often recorded in terms of two size groups: small ($< 63$ cm
FL) and large ($\geq 63$ cm FL) salmon, as these closely represent different
life-history strategies (i.e. 1SW and 2SW), but can be conflated with repeat
spawners of different sizes. To correct for this, we estimated the abundance
of 1SW and 2SW returns using yearly scale data of a subsample of returns:

\begin{equation}
    p_{r,y,a} = \frac{\sum_{s}{(\frac{n_{r,y,s,a}}{n_{r,y,s}} * N_{r,y,s})}}{\sum_{s}{N_{r,y,s}}}
\end{equation}

where $p_{r,y,a}$ is the proportion of annual returns in river $r$, year $y$,
and of spawning history $a$ (either 1SW or 2SW returns); $n_{r,y,s,a}$ is the
number of samples in river $r$, year $y$, of spawning history $a$, and of size
group $s$; $n_{r,y,s}$ is the total number of samples in river $r$, year $y$,
and of size group $s$; and $N_{r,y,s}$ is the returns of salmon
in river $r$, year $y$, and of size group $s$.

\subsection*{Bayesian model}

We developed a hierarchical Bayesian model that uses Murphy's maturity
schedule method, in conjunction with informative priors, to estimate yearly
marine survival in seven populations of Atlantic salmon. We account for
observation error in smolt and return estimates, as well as estimating the
proportion returning after one winter (i.e. \Pg) hierarchically.
There is an identifiability problem in the maturity schedule equations where
the parameter estimates cannot be optimally solved \citep{Chaput2003a}.
However, this issue can be mathematically overcome, at least partially, by
using informative priors for all three marine survival parameters in a
Bayesian framework.
This method requires abundance estimates of smolts as well as abundance estimates
of returning one-sea-winter (1SW) and two-sea-winter adults. With these data,
it estimates three parameters: survival in the first year at sea (\So), survival
in the second year at sea (\St), and the proportion of fish returning after one
year at sea (\Pg). 
Our model does not include repeat spawners and assumes that no fish spend
three or more winters at sea before returning to spawn for the first time.

The model also assumes that mortality in the second winter at sea 
is additional to mortality in the first winter at sea in the previous year, 
and therefore does not account for differences
the returning (Pr) occurs just before
actually being counted as returns and that S2 is any additional mortality in
the subsequent year. I agree this is far from ideal, given that we know there
is a period of a few months where 1SW returns are subject to a different
environment than those 2SW+ fishes.

%\comment{Other assumptions...}

Observed smolt estimates were modelled hierarchically and included
observation error:

\begin{equation}
log(smolts_{obs,t,r}) = log(smolts_{true,t,r}) + \epsilon_{t,r}
\end{equation}

where $smolts_{true,t,r}$ are the true smolt abundances for year $t$ and river
$r$, and $\epsilon_{t,r}$ is the error term, which is empirically derived from
by calculating the yearly coefficient of variation in the empirically derived
smolt estimates (see Table S1 in the Supplementary material). 
The log-transformed true smolt abundances are
normally distributed around a population-level mean and standard deviation:

\begin{equation}
log(smolts_{true,t,r}) \sim Normal(\mu_{smolts,r}, \sigma_{smolts,r})
\end{equation}

where $\mu_{smolts,r}$ and $\sigma_{smolts,r}$ are parameters estimated by the
model for each river.

Once we have yearly estimates of smolt, 1SW, and 2SW abundances, we estimate
marine survival parameters using Murphy's maturity schedule method
\citep{Murphy1952, Ricker1975}:

\begin{align}
    R_{1,t} &= smolts_{t-1} * S1_t * Pr_t \label{eq:1}, \\
    R_{2,t+1} &= smolts_{t-1} * S1_t * (1 - Pr_t) * S2_{t+1} \label{eq:2}
\end{align}

where $R_{1,t}$ and $R_{2,t+1}$ are the estimated abundances of 1SW and 2SW
salmon returning in years $t$ and $t+1$, respectively, $smolts_{t-1}$ is the
number of outmigrating smolts in year $t-1$, $S1_t$ is the proportion of
salmon surviving in their first year ($t$) at sea, $Pr_t$ is the proportion of
salmon that return to spawn at year $t$, $S2_{t+1}$ is the survival in their
second year at sea of the same cohort of salmon who did not return to spawn at
year $t$.

However, to allow for normally and log-normally-distributed parameters we
log-transform equations~\ref{eq:1} and~\ref{eq:2} to obtain

\begin{align}
    log(R_{1,r,t}) &= log(smolts_{r,t-1}) + log(Pr_{r,t}) - Z_{1,r,t} + \epsilon_{1,r} \label{eq:3}, \\
    log(R_{2,r,t+1}) &= log(smolts_{r,t-1}) - Z_{1,r,t} + log(1 - Pr_{r,t})  - Z_{2,r,t+1} + \epsilon_{2,r} \label{eq:4} 
\end{align}

where $Z_{1,r,t}$ and $Z_{2,r,t+1}$ are the instantaneous mortality rates and $\epsilon_{1,r}$ and $\epsilon_{2,r}$
are the process error terms. 
These error terms are normally distributed with a standard deviation
of 0.01, which are almost equivalent to using a coefficient of variation in
return estimates 1\%, given that return estimates in equations~\ref{eq:3}
and~\ref{eq:4} are log-transformed:

\begin{align}
\epsilon_{1,r} &\sim Normal(0, 0.01) \\
\epsilon_{2,r} &\sim Normal(0, 0.01)
\end{align}

Furthermore, we use instantaneous mortality rates in the model instead of survival probabilities
as the model is more efficient in its parameter search in log-space, and instantaneous rates
are easy to interpret. Instantaneous rates are easily converted to survival probabilities by

\begin{align}
 S_{1} &= e^{-Z_1}, \\
 S_{2} &= e^{-Z_2} 
\end{align}

We estimate population-level mean \Pg values around which the yearly \Pg
values are normally distributed. We specify different informative hyperpriors
for \prmu and \prsig based on whether the population is 1SW-dominated or not:

\begin{align}
    logit(P_{r,t}) &\sim Normal(\mu_r, \sigma_r) \\
    \mu_r &\sim 
    \begin{cases}
       Normal(2.3, 0.4),  &\text{for 1SW-dominated populations} \\
       Normal(0, 2.8), &\text{for non-1SW-dominated populations} \\
   \end{cases} \\
    \sigma_r &\sim halfNormal(0, 1)
\end{align}

The priors for \So and \St are specified as log-normal distributions around
the instantaneous mortality rates $Z_1$ and $Z_2$:

\begin{align}
Z_1 &\sim logNormal(1, 0.22),   \\ 
Z_2 &\sim logNormal(0.2, 0.3)
\end{align}

The model was written in Stan \citep{Carpenter2017} and run in R version 3.6.1
\citep{RCoreTeam2019} using the \texttt{rstan} package version 2.19.2
\citep{StanDevelopmentTeam2019}.
The model was run with three chains and 3,000 iterations, with the first 1,500
discarded as a burn-in. The models were considered to have converged when the
$\hat R$ of all parameters were lower than 1.03 and the effective sample size 
were higher than 500.

\subsection*{Correlations among trends in survival}

%\comment{How do were quantitatively compare trends among rivers? DFA is an
%option, but so is a t-test of averages before and after a certain year}

We looked at the correlation of trends in \So by calculating the Pearson's
correlation coefficient for the Z-scores of these trends. Given that the time
series do not cover the same years, and that some rivers have missing years in
the middle of the time series, pair-wise Pearson's tests were done using only
the years where there is data for both rivers.

\section*{Results}

%\subsubsection*{Model convergence}

\subsubsection*{Trends in marine survival parameters}

Trends in estimates of \So were highly variable within among rivers
(Fig.~\ref{fig:s1-dual}). The highest median posterior estimates of \So
were for the Nashwaak River in 2006 and 2008, with values of 0.18 and 0.21,
respectively. The lowest median \So estimate was in the Trinit\'{e} in 2001,
with an estimate of \So of 0.007, while the Conne, LaHave, and Trinit\'{e} had
years of estimates of \So between 0.01 and 0.02 (Fig.~\ref{fig:s1-faceted}).

Trends among populations were variable as well: roughly, the Campbellton,
Saint-Jean, and Western Arm Brook populations showed increases in \So
over time, the Trinit\'{e}, Conne, and La Have populations showing decreases,
and the Nashwaak population showing an increase in median \So during the early
2000s but a decrease in the 2010s. Yearly estimates of \So had very little
variability for one sea-winter dominated populations (Conne, Campbellton, and
WAB), but were more variable (i.e. wider credible intervals) in the other
populations.

\begin{figure}[htbp] \centering
    \includegraphics[width=0.95\linewidth]{figures/s1-trends-dual.png}
    \caption{Trends in survival in the first year at sea (\So). a) Posterior
        estimates of \So for the seven rivers examined, error bars indicate
        the 90\% credible intervals. b) Z-scores of median posterior
        estimates.} \label{fig:s1-dual} 
\end{figure}

\begin{figure}[htbp] \centering
    \includegraphics[width=0.95\linewidth]{figures/s1-trends-faceted.png}
    \caption{Posterior estimates of \So for the seven rivers examined, error
        bars indicate the 90\% credible intervals.} \label{fig:s1-faceted}
\end{figure}

Estimates of \St were highly uncertain in all rivers, and trends of any kind
were not apparent in most rivers given the large range of the credible
intervals in the yearly estimates (Fig.~\ref{fig:s2-faceted}). The estimates
of \St for the Saint-Jean and Trinit\'{e} were considerably less uncertain
than for the other populations, but showed no apparent trends through time.

\begin{figure}[htbp] \centering
    \includegraphics[width=0.95\linewidth]{figures/s2-trends-faceted.png}
    \caption{Posterior estimates of \St for the seven rivers examined, error
        bars indicate the 90\% credible intervals.} \label{fig:s2-faceted}
\end{figure}

Estimates of \Pg were mostly stable across time, except for the LaHave and
Nashwaak Rivers; The estimates of \Pg were slightly lower in the last four
years than in the previous ones, while in the Nashwaak the posterior estimates
of \Pg in 2012 were much lower than in all other years
(Fig.~\ref{fig:s2-faceted}). Uncertainty in yearly estimates was highest in
the LaHave, Nashwaak, and Trinit\'{e}, and lowest in the 1SW-dominated
populations.

\begin{figure}[htbp] \centering
    \includegraphics[width=0.95\linewidth]{figures/pr-trends-faceted.png}
    \caption{Posterior estimates of \Pg for the seven rivers examined, error
        bars indicate the 90\% credible intervals.} \label{fig:pr-faceted}
\end{figure}

Population-level estimates of \prmu and \prsig varied considerably among rivers. For all
three 1SW-dominated rivers (Campbellton, Conne, and WAB), estimates of \prmu
were very close to 1.0 and had little variability in \prsig (Fig~\ref{fig:prmu-post}).
Estimates of \prmu were the lowest for the two QC rivers, particularly the Saint-Jean (median \prmu = 0.11).
Estimates of \prmu for the Nashwaak and the LaHave Rivers were close to 0.5, with these two rivers having 
the highest estimated values of \prsig, particularly the Nashwaak (Fig~\ref{fig:prmu-post}).

\begin{figure}[htbp] \centering
    \includegraphics[width=1.0\linewidth]{figures/pr-mu-posteriors.png}
    \caption{Posterior estimates of the population-level parameters $Pr_{\mu}$
       $logit(Pr_{\mu})$, and $logit(Pr_{\sigma})$. Dots denote median estimates, while the thick and thing error bars indicate
       the 50\% and 90\% credible intervals, respectively.} 
   \label{fig:prmu-post} 
\end{figure}

\subsubsection*{Correlations}

Most correlations between z-scored trends were not significant, with only 
three out of the 21 pair-wise comparisons having a p-value below 0.01 (Fig.~\ref{fig:s1-corr}a).
When looking at the direction of the correlation, regardless whether they were
significant or not, these spanned both positive and negative coefficients
(Fig.~\ref{fig:s1-corr}b).

\begin{figure}[htbp] \centering
    \includegraphics[width=0.95\linewidth]{figures/corr-s1.png} \caption{
        Correlation among Z-scores of median estimated trends in \So among
        rivers. a) Pearson's correlation coefficients are shown in each square,
        while colouring denotes significance of the correlation ($p \leq 0.01$), b)
        colours denote direction and magnitude of correlation, while asterisks denote significance.}
\label{fig:s1-corr} 
\end{figure}

\section*{Discussion}

% 1. Main findings
% 2. How they compare to other publications
% 3. Potential reasons
% 4. Caveats
% 5. Future directions

% 1. Main findings
Our findings challenge the widespread idea that marine survival is declining
uniformly across the range of Atlantic salmon in the northwest Atlantic. We
find that there is no apparent shared trends among the seven salmon
populations examined; some rivers show positive trends in survival in the
first winter at sea (\So) while other show stable trends, and some even show
declines. We could not assess trends in the second winter at sea (\So) or
proportion returning as grilse (\Pg) as these parameter estimates were highly
uncertain and were strongly influenced by the priors used.
Perhaps there is a need to rethink our understanding of Atlantic salmon
population dynamics in light that any real decline in marine survival was
experienced by some populations, that reductions in survival might
have occurred over a brief period of time, and that since that
reduction occurred (for a subset of populations) marine survival has actually
been stable, or even increasing, since that point.

% 2. How they compare to other publications
Our findings are contrary to those of \citet{Olmos2019}, who detected positive
correlations in post-smolt survival among broad stock units. These differences
could be due to a number of reasons: different model specifications and
structure, different methods for estimating covariance, difference in the
spatial scales of data sources, and the use stock-recruitment relationships
rather than empirical smolt observations to estimate marine survival.
Interestingly, our estimates of \So and \St are very similar to those produced
by \citet{Chaput2003b}, and the trends are almost identical for the
overlapping time period that marine survival was estimated for in that study
(1984-1998). While that study separated abundance data for males and females
and assumed their survival rates were the same (to be able to reach an
analytical solution), our study reached almost the same results (albeit with
slightly higher uncertainty) using a Bayesian approach with informative
priors. These overlapping trends obtained with two different methods are
indicative that our method is effective at estimating marine survival.
While \citet{Chaput2003b} did not compare trends in marine survival
among populations, \citet{Chaput2012a} did, using adult return rates, and
found that for four out of the six population examined, return rates in the 1990s 
were less than those during the 1970s.
\citet{Friedland1993} compared return rates for a number rivers in eastern
North America and suggested there are similar trends among these. The
similarity in trends is driven primarily by two years, 1977 and 1978, which
show concurrent low and high relative return rates across rivers,
respectively. Others years are much more variable relative to each other.
Their time series stops at 1988, thus there are only a few years for which 
to assess overlap with the time series in our study.
However, using pooled adult return rates, as done by \citet{Chaput2012a} and
\citet{Friedland1993} can mask inter-annual variability in marine survival,
and hence might not produce an accurate depiction of marine survival trends.
\citet{Dempson2003} describe a general declining trend in marine survival for
Newfoundland rivers (except the WAB); which in our study we observe in the
Conne but not the Campbellton or the WAB. It is not possible to draw conclusions
with data from only three Newfoundland rivers, but it seems that among index rivers,
those in Newfoundland are among those with the highest marine survival rates.

% 3. Potential reasons
There are multiple potential explanations as to why there are no synchronous
trends in estimates of \So. 
Marine survival in the first winter at sea could be highly variable between
rivers from environmental factors (e.g., temperature, predation) being
variable at localized spatial scales rather than ocean-wide ones.
Given that synchrony is being observed for marine survival trends estimated using
stock-recruitment relationships and broader spatial scales \citep{Olmos2019},
these overlapping declining trends in marine
survival estimated using stock-recruitment relationships have been confounded
by changes in recruitment dynamics.
There is some evidence of a correlation between return rate and growth (as
indicated by intercirculi spacing in scales), where years of poor growth
tended to also be years of poor survival \citep{Friedland1993}, supporting the
idea that environmental variability can affect marine survival.
Furthermore, among European salmon, there is evidence of a positive correlation
between spring temperature and population abundance, suggesting warmer
conditions favour post-smolts \citep{Friedland1998} based on mapping the
extent of area of suitable temperature (7-13 \textdegree C) in the North Atlantic.
Nonetheless, how warming impacts survival of post-smolts can differ based
whether it results in temperatures coming into or out of their suitable range;
which could be part of why populations in eastern North America are declining
in the southern part of their range while potentially increasing in the
northern part. The relationship between temperature and survival warrants 
further study.
While there is very little evidence that marine survival is density-dependent
in Atlantic salmon, and it is broadly considered to be a density-independent
process \citep{Jonsson1998,Gibson2006}, there could potentially be some
density-dependent relationship, particularly for populations that are severely
depleted (i.e. Allee effects). Exploring relationships between survival
and population size could potentially shed light about the processes that
have caused many of the population declines we have observed.

% 4. Caveats
Nonetheless, there are caveats to consider in our modelling approach.
While there are analytical issues with estimation of \So, \St, and \Pg,
The assumption that \St is additive to \So could produce unrealistic results.
Overcoming this assumption would require an additional parameter to be
estimated, or an additional assumption as to what proportion of \So is not
additive to \St (as the returning 1SW adults do not experience the same
environment when they return to their natal streams as those fish who stayed
at sea for an additional winter).
Furthermore, the assumed hierarchical structure might not be the most adequate
for modelling smolt abundances. This approach results in shrinkage to the
mean, which means that the variability of yearly smolt estimates is less than
it would be if not modelled hierarchically. Nonetheless, this assumption
likely results in more conservative trends in marine survival, as the
variation of smolt estimates is shrunk.
It is important to note that the hierarchical structure in the estimation of
\Pg seems like a reasonable assumption given that the probability of returning
as grilse has a genetic component associated to it \citep{Aykanat2019} and is
not expected to vary much, within a population, among years.

% 5. Future directions
Perhaps reframing the issue of marine survival is key to further our
understanding of Atlantic salmon population dynamics: it is not that marine
survival has declined across rivers, but the fact that it has stayed at
roughly similar levels as previously \emph{despite} their reduced abundance
what we should focus into understanding better.
Given the literature of correlates on marine survival in the past, if survival
has been stable or increasing, this clearly warrants reexamination of the
correlates and causes of temporal and spatial shifts in marine survival.

\section*{Acknowledgements}

% We would like to thank Geir Bolstad, G\'{e}rald Chaput, Brian Dempson, and Martha
% Robertson for their useful discussions on estimating marine survival in
% Atlantic salmon, and Amanda Kissel for her helpful comments on the manuscript.
% Brian Dempson and Geoff Venoit for providing the data Conne River data.
We would like to thank G\'{e}rald Chaput for his useful discussions on
estimating marine survival in Atlantic salmon, and Sean Anderson for his help
with implementing the non-centered parameterization of the model.
This research was supported by the Atlantic Salmon Conservation Foundation and
the Atlantic Salmon Research Joint Venture.

\section*{Conflicts of Interest}

The authors declare no conflicts of interest.

\bibliography{subset}

%\input{ms.bbl}

\end{document}

### 2020-01-15 analysis/00-campbellton-wrangling.R

@@ -1,0 +1,66 @@
+### WAB data
+
+library(tidyverse)
+
+campsmolts <- read_csv("raw-data/NL-smolt-estimates-Kelly-et-al-2017-ResDoc.csv",
+         na = c("-", "NA")) %>%
+  rename(year = Year) %>%
+  gather(river_name, smolt, -year) %>%
+  filter(river_name == "Campbellton River") %>%
+  na.omit()
+
+# Loading estimated returns of 1SW, 2Sw, and RS
+camp <- readRDS(file = "~/Dropbox/salmon-lh-variation/output/NL/campbellton-estimated-returns.rds")
+
+campall <- right_join(camp, campsmolts, by = "year") %>%
+  select(year, smolt, estM1, estM2p, estRS, totalreturns) %>%
+  mutate(river_name = "Campbellton River")
+  
+
+
+campreturns <- campall %>%
+  mutate(smoltlag = lag(smolt),
+         #smolt025lag = lag(smolt_025),
+         #smolt975lag = lag(smolt_975),
+         est_1SW = estM1,
+         est_2SWlead = lead(estM2p),
+         est_RSlead = lead(estRS)
+         #logsmoltsdlag = sqrt(300) * (log(smolt975lag) - log(smolt025lag))/3.92
+  ) %>%
+  filter(year %in% 1994:2015)
+
+campreturns %>%
+  filter(est_2SWlead != 0) %>%
+  summarise(min2SW = min(est_2SWlead)/2)
+
+ggplot(campreturns, aes(year, smolt)) + geom_point() + geom_line() # +
+
+campreturns[campreturns$est_2SWlead == 0, "est_2SWlead"] <- 1.351 # roughly min(est_2SW)/2
+
+campreturn_rate <- campreturns %>%
+  mutate(return_rate = est_1SW/smoltlag) %>%
+  select(year, return_rate) %>%
+  mutate(river_name = "Campbellton River")
+
+ggplot(campreturn_rate, aes(year, return_rate)) +
+  geom_point() + geom_line() + ggtitle("Campbellton return rates")
+
+saveRDS(campreturn_rate, file = "data/campbelltonreturnrate.rds")
+
+# Creating list for use in TMB/Stan
+campdata <- list(years = campreturns$year,
+                  logsmolts = log(campreturns$smoltlag),
+                  #logsmolts_SE = nashwaakreturns$logsmoltsdlag,
+                  logsmolts_cv =  rep(0.05, length(campreturns$smoltlag)), # should be replaced with actual uncertainty, if available
[…]

### 2020-01-15 analysis/00-conne-wrangling.R

@@ -6,11 +6,20 @@ library(tidyverse)
 conne <- readRDS(file = "~/Dropbox/salmon-lh-variation/output/NL/conne-estimated-returns.rds")
 
 # Loading small, large, and smolt data
+# Smolt years missing: up to 1988, 2016 and 2017, which
+# is the duration of the dataset, so no need to remove years in the middle
 connewithsmolts <- readRDS(file = "~/Dropbox/salmon-inverse-matrix/data/connereturns.rds")
 
+# Adding missing years based on Dempson et al 2003 (Ch 1)
+connewithsmolts[connewithsmolts$year == 1987, "smolt"] <- 74585
+connewithsmolts[connewithsmolts$year == 1988, "smolt"] <- 65692
+
+
 # joining by year
-conneall <- left_join(conne, connewithsmolts, by = "year") %>% print(n = 32)
-  select(year, smolt, estM1, estM2p, estRS, totalreturns)
+conneall <- left_join(conne, connewithsmolts, by = "year") %>%
+  select(year, smolt, estM1, estM2p, estRS, totalreturns) %>%
+  mutate(river_name = "Conne River")
+  
 
 connereturns <- conneall %>%
   mutate(smoltlag = lag(smolt),
@@ -21,29 +30,47 @@ connereturns <- conneall %>%
          est_RSlead = lead(estRS),
          #logsmoltsdlag = sqrt(300) * (log(smolt975lag) - log(smolt025lag))/3.92
          ) %>% 
- filter(year %in% 1999:2015) #  2017) # NEED 2017 SCALE DATA FOR 1SW/2SW/RS ESTIMATION IN BOTH 2016 AND 2017, ASK BRIAN
+ filter(year %in% 1988:2015) #  2017) # NEED 2017 SCALE DATA FOR 1SW/2SW/RS ESTIMATION IN BOTH 2016 AND 2017, ASK BRIAN
 
 ggplot(connereturns, aes(year, smolt)) + geom_point() + geom_line() # +
  # geom_errorbar(aes(ymin = smolt_025, ymax = smolt_975))
 
 # Removing zeros in 2SW-RS estimates as they cannot be log transformed
-connereturns$est_2SWlead[connereturns$est_2SWlead == 0] <- 0.0001
+# not needed after fixing years with very few large scales (and 0s in estimated 2SW returns)
+# connereturns$est_2SWlead[connereturns$est_2SWlead == 0] <- 0.0001
[…]

### 2020-01-15 analysis/00-lahave-wrangling.R

@@ -1,0 +1,136 @@
+### LaHave
+
+library(tidyverse)
+library(readxl)
+library(lubridate)
+
+# load biological characteristic data for returns (to assign 1SW-2SW)
+lahave_bio_adult <- read_csv("raw-data/lahave/Lahave_TRadult.csv") %>%
+  mutate(RCDATE = mdy(RCDATE))
+
+# Reading additional 2018 data not needed because there aren't any smolt 
+# abundance estimates for that year
+
+lhbiochar <- lahave_bio_adult %>%
+  select(-X1, -LOGID, -SEXID, -METHID, -CSITE, -GEARID, 
+         -ORIGINID, -OTYPEID, -SCALE, -TAG, -EST, -BASIN) %>%
+  mutate(RS = grepl("[Xx]", paste0(SP1, SP2, SP3, SP4, SP5))) %>%
+  select(-SALMON) %>%
+  filter(!is.na(POST)) %>% # removing those with no sea ages
+  mutate(LS = ifelse(RS == TRUE, "RS", paste0(POST, "SW"))) 
+
+# Test that only Xs are present
+lhbiochar %>% 
+  mutate(RS2 = paste0(SP1, SP2, SP3, SP4, SP5)) %>%
+  pull(RS2) %>% 
+  gsub("[NA]", "", .) %>% table
+
+lhbiochar %>% filter(LS == "NASW")
+
+# test that there are no false positives
+lhbiochar %>% filter(RS == FALSE & !is.na(SP1))
+lhbiochar %>% filter(RS == FALSE & !is.na(SP2))
+lhbiochar %>% filter(RS == FALSE & !is.na(SP3))
+lhbiochar %>% filter(RS == FALSE & !is.na(SP4))
+lhbiochar %>% filter(RS == FALSE & !is.na(SP5))
+lhbiochar %>% filter(RS == FALSE & !is.na(SP6))
+
+lhbiowide <- lhbiochar %>% 
+  select(-(SP1:SP6)) %>%
+  rename(year = RCYEAR) %>%
+  mutate(smla = ifelse(FORK >= 630, "large", "small")) %>%
+  group_by(year, smla, LS) %>%
+  summarise(number = n()) %>%
+  spread(LS, number) %>%
+  mutate(total = sum(`1SW`, `2SW`, `3SW`, `4SW`, `5SW`, RS, na.rm = TRUE)) %>%
[…]

### 2020-01-15 analysis/00-nashwaak-wrangling.R

@@ -6,7 +6,10 @@ library(tidyverse)
 load("~/Dropbox/salmon-inverse-matrix/data/nashwaak.rda")
 
 # We use the return data frame that includes wild salmon only, as it
+nashwaak_wild_returns
 nashwaak_returns_wild_hatch
+nashwaak_smolts
+# Smolt data from 1998 to 2016
 
 # Saving filtered data frame in new object
 nashwaakdat <- right_join(
@@ -14,19 +17,26 @@ nashwaakdat <- right_join(
   select(nashwaak_wild_returns, year, est_1SW, est_2SW, est_RS), 
   by = "year")
 
+
 # Missing data:
 #
 # Smolt data missing for 2017 and 2018 
 # so after lag in smolts just need to exclude year 2018
 
 nashwaakreturns <- nashwaakdat %>%
-  mutate(smoltlag = lag(smolt_mode),
+  mutate(smoltdiff =  ((smolt_mode-smolt_025)+(smolt_975-smolt_mode))/2,
+         smoltdiff_upper =  smolt_975-smolt_mode,
+         smoltdiff_lower =  smolt_mode-smolt_025,
+         smoltsd = smoltdiff/1.96,
+         smoltscv = smoltsd/smolt_mode,
+         smoltlag = lag(smolt_mode),
          smolt025lag = lag(smolt_025),
          smolt975lag = lag(smolt_975),
          est_2SWlead = lead(est_2SW),
          est_RSlead = lead(est_RS),
-         logsmoltsdlag = sqrt(300) * (log(smolt975lag) - log(smolt025lag))/3.92) %>% 
+         smoltscvlag = lag(smoltscv)) %>% 
   filter(year %in% 1999:2017)
+
 
 nashwaakreturns
 
@@ -36,15 +46,16 @@ ggplot(nashwaakreturns, aes(year, smolt_mode)) + geom_point() + geom_line() +
 # Creating list for use in TMB
 nashwaakdata <- list(years = nashwaakreturns$year,
   logsmolts = log(nashwaakreturns$smoltlag),
-  logsmolts_SE = nashwaakreturns$logsmoltsdlag,
-  #logsmolts_SE =  rep(0.05, length(nashwaakreturns$smoltlag)), # should be replaced with actual uncertainty
+  logsmolts_cv = nashwaakreturns$smoltscvlag,
   loggrilse = log(nashwaakreturns$est_1SW),
[…]

### 2020-01-15 analysis/00-saintjean-wrangling.R

@@ -1,12 +1,33 @@
 ### Saint-Jean Data
 
 library(tidyverse)
+library(readxl)
+
 
 saintjean <- readRDS(file = "~/Dropbox/salmon-bayesian-survival/qc-stjean-trinite-all.rds") %>%
   filter(river == "Saint-Jean", year >= 1981) %>%
   spread(stage, numbers) %>%
   select(year, smolts, returns_grilse, returns_2sw_est, returns_3sw_est, returns_rs_est)# %>%
   #filter(year %in% 1998:2014)
+
+# smolt data from 1989 onwards, missing 1997 and 2015 
+
+
+sjsmolts <- read_excel("raw-data/Estimation dévalaison St-Jean-Trinité 1984-2018.xlsx", sheet = "Saint-Jean") %>%
+  select("Year", "N Min (95%)":"Smolt migrate to sea") %>%
+  rename(year = Year,
+         n_min = `N Min (95%)`,
+         n = `N Estimation`,
+         n_max = `N max (95%)`,
+         smolt_prod = `Smolt produce`,
+         smolt_tosea = `Smolt migrate to sea`) %>%
+  mutate(n_sea_diff = n - smolt_tosea,
+         smolt_est = smolt_tosea,
+         smolt_min = n_min - n_sea_diff,
+         smolt_max = n_max - n_sea_diff)
+
+
+excel_sheets("raw-data/Estimation dévalaison St-Jean-Trinité 1984-2018.xlsx")
 
 
 saintjeanreturns <- saintjean %>%
@@ -18,33 +39,34 @@ saintjeanreturns <- saintjean %>%
          est_RSlead = lead(returns_rs_est),
          #logsmoltsdlag = sqrt(300) * (log(smolt975lag) - log(smolt025lag))/3.92
   ) %>% 
-  filter(year %in% 1990:2014)
-
+  filter(year %in% 1990:2015) %>% # updated from 2014 to 2015
+  filter(year != 1998) # removing smotl year 1997 as its missing smolt data
 
 # Removing zeros in 2SW-RS estimates as they cannot be log transformed
-saintjeanreturns$est_RSlead[saintjeanreturns$est_RSlead == 0] <- 0.0001
+filter(saintjeanreturns, est_RSlead != 0) %>% pull(est_RSlead) %>% min(.)/2 # picking min(RS)/2
+saintjeanreturns$est_RSlead[saintjeanreturns$est_RSlead == 0] <- 4.678571
+
[…]

### 2020-01-15 analysis/00-trinite-wrangling.R

@@ -26,15 +26,18 @@ trinitereturns <- trinite %>%
          est_2SWlead = lead(returns_2sw_est),
          est_RSlead = lead(returns_rs_est),
          #logsmoltsdlag = sqrt(300) * (log(smolt975lag) - log(smolt025lag))/3.92
-  ) %>% 
-filter(year %in% 1985:2016) # lead 2017 is missing 2SW and RS
+        ) %>% 
+  filter(year %in% 1985:2016) %>% # lead 2017 is missing 2SW and RS
+  filter(year != 2007) # Removing smolt year 2006 as its missing smolt estimates 
 
-# YEAR WITH NO SMOLT DATA
+# YEAR WITH NO SMOLT DATA (NOT NEEDED ANYMORE AS AM EXCLUDING THOSE YEARS)
 # trinitedata$smolts[trinitedata$year == 2007] <- 1
 # Can't use a dummy value since it will screw up with resulting 
 # distributions for the population
 # using average of contiguous 4 years (+-2) instead
-trinitereturns[is.na(trinitereturns$smoltlag), "smoltlag"] <-  mean(trinitereturns$smoltlag[trinitereturns$year %in% 2005:2009], na.rm = TRUE)
+# trinitereturns[is.na(trinitereturns$smoltlag), "smoltlag"] <-  mean(trinitereturns$smoltlag[trinitereturns$year %in% 2005:2009], na.rm = TRUE)
+
+
 
 
 ggplot(trinitereturns, aes(year, smolts)) + geom_point() + geom_line() +
@@ -45,15 +48,16 @@ ggplot(trinitereturns, aes(year, smolts)) + geom_point() + geom_line() +
 trinitedata <- list(years = trinitereturns$year,
                   logsmolts = log(trinitereturns$smoltlag),
                   #logsmolts_SE = nashwaakreturns$logsmoltsdlag,
-                  logsmolts_SE =  rep(1, length(trinitereturns$smoltlag)), # should be replaced with actual uncertainty
+                  logsmolts_cv =  rep(0.1, length(trinitereturns$smoltlag)), # should be replaced with actual uncertainty
                   loggrilse = log(trinitereturns$est_1SW),
                   logSW2 = log(trinitereturns$est_2SWlead),
                   logRS = log(trinitereturns$est_RSlead),
                   #N =  length(connereturns$small[-1]),
                   #Pr = rep(0.8, nrow(connereturns)))
-                  cv = 0.01, # ASSUMPTION, CV of return abundance counts
[…]

### 2020-01-15 analysis/00-wab-wrangling.R

@@ -1,0 +1,65 @@
+### WAB data
+
+library(tidyverse)
+
+wabsmolts <- read_csv("raw-data/NL-smolt-estimates-Kelly-et-al-2017-ResDoc.csv",
+         na = c("-", "NA")) %>%
+  rename(year = Year) %>%
+  gather(river_name, smolt, -year) %>%
+  filter(river_name == "Western Arm Brook")
+
+# Loading estimated returns of 1SW, 2Sw, and RS
+wab <- readRDS(file = "~/Dropbox/salmon-lh-variation/output/NL/wab-estimated-returns.rds")
+
+waball <- right_join(wab, wabsmolts, by = "year") %>%
+  select(year, smolt, estM1, estM2p, estRS, totalreturns)
+
+
+wabreturns <- waball %>%
+  mutate(smoltlag = lag(smolt),
+         #smolt025lag = lag(smolt_025),
+         #smolt975lag = lag(smolt_975),
+         est_1SW = estM1,
+         est_2SWlead = lead(estM2p),
+         est_RSlead = lead(estRS)
+         #logsmoltsdlag = sqrt(300) * (log(smolt975lag) - log(smolt025lag))/3.92
+  ) %>%
+  filter(year %in% 1974:2015)
+
+wabreturns %>%
+  filter(est_2SWlead != 0) %>%
+  summarise(min2SW = min(est_2SWlead))
+
+ggplot(wabreturns, aes(year, smolt)) + geom_point() + geom_line() # +
+
+wabreturns[wabreturns$est_2SWlead == 0, "est_2SWlead"] <- 0.1 # roughly min(est_2SW)/2
+
+wabreturn_rate <- wabreturns %>%
+  mutate(return_rate = est_1SW/smoltlag) %>%
+  select(year, return_rate) %>%
+  mutate(river_name = "Western Arm Brook")
+
+ggplot(wabreturn_rate, aes(year, return_rate)) + 
+  geom_point() + geom_line() + ggtitle("WAB return rates")
+
+
+saveRDS(wabreturn_rate, file = "data/wabreturnrate.rds")
+
+# Creating list for use in TMB
+wabdata <- list(years = wabreturns$year,
+                  logsmolts = log(wabreturns$smoltlag),
+                  #logsmolts_SE = nashwaakreturns$logsmoltsdlag,
+                  logsmolts_cv =  rep(0.05, length(wabreturns$smoltlag)), # should be replaced with actual uncertainty, if available
+                  loggrilse = log(wabreturns$est_1SW),
[…]

### 2020-01-15 analysis/01-merge-populations.R

@@ -1,4 +1,5 @@
-# Merge data from all populations into structure to input to TMB hierachical model
+# Merge data from all populations into structure to input to TMB hierachical models
+# as well as create object for Stan models
 
 library(tidyverse)
 library(rlist)
@@ -11,17 +12,76 @@ saintjeandata <- readRDS(file = "data/saintjeandata.rds")
 trinitedata <- readRDS(file = "data/trinitedata.rds")
 connedata <- readRDS(file = "data/connedata.rds")
 nashwaakdata <- readRDS(file = "data/nashwaakdata.rds")
+wabdata <- readRDS(file = "data/wabdata.rds")
+campdata <- readRDS(file = "data/campbelltondata.rds")
+lahavedata <- readRDS(file = "data/lahavedata.rds")
 
 # Converting lists for separate population into single list with  matrices for each data type 
 # where each population is a separate row
 # based on answer in 
 # https://stackoverflow.com/questions/57401290/bind-vectors-across-lists-to-single-list-of-matrices
 
-alldata <- list(connedata, saintjeandata, trinitedata, nashwaakdata) %>% 
+alldata <- list(saintjeandata, trinitedata, connedata, nashwaakdata, wabdata, campdata, lahavedata) %>% 
   purrr::transpose() %>%
   map(~ reduce(.x, cbind.fill, fill = NA) %>% 
         t %>% 
-        `row.names<-`(NULL))
-
+        `row.names<-`(NULL)) %>%
+  list.remove("allreturns")
 
 saveRDS(alldata, file = "data/alldata.rds")
+
+# Creating list for Stan models
+
+# listnames <- names(alldata)
+
+alldatastan <- alldata %>%
+  lapply(t) %>%
+  lapply(as.vector) %>%
+  lapply(na.omit) %>%
+  lapply(as.vector) # To remove na.omit class and subsequent join warnings
+
+# names(alldatastan) <- listnames
+
+# Saving filtered data frame in new object (connereturns)
+alldatastan$N <- length(alldatastan$nyears)
+alldatastan$K <- sum(alldatastan$nyears)
+
+# re-adding matrix of years
+alldatastan$yearmat <- alldata$year
+
+grilse_rivers <- c("Conne River", "Western Arm Brook", "Campbellton River")
+
[…]

### 2020-01-15 analysis/03-stan-run-models.R

@@ -1,0 +1,38 @@
+### Running Stan models, both hierarchical and not (simple)
+### Both of which have non-centered prameterizations to improve
+### sampling independence
+
+library(tidyverse)
+library(rstan)
+library(bayesplot)
+
+options(mc.cores = parallel::detectCores())
+rstan_options(auto_write = TRUE)
+
+# Loading data
+alldatastan <- readRDS("data/alldatastan.rds")
+
+# Loading and running hierarchical model with non-centered parameterization
+modhierncp <- stan_model("analysis/survival_hier_ncp.stan")
+fithierncp <- sampling(object = modhierncp, data = alldatastan, iter = 5000, chains = 3,
+                       control = list(adapt_delta = 0.95, max_treedepth = 15))
+saveRDS(fithierncp, "data/stanfit_hier_ncp.rds")
+# fithierncp <- readRDS("data/stanfit_hier_ncp.rds")
+
+# Investigating model object
+print(fithierncp, pars = "S1")
+print(fithierncp, pars = paste0("S1[",101:186,"]"))
+
+mcmc_intervals(fithierncp, pars = paste0("Pr_mu[",1:7,"]")) + 
+  scale_y_discrete(labels = alldatastan$river_name)
+traceplot(fithierncp, pars = paste0("S1[",170:186,"]"), inc_warmup = TRUE)
+
+
+# Fitting non-hierarchical model (still with ncp)
+modsimplencp <- stan_model("analysis/survival_simple_ncp.stan")
+fitsimplencp <- sampling(object = modsimplencp, data = alldatastan, iter = 1000, chains = 3, 
+                         control = list(adapt_delta = 0.95, max_treedepth = 15))
+saveRDS(fitsimplencp, "data/stanfit_simple_ncp.rds")
+#fitopt <- readRDS("data/stanfit_simple_ncp.rds")
+
+############## RUN FILE SHOULD END HERE

### 2020-01-15 analysis/05-convergence-tests.R

@@ -1,0 +1,29 @@
+library(tidyverse)
+library(rstan)
+library(bayesplot)
+
+# loading data
+alldatastan <- readRDS("data/alldatastan.rds")
+
+# loading stanfit objects
+fithierncp <- readRDS("data/stanfit_hier_ncp.rds") # hierarchical
+#fitsimplencp <- readRDS("data/stanfit_simple_ncp.rds")   # non-hierarchical
+
+summary(fithierncp)$summary[,"Rhat"]
+range(summary(fithierncp)$summary[,"Rhat"])
+
+neffs <- summary(fithierncp)$summary[,"n_eff"]
+neffs[neffs < 1000]
+
+range(summary(fithierncp)$summary[,"n_eff"])
+
+print(fithierncp, pars = "Pr_mu")
+print(fithierncp, pars = "logitPr_sigma")
+
+print(fithierncp, pars = "Z1")
+
+ratios_ncp <- neff_ratio(fithierncp)
+mcmc_neff(ratios_ncp, size = 2)
+
+500/7500
+7

### 2020-01-15 analysis/05-correlation-tests.R

@@ -1,0 +1,161 @@
+# Exploring correlations among trends of S1, smolt, and return abundances
+
+library(tidyverse)
+library(corrplot)
+source("99-helper-functions.R")
+
+alldatastan <- readRDS("data/alldatastan.rds")
+s1quant <- readRDS("data/s1quant.rds")
+allreturns <- readRDS("data/allreturns.rds") %>% 
+  group_by(river_name) %>%
+  mutate(zscoresmolts = scale(log(smolts)), zscoretotal = scale(log(totalreturns))) %>%
+  ungroup()
+  
+
+
+river_names <- s1quant %>%
+  ungroup() %>%
+  select(year, zscore, river_name) %>%
+  spread(year, zscore)  %>%
+  pull(river_name)
+
+s1mat <- s1quant %>%
+  #filter(year <= 2013) %>%
+  ungroup() %>%
+  select(year, zscore, river_name) %>%
+  spread(year, zscore) %>%
+  select(-river_name) %>%
+  t %>% as.matrix
+
+colnames(s1mat) <- c(word(river_names)[1:6], "WAB")
+
+M <- cor(s1mat, use = "complete.obs", method = c("pearson"))
+M
+
+# matrix of the p-value of the correlation
+# Pearson's product moment correlation coefficient
+p.mat <- cor.mtest(s1mat)
+
+col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
+
+oldpar <- par() 
+
+png("figures/corr-s1.png", width = 7.5, height = 3, units = "in", res = 300, pointsize = 9.5)
+par(mfrow = c(1,2), mar = c(0,0,0,0), oma = c(0,0,0,0))
+corrplot(M, method = "color", col = col(200),  
+         type = "upper", order = "hclust",  diag = FALSE, 
+          number.cex = 0.8,
+         addCoef.col = "black", # Add coefficient of correlation
+         tl.col = "black", tl.srt = 45, #Text label color and rotation
+         # Combine with significance
+         p.mat = p.mat, sig.level = 0.01, insig = "blank", # only coloured boxes have significant pairwise p-values.
+         # hide correlation coefficient on the principal diagonal
+         # diag=FALSE,
+       #  title = expression(paste("Correlation of standardized Z-scores in ", S[1])),
+         mar = c(0, 0, 0, 0)
+)
+
+mtext("(a)", line = -2, adj =0.1, cex = 1.6)
+
+corrplot(M, method = "color", col = col(200),  
[…]

### 2020-01-15 analysis/05-figures-for-trends-paper.R

@@ -1,0 +1,124 @@
+# plots for trends paper
+
+library(tidyverse)
+library(cowplot)
+
+s1post <- readRDS(file = "data/s1-posteriors.rds")
+s1quant <- readRDS(file = "data/s1quant.rds")
+
+# Add rows with NA values as to stop geom_line to connect years
+# with data gaps
+addnas <- tibble(pos = NA,
+                 river_name = c("Saint-Jean River", "Trinité River", rep("LaHave River", 2)),
+                 year = c(1998, 2007, 2012:2013),
+                 median = c(NA),
+                 q05 = c(NA),
+                 q25 = c(NA),
+                 q75 = c(NA),
+                 q95 = c(NA),
+                 zscore = NA
+)
+
+s1quant <- bind_rows(s1quant, addnas)
+
+ggplot(s1quant, aes(year, median)) + 
+  geom_line(alpha = 0.5) + geom_point(size = 1, alpha = 0.7) + 
+  ylab(expression(paste(S[1]~"estimates", sep = ""))) +
+  geom_errorbar(aes(ymin = q05, ymax = q95),width= 0.1) +
+  facet_wrap(~river_name, ncol = 2, scales = "free_y") + theme_bw() +
+  xlab("Year") +
+  labs(color = "River") +
+  theme(legend.position = "none", 
+        legend.text =  element_text(size = 12))
+ggsave("figures/s1-trends-faceted.png", height = 6, width = 5)
+
+s1trends <- ggplot(s1quant, aes(year, median, color = river_name)) + 
+  geom_line(alpha = 0.5) + geom_point(size = 2, alpha = 0.8) +
+  ylab(expression(paste(S[1]~"posterior estimates", sep = ""))) +
+  geom_errorbar(aes(ymin = q05, ymax = q95), alpha = 0.5, width = 0.1) +
+  theme_cowplot() + xlab("Year") + labs(color = "River") +
+  theme(legend.position = "bottom") + guides(color = guide_legend(nrow = 3))
+s1trends
+ggsave("figures/s1-trends.png", height = 5, width = 7)
+
+s1zscore <- ggplot(s1quant, aes(year, zscore, color = river_name)) + 
+  geom_line(alpha = 1) + #geom_point(size = 2) + 
+  ylab(expression("Z-scores of median"~paste(S[1]~"estimates", sep = ""))) +
[…]

### 2020-01-15 analysis/lc_model_hierarchical.cpp

@@ -73,8 +73,8 @@ Type objective_function<Type>::operator() () {
       nll -= dnorm(logsmolts_true(j,i), logsmolts_mean(j), exp(logsmolts_logsd(j)), true); // distribution of x_true
       nll -= dnorm(logsmolts(j,i), logsmolts_true(j,i), logsmolts_SE(j,i), true); // oibservation error in x (dist of error around x_true)
       
-      nll -= dnorm(loggrilse(j,i), mu(j,i), cv(j), true); // observation error in grilse (based on data cv)
-      nll -= dnorm(logSW2(j,i), mu2(j,i), cv(j), true); // observation error in 2SW (based on data cv)
+      nll -= dnorm(loggrilse(j,i), mu(j,i), mu(j,i) * cv(j), true); // observation error in grilse (based on data cv)
+      nll -= dnorm(logSW2(j,i), mu2(j,i), mu2(j,i) * cv(j), true); // observation error in 2SW (based on data cv)
       
     }
     // NLL DECLARATION FOR GLOBAL PARAMETER ESTIMATES

### 2020-01-15 ms/ms.tex
Incorporate Jeff's edits

@@
 least known yet one of the most crucial with regards to population
-persistence.
-Recently, declines in many Atlantic salmon populations in eastern Canada have
-been attributed to changes in the conditions at sea, thus affecting their
-survival.
-However, marine survival estimates are difficult to obtain given that many
-individuals spend multiple winters in the ocean before returning to freshwater to
-spawn; therefore multiple parameters need to be estimated.
-We used smolt and adult abundance time series for seven populations, ranging between 19 and 42 years,
-to fit a hierarchical Bayesian maturity schedule model and estimate three components of
-marine survival: survival in the first year at sea (\So), survival in the second year
-at sea (\St), and proportion returning as one sea-winter (\Pg).
-Trends in estimates of \So were not consistent among rivers, with some population showing increases in \So over
-time, other showing decreases, and some being relatively stable.
-Estimates of \St were highly uncertain and therefore trends could not be assessed, while 
-estimates of \Pg were mostly stable across time, except for the LaHave and Nashwaak Rivers.
-Our findings challenge the assumption that marine survival has decreased for
-all salmon populations, and suggest that trends can be river-specific. 
-Further work exploring potential correlates of these trends in marine survival
-is warranted.
+persistence. Recently, declines in many Atlantic salmon populations in eastern
+Canada have been attributed to changes in conditions at sea, negatively
+affecting their survival. However, marine survival estimates are difficult to
+obtain given that many individuals spend multiple winters in the ocean before
+returning to freshwater to spawn, necessitating the estimation of multiple
+parameters. To do so, we fit a hierarchical Bayesian maturity schedule model
+to smolt and adult abundance time series for seven populations located in
+Newfoundland and Labrador (NL), Qu\'{e}bec (QC), New Brunswick (NB), and Nova
+Scotia (NS). The datasets ranged between 19 and 42 years in length. We
+estimated three components of marine survival: survival in the first year at
+sea (\So), survival in the second year at sea (\St), and proportion returning
+as one sea-winter adults (\Pg). Controlling for time frame, trends in
+estimates of \So were not consistent among rivers.
+In the four populations for which data extended to 1990, marine survival
+during the first year at sea predominantly increased over time in Western Arm
+Brook (NL) and Rivi\`{e}re Saint-Jean (QC) but largely declined in Conne River
+(NL) and Rivi\`{e}re Trinit\'{e} (QC). In the three other populations, \So
+exhibited an increase in Campbellton River (NL), a decline in LaHave River
+(NS), and fluctuating stability in Nashwaak River (NB) since approximately the
+mid-1990s. 
+Estimates of \St were highly uncertain and trends could not be assessed. \Pg
+was temporally stable, except for LaHave and Nashwaak Rivers. Our findings
+challenge the assumption that marine survival has changed in a consistent
+manner among populations across a broad geographic scale, and suggest that
+trends can be river-specific. Further work exploring potential correlates of
+these trends in marine survival is warranted.
 
@@
 %1. Declining salmon numbers
-Across eastern Canada, many wild populations of Atlantic salmon (\emph{Salmo salar})
-have declined to the point of being considered Threatened by the 
-Committee on the Status of Endangered Wildlife in Canada \citep[COSEWIC, ][]{Cosewic2010}.
-While it is not fully understood what is driving these declines, the potential
+
+Reductions in fishing mortality, albeit necessary, are not always sufficient
+to facilitate population recovery. Experience with numerous commercially
+exploited marine fisheries since the early 1990s has shown that not all
+populations respond as favourably as anticipated to considerable reductions in
+exploitation \citep{Hutchings2017}. Efforts to close commercial
+Atlantic salmon (\emph{Salmo salar}) fisheries in eastern Canada culminated in
+moratoria in all regions, beginning in the Maritime provinces (1984) and
+following in Newfoundland (1992), Labrador (1998), and Québec (2000). Since
+these closures, many populations have not increased as expected (REF): some 
+have been assessed as considered threatened or endangered by the 
+Committee on the Status of Endangered Wildlife in Canada \citep[COSEWIC, ][]{Cosewic2010}, 
+Canada’s national science advisory body (to the national government) on species risk of extinction.
+While it is not fully understood what is driving population declines, the potential
 drivers of these are many \citep[see ][for a detailed discussion of possible
@@
 % 2.
-Theres a common perception that marine survival of Atlantic salmon has
-declined throughout the North Atlantic \citep{OMaoileidigh2003, Hansen1998,
-    Chaput2012a}. This perception is widespread both in the scientific
-literature as well as federal reports. The latest status report on Atlantic
-salmon by COSEWIC states that ``While the mechanism(s) of marine mortality is
-uncertain, what is clear is that the recent period of poor sea survival is
-occurring in parallel with many widespread changes in the North Atlantic
-ecosystem.'' \citep{Cosewic2010}. Consequently, there have several lines of
-inquiry as to what the causes behind these declines might be attributable to
-\citep{Friedland1993, Friedland1998}.
+Over the past three decades, a narrative has emerged that marine survival of Atlantic salmon has
+declined throughout the North Atlantic \citep{Hansen1998,OMaoileidigh2003,Chaput2012a}. 
+An implicit assumption, albeit infrequently acknowledged, is that any trend in
+survival in ocean habitat that is shared by multiple populations during their
+seaward migration period will be experienced similarly by multiple
+populations. Put another way, given that salmon from different river systems
+are hypothesized to share marine habitat during some of their time at sea, it
+has been implicitly assumed that populations share similar temporal trends in
+at-sea mortality \citep{Friedland1993, Friedland1998}.  
+A recent study by \citet{Olmos2019} suggested that trends in post-smolt
+survival, when estimated at the stock unit level, are synchronously declining
+for all Atlantic salmon in eastern North America.
 
-Nonetheless, most of the populations declines assessed by COSEWIC have
-occurred in populations in the southern extent of its distribution:
-populations in the Bay of Fundy, Anticosti Island, and the Atlantic coast of
-Nova Scotia being assessed as Endangered, populations in the south coast of
-Newfoundland assessed as Threatened, And the populations in the New Brunswick
-and Qu\`{e}bec coasts of the Gulf of St. Lawrence assessed as Special Concern
-\citep{Cosewic2010}. On the other hand, northermost populations have shown
-stable, or even increasing population trends, suggesting that, if marine
-survival were to be a factor in many of these declines, these changes are not
-uniformly distributed across all populations.
+% This perception is widespread both in the scientific
+% literature as well as federal reports. The latest status report on Atlantic
+% salmon by COSEWIC states that ``While the mechanism(s) of marine mortality is
+% uncertain, what is clear is that the recent period of poor sea survival is
+% occurring in parallel with many widespread changes in the North Atlantic
+% ecosystem.'' \citep{Cosewic2010}. Consequently, there have several lines of
+% inquiry as to what the causes behind these declines might be attributable to
+% \citep{Friedland1993, Friedland1998}.
 
-% 3. Nonetheless, this premise has never been examined in detailed.
-While the purported decline of marine survival is mentioned widely in the 
-scientific literature, only relatively few studies have quantified it in detail.
-The basis for this premise is a limited number of time series that exhibit
-temporal declines in a proxy of marine survival (i.e. return rates or
-post-smolt survival).
-\citet{Chaput2012a} looked at return rates of smolts to adult salmon to assess
+In contrast to the narrative of widespread, demographically similar increases
+in at-sea mortality, the conservation status of Canadian populations differs
+considerably. Populations in the southern part of the Canadian range are more
+likely to be assessed as being below conservation targets than those in more
+northerly regions \citep{Cosewic2010}. This geographical disparity in status
+suggests that if marine survival has been, or is, a key factor responsible for
+most population declines, these changes are not uniformly distributed across
+all populations.
+
+% Nonetheless, most of the populations declines assessed by COSEWIC have
+% occurred in populations in the southern extent of its distribution:
+% populations in the Bay of Fundy, Anticosti Island, and the Atlantic coast of
+% Nova Scotia being assessed as Endangered, populations in the south coast of
+% Newfoundland assessed as Threatened, And the populations in the New Brunswick
+% and Qu\`{e}bec coasts of the Gulf of St. Lawrence assessed as Special Concern
+% \citep{Cosewic2010}. On the other hand, northermost populations have shown
+% stable, or even increasing population trends, suggesting that, if marine
+% survival were to be a factor in many of these declines, these changes are not
+% uniformly distributed across all populations.
+
+% While the purported decline of marine survival is mentioned widely in the 
+% scientific literature, only relatively few studies have quantified it in detail.
+% The basis for this premise is a limited number of time series that exhibit
+% temporal declines in a proxy of marine survival (i.e. return rates or
+% post-smolt survival).
+
+% 3. Nonetheless, this premise has never been examined in detail
+Given the logistical challenges associated with estimating at-sea survival, it
+is not surprising that the number of studies that have estimated temporal
+trends has been limited. An additional limitation has been the derivation of
+proxies (e.g., return rates), rather than direct model-based estimates, of
+marine survival.
+\citet{Chaput2012a}, for example, examined return rates of smolts to adult salmon to assess
 declines in marine survival and found that most populations from Canadian
@@
 can mask changes in differential survival in different years at sea, as well
-as changes in proportion of adults returning after one or two years at sea.
+as changes in the proportion of adults returning after one or two years at sea.
 A more recent study by \citet{Olmos2019} suggested that trends in post-smolt
@@
 for all Atlantic salmon in Eastern North America and Canada. Nonetheless,
-their analysis relies on stock-recruitment relationships rather than smolt
-abundance indices.
+the aforementioned study by \citet{Olmos2019} relied on the veracity of stock-recruitment relationships.
 
 %4. 
-In this study, we compile smolt and returns times series data from seven wild
-populations of Atlantic salmon in eastern Canada to model trends in marine survival.
-While previous studies have used maturity schedule models to estimate marine
-survival for a couple of Atlantic salmon populations \citep{Chaput2003b}, none
-have incorporated data from the last decades and they have not examined
-trends among more than a couple of populations.
+In the present study, we compile smolt and adult returns times series data for seven wild
+Canadian populations of Atlantic salmon to model trends in marine survival.
+While some studies have used maturity-schedule models to estimate marine
+survival for a limited number of salmon populations \citep{Chaput2003b}, none
+have incorporated data extending over decates, nor have they examined
+trends among more than two or three populations. 
+Here, we develop a hierarchical Bayesian model that uses Murphy's maturity
+schedule method, in conjunction with informative priors, to estimate yearly
+marine survival in salmon. In addition to accounting for observation error in
+smolt and return estimates, we estimate the proportion of salmon returning
+after one winter hierarchically.
 
@@
 We obtained time series data of outmigrating smolt and returning adult
-abundances for seven Atlantic salmon populations in eastern Canada, covering a
+abundances for seven Atlantic salmon populations in eastern Canada, encompassing a
 wide range of the species' distribution (Fig.~\ref{fig:map}). These
-populations included the LaHave River in the Southern Uplands region of Nova
-Scotia (NS), The Nashwaak River in New Brunswick (NS) which connects to the
-Bay of Fundy, The Trinit\'{e} and Saint-Jean Rivers in Qu\`{e}bec (QC), which
-flow to the St Lawrence River and Bay of St Lawrence, respectively, and the
-Western Arm Brook, Campbellton, and Conne in Newfoundland, which flow to the
-Bay of St Lawrence, Labrador Sea, and Atlantic Ocean, respectively.
+populations included LaHave River in the Southern Uplands region of Nova
+Scotia (NS), Nashwaak River in New Brunswick (NS) which connects to the
+Bay of Fundy, Trinit\'{e} and Saint-Jean Rivers in Qu\'{e}bec (QC), which
+flow to the St. Lawrence River and Bay of St. Lawrence, respectively, and
+Western Arm Brook (also referred to as WAB), Campbellton, and Conne Rivers in Newfoundland, which flow to the
+Strait of Belle Isle, Labrador Sea, and Gulf of St. Lawrence, respectively.
 Data were collected in NS, NB, and Newfoundland
 and Labrador (NL) by Fisheries and Oceans Canada (DFO) and in QC
-by the Minist\`{e}re des For\^{e}ts, de la Faune et des Parcs, Qu\`{e}bec.
+by the Minist\`{e}re des For\^{e}ts, de la Faune et des Parcs, Qu\'{e}bec.
 
@@
 % 1. Main findings
-Our findings challenge the widespread idea that marine survival is declining
-uniformly across the range of Atlantic salmon in the northwest Atlantic. We
-find that there is no apparent shared trends among the seven salmon
-populations examined; some rivers show positive trends in survival in the
+Our findings challenge the narrative that marine survival is declining
+uniformly throughout the range of Atlantic salmon in the northwest Atlantic.
+Temporal trends are not consistent among populations 
+. Over the time periods for which data were available, some rivers show positive trends in survival in the
 first winter at sea (\So) while other show stable trends, and some even show
 declines. We could not assess trends in the second winter at sea (\So) or
-proportion returning as grilse (\Pg) as these parameter estimates were highly
+proportion returning as grilse (\Pg), as these parameter estimates were highly
 uncertain and were strongly influenced by the priors used.
 Perhaps there is a need to rethink our understanding of Atlantic salmon
-population dynamics in light that any real decline in marine survival was
-experienced by some populations, that reductions in survival might
-have occurred over a brief period of time, and that since that
-reduction occurred (for a subset of populations) marine survival has actually
-been stable, or even increasing, since that point.
+population dynamics in light of the possibility that any real and persistent decline in marine survival was
+experienced by some but not all populations, that reductions in survival might
+have occurred over a relatively brief period of time and may not longer be ongoing, and that since that
+reduction occurred (for a subset of populations) marine survival has
+been stable, or even increasing.
 
@@
 Our findings are contrary to those of \citet{Olmos2019}, who detected positive
-correlations in post-smolt survival among broad stock units. These differences
+correlations in post-smolt survival among spatially broad stock units. These differences
 could be due to a number of reasons: different model specifications and
@@
 Interestingly, our estimates of \So and \St are very similar to those produced
-by \citet{Chaput2003b}, and the trends are almost identical for the
-overlapping time period that marine survival was estimated for in that study
-(1984-1998). While that study separated abundance data for males and females
+by \citet{Chaput2003b}, and our trends are almost identical for the
+overlapping time period that marine survival was estimated for in their study
+(1984-1998). 
+While \citet{Chaput2003b} separated abundance data for males and females
 and assumed their survival rates were the same (to be able to reach an
@@
 slightly higher uncertainty) using a Bayesian approach with informative
-priors. These overlapping trends obtained with two different methods are
-indicative that our method is effective at estimating marine survival.
-While \citet{Chaput2003b} did not compare trends in marine survival
-among populations, \citet{Chaput2012a} did, using adult return rates, and
-found that for four out of the six population examined, return rates in the 1990s 
+priors. These overlapping trends obtained with two different methods 
+suggest that our method is effective at estimating marine survival.
+
+Trends in marine survival
+among populations were compared by \citet{Chaput2012a} using adult return rates, 
+and found that for four out of the six population examined, return rates in the 1990s 
 were less than those during the 1970s.
 \citet{Friedland1993} compared return rates for a number rivers in eastern
-North America and suggested there are similar trends among these. The
-similarity in trends is driven primarily by two years, 1977 and 1978, which
+North America between 1973 and 1988, and suggested there are similar trends among these. 
+However, the similarity in trends is driven primarily by two years, 1977 and 1978, which
 show concurrent low and high relative return rates across rivers,
 respectively. Others years are much more variable relative to each other.
-Their time series stops at 1988, thus there are only a few years for which 
-to assess overlap with the time series in our study.
-However, using pooled adult return rates, as done by \citet{Chaput2012a} and
+\citeauthor{Friedland1993}'s \citeyear{Friedland1993} time series stops at
+1988, thus there are only a few years for which to assess overlap with the
+time series in our study.
+However, the pooling of adult return rates, as done by \citet{Chaput2012a} and
 \citet{Friedland1993} can mask inter-annual variability in marine survival,
@@
 \citet{Dempson2003} describe a general declining trend in marine survival for
-Newfoundland rivers (except the WAB); which in our study we observe in the
-Conne but not the Campbellton or the WAB. It is not possible to draw conclusions
+Newfoundland rivers (except WAB); which in our study we observe in the
+Conne but not in Campbellton River or WAB. It is not possible to draw conclusions
 with data from only three Newfoundland rivers, but it seems that among index rivers,
@@
 % 3. Potential reasons
-There are multiple potential explanations as to why there are no synchronous
+There are multiple potential explanations for the lack of synchronous
 trends in estimates of \So. 
 Marine survival in the first winter at sea could be highly variable between
-rivers from environmental factors (e.g., temperature, predation) being
-variable at localized spatial scales rather than ocean-wide ones.
-Given that synchrony is being observed for marine survival trends estimated using
-stock-recruitment relationships and broader spatial scales \citep{Olmos2019},
-these overlapping declining trends in marine
-survival estimated using stock-recruitment relationships have been confounded
-by changes in recruitment dynamics.
+populations because of the predominance of spatially local environmental drivers of survival (e.g., temperature, predation) 
+relative to broader-scale, even ocean-wide, drivers.
+The synchrony reported for marine survival trends at broader spatial scales \citep{Olmos2019}
+might be attributable to the use of stock-recruitment relationships to estimate survival,
+relationships thay may have been confounded by changes in recruitment dynamics.
+
 There is some evidence of a correlation between return rate and growth (as
-indicated by intercirculi spacing in scales), where years of poor growth
+indicated by intercirculi spacing on scales), where years of poor growth
 tended to also be years of poor survival \citep{Friedland1993}, supporting the
@@
 Furthermore, among European salmon, there is evidence of a positive correlation
-between spring temperature and population abundance, suggesting warmer
+between spring temperature in the Norwegian and North Seas and population abundance, suggesting warmer
 conditions favour post-smolts \citep{Friedland1998} based on mapping the
-extent of area of suitable temperature (7-13 \textdegree C) in the North Atlantic.
-Nonetheless, how warming impacts survival of post-smolts can differ based
-whether it results in temperatures coming into or out of their suitable range;
-which could be part of why populations in eastern North America are declining
-in the southern part of their range while potentially increasing in the
-northern part. The relationship between temperature and survival warrants 
+extent of area of suitable temperature (7-13 \textdegree C).
+
+Nonetheless, the causal mechanisms for why warming should affect post-smolt
+survival almost certainly differs depending on the difference between
+temperature experienced by the post-smolts and their respective
+population-specific thermal optima, which could explain why populations in eastern North America are declining
+in the southern part of their range but potentially increasing further north.
+
+A potential association between temperature and survival warrants
 further study.
-While there is very little evidence that marine survival is density-dependent
-in Atlantic salmon, and it is broadly considered to be a density-independent
-process \citep{Jonsson1998,Gibson2006}, there could potentially be some
-density-dependent relationship, particularly for populations that are severely
-depleted (i.e. Allee effects). Exploring relationships between survival
-and population size could potentially shed light about the processes that
-have caused many of the population declines we have observed.
+
+While there is little evidence that marine survival is density-dependent in
+Atlantic salmon \citep{Jonsson1998,Gibson2006}, there could potentially be
+some density-dependent processes during parts of the post-smolt migration
+period, particularly for populations that are likely to be subjected to
+declining per capita population growth rates generated by Allee effects.
+Exploring relationships between survival and population size could potentially
+shed light about the processes that have caused many of the population
+declines that have been documented.
 
 % 4. Caveats
-Nonetheless, there are caveats to consider in our modelling approach.
+As with all novel modelling approaches, there are caveats to acknowledge.
 While there are analytical issues with estimation of \So, \St, and \Pg,
-The assumption that \St is additive to \So could produce unrealistic results.
+the assumption that \St is additive to \So could produce unrealistic results.
 Overcoming this assumption would require an additional parameter to be
@@
 environment when they return to their natal streams as those fish who stayed
-at sea for an additional winter).
-Furthermore, the assumed hierarchical structure might not be the most adequate
+at sea for an additional winter before returning to spawn).
+Secondly, the assumed hierarchical structure might not be the most appropriate
 for modelling smolt abundances. This approach results in shrinkage to the
 mean, which means that the variability of yearly smolt estimates is less than
-it would be if not modelled hierarchically. Nonetheless, this assumption
-likely results in more conservative trends in marine survival, as the
-variation of smolt estimates is shrunk.
+it would be if not modelled hierarchically. 
+That said, this assumption is likely to result in more conservative trends in marine survival, as the
+variation is smolt estimates is reduced.
 It is important to note that the hierarchical structure in the estimation of
@@
 % 5. Future directions
-Perhaps reframing the issue of marine survival is key to further our
-understanding of Atlantic salmon population dynamics: it is not that marine
-survival has declined across rivers, but the fact that it has stayed at
-roughly similar levels as previously \emph{despite} their reduced abundance
-what we should focus into understanding better.
+Perhaps a reframing of the issue of marine survival is key to furthering our
+understanding of Atlantic salmon population dynamics. Marine
+survival may not have declined consistently, and over the same time periods, across across populations, but the fact that it has stayed at
+roughly similar levels as it was previously \emph{despite} reduced commercial fishing mortalities,
+and that many populations are smaller in size compared to what they once were, is
+what we should focus on understanding better.
 Given the literature of correlates on marine survival in the past, if survival

### 2020-01-17 analysis/05-river-location-map.R
Add R code for map figure

@@ -1,0 +1,107 @@
+library(tidyverse)
+library(ggmap)
+library(mapdata)
+library(maps)
+library(rgdal)
+library(sf)
+library(rgeos)
+library(ggsn)
+library(ggrepel)
+
+# from https://timogrossenbacher.ch/2016/12/beautiful-thematic-maps-with-ggplot2-only/
+theme_map <- function(...) {
+  theme_minimal() +
+    theme(
+      text = element_text(family = "Ubuntu Regular", color = "#22211d"),
+      #axis.line = element_blank(),
+      #axis.text.x = element_blank(),
+      #axis.text.y = element_blank(),
+      #axis.ticks = element_blank(),
+      # panel.grid.minor = element_line(color = "#ebebe5", size = 0.2),
+      panel.grid.major = element_line(color = "#ebebe5", size = 0.2),
+      panel.grid.minor = element_blank(),
+      # plot.background = element_rect(fill = "#f5f5f2", color = NA), 
+      panel.background = element_rect(fill = "#f5f5f2", color = NA), 
+      legend.background = element_rect(fill = "#f5f5f2", color = NA),
+      panel.border = element_blank(),
+      ...
+    )
+}
+
+# Based on https://semba-blog.netlify.com/10/20/2018/genetic-connectivity-in-western-indian-ocean-region/
+theme_map2 <- function(...) {
+  theme_bw() + 
+    theme(panel.background = element_rect(fill = "lightblue"),
+          axis.text = element_text(size = 11, colour = 1),
+          panel.grid = element_line(colour = NA),
+          axis.title.x = element_blank(),
+          axis.title.y = element_blank(),
+          ...
+    )
+}
+
+if (!file.exists("./data/mapdata/ne_50m_admin_1_states_provinces_lakes.dbf")){
+  download.file(file.path('http://www.naturalearthdata.com/http/',
+                          'www.naturalearthdata.com/download/50m/cultural',
+                          'ne_50m_admin_1_states_provinces_lakes.zip'), 
+                f <- tempfile())
+  unzip(f, exdir = "./data/mapdata")
+  rm(f)
+}
+
+region <- readOGR("./data/mapdata", 'ne_50m_admin_1_states_provinces_lakes', encoding='UTF-8')
+
+
[…]

### 2020-01-17 analysis/05-figures-for-trends-paper.R
Add moratoria lines to trends figure

@@ -32,12 +32,18 @@ ggplot(s1quant, aes(year, median)) +
         legend.text =  element_text(size = 12))
 ggsave("figures/s1-trends-faceted.png", height = 6, width = 5)
 
+morat <- tibble(year = c(1984, 1992, 2000),
+                region = c("NS+\nNB", "NL", "QC"),
+                river_name = c("Nashwaak River", "Western Arm Brook", "Trinité River"))
+
 s1trends <- ggplot(s1quant, aes(year, median, color = river_name)) + 
   geom_line(alpha = 0.5) + geom_point(size = 2, alpha = 0.8) +
   ylab(expression(paste(S[1]~"posterior estimates", sep = ""))) +
   geom_errorbar(aes(ymin = q05, ymax = q95), alpha = 0.5, width = 0.1) +
   theme_cowplot() + xlab("Year") + labs(color = "River") +
-  theme(legend.position = "bottom") + guides(color = guide_legend(nrow = 3))
+  theme(legend.position = "bottom") + guides(color = guide_legend(nrow = 3)) +
+  geom_vline(data = morat, aes(xintercept = year, color = river_name), alpha = 0.4, linetype = 2) +
+  geom_text(data = morat, aes(year-1.5, 0.23, label = region, color = river_name))
 s1trends
 ggsave("figures/s1-trends.png", height = 5, width = 7)
 

### 2020-01-17 ms/ms.tex
Incorporate Brian's comments and improve discussion on potential mechanisms of climate effects

@@
 \bibpunct{(}{)}{;}{a}{}{;}
-\bibliographystyle{mee}
+\bibliographystyle{fishfishnourl}
 
@@
 \small{\textsuperscript{2} Norwegian Institute for Nature Research (NINA), Trondheim, Norway\\}
-\small{\textsuperscript{3} Department of Fisheries and Oceans Canada, St. John's, NL, Canada\\}
-\small{\textsuperscript{4} Minist\`{e}re des For\^{e}ts, de la Faune et des Parcs, Qu\`{e}bec, QC, Canada\\}
-\small{\textsuperscript{5} Department of Fisheries and Oceans Canada, Moncton, NB, Canada\\}
-\small{\textsuperscript{6} Department of Fisheries and Oceans Canada, Dartmouth, NS, Canada\\}
+\small{\textsuperscript{3} Fisheries and Oceans Canada, St. John's, NL, Canada\\}
+\small{\textsuperscript{4} Minist\`{e}re des For\^{e}ts, de la Faune et des Parcs, Qu\'{e}bec, QC, Canada\\}
+\small{\textsuperscript{5} Fisheries and Oceans Canada, Moncton, NB, Canada\\}
+\small{\textsuperscript{6} Fisheries and Oceans Canada, Dartmouth, NS, Canada\\}
 \small{\textsuperscript{*} Corresponding author: spardo@dal.ca}
@@
 least known yet one of the most crucial with regards to population
-persistence. Recently, declines in many Atlantic salmon populations in eastern
-Canada have been attributed to changes in conditions at sea, negatively
+persistence. Declines in many Atlantic salmon populations in eastern
+Canada have often been attributed to changes in conditions at sea, negatively
 affecting their survival. However, marine survival estimates are difficult to
@@
 mid-1990s. 
-Estimates of \St were highly uncertain and trends could not be assessed. \Pg
-was temporally stable, except for LaHave and Nashwaak Rivers. Our findings
-challenge the assumption that marine survival has changed in a consistent
+Estimates of \St were highly uncertain, particularly for 1SW-dominated rivers
+(Conne, Campbellton and WAB, where the abundance of 2SW returns is very low)
+and thus the posterior distributions matched our choice of prior and trends
+could not be assessed. 
+\Pg was temporally stable within rivers, except for LaHave and Nashwaak Rivers. 
+Our findings challenge the assumption that marine survival has changed in a consistent
 manner among populations across a broad geographic scale, and suggest that
-trends can be river-specific. Further work exploring potential correlates of
-these trends in marine survival is warranted.
+trends can be river-specific. 
+The well established correlations between climate variables and abundance are
+not being mediated solely by marine survival, and there can potentially be
+some important indirect effects on fecundity.
+Further work exploring potential correlates of marine survival and the
+potential non-lethal effects of climate effects is warranted.
 
@@
 moratoria in all regions, beginning in the Maritime provinces (1984) and
-following in Newfoundland (1992), Labrador (1998), and Québec (2000). Since
-these closures, many populations have not increased as expected (REF): some 
+following in Newfoundland (1992), Labrador (1998), and Qu\'{e}bec (2000). Since
+these closures, many populations have not increased as 
+expected \citep{Dempson2004, ICES2019}; some 
 have been assessed as considered threatened or endangered by the 
 Committee on the Status of Endangered Wildlife in Canada \citep[COSEWIC, ][]{Cosewic2010}, 
-Canada’s national science advisory body (to the national government) on species risk of extinction.
+Canada's national science advisory body (to the national government) on
+species risk of extinction.
 While it is not fully understood what is driving population declines, the potential
 drivers of these are many \citep[see ][for a detailed discussion of possible
-causes]{Cairns2001}: fishing mortality, damming of waterways
-and changes in their freshwater habitats \citep{Dunfield1985}, acidification
+causes]{Cairns2001}, including but not limited to: fishing mortality \citep{Dempson2004}, 
+damming of waterways and changes in their freshwater habitats \citep{Dunfield1985}, acidification
 of freshwater habitats \citep[particularly in the Southern Uplands region of
@@
 effects of interbreeding or interactions with escaped farmed salmon
-\citep{Keyser2018}, and changes in their marine survival.
+\citep{Keyser2018}, and climate-driven changes in survival and productivity \citep{Mills2013}.
 
 % 2.
-Over the past three decades, a narrative has emerged that marine survival of Atlantic salmon has
-declined throughout the North Atlantic \citep{Hansen1998,OMaoileidigh2003,Chaput2012a}. 
+Over the past three decades, a narrative has emerged that marine survival of
+Atlantic salmon has declined throughout the North Atlantic \citep{ICES2019}.
+%\citep{Hansen1998,OMaoileidigh2003,Chaput2012a}.
+As there are multiple lines of evidence that climate conditions directly and
+indirectly influence the abundance and productivity of Atlantic salmon
+populations \citep{Mills2013,Almodovar2019}, it is expected that these oceanic climate effects are mostly
+driving population dynamics through changes in marine survival.
 An implicit assumption, albeit infrequently acknowledged, is that any trend in
@@
 has been implicitly assumed that populations share similar temporal trends in
-at-sea mortality \citep{Friedland1993, Friedland1998}.  
+at-sea mortality \citep{Friedland1993, Friedland1998, Russell2012}.  
 A recent study by \citet{Olmos2019} suggested that trends in post-smolt
@@
 considerably. Populations in the southern part of the Canadian range are more
-likely to be assessed as being below conservation targets than those in more
+likely to be assessed as being of conservation concern than those in more
 northerly regions \citep{Cosewic2010}. This geographical disparity in status
@@
 Canadian populations of Atlantic salmon to model trends in marine survival.
-While some studies have used maturity-schedule models to estimate marine
+While some studies have previously used maturity-schedule models to estimate marine
 survival for a limited number of salmon populations \citep{Chaput2003b}, none
-have incorporated data extending over decates, nor have they examined
+have incorporated data extending over decades, nor have they examined
 trends among more than two or three populations. 
@@
 abundances for seven Atlantic salmon populations in eastern Canada, encompassing a
-wide range of the species' distribution (Fig.~\ref{fig:map}). These
-populations included LaHave River in the Southern Uplands region of Nova
-Scotia (NS), Nashwaak River in New Brunswick (NS) which connects to the
-Bay of Fundy, Trinit\'{e} and Saint-Jean Rivers in Qu\'{e}bec (QC), which
-flow to the St. Lawrence River and Bay of St. Lawrence, respectively, and
-Western Arm Brook (also referred to as WAB), Campbellton, and Conne Rivers in Newfoundland, which flow to the
-Strait of Belle Isle, Labrador Sea, and Gulf of St. Lawrence, respectively.
+wide range of the species' distribution (Fig.~\ref{fig:map}). 
+Populations included the LaHave River in the Southern Uplands region of Nova
+Scotia (NS), Nashwaak River, New Brunswick (NB), Rivi\`{e}re de la Trinit\'{e} and
+Saint-Jean rivers in Qu\'{e}bec (QC), and  Western Arm Brook (also referred to asalso referred to as  WAB), Campbellton, and
+Conne River, Newfoundland (NL). 
 Data were collected in NS, NB, and Newfoundland
@@
 
-\subsubsection*{Smolt data}
+\subsubsection*{Smolt and adult return abundance data}
 
-\comment{Co-authors: please check data sources}
-
-Smolt estimates originate from a variety of sources.
+Smolt and adult return abundance estimates originate from a variety of sources.
 Estimates from the XXXX populations were obtained using a capture-mark-recapture approach, while
 Estimates from the XXXX populations were obtained by direct counts using a XXXX.
-Refer to ... for more details on the data collection details.
-
-\subsubsection*{Returns data}
+... for more details on the . 
+For data collection details refer to 
+\citet{Venoitt2018} for NL populations, 
+\citet{April2018}  for QC populations,
+\citet{Jones2014} for NB populations,
+and \textbf{REF} for NS populations.
 
@@
 year at sea (\Pg).
+
 Our model does not include repeat spawners and assumes that no fish spend
 three or more winters at sea before returning to spawn for the first time.
-
-The model also assumes that mortality in the second winter at sea 
+The model also assumes that mortality in the second winter at sea (\St)
 is additional to mortality in the first winter at sea in the previous year, 
-and therefore does not account for differences
-the returning (Pr) occurs just before
-actually being counted as returns and that S2 is any additional mortality in
-the subsequent year. I agree this is far from ideal, given that we know there
-is a period of a few months where 1SW returns are subject to a different
-environment than those 2SW+ fishes.
+and therefore does not account for differences in environmental conditions experienced
+between 1SW and 2SW fish of the same smolt cohort during their overlapping first year at sea.
+In other words, our model assumes that the decision of returning occurs just before
+actually being counted as returns and that \St is any additional mortality in
+the subsequent year.
 
@@
 \epsilon_{1,r} &\sim Normal(0, 0.01) \\
-\epsilon_{2,r} &\sim Normal(0, 0.01)
+\epsilon_{2,r} &\sim Normal(0, 0.01).
 \end{align}
@@
  S_{1} &= e^{-Z_1}, \\
- S_{2} &= e^{-Z_2} 
+ S_{2} &= e^{-Z_2}. 
 \end{align}
@@
    \end{cases} \\
-    \sigma_r &\sim halfNormal(0, 1)
+    \sigma_r &\sim halfNormal(0, 1).
 \end{align}
@@
 Z_1 &\sim logNormal(1, 0.22),   \\ 
-Z_2 &\sim logNormal(0.2, 0.3)
+Z_2 &\sim logNormal(0.2, 0.3).
 \end{align}
@@
         estimates of \So for the seven rivers examined, error bars indicate
-        the 90\% credible intervals. b) Z-scores of median posterior
-        estimates.} \label{fig:s1-dual} 
-\end{figure}
+        the 90\% credible intervals, and dashed lines denote years of commercial
+        fishing moratoria for each province. b) Z-scores of median posterior
+        estimates.} \label{fig:s1-dual} \end{figure}
 
@@
 % 1. Main findings
-Our findings challenge the narrative that marine survival is declining
+Our findings challenge the narrative that marine survival, specifically survival in the first year at sea, is declining
 uniformly throughout the range of Atlantic salmon in the northwest Atlantic.
@@
 . Over the time periods for which data were available, some rivers show positive trends in survival in the
-first winter at sea (\So) while other show stable trends, and some even show
+first winter at sea (\So) while other show highly variable yet stable trends, and some even show
 declines. We could not assess trends in the second winter at sea (\So) or
@@
 might be attributable to the use of stock-recruitment relationships to estimate survival,
-relationships thay may have been confounded by changes in recruitment dynamics.
-
+relationships that may have been confounded by changes in recruitment dynamics.
 There is some evidence of a correlation between return rate and growth (as
@@
 temperature experienced by the post-smolts and their respective
-population-specific thermal optima, which could explain why populations in eastern North America are declining
-in the southern part of their range but potentially increasing further north.
-
-A potential association between temperature and survival warrants
-further study.
+population-specific thermal optima. 
+This difference could explain why populations in eastern North America are
+declining in the southern part of their range but potentially increasing
+further north, and also why some studies find possitive correlations between
+temprature and abundance \citep{Friedland1998, Friedland1998b, Jonsson2004}
+while other find negative ones \citep{Friedland1993, Todd2008}.
+A potential association between temperature and direct estimates of marine
+survival warrants further study.
 
@@
 
+Oceanic conditions have been correlated with abundance trends and growth
+\citep{Todd2008}, however, the mechanism by which bottom-um effects, mediated by changes in food availability,
+population dynamics beyond marine survival needs to
+be thoroughly reassessed. If marine survival on its own cannot fully explain
+trends in abundance, then there are potential carry-on effects of oceanic
+conditions that manifest with regards to fresh production. 
+For example, adults
+that return to spawn after spending suboptimal conditions at sea might be less
+likely to make it to their spawning grounds, successfully secure a mate,
+produce fewer eggs, or produce eggs with lower per capita fitness than those
+produced by adults which reared in optimal oceanic conditions.
+As larger females tend to be more productive, in terms of fecundity and total
+reproductive energy, than the same weight's worth of smaller females
+\citep{Barneche2018}, a small decrease in body condition resulting from bottom-up
+impacts on food availability, could potentially have disproportionate effects on fecundity
+and fitness of the offspring.
+Egg-to-smolt survival in Atlantic salmon is highly variable \citep{Chaput2015}
+and changes in the oceanic conditions that spawners experiences could be
+contributing to this variability.
+Obviously, there would be a delay in these effects to manifest as adult
+trends, as it would take a whole life cycle for fecundity effects to manifest
+as abundance of returning adults.
+However, given that most correlations are between relatively monotonic declines
+in abundance coupled also monotonic increases in climatic indices
+over decadal time scales \citep[e.g.,][]{Friedland1998, Todd2008,
+    Beaugrand2012}, it would be expected that this correlations would be
+maintained even if salmon abundances were lagged by a generation length.
+
 % 4. Caveats
 As with all novel modelling approaches, there are caveats to acknowledge.
+The seven populations explored in this study might not be representative
+of the regional trends in marine survival. However, there are no other
+long-term time series of smolts and adult returns to draw inferences from.
 While there are analytical issues with estimation of \So, \St, and \Pg,
 the assumption that \St is additive to \So could produce unrealistic results.
-Overcoming this assumption would require an additional parameter to be
+We know there is a period of a few months where 1SW
+returns are subject to a different environment than those salmon that will
+return as 2SW the next year. 
+While this is not ideal,
+overcoming this assumption would require an additional parameter to be
 estimated, or an additional assumption as to what proportion of \So is not
@@
 
-% We would like to thank Geir Bolstad, G\'{e}rald Chaput, Brian Dempson, and Martha
-% Robertson for their useful discussions on estimating marine survival in
-% Atlantic salmon, and Amanda Kissel for her helpful comments on the manuscript.
-% Brian Dempson and Geoff Venoit for providing the data Conne River data.
+% We would like to thank Geir Bolstad, G\'{e}rald Chaput, Brian Dempson, and
+% Martha Robertson for their useful discussions on estimating marine survival
+% in Atlantic salmon, and Amanda Kissel for her helpful comments on the
+% manuscript. Brian Dempson and Geoff Venoit for providing the data Conne
+% River data.
 We would like to thank G\'{e}rald Chaput for his useful discussions on
-estimating marine survival in Atlantic salmon, and Sean Anderson for his help
-with implementing the non-centered parameterization of the model.
-This research was supported by the Atlantic Salmon Conservation Foundation and
-the Atlantic Salmon Research Joint Venture.
+estimating marine survival in Atlantic salmon, Carmen David for her comments
+on the manuscript, and Sean Anderson for his help with implementing the
+non-centered parameterization of the model. This research was supported by the
+Atlantic Salmon Conservation Foundation and the Atlantic Salmon Research Joint
+Venture.
 

### 2020-01-20 ms/ms.tex
Update map R code and minor ms edits

@@
 
-\noindent\small{\textsuperscript{1} Department of Biology, Dalhousie University, Halifax, NS, B3H 4R2, Canada\\}
+\noindent\small{\textsuperscript{1} Department of Biology, Dalhousie University, Halifax, NS, Canada\\}
 \small{\textsuperscript{2} Norwegian Institute for Nature Research (NINA), Trondheim, Norway\\}
@@
 \citet{Jones2014} for NB populations,
-and \textbf{REF} for NS populations.
+and \citet{Gibson2009} for NS populations.
 
@@
 uniformly throughout the range of Atlantic salmon in the northwest Atlantic.
-Temporal trends are not consistent among populations 
-. Over the time periods for which data were available, some rivers show positive trends in survival in the
+Temporal trends are not consistent among populations. 
+Over the time periods for which data were available, some rivers show positive trends in survival in the
 first winter at sea (\So) while other show highly variable yet stable trends, and some even show

### 2020-01-20 analysis/05-river-location-map.R

@@ -62,8 +62,8 @@ r2sf <- st_as_sf(regions2)
 regions_sf <- st_as_sf(region) %>%
   filter(region_sub %in% c("Atlantic Canada", "New England", "Québec"))
 
-provlabels <- read_csv("prov-labels.csv") 
-riverlocs <- read_csv("river-locs.csv") 
+provlabels <- read_csv("raw-data/prov-labels.csv") 
+riverlocs <- read_csv("raw-data/river-locs.csv") 
 dput(riverlocs)
 sealoc <- tibble(sea = c("Gulf of\nSt. Lawrence", "Gulf of\nMaine", "ATLANTIC\nOCEAN"),
                  long = c(-60.7, -68, -56),

### 2020-01-27 ms/ms.tex
Fix smolt uncertainty not being incorporated in maturity schedule model (logsmolts instead of logsmolts_true)

@@
 
-The marine phase of anadromous Atlantic salmon (\emph{Salmo salar}) is the
-least known yet one of the most crucial with regards to population
-persistence. Declines in many Atlantic salmon populations in eastern
-Canada have often been attributed to changes in conditions at sea, negatively
-affecting their survival. However, marine survival estimates are difficult to
-obtain given that many individuals spend multiple winters in the ocean before
-returning to freshwater to spawn, necessitating the estimation of multiple
-parameters. To do so, we fit a hierarchical Bayesian maturity schedule model
-to smolt and adult abundance time series for seven populations located in
+Declines in wild Atlantic salmon (\emph{Salmo salar}) throughout the north
+Atlantic are primarily attributed to declining survival at sea. This
+hypothesis has proven challenging to test on a river-by-river basis because of
+the need to model data on both migrating smolts and returning adults and to
+simultaneously estimate multiple parameters, especially for salmon spending
+more than one sea-winter (1SW) before spawning. We fit a hierarchical Bayesian
+maturity schedule model to data (19-42 years) for seven populations in Canada's
 Newfoundland and Labrador (NL), Qu\'{e}bec (QC), New Brunswick (NB), and Nova
-Scotia (NS). The datasets ranged between 19 and 42 years in length. We
-estimated three components of marine survival: survival in the first year at
-sea (\So), survival in the second year at sea (\St), and proportion returning
-as one sea-winter adults (\Pg). Controlling for time frame, trends in
-estimates of \So were not consistent among rivers.
-In the four populations for which data extended to 1990, marine survival
-during the first year at sea predominantly increased over time in Western Arm
-Brook (NL) and Rivi\`{e}re Saint-Jean (QC) but largely declined in Conne River
-(NL) and Rivi\`{e}re Trinit\'{e} (QC). In the three other populations, \So
-exhibited an increase in Campbellton River (NL), a decline in LaHave River
-(NS), and fluctuating stability in Nashwaak River (NB) since approximately the
-mid-1990s. 
-Estimates of \St were highly uncertain, particularly for 1SW-dominated rivers
-(Conne, Campbellton and WAB, where the abundance of 2SW returns is very low)
-and thus the posterior distributions matched our choice of prior and trends
-could not be assessed. 
-\Pg was temporally stable within rivers, except for LaHave and Nashwaak Rivers. 
-Our findings challenge the assumption that marine survival has changed in a consistent
-manner among populations across a broad geographic scale, and suggest that
-trends can be river-specific. 
-The well established correlations between climate variables and abundance are
-not being mediated solely by marine survival, and there can potentially be
-some important indirect effects on fecundity.
-Further work exploring potential correlates of marine survival and the
-potential non-lethal effects of climate effects is warranted.
+Scotia (NS). We estimate survival in the first (\So) and second year at sea (\St),
+and the proportion returning as 1SW adults (\Pg). Trends in  were not consistent
+among rivers. Since 1990, \So predominantly increased in Western Arm Brook (NL)
+and Rivi\`{e}re Saint-Jean (QC), declining in Conne River (NL) and Rivière
+Trinité (QC). Since the mid-1990s, \So has increased in Campbellton River (NL),
+declined in LaHave River (NS), and exhibited fluctuating stability in Nashwaak
+River (NB). Estimates of \St were highly uncertain, particularly for
+1SW-dominated populations; \Pg was generally stable. These results challenge the
+narrative that marine survival has changed in a temporally consistent manner
+among spatially disparate populations. Our findings suggest that, at the
+population level, changes in abundance are attributable to temporal shifts in
+multiple components of individual fitness, including at-sea survival. If
+salmon populations do not respond in a consistent, uniform manner to changing
+environmental conditions throughout their range, future research initiatives
+should explore why.
+
+% The marine phase of anadromous Atlantic salmon (\emph{Salmo salar}) is the
+% least known yet one of the most crucial with regards to population
+% persistence. Declines in many Atlantic salmon populations in eastern
+% Canada have often been attributed to changes in conditions at sea, negatively
+% affecting their survival. However, marine survival estimates are difficult to
+% obtain given that many individuals spend multiple winters in the ocean before
+% returning to freshwater to spawn, necessitating the estimation of multiple
+% parameters. To do so, we fit a hierarchical Bayesian maturity schedule model
+% to smolt and adult abundance time series for seven populations located in
+% Newfoundland and Labrador (NL), Qu\'{e}bec (QC), New Brunswick (NB), and Nova
+% Scotia (NS). The datasets ranged between 19 and 42 years in length. We
+% estimated three components of marine survival: survival in the first year at
+% sea (\So), survival in the second year at sea (\St), and proportion returning
+% as one sea-winter adults (\Pg). Controlling for time frame, trends in
+% estimates of \So were not consistent among rivers.
+% In the four populations for which data extended to 1990, marine survival
+% during the first year at sea predominantly increased over time in Western Arm
+% Brook (NL) and Rivi\`{e}re Saint-Jean (QC) but largely declined in Conne River
+% (NL) and Rivi\`{e}re Trinit\'{e} (QC). In the three other populations, \So
+% exhibited an increase in Campbellton River (NL), a decline in LaHave River
+% (NS), and fluctuating stability in Nashwaak River (NB) since approximately the
+% mid-1990s. 
+% Estimates of \St were highly uncertain, particularly for 1SW-dominated rivers
+% (Conne, Campbellton and WAB, where the abundance of 2SW returns is very low)
+% and thus the posterior distributions matched our choice of prior and trends
+% could not be assessed. 
+% \Pg was temporally stable within rivers, except for LaHave and Nashwaak Rivers. 
+% Our findings challenge the assumption that marine survival has changed in a consistent
+% manner among populations across a broad geographic scale, and suggest that
+% trends can be river-specific. 
+% The well established correlations between climate variables and abundance are
+% not being mediated solely by marine survival, and there can potentially be
+% some important indirect effects on fecundity.
+% Further work exploring potential correlates of marine survival and the
+% potential non-lethal effects of climate effects is warranted.
 
@@
 causes]{Cairns2001}, including but not limited to: fishing mortality \citep{Dempson2004}, 
-damming of waterways and changes in their freshwater habitats \citep{Dunfield1985}, acidification
-of freshwater habitats \citep[particularly in the Southern Uplands region of
+damming of waterways and changes in the freshwater habitat \citep{Dunfield1985}, acidification
+\citep[particularly in the Southern Uplands region of
 NS, see][]{Gibson2010}, predation by seals and birds \citep{Cairns2000}, negative
@@
 %\citep{Hansen1998,OMaoileidigh2003,Chaput2012a}.
-As there are multiple lines of evidence that climate conditions directly and
+Based on multiple lines of evidence that climate conditions can directly and
 indirectly influence the abundance and productivity of Atlantic salmon
-populations \citep{Mills2013,Almodovar2019}, it is expected that these oceanic climate effects are mostly
-driving population dynamics through changes in marine survival.
-An implicit assumption, albeit infrequently acknowledged, is that any trend in
+populations \citep{Mills2013,Almodovar2019}, it has been presumed that oceanic climate effects are
+driving population dynamics primarily through changes in marine survival.
+An implicit assumption is that any trend in
 survival in ocean habitat that is shared by multiple populations during their
@@
 are hypothesized to share marine habitat during some of their time at sea, it
-has been implicitly assumed that populations share similar temporal trends in
+has been presumed that populations share similar temporal trends in
 at-sea mortality \citep{Friedland1993, Friedland1998, Russell2012}.  
-A recent study by \citet{Olmos2019} suggested that trends in post-smolt
+For example, a recent study by \citet{Olmos2019} suggested that trends in post-smolt
 survival, when estimated at the stock unit level, are synchronously declining
@@
 In contrast to the narrative of widespread, demographically similar increases
-in at-sea mortality, the conservation status of Canadian populations differs
-considerably. Populations in the southern part of the Canadian range are more
+in at-sea mortality, the conservation status of Canadian salmon populations differs
+considerably. Populations in the southern part of their range are more
 likely to be assessed as being of conservation concern than those in more
@@
 marine survival.
-\citet{Chaput2012a}, for example, examined return rates of smolts to adult salmon to assess
-declines in marine survival and found that most populations from Canadian
-rivers had declining return rates. 
-However, looking at trends in return rates
-can mask changes in differential survival in different years at sea, as well
+\citet{Chaput2012a}, for example, examined the return rate of smolts to adult salmon 
+as a metric of marine survival, finding that most Canadian populations 
+had experienced declining return rates. 
+However, examination of trends in return rates alone
+can mask changes in differential survival during different years at sea, as well
 as changes in the proportion of adults returning after one or two years at sea.
-A more recent study by \citet{Olmos2019} suggested that trends in post-smolt
+Recently \citet{Olmos2019} suggested that trends in post-smolt
 survival, when estimated at the stock unit level, are synchronously declining
-for all Atlantic salmon in Eastern North America and Canada. Nonetheless,
-the aforementioned study by \citet{Olmos2019} relied on the veracity of stock-recruitment relationships.
+for all Atlantic salmon in Eastern North America and Canada, a conclusion ultimately 
+grounded on the veracity of highly variable stock-recruitment relationships.
 
 %4. 
-In the present study, we compile smolt and adult returns times series data for seven wild
-Canadian populations of Atlantic salmon to model trends in marine survival.
+In the present study, we compile data on the number of migrating smolts and number of returning adults 
+for seven wild Canadian populations of Atlantic salmon to model trends in marine survival.
 While some studies have previously used maturity-schedule models to estimate marine
 survival for a limited number of salmon populations \citep{Chaput2003b}, none
-have incorporated data extending over decades, nor have they examined
+have incorporated data extending over multiple decades, nor have they examined
 trends among more than two or three populations. 
@@
 Scotia (NS), Nashwaak River, New Brunswick (NB), Rivi\`{e}re de la Trinit\'{e} and
-Saint-Jean rivers in Qu\'{e}bec (QC), and  Western Arm Brook (also referred to asalso referred to as  WAB), Campbellton, and
+Saint-Jean rivers in Qu\'{e}bec (QC), and  Western Arm Brook (also referred to as  WAB), Campbellton, and
 Conne River, Newfoundland (NL). 
@@
 
-Smolt and adult return abundance estimates originate from a variety of sources.
-Estimates from the XXXX populations were obtained using a capture-mark-recapture approach, while
-Estimates from the XXXX populations were obtained by direct counts using a XXXX.
-... for more details on the . 
-For data collection details refer to 
+Smolt and adult return abundance estimates originate from a variety of
+sources. Smolt estimates from the XXXX populations were obtained using a
+capture-mark-recapture approach, while Estimates from the XXXX populations
+were obtained by direct counts using a XXXX.
+For further details on the data collection methodologies refer to 
 \citet{Venoitt2018} for NL populations, 
@@
 \begin{equation}
-    p_{r,y,a} = \frac{\sum_{s}{(\frac{n_{r,y,s,a}}{n_{r,y,s}} * N_{r,y,s})}}{\sum_{s}{N_{r,y,s}}}
+    p_{r,t,a} = \frac{\sum_{s}{(\frac{n_{r,t,s,a}}{n_{r,t,s}} * N_{r,y,s})}}{\sum_{s}{N_{r,t,s}}}
 \end{equation}
 
-where $p_{r,y,a}$ is the proportion of annual returns in river $r$, year $y$,
-and of spawning history $a$ (either 1SW or 2SW returns); $n_{r,y,s,a}$ is the
-number of samples in river $r$, year $y$, of spawning history $a$, and of size
-group $s$; $n_{r,y,s}$ is the total number of samples in river $r$, year $y$,
-and of size group $s$; and $N_{r,y,s}$ is the returns of salmon
-in river $r$, year $y$, and of size group $s$.
+where $p_{r,t,a}$ is the proportion of annual returns in river $r$, year $t$,
+and of spawning history $a$ (either 1SW or 2SW returns); $n_{r,t,s,a}$ is the
+number of samples in river $r$, year $t$, of spawning history $a$, and of size
+group $s$; $n_{r,t,s}$ is the total number of samples in river $r$, year $t$,
+and of size group $s$; and $N_{r,t,s}$ is the returns of salmon
+in river $r$, year $t$, and of size group $s$.
 
@@
 smolt estimates (see Table S1 in the Supplementary material). 
+Where available, we used population-specific measurement error estimates for smolt abundances; if not 
+available, we set measurement error at 5\%. 
 The log-transformed true smolt abundances are
@@
 \begin{align}
-    R_{1,t} &= smolts_{t-1} * S1_t * Pr_t \label{eq:1}, \\
-    R_{2,t+1} &= smolts_{t-1} * S1_t * (1 - Pr_t) * S2_{t+1} \label{eq:2}
+    R_{1,t} &= smolts_{true,t-1} * S_{1,t} * Pr_t \label{eq:1}, \\
+    R_{2,t+1} &= smolts_{true,t-1} * S_{1,t} * (1 - Pr_t) * S_{2,t+1} \label{eq:2}
 \end{align}
@@
 where $R_{1,t}$ and $R_{2,t+1}$ are the estimated abundances of 1SW and 2SW
-salmon returning in years $t$ and $t+1$, respectively, $smolts_{t-1}$ is the
-number of outmigrating smolts in year $t-1$, $S1_t$ is the proportion of
+salmon returning in years $t$ and $t+1$, respectively, $smolts_{true,t-1}$ is the
+estimated number of outmigrating smolts in year $t-1$, $S_{1,t}$ is the proportion of
 salmon surviving in their first year ($t$) at sea, $Pr_t$ is the proportion of
-salmon that return to spawn at year $t$, $S2_{t+1}$ is the survival in their
+salmon that return to spawn at year $t$, $S_{2,t+1}$ is the survival in their
 second year at sea of the same cohort of salmon who did not return to spawn at
@@
 Over the time periods for which data were available, some rivers show positive trends in survival in the
-first winter at sea (\So) while other show highly variable yet stable trends, and some even show
+first winter at sea (\So) while other exhibit highly variable yet stable trends, and some show
 declines. We could not assess trends in the second winter at sea (\So) or
 proportion returning as grilse (\Pg), as these parameter estimates were highly
-uncertain and were strongly influenced by the priors used.
+uncertain and were strongly influenced by the priors.
 Perhaps there is a need to rethink our understanding of Atlantic salmon
-population dynamics in light of the possibility that any real and persistent decline in marine survival was
-experienced by some but not all populations, that reductions in survival might
-have occurred over a relatively brief period of time and may not longer be ongoing, and that since that
-reduction occurred (for a subset of populations) marine survival has
-been stable, or even increasing.
+population dynamics in light of the possibilities that (1) any real and persistent decline in marine survival was
+experienced by some but not necessarily all populations, (2) reductions in survival might
+have occurred over a relatively brief period of time and have not persisted, and (3) marine survival has
+been relatively stable, or increasing in some populations for one or more decades.
 
@@
 structure, different methods for estimating covariance, difference in the
-spatial scales of data sources, and the use stock-recruitment relationships
-rather than empirical smolt observations to estimate marine survival.
+spatial scales of data sources, and perhaps most importantly the use stock-recruitment relationships
+rather than empirical smolt count data to estimate marine survival.
 Interestingly, our estimates of \So and \St are very similar to those produced
@@
 analytical solution), our study reached almost the same results (albeit with
-slightly higher uncertainty) using a Bayesian approach with informative
+slightly higher uncertainty), using a Bayesian approach with informative
 priors. These overlapping trends obtained with two different methods 
@@
 Trends in marine survival
-among populations were compared by \citet{Chaput2012a} using adult return rates, 
-and found that for four out of the six population examined, return rates in the 1990s 
-were less than those during the 1970s.
-\citet{Friedland1993} compared return rates for a number rivers in eastern
+among populations were compared by \citet{Chaput2012a} using adult return rates.
+He found that for 4 of 6 populations examined, return rates in the 1990s 
+were lower than those during the 1970s.
+\citet{Friedland1993} compared return rates for a number of rivers in eastern
 North America between 1973 and 1988, and suggested there are similar trends among these. 
-However, the similarity in trends is driven primarily by two years, 1977 and 1978, which
+However, the similarity in these trends was driven primarily by two years, 1977 and 1978, which
 show concurrent low and high relative return rates across rivers,
-respectively. Others years are much more variable relative to each other.
-\citeauthor{Friedland1993}'s \citeyear{Friedland1993} time series stops at
-1988, thus there are only a few years for which to assess overlap with the
+respectively. Other years are much more variable relative to each other.
+\citeauthor{Friedland1993}'s \citeyear{Friedland1993} time series ends in  
+1988; thus there are only a few years for which to assess overlap with the
 time series in our study.
-However, the pooling of adult return rates, as done by \citet{Chaput2012a} and
-\citet{Friedland1993} can mask inter-annual variability in marine survival,
+In any event, we caution that the pooling of adult return rates \citep{Chaput2012a, Friedland1993} 
+can mask inter-annual variability in marine survival,
 and hence might not produce an accurate depiction of marine survival trends.
-\citet{Dempson2003} describe a general declining trend in marine survival for
-Newfoundland rivers (except WAB); which in our study we observe in the
-Conne but not in Campbellton River or WAB. It is not possible to draw conclusions
+\citet{Dempson2003} described a general declining trend in marine survival for
+Newfoundland rivers (except WAB); we drew the same conclusion for 
+Conne River but not Campbellton River or WAB. It is not possible to draw broader conclusions
 with data from only three Newfoundland rivers, but it seems that among index rivers,
@@
 There is some evidence of a correlation between return rate and growth (as
-indicated by intercirculi spacing on scales), where years of poor growth
+indicated by inter-circuli spacing on scales), where years of poor growth
 tended to also be years of poor survival \citep{Friedland1993}, supporting the
@@
 between spring temperature in the Norwegian and North Seas and population abundance, suggesting warmer
-conditions favour post-smolts \citep{Friedland1998} based on mapping the
+conditions favour post-smolts \citep{Friedland1998}, based on mapping the
 extent of area of suitable temperature (7-13 \textdegree C).
@@
 declining in the southern part of their range but potentially increasing
-further north, and also why some studies find possitive correlations between
-temprature and abundance \citep{Friedland1998, Friedland1998b, Jonsson2004}
-while other find negative ones \citep{Friedland1993, Todd2008}.
-A potential association between temperature and direct estimates of marine
-survival warrants further study.
+further north, and also why some studies find positive correlations between
+temperature and abundance \citep{Friedland1998, Friedland1998b, Jonsson2004}
+while others find negative ones \citep{Friedland1993, Todd2008}.
+Putative associations between temperature and direct estimates of marine
+survival warrants further study at the population level.
 
@@
 period, particularly for populations that are likely to be subjected to
-declining per capita population growth rates generated by Allee effects.
+declining per capita population growth rates ($r$) generated by Allee effects.
 Exploring relationships between survival and population size could potentially
@@
 Oceanic conditions have been correlated with abundance trends and growth
-\citep{Todd2008}, however, the mechanism by which bottom-um effects, mediated by changes in food availability,
-population dynamics beyond marine survival needs to
+\citep{Todd2008}, however, the mechanism by which such bottom-up effects, \
+mediated by changes in food availability,
+affect population dynamics beyond marine survival needs to
 be thoroughly reassessed. If marine survival on its own cannot fully explain
@@
 produce fewer eggs, or produce eggs with lower per capita fitness than those
-produced by adults which reared in optimal oceanic conditions.
+produced by adults which grew in optimal oceanic conditions.
 As larger females tend to be more productive, in terms of fecundity and total
@@
 \citep{Barneche2018}, a small decrease in body condition resulting from bottom-up
-impacts on food availability, could potentially have disproportionate effects on fecundity
+impacts on food availability could potentially have disproportionate effects on fecundity
 and fitness of the offspring.
+
 Egg-to-smolt survival in Atlantic salmon is highly variable \citep{Chaput2015}
@@
 contributing to this variability.
-Obviously, there would be a delay in these effects to manifest as adult
-trends, as it would take a whole life cycle for fecundity effects to manifest
-as abundance of returning adults.
+Obviously, there would be a time lag (perhaps as much as a generation) in how such effects
+might be manifest at the adult stage.
 However, given that most correlations are between relatively monotonic declines
@@
 As with all novel modelling approaches, there are caveats to acknowledge.
-The seven populations explored in this study might not be representative
-of the regional trends in marine survival. However, there are no other
+The seven populations explored in the present study might not be representative
+of regional trends in marine survival. However, there are no other
 long-term time series of smolts and adult returns to draw inferences from.
-While there are analytical issues with estimation of \So, \St, and \Pg,
+While there are analytical issues associated with the estimation of \So, \St, and \Pg,
 the assumption that \St is additive to \So could produce unrealistic results.
@@
 at sea for an additional winter before returning to spawn).
+
 Secondly, the assumed hierarchical structure might not be the most appropriate
@@
 Perhaps a reframing of the issue of marine survival is key to furthering our
-understanding of Atlantic salmon population dynamics. Marine
-survival may not have declined consistently, and over the same time periods, across across populations, but the fact that it has stayed at
-roughly similar levels as it was previously \emph{despite} reduced commercial fishing mortalities,
-and that many populations are smaller in size compared to what they once were, is
-what we should focus on understanding better.
-Given the literature of correlates on marine survival in the past, if survival
-has been stable or increasing, this clearly warrants reexamination of the
-correlates and causes of temporal and spatial shifts in marine survival.
+understanding of Atlantic salmon population dynamics. Marine survival may not
+have declined consistently, and over the same time periods, across all
+populations. 
+But the fact that it has remained at roughly similar levels as it was
+previously \emph{despite} reduced commercial fishing mortalities, suggests
+that there may well be an interaction between small population size (small
+relative to unfished population size or carrying capacity), recovery
+potential, and environmental stochasticity that has not been fully explored in
+Atlantic salmon. 
+All else being equal, relatively small populations are more vulnerable to
+demographic, environmental, and genetic stochasticity than large populations
+\citep{Lande1993, Hutchings2015}. Interactions between population size and the
+demographic consequences of environmental stochasticity appear to have
+affected recovery in many marine fishes that have exhibited impaired recovery
+since mitigation of the threat posed by fishing mortality
+\citep{Hutchings2017, Hutchings2020}. The possibility that similar
+interactions may be impairing the recovery of wild Atlantic salmon merits
+study.
 

### 2020-01-27 analysis/04-extract-posteriors.R

@@ -1,0 +1,163 @@
+library(tidyverse)
+library(rstan)
+library(bayesplot)
+
+# loading stanfit objects
+fithierncp <- readRDS("data/stanfit_hier_ncp.rds") # hierarchical
+#fitsimplencp <- readRDS("data/stanfit_simple_ncp.rds")   # non-hierarchical
+
+# loading data
+alldatastan <- readRDS("data/alldatastan.rds")
+
+# Indexing positions of years for each river
+indexdf <- tibble(pos = as.character(1:length(alldatastan$years)),
+                  river_name = rep(alldatastan$river_name, 
+                                   times = alldatastan$nyears), 
+                  year = alldatastan$years)
+
+# For indexing population level posteriors
+riverindexdf <- tibble(pos = as.character(1:length(alldatastan$river_name)),
+                       river_name = alldatastan$river_name)
+
+# Extracting posterior distributions of S1
+s1post <- as.matrix(fithierncp, pars = "S1") %>%
+  as_tibble %>%
+  gather(parameter, value) %>%
+  mutate(pos = str_extract(parameter, "(?<=\\[).+?(?=\\])"),
+         param = str_extract(parameter, ".+?(?=\\[)")) %>%
+  left_join(indexdf, by = "pos")
+
+# These two should fail now as these years have been removed upstream
+s1post %>%
+  filter((river_name == "Saint-Jean River" & year == 1998))# %>%
+#summarise(parameter = unique(parameter))# S1[9]
+s1post %>%  
+  filter((river_name == "Trinité River" & year == 2007)) # %>%
+# summarise(parameter = unique(parameter)) # S1[48]
+
+saveRDS(s1post, file = "data/s1-posteriors.rds")
+
+s1quant <- s1post %>%
+  group_by(pos, river_name, year) %>%
+  summarise(median = median(value),
+            q05 = quantile(value, 0.05),
+            q25 = quantile(value, 0.25),
+            q75 = quantile(value, 0.75),
+            q95 = quantile(value, 0.95))
+
+# Removing years with missing smolt data (which was filled with average smolt counts) 
+s1quant <- s1quant %>%
+  #filter(!(river_name == "Saint-Jean River" & year == 1998))  %>%
[…]

### 2020-01-27 analysis/05-correlation-tests.R

@@ -2,7 +2,7 @@
 
 library(tidyverse)
 library(corrplot)
-source("99-helper-functions.R")
+source("analysis/99-helper-functions.R")
 
 alldatastan <- readRDS("data/alldatastan.rds")
 s1quant <- readRDS("data/s1quant.rds")
@@ -11,8 +11,6 @@ allreturns <- readRDS("data/allreturns.rds") %>%
   mutate(zscoresmolts = scale(log(smolts)), zscoretotal = scale(log(totalreturns))) %>%
   ungroup()
   
-
-
 river_names <- s1quant %>%
   ungroup() %>%
   select(year, zscore, river_name) %>%

### 2020-03-09 ms/ms.tex
Incorporate Geir's suggestions

@@
 \begin{align}
-    log(R_{1,r,t}) &= log(smolts_{r,t-1}) + log(Pr_{r,t}) - Z_{1,r,t} + \epsilon_{1,r} \label{eq:3}, \\
-    log(R_{2,r,t+1}) &= log(smolts_{r,t-1}) - Z_{1,r,t} + log(1 - Pr_{r,t})  - Z_{2,r,t+1} + \epsilon_{2,r} \label{eq:4} 
+    log(R_{1,r,t}) &= log(smolts_{r,t-1}) + log(Pr_{r,t}) - Z_{1,r,t} \label{eq:3}, \\
+    log(R_{2,r,t+1}) &= log(smolts_{r,t-1}) - Z_{1,r,t} + log(1 - Pr_{r,t})  - Z_{2,r,t+1} \label{eq:4} 
 \end{align}
 
-where $Z_{1,r,t}$ and $Z_{2,r,t+1}$ are the instantaneous mortality rates and $\epsilon_{1,r}$ and $\epsilon_{2,r}$
-are the process error terms. 
+where $Z_{1,r,t}$ and $Z_{2,r,t+1}$ are the instantaneous mortality rates.
+Process error was included as the standard deviation of the log-transformed
+return estimates from equation~\ref{eq:4}:
+
+\begin{align}
+log(R_{obs,1,t,r}) &\sim Normal(log(R_{1,t,r}), \epsilon_{1,r}), \\
+log(R_{obs,2,t,r}) &\sim Normal(log(R_{2,t,r}), \epsilon_{2,r}) \label{eq:5} 
+\end{align}
+
+where $R_{obs,1,t,r}$ and $R_{obs,2,t,r}$ are the observed return estimates
+for year $t$ and river $r$ of 1SW and 2SW fish, respectively, $\epsilon_{1,r}$
+and $\epsilon_{2,r}$ are the process error terms. 
 These error terms are normally distributed with a standard deviation
@@
 
-The priors for \So and \St are specified as log-normal distributions around
-the instantaneous mortality rates $Z_1$ and $Z_2$:
+The priors for $Z_1$ and $Z_2$ are specified as log-normal distributions:
 
@@
 first winter at sea (\So) while other exhibit highly variable yet stable trends, and some show
-declines. We could not assess trends in the second winter at sea (\So) or
+declines. We could not assess trends in the second winter at sea (\St) or
 proportion returning as grilse (\Pg), as these parameter estimates were highly
@@
 rather than empirical smolt count data to estimate marine survival.
+As trends in marine survival on the first winter at sea are highly independent
+among rivers on relatively small spatial scale, trends from broader
+geographical areas (i.e. province, state, or country-wide estimates) may not
+be representative.
 Interestingly, our estimates of \So and \St are very similar to those produced

### 2020-03-09 ms/ms.tex
Incorporated Brian's edits

@@
 
-\section*{Trends in marine survival of Altantic salmon in eastern Canada}
+\section*{Trends in marine survival of Atlantic salmon in eastern Canada}
 
@@
 simultaneously estimate multiple parameters, especially for salmon spending
-more than one sea-winter (1SW) before spawning. We fit a hierarchical Bayesian
-maturity schedule model to data (19-42 years) for seven populations in Canada's
+more than one winter at sea (1SW) before spawning. We fit a hierarchical Bayesian
+maturity schedule model to data (19-42 years) for seven populations in  
 Newfoundland and Labrador (NL), Qu\'{e}bec (QC), New Brunswick (NB), and Nova
-Scotia (NS). We estimate survival in the first (\So) and second year at sea (\St),
+Scotia (NS), Canada. We estimate survival in the first (\So) and second year at sea (\St),
 and the proportion returning as 1SW adults (\Pg). Trends in  were not consistent
-among rivers. Since 1990, \So predominantly increased in Western Arm Brook (NL)
-and Rivi\`{e}re Saint-Jean (QC), declining in Conne River (NL) and Rivière
-Trinité (QC). Since the mid-1990s, \So has increased in Campbellton River (NL),
-declined in LaHave River (NS), and exhibited fluctuating stability in Nashwaak
+among rivers. Since 1990, \So increased at Western Arm Brook (NL)
+and Rivi\`{e}re Saint-Jean (QC), but declined at Conne River (NL) and Rivière
+de la Trinité (QC). Since the mid-1990s, \So increased at Campbellton River (NL),
+declined at LaHave River (NS), and fluctuated at Nashwaak
 River (NB). Estimates of \St were highly uncertain, particularly for
@@
 exploited marine fisheries since the early 1990s has shown that not all
-populations respond as favourably as anticipated to considerable reductions in
+populations respond as favourably as anticipated to major reductions in
 exploitation \citep{Hutchings2017}. Efforts to close commercial
@@
 have been assessed as considered threatened or endangered by the 
-Committee on the Status of Endangered Wildlife in Canada \citep[COSEWIC, ][]{Cosewic2010}, 
+Committee on the Status of Endangered Wildlife in Canada \citep[][]{Cosewic2010}, 
 Canada's national science advisory body (to the national government) on
@@
 survival in ocean habitat that is shared by multiple populations during their
-seaward migration period will be experienced similarly by multiple
-populations. Put another way, given that salmon from different river systems
+seaward migration period will be experienced similarly. 
+Put another way, given that salmon from different rivers 
 are hypothesized to share marine habitat during some of their time at sea, it
@@
 at-sea mortality \citep{Friedland1993, Friedland1998, Russell2012}.  
-For example, a recent study by \citet{Olmos2019} suggested that trends in post-smolt
-survival, when estimated at the stock unit level, are synchronously declining
+A recent study by \citet{Olmos2019} suggested that trends in post-smolt
+survival, when estimated at the stock level, are synchronously declining
 for all Atlantic salmon in eastern North America.
@@
 Populations included the LaHave River in the Southern Uplands region of Nova
-Scotia (NS), Nashwaak River, New Brunswick (NB), Rivi\`{e}re de la Trinit\'{e} and
+Scotia (NS), Nashwaak River, New Brunswick (NB), Rivi\`{e}re de la Trinit\'{e} (Trinit\'{e}) and
 Saint-Jean rivers in Qu\'{e}bec (QC), and  Western Arm Brook (also referred to as  WAB), Campbellton, and
@@
 Smolt and adult return abundance estimates originate from a variety of
-sources. Smolt estimates from the XXXX populations were obtained using a
-capture-mark-recapture approach, while Estimates from the XXXX populations
-were obtained by direct counts using a XXXX.
+sources. Smolt estimates from the ... and Conne populations were obtained using a
+mark-recapture approach, while estimates from the ..., WAB, and Campbellton populations
+were obtained by direct counts using fish counting fences.
 For further details on the data collection methodologies refer to 
-\citet{Venoitt2018} for NL populations, 
+\citet{Dempson1991, Venoitt2018} for NL populations, 
 \citet{April2018}  for QC populations,
@@
 
-Yearly return data is often recorded in terms of two size groups: small ($< 63$ cm
+Yearly return data are recorded in terms of two size groups: small ($< 63$ cm
 FL) and large ($\geq 63$ cm FL) salmon, as these closely represent different
-life-history strategies (i.e. 1SW and 2SW), but can be conflated with repeat
+life-history strategies (i.e. 1SW and 2SW), but can be confounded with repeat
 spawners of different sizes. To correct for this, we estimated the abundance
-of 1SW and 2SW returns using yearly scale data of a subsample of returns:
+of 1SW and 2SW returns using yearly scale age data of a subsample of returns:
 
@@
 
-Trends in estimates of \So were highly variable within among rivers
+Trends in estimates of \So were highly variable within and among rivers
 (Fig.~\ref{fig:s1-dual}). The highest median posterior estimates of \So
@@
 with an estimate of \So of 0.007, while the Conne, LaHave, and Trinit\'{e} had
-years of estimates of \So between 0.01 and 0.02 (Fig.~\ref{fig:s1-faceted}).
+years where estimates of \So varied between 0.01 and 0.02 (Fig.~\ref{fig:s1-faceted}).
 
-Trends among populations were variable as well: roughly, the Campbellton,
+Trends among populations also varied: Campbellton,
 Saint-Jean, and Western Arm Brook populations showed increases in \So
-over time, the Trinit\'{e}, Conne, and La Have populations showing decreases,
-and the Nashwaak population showing an increase in median \So during the early
+over time, Trinit\'{e}, Conne, and La Have populations showed decreases,
+while at Nashwaak there was an increase in median \So during the early
 2000s but a decrease in the 2010s. Yearly estimates of \So had very little
@@
 
-Estimates of \St were highly uncertain in all rivers, and trends of any kind
+Estimates of \St were highly uncertain in all rivers, and trends 
 were not apparent in most rivers given the large range of the credible
@@
 
-Most correlations between z-scored trends were not significant, with only 
+Most correlations between z-scored trends were not statistically significant, with only 
 three out of the 21 pair-wise comparisons having a p-value below 0.01 (Fig.~\ref{fig:s1-corr}a).
@@
 % 1. Main findings
-Our findings challenge the narrative that marine survival, specifically survival in the first year at sea, is declining
+Our results challenge the narrative that marine survival, specifically survival in the first year at sea, is declining
 uniformly throughout the range of Atlantic salmon in the northwest Atlantic.
@@
 % 2. How they compare to other publications
-Our findings are contrary to those of \citet{Olmos2019}, who detected positive
+Our results are contrary to those of \citet{Olmos2019}, who detected positive
 correlations in post-smolt survival among spatially broad stock units. These differences
@@
 rather than empirical smolt count data to estimate marine survival.
-As trends in marine survival on the first winter at sea are highly independent
+As trends in marine survival during the first winter at sea are highly independent
 among rivers on relatively small spatial scale, trends from broader
@@
 % 3. Potential reasons
-There are multiple potential explanations for the lack of synchronous
+There are a variety of potential explanations for the lack of synchronous
 trends in estimates of \So. 
@@
 
-Egg-to-smolt survival in Atlantic salmon is highly variable \citep{Chaput2015}
-and changes in the oceanic conditions that spawners experiences could be
+Egg-to-smolt survival in Atlantic salmon is highly variable \citep{Klemetsen2003,Chaput2015}
+and changes in the oceanic conditions that spawners experience could be
 contributing to this variability.

### 2020-03-09 ms/ms.tex
Incorporate Julien's edits

@@
 populations respond as favourably as anticipated to major reductions in
-exploitation \citep{Hutchings2017}. Efforts to close commercial
+exploitation \citep{Hutchings2017}. Gradual efforts to close commercial
 Atlantic salmon (\emph{Salmo salar}) fisheries in eastern Canada culminated in
-moratoria in all regions, beginning in the Maritime provinces (1984) and
+full moratoria in all regions, beginning in the Maritime provinces (1984) and
 following in Newfoundland (1992), Labrador (1998), and Qu\'{e}bec (2000). Since
@@
 Smolt and adult return abundance estimates originate from a variety of
-sources. Smolt estimates from the ... and Conne populations were obtained using a
+sources. Smolt estimates from the Trinit\'{e}, Saint-Jean, and Conne populations were obtained using a
 mark-recapture approach, while estimates from the ..., WAB, and Campbellton populations
@@
 
-Yearly return data are recorded in terms of two size groups: small ($< 63$ cm
+Yearly return data are often recorded in terms of two size groups: small ($< 63$ cm
 FL) and large ($\geq 63$ cm FL) salmon, as these closely represent different
 life-history strategies (i.e. 1SW and 2SW), but can be confounded with repeat
-spawners of different sizes. To correct for this, we estimated the abundance
+spawners of different sizes. To correct for this in returns 
+reported as small and large salmon, we estimated the abundance
 of 1SW and 2SW returns using yearly scale age data of a subsample of returns:
@@
 structure, different methods for estimating covariance, difference in the
-spatial scales of data sources, and perhaps most importantly the use stock-recruitment relationships
+spatial scales of data sources (i.e. river vs province scales), and perhaps most importantly the use stock-recruitment relationships
 rather than empirical smolt count data to estimate marine survival.

### 2020-03-12 analysis/04-extract-posteriors.R
Adding logsmolt estimates to posterior extraction

@@ -160,4 +160,31 @@ saveRDS(logitprcvquant, file = "data/logitprcvquant.rds")
 
 ############################
 
+# extracting estimates of logsmolts_true
+logsmoltspost <- as.matrix(fithierncp, pars = "logsmolts_true") %>%
+  as_tibble %>%
+  gather(parameter, value) %>%
+  mutate(pos = str_extract(parameter, "(?<=\\[).+?(?=\\])"),
+         param = str_extract(parameter, ".+?(?=\\[)")) %>%
+  left_join(indexdf, by = "pos")
+
+saveRDS(logsmoltspost, file = "data/logsmoltstrue-posteriors.rds")
+
+
+z1post <- as.matrix(fithierncp, pars = "Z1") %>%
+  as_tibble %>%
+  gather(parameter, value) %>%
+  mutate(pos = str_extract(parameter, "(?<=\\[).+?(?=\\])"),
+         param = str_extract(parameter, ".+?(?=\\[)")) %>%
+  left_join(indexdf, by = "pos")
+
+z2post <- as.matrix(fithierncp, pars = "Z2") %>%
+  as_tibble %>%
+  gather(parameter, value) %>%
+  mutate(pos = str_extract(parameter, "(?<=\\[).+?(?=\\])"),
+         param = str_extract(parameter, ".+?(?=\\[)")) %>%
+  left_join(indexdf, by = "pos")
+
+saveRDS(z1post, file = "data/z1-posteriors.rds")
+saveRDS(z2post, file = "data/z2-posteriors.rds")
 

### 2020-03-12 analysis/05-figures-for-trends-paper.R
Update moratoria years

@@ -21,7 +21,7 @@ addnas <- tibble(pos = NA,
 
 s1quant <- bind_rows(s1quant, addnas)
 
-ggplot(s1quant, aes(year, median)) + 
+s1trends <- ggplot(s1quant, aes(year, median)) + 
   geom_line(alpha = 0.5) + geom_point(size = 1, alpha = 0.7) + 
   ylab(expression(paste(S[1]~"estimates", sep = ""))) +
   geom_errorbar(aes(ymin = q05, ymax = q95),width= 0.1) +
@@ -30,9 +30,16 @@ ggplot(s1quant, aes(year, median)) +
   labs(color = "River") +
   theme(legend.position = "none", 
         legend.text =  element_text(size = 12))
+s1trends
 ggsave("figures/s1-trends-faceted.png", height = 6, width = 5)
 
-morat <- tibble(year = c(1984, 1992, 2000),
+s1trends + facet_wrap(~river_name, ncol = 3, scales = "free_y") 
+ggsave("figures/s1-trends-presentation.png", height = 4, width = 8)
+
+
+
+
+morat <- tibble(year = c(1984, 1998, 2000),
                 region = c("NS+\nNB", "NL", "QC"),
                 river_name = c("Nashwaak River", "Western Arm Brook", "Trinité River"))
 
@@ -68,7 +75,7 @@ s2quant <- readRDS(file = "data/s2quant.rds") %>%
 prquant <- readRDS(file = "data/prquant.rds") %>%
   bind_rows(addnas)
 
-ggplot(s2quant, aes(year, median)) + 
+s2trends <- ggplot(s2quant, aes(year, median)) + 
   geom_line(alpha = 0.5) + geom_point(size = 1, alpha = 0.7) + 
   ylab(expression(paste(S[2]~"estimates", sep = ""))) +
   geom_errorbar(aes(ymin = q05, ymax = q95),width= 0.1) +
@@ -77,9 +84,13 @@ ggplot(s2quant, aes(year, median)) +
   labs(color = "River") +
   theme(legend.position = "none", 
         legend.text =  element_text(size = 12))
+s2trends
 ggsave("figures/s2-trends-faceted.png", height = 6, width = 5)
+s2trends + facet_wrap(~river_name, ncol = 3, scales = "free_y") 
+ggsave("figures/s2-trends-presentation.png", height = 4, width = 8)
 
-ggplot(prquant, aes(year, median)) + 
+
[…]

### 2020-03-12 ms/ms.tex
Add Ross's edits

@@
 
-Declines in wild Atlantic salmon (\emph{Salmo salar}) throughout the north
+Declines in wild Atlantic salmon (\emph{Salmo salar}) abundance throughout the north
 Atlantic are primarily attributed to declining survival at sea. This
@@
 Scotia (NS), Canada. We estimate survival in the first (\So) and second year at sea (\St),
-and the proportion returning as 1SW adults (\Pg). Trends in  were not consistent
+and the proportion returning as 1SW adults (\Pg). Trends were not consistent
 among rivers. Since 1990, \So increased at Western Arm Brook (NL)
-and Rivi\`{e}re Saint-Jean (QC), but declined at Conne River (NL) and Rivière
-de la Trinité (QC). Since the mid-1990s, \So increased at Campbellton River (NL),
+and Rivi\`{e}re Saint-Jean (QC), but declined at Conne River (NL) and Rivi\'{e}re
+de la Trinit\'{e} (QC). Since the mid-1990s, \So increased at Campbellton River (NL),
 declined at LaHave River (NS), and fluctuated at Nashwaak
@@
 expected \citep{Dempson2004, ICES2019}; some 
-have been assessed as considered threatened or endangered by the 
+have been assessed as threatened or endangered by the 
 Committee on the Status of Endangered Wildlife in Canada \citep[][]{Cosewic2010}, 
-Canada's national science advisory body (to the national government) on
+Canada's science advisory body (to the national government) on
 species risk of extinction.
@@
 causes]{Cairns2001}, including but not limited to: fishing mortality \citep{Dempson2004}, 
-damming of waterways and changes in the freshwater habitat \citep{Dunfield1985}, acidification
+damming of waterways and changes in the freshwater habitat \citep{Dunfield1985,Clarke2014a}, acidification
 \citep[particularly in the Southern Uplands region of
@@
 \citet{April2018}  for QC populations,
-\citet{Jones2014} for NB populations,
-and \citet{Gibson2009} for NS populations.
+\citet{Jones2014} for NB population,
+and \citet{Gibson2009} for NS population.
 
-Yearly return data are often recorded in terms of two size groups: small ($< 63$ cm
+Annual return data are often recorded in terms of two size groups: small ($< 63$ cm
 FL) and large ($\geq 63$ cm FL) salmon, as these closely represent different
@@
 spawners of different sizes. To correct for this in returns 
-reported as small and large salmon, we estimated the abundance
+reported as small and large salmon, rather than 1SW and 2SW fish, we estimated the abundance
 of 1SW and 2SW returns using yearly scale age data of a subsample of returns:
@@
 We developed a hierarchical Bayesian model that uses Murphy's maturity
-schedule method, in conjunction with informative priors, to estimate yearly
+schedule method, in conjunction with informative priors, to estimate annual
 marine survival in seven populations of Atlantic salmon. We account for
@@
 Saint-Jean, and Western Arm Brook populations showed increases in \So
-over time, Trinit\'{e}, Conne, and La Have populations showed decreases,
+over time, Trinit\'{e}, Conne, and LaHave populations showed decreases,
 while at Nashwaak there was an increase in median \So during the early
@@
 uniformly throughout the range of Atlantic salmon in the northwest Atlantic.
-Temporal trends are not consistent among populations. 
+Temporal trends are not consistent among all populations. 
 Over the time periods for which data were available, some rivers show positive trends in survival in the
@@
 % 2. How they compare to other publications
-Our results are contrary to those of \citet{Olmos2019}, who detected positive
+Our results are contrary to those of \citet{Olmos2019}, who detected mostly positive
 correlations in post-smolt survival among spatially broad stock units. These differences
@@
 spatial scales of data sources (i.e. river vs province scales), and perhaps most importantly the use stock-recruitment relationships
-rather than empirical smolt count data to estimate marine survival.
+rather than empirical smolt count estimates to estimate marine survival.
 As trends in marine survival during the first winter at sea are highly independent
@@
 were lower than those during the 1970s.
+\citet{Gibson2016} calculated higher return rates of 2SW for the Nashwaak in
+the 1970s than in the period since, with return rates of 1SW in the 1970s
+being comparable to those in the late 2000s.
 \citet{Friedland1993} compared return rates for a number of rivers in eastern
@@
 time series in our study.
-In any event, we caution that the pooling of adult return rates \citep{Chaput2012a, Friedland1993} 
+While declines in return rates since the 1970s seem to be consistent among populations, 
+we were unable to assess if marine survival estimates were also higher in the 1970s as we lacked 
+data from this decade.
+In any event, we caution that the pooling of adult return rates \citep{Chaput2012a, Friedland1993,Gibson2016} 
 can mask inter-annual variability in marine survival,
@@
 Marine survival in the first winter at sea could be highly variable between
-populations because of the predominance of spatially local environmental drivers of survival (e.g., temperature, predation) 
+populations because of the predominance of spatially local environmental drivers of survival (e.g., temperature, predation, prey abundance) 
 relative to broader-scale, even ocean-wide, drivers.
@@
 trends in abundance, then there are potential carry-on effects of oceanic
-conditions that manifest with regards to fresh production. 
+conditions that manifest with regards to freshwater production. 
 For example, adults
@@
 \citep{Barneche2018}, a small decrease in body condition resulting from bottom-up
-impacts on food availability could potentially have disproportionate effects on fecundity
+impacts on food availability and quality could potentially have disproportionate effects on fecundity
 and fitness of the offspring.
 
-Egg-to-smolt survival in Atlantic salmon is highly variable \citep{Klemetsen2003,Chaput2015}
-and changes in the oceanic conditions that spawners experience could be
+Egg-to-smolt survival in Atlantic salmon is highly variable \citep{Klemetsen2003,Chaput2015}.
+
+While this variability is likely attributable to changes in freshwater
+conditions (e.g. discharge, temperature, water quality) and uncertainty in
+spawner and smolt estimates, 
+changes in the oceanic conditions that spawners experience could also be
 contributing to this variability.
@@
 over decadal time scales \citep[e.g.,][]{Friedland1998, Todd2008,
-    Beaugrand2012}, it would be expected that this correlations would be
+    Beaugrand2012}, it would be expected that these correlations would be
 maintained even if salmon abundances were lagged by a generation length.
@@
 for modelling smolt abundances. This approach results in shrinkage to the
-mean, which means that the variability of yearly smolt estimates is less than
+mean, which means that the variability of annual smolt estimates is less than
 it would be if not modelled hierarchically. 

### 2020-03-12 analysis/05-figures-for-trends-paper.R
Fix moratoria label position

@@ -50,7 +50,7 @@ s1trends <- ggplot(s1quant, aes(year, median, color = river_name)) +
   theme_cowplot() + xlab("Year") + labs(color = "River") +
   theme(legend.position = "bottom") + guides(color = guide_legend(nrow = 3)) +
   geom_vline(data = morat, aes(xintercept = year, color = river_name), alpha = 0.4, linetype = 2) +
-  geom_text(data = morat, aes(year-1.5, 0.23, label = region, color = river_name))
+  geom_text(data = morat, aes(year-c(1.5, 1.5, -1.5), 0.23, label = region, color = river_name), show.legend = FALSE)
 s1trends
 ggsave("figures/s1-trends.png", height = 5, width = 7)
 

### 2020-03-12 analysis/05-corrected-covariances.R
Add corrected covariance and correlation analyses, helper functions, prior figures

@@ -1,0 +1,215 @@
+
+mu1var <- tibble(river_name = rep(unique(river_names), each = 7),
+                 correlation = rep(c("Var(logsmolts)/Var(mu)", 
+                                     "Var(Z1)/Var(mu)", 
+                                     "Var(logPr)/Var(mu)",
+                                     "2*Cov(logsmolts,-Z1)/Var(mu)",
+                                     "2*Cov(logsmolts,logPr)/Var(mu)",
+                                     "2*Cov(-Z1,logPr)/Var(mu)",
+                                     "mu sum"), 7),
+                 median = NaN,
+                 lowerCI = NaN,
+                 upperCI = NaN)
+
+mu1errorvar <- tibble(river_name = rep(unique(river_names), each = 7),
+                      correlation = rep(c("Var(logsmolts)", 
+                                          "Var(Z1)", 
+                                          "Var(logPr)",
+                                          "Var(mu)",
+                                          "Cov(logsmolts,-Z1)",
+                                          "Cov(logsmolts,logPr)",
+                                          "Cov(-Z1,logPr)"), 7),
+                      mean = NaN)
+
+mu1correctedcov <- tibble(river_name = rep(unique(river_names), each = 3),
+                 correlation = rep(c("Corrected cov(logsmolts,-Z1)",
+                                     "Corrected cov(logsmolts,logPr)",
+                                     "Corrected cov(-Z1,logPr)"), 7),
+                 median = NaN,
+                 lowerCI = NaN,
+                 upperCI = NaN)
+
+mu1correctedrsq <- tibble(river_name = rep(unique(river_names), each = 4),
+                          correlation = rep(c("R2(mu1,-Z1)",
+                                              "R2(mu1,logsmolts)",
+                                              "R2(mu1,logPr)",
[…]

### 2020-03-12 analysis/05-correlation-tests.R

@@ -2,7 +2,10 @@
 
 library(tidyverse)
 library(corrplot)
+library(cowplot)
 source("analysis/99-helper-functions.R")
+
+fithierncp <- readRDS("data/stanfit_hier_ncp.rds") # hierarchical
 
 alldatastan <- readRDS("data/alldatastan.rds")
 s1quant <- readRDS("data/s1quant.rds")
@@ -11,6 +14,440 @@ allreturns <- readRDS("data/allreturns.rds") %>%
   mutate(zscoresmolts = scale(log(smolts)), zscoretotal = scale(log(totalreturns))) %>%
   ungroup()
   
+estreturns <- tibble(river_name = rep(alldatastan$river_name, times = alldatastan$nyears),
+       year = alldatastan$years,
+       #logsmolts = alldatastan$logsmolts, 
+       log1SW = alldatastan$loggrilse,
+       log2SW = alldatastan$logSW2)
+
+river_names <- rep(alldatastan$river_name, times = alldatastan$nyears)
+
+propgrilse <- estreturns %>%
+  mutate(propgrilse = exp(log1SW)/(exp(log1SW)+exp(log2SW))) %>%
+  group_by(river_name) %>%
+  summarise(propgrilse = mean(propgrilse)) %>%
+  arrange(desc(propgrilse))
+
+s1post <- readRDS(file = "data/s1-posteriors.rds")
+s2post <- readRDS(file = "data/s2-posteriors.rds")
+z1post <- readRDS(file = "data/z1-posteriors.rds")
+z2post <- readRDS(file = "data/z2-posteriors.rds")
+prpost <- readRDS(file = "data/pr-posteriors.rds")
+logsmoltspost <- readRDS(file = "data/logsmoltstrue-posteriors.rds")
+
+# estimating mu1 and mu2 from posteriors
+mu1 <- logsmoltspost$value + log(prpost$value) - z1post$value
+mu2 <- logsmoltspost$value - z1post$value + log(1 - prpost$value)  - z2post$value
+
+z1mat <- as.matrix(fithierncp, pars = "Z1") 
+s1mat <- as.matrix(fithierncp, pars = "S1") 
+z2mat <- as.matrix(fithierncp, pars = "Z2") 
+prmat <- as.matrix(fithierncp, pars = "Pr")
+logsmoltsmat <- as.matrix(fithierncp, pars = "logsmolts_true")
+
[…]

### 2020-03-12 analysis/05-river-correlations.R

@@ -1,0 +1,184 @@
+library(tidyverse)
+library(corrplot)
+#source("analysis/99-helper-functions.R")
+
+fithierncp <- readRDS("data/stanfit_hier_ncp.rds") # hierarchical
+alldatastan <- readRDS("data/alldatastan.rds")
+river_names <- rep(alldatastan$river_name, times = alldatastan$nyears)
+unique_rivers <- unique(river_names) %>% sort
+
+z1mat <- as.matrix(fithierncp, pars = "Z1") 
+s1mat <- as.matrix(fithierncp, pars = "S1") 
+
+z1tib2 <- z1mat %>% t %>% as_tibble %>% mutate(river_name = river_names, year = alldatastan$years) %>%
+  select(river_name, year, everything())
+s1tib2 <- s1mat %>% t %>% as_tibble %>% mutate(river_name = river_names, year = alldatastan$years) %>%
+  select(river_name, year, everything())
+
+
+riverpos <- tibble(year = alldatastan$years, river_name = river_names) %>%
+  mutate(pos = row_number()) %>% select(pos, everything())
+
+# matchingyears <- intersect(filter(riverpos, river_name == "Saint-Jean River") %>% pull(year), 
+                           # filter(riverpos, river_name == "Western Arm Brook") %>% pull(year))
+
+riverpairs <- combn(unique_rivers %>% sort, 2, simplify = FALSE)
+
+river_cor <- function(river1, river2) {
+# river1 <- "Saint-Jean River"
+# river2 <- "Western Arm Brook"
+  
+  matchingyears <- intersect(filter(riverpos, river_name == river1) %>% pull(year), 
+                             filter(riverpos, river_name == river2) %>% pull(year))
+  
+  #browser()
+  z1river1 <- filter(z1tib2, river_name == river1 & year %in% matchingyears) %>% 
+    dplyr::select(-river_name, -year) 
+  z1river2 <- filter(z1tib2, river_name == river2 & year %in% matchingyears) %>% 
+    dplyr::select(-river_name, -year) 
+  
+  # pivot_longer(z1river1, cols = V1:V7500, names_to = "pos", names_prefix = "^V", values_to = "value") %>%
+  #   mutate(year = as.character(year), pos = as.numeric(pos)) %>% ungroup() %>%
+  #   ggplot(aes(pos, value, color = year)) + geom_line() + facet_wrap(~year)
+  
+  var1 <- mapply(var, z1river1)
[…]

### 2020-03-12 analysis/07-prior-figs.R

@@ -1,0 +1,63 @@
+library(tidyverse)
+library(cowplot)
+
+
+# S1 and S2 priors
+x <- seq(0.01,2,by=0.01)
+Z1 <- dlnorm(x, 1, 0.22)
+exp(-Z1)
+plot(Z1~exp(-Z1))
+
+Z2 ~ lognormal(0.2, 0.3);
+
+S1 <- seq(0.000,0.7, by = 0.001)
+
+Z1 <- dlnorm(-log(S1), 1, 0.22)
+Z2 <- dlnorm(-log(S1), 0.2, 0.3)
+plot(Z1~S1, type = "l", lwd = 2)
+lines(Z2/2~S1, lwd =2, col = "blue")
+
+s1df <- data.frame(Survival = S1,
+           Probability = Z1,
+           Parameter = "S1")
+s2df <- data.frame(Survival = S1,
+                   Probability = Z2/2,
+                   Parameter = "S2")
+
+df1 <- bind_rows(s1df, s2df) %>% as_tibble()
+
+# Pr priors
+
+x <- seq(0,1, by = 0.001)
+
+# Prior for Conne
+osw <- dlogis(qlogis(x), 2.4, 0.35)
+plot(osw/2~x, type = "l", col ="blue")
+
+# Prior for others
+tsw <- dlogis(qlogis(x), 0, 0.8)
+lines(tsw~x, col = "green")
+
+pr1 <- data.frame(Proportion = x,
+                  Probability = osw/2,
+                  Population = "1SW rivers")
+
+pr2 <- data.frame(Proportion = x,
+                  Probability = tsw,
+                  Population = "Other rivers")
+
+df2 <- bind_rows(pr1, pr2) %>% as_tibble() 
+
+
+prpriors <- ggplot(df2, aes(Proportion, Probability, linetype = Population)) + geom_line(size = 1.5) + 
+  theme_cowplot() + xlab("Proportion returning after first winter at sea") 
+prpriors
+ggsave("~/Dropbox/2020-ASCF-webinar/pr-priors.png", height = 4, width = 6)
+
+s1priors <- ggplot(df1, aes(Survival, Probability, color = Parameter)) + geom_line(size = 1.5) + 
+  theme_cowplot()
+s1priors
+ggsave("~/Dropbox/2020-ASCF-webinar/surv-priors.png", height = 4, width = 6)
+
[…]

### 2020-03-12 analysis/99-helper-functions.R

@@ -1,0 +1,28 @@
+
+# Non-centered parameterization comparison
+compare_cp_ncp <- function(cp_plot, ncp_plot, ncol = 2, ...) {
+  bayesplot_grid(
+    cp_plot, ncp_plot,
+    grid_args = list(ncol = ncol),
+    subtitles = c("Centered parameterization",
+                  "Non-centered parameterization"),
+    ...
+  )
+}
+
+# function from http://www.sthda.com/english/wiki/visualize-correlation-matrix-using-correlogram
+# to estimate p-values for all correlations in matrix
+cor.mtest <- function(mat, ...) {
+  mat <- as.matrix(mat)
+  n <- ncol(mat)
+  p.mat <- matrix(NA, n, n)
+  diag(p.mat) <- 0
+  for (i in 1:(n - 1)) {
+    for (j in (i + 1):n) {
+      tmp <- cor.test(mat[, i], mat[, j], ...)
+      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
+    }
+  }
+  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
+  p.mat
+}

### 2020-03-28 ms/ms.tex
Add Dustin's comments

@@
 causes]{Cairns2001}, including but not limited to: fishing mortality \citep{Dempson2004}, 
-damming of waterways and changes in the freshwater habitat \citep{Dunfield1985,Clarke2014a}, acidification
-\citep[particularly in the Southern Uplands region of
-NS, see][]{Gibson2010}, predation by seals and birds \citep{Cairns2000}, negative
+damming of waterways and changes in the freshwater habitat \citep{Dunfield1985,Clarke2014a}, 
+acidification \citep[particularly in Nova Scotia's Southern Uplands, see][]{Gibson2010}, 
+predation by seals and birds \citep{Cairns2000}, negative
 effects of interbreeding or interactions with escaped farmed salmon
@@
 driving population dynamics primarily through changes in marine survival.
-An implicit assumption is that any trend in
-survival in ocean habitat that is shared by multiple populations during their
-seaward migration period will be experienced similarly. 
+Thus, an implicit assumption is that populations sharing a particular oceanic
+route during seaward migration will also show similar trends in marine
+survival. 
 Put another way, given that salmon from different rivers 
@@
 likely to be assessed as being of conservation concern than those in more
-northerly regions \citep{Cosewic2010}. This geographical disparity in status
+northerly regions \citep{Cosewic2010}. This latitudinal disparity in status
 suggests that if marine survival has been, or is, a key factor responsible for
@@
 had experienced declining return rates. 
-However, examination of trends in return rates alone
-can mask changes in differential survival during different years at sea, as well
-as changes in the proportion of adults returning after one or two years at sea.
+However, examination of trends in return rates alone can mask changes in
+differential survival during different years at sea, as well as changes in the
+proportion of adults returning after one or two years at sea.
 Recently \citet{Olmos2019} suggested that trends in post-smolt
@@
 wide range of the species' distribution (Fig.~\ref{fig:map}). 
-Populations included the LaHave River in the Southern Uplands region of Nova
-Scotia (NS), Nashwaak River, New Brunswick (NB), Rivi\`{e}re de la Trinit\'{e} (Trinit\'{e}) and
-Saint-Jean rivers in Qu\'{e}bec (QC), and  Western Arm Brook (also referred to as  WAB), Campbellton, and
-Conne River, Newfoundland (NL). 
+Populations included the LaHave River in Nova Scotia's (NS) Southern Uplands, 
+Nashwaak River, New Brunswick (NB), Rivi\`{e}re de la Trinit\'{e} (Trinit\'{e}) and
+Saint-Jean rivers in Qu\'{e}bec (QC), and  Western Arm Brook (also referred to as  WAB), 
+Campbellton, and Conne River, Newfoundland (NL). 
 Data were collected in NS, NB, and Newfoundland
@@
 Smolt and adult return abundance estimates originate from a variety of
-sources. Smolt estimates from the Trinit\'{e}, Saint-Jean, and Conne populations were obtained using a
-mark-recapture approach, while estimates from the ..., WAB, and Campbellton populations
-were obtained by direct counts using fish counting fences.
+sources. Smolt estimates from the Trinit\'{e}, Saint-Jean, LaHave, Nashwaak,
+and Conne populations were obtained using a mark-recapture approach, while
+estimates from the WAB and Campbellton populations were obtained by direct
+counts using fish counting fences.
 For further details on the data collection methodologies refer to 
@@
     \caption{Posterior estimates of the population-level parameters $Pr_{\mu}$
-       $logit(Pr_{\mu})$, and $logit(Pr_{\sigma})$. Dots denote median estimates, while the thick and thing error bars indicate
+       $logit(Pr_{\mu})$, and $logit(Pr_{\sigma})$. Dots denote median estimates, while the thick and thin error bars indicate
        the 50\% and 90\% credible intervals, respectively.} 
@@
 and hence might not produce an accurate depiction of marine survival trends.
+While river-specific return rate estimates are available for the populations
+used in this study \citep{ICES2019}, these estimates are only an approximation
+of marine survival with varying degrees of similarity depending on each
+population's life histories. The estimates of return rates of one sea-winter
+should approximate marine survival in the first year at sea for one
+sea-winter-dominated populations. However, as the proportion of fish that
+return as one sea-winter decreases, marine survival in the first year at sea
+can increasingly diverge from return rates of fish, and trends in return rate
+would be particularly misleading if the proportion returning as one sea-winter
+varies in time. 
+Furthermore, return rates of two sea-winter fish also include survival in the
+first winter at sea and the proportion returning as two sea-winter, thus
+providing a very coarse estimate of marine survival in the second year at sea. 
+In fact, changes in return rates of two sea-winter fish are a result of a
+combination of changes in survival in the first year at sea in the year prior,
+survival in the second year at sea, and in the proportion returning as two
+sea-winter, and an increase in any of those parameters while the others remain
+the same will result in an increase in return rates. Our model attempts to
+improve some of the shortcomings of using return rates as a proxy of marine
+survival by directly estimating marine survival in the first and second years
+at sea, as well as the proportion returning as 1SW.
+
 \citet{Dempson2003} described a general declining trend in marine survival for
@@
 Oceanic conditions have been correlated with abundance trends and growth
-\citep{Todd2008}, however, the mechanism by which such bottom-up effects, \
+\citep{Todd2008}, however, the mechanism by which such bottom-up effects,
 mediated by changes in food availability,
@@
 \citep{Barneche2018}, a small decrease in body condition resulting from bottom-up
-impacts on food availability and quality could potentially have disproportionate effects on fecundity
-and fitness of the offspring.
+impacts on food availability and quality could potentially have
+disproportionate effects on fecundity and fitness of the offspring.
+
+% Furthermore, freshwater carryover effects acting on smolts, such as pH effects
+% on prey-avoidance behaviour, could potentially mediate post-smolt survival
+% \citep{Halfyard2012}.
 
 Egg-to-smolt survival in Atlantic salmon is highly variable \citep{Klemetsen2003,Chaput2015}.
-
 While this variability is likely attributable to changes in freshwater

### 2020-03-30 ms/ms.tex
Incorporate Gerald's comments

@@
 had experienced declining return rates. 
-However, examination of trends in return rates alone can mask changes in
+However, return rates cannot directly be interpreted as marine survival rates and 
+examination of trends in return rates alone can mask changes in
 differential survival during different years at sea, as well as changes in the
-proportion of adults returning after one or two years at sea.
+proportion of adults returning after one or two years at sea \citep{Hubley2011}.
 Recently \citet{Olmos2019} suggested that trends in post-smolt
@@
 While some studies have previously used maturity-schedule models to estimate marine
-survival for a limited number of salmon populations \citep{Chaput2003b}, none
+survival for a limited number of salmon populations \citep{Chaput2003b,Hubley2011}, none
 have incorporated data extending over multiple decades, nor have they examined
@@
 Here, we develop a hierarchical Bayesian model that uses Murphy's maturity
-schedule method, in conjunction with informative priors, to estimate yearly
+schedule method \citep{Murphy1952}, in conjunction with informative priors, to estimate yearly
 marine survival in salmon. In addition to accounting for observation error in
@@
 and of spawning history $a$ (either 1SW or 2SW returns); $n_{r,t,s,a}$ is the
-number of samples in river $r$, year $t$, of spawning history $a$, and of size
+number of samples in river $r$, year $t$, of returning age $a$, and of size
 group $s$; $n_{r,t,s}$ is the total number of samples in river $r$, year $t$,
@@
 in river $r$, year $t$, and of size group $s$.
+While, ideally, we would account the uncertainty in the conversion between
+size group and returning age, this is being ignored by doing this conversion
+deterministically as shown above. However, sampling a parameter from a
+hypergeometric distribution is not possible in programming the model was
+written in (i.e. Stan).
 
@@
 However, this issue can be mathematically overcome, at least partially, by
-using informative priors for all three marine survival parameters in a
-Bayesian framework.
+using informative priors for the two marine survival parameters and the maturation
+parameter in a Bayesian framework.
 This method requires abundance estimates of smolts as well as abundance estimates
@@
 three or more winters at sea before returning to spawn for the first time.
-The model also assumes that mortality in the second winter at sea (\St)
-is additional to mortality in the first winter at sea in the previous year, 
+The model also assumes that mortality in the second winter at sea (\St) for 2SW returns
+is additive to mortality in the first winter at sea in the previous year, 
 and therefore does not account for differences in environmental conditions experienced
@@
 where $Z_{1,r,t}$ and $Z_{2,r,t+1}$ are the instantaneous mortality rates.
-Process error was included as the standard deviation of the log-transformed
+Observation error was included as the standard deviation of the log-transformed
 return estimates from equation~\ref{eq:4}:
@@
 
-Furthermore, we use instantaneous mortality rates in the model instead of survival probabilities
-as the model is more efficient in its parameter search in log-space, and instantaneous rates
-are easy to interpret. Instantaneous rates are easily converted to survival probabilities by
+Furthermore, we use instantaneous mortality rates in the model instead of
+survival probabilities as the model is more efficient in its parameter search
+in log-space. Instantaneous rates are easily converted to survival
+probabilities by
 
@@
 
-We estimate population-level mean \Pg values around which the yearly \Pg
+The priors for $Z_1$ and $Z_2$ are specified as log-normal distributions:
+
+\begin{align}
+Z_1 &\sim logNormal(1, 0.22),   \\ 
+Z_2 &\sim logNormal(0.2, 0.3).
+\end{align}
+
+These priors, when converted to yearly survival, are roughly constrained between
+0 and 0.2 for \So and between 0.2 and 0.5 for \St (Fig.~\ref{fig:priors}a).
+
+We estimate population-level mean logit \Pg values around which the yearly \Pg
 values are normally distributed. We specify different informative hyperpriors
-for \prmu and \prsig based on whether the population is 1SW-dominated or not:
+for \prmu and \prsig based on whether the population is 1SW-dominated (i.e.
+with a proportion of 1SW fish in the total returns greater than 0.9) or not:
 
@@
 
-The priors for $Z_1$ and $Z_2$ are specified as log-normal distributions:
+These priors, when converted to proportion returning, are narrowly constrained
+for 1SW-dominated populations, but relatively wide for all other populations
+(Fig.~\ref{fig:priors}b).
 
-\begin{align}
-Z_1 &\sim logNormal(1, 0.22),   \\ 
-Z_2 &\sim logNormal(0.2, 0.3).
-\end{align}
+\begin{figure}[htbp] \centering
+    \includegraphics[width=0.95\linewidth]{figures/priors.png} \caption{Priors
+        for a) marine survival of one sea-winter (\So) and two sea-winter
+        (\St) returns, and b) probability of returning as 1SW for
+        1SW-dominated rivers (proportion 1SW > 0.9) and other rivers.}
+    \label{fig:priors} \end{figure}
 
@@
     \caption{Posterior estimates of \Pg for the seven rivers examined, error
-        bars indicate the 90\% credible intervals.} \label{fig:pr-faceted}
-\end{figure}
+        bars indicate the 90\% credible intervals. Year corresponds to the
+        year of 1SW return.} \label{fig:pr-faceted} \end{figure}
 

### 2020-04-25 ms/ms.tex
Add Methods and Results for correlation estimation

@@
 
-where $Z_{1,r,t}$ and $Z_{2,r,t+1}$ are the instantaneous mortality rates.
-Observation error was included as the standard deviation of the log-transformed
-return estimates from equation~\ref{eq:4}:
+where $Z_{1,r,t}$ and $Z_{2,r,t+1}$ are the instantaneous mortality rates and
+$R_{1,r,t}$ and $R_{2,r,t+1}$ are the estimated 1SW and 2SW returns in
+consecutive years, respectively. Observation error was included as the
+standard deviation of the log-transformed return estimates from
+equation~\ref{eq:4}:
 
@@
 \begin{figure}[htbp] \centering
-    \includegraphics[width=0.95\linewidth]{figures/priors.png} \caption{Priors
+    \includegraphics[width=0.55\linewidth]{figures/priors.png} \caption{Priors
         for a) marine survival of one sea-winter (\So) and two sea-winter
@@
         1SW-dominated rivers (proportion 1SW > 0.9) and other rivers.}
-    \label{fig:priors} \end{figure}
+    \label{fig:priors} 
+\end{figure}
 
@@
 
-\subsection*{Correlations among trends in survival}
+\subsection*{Correlations among trends in survival and parameters}
 
@@
 
-We looked at the correlation of trends in \So by calculating the Pearson's
-correlation coefficient for the Z-scores of these trends. Given that the time
-series do not cover the same years, and that some rivers have missing years in
-the middle of the time series, pair-wise Pearson's tests were done using only
-the years where there is data for both rivers.
+We looked at the correlation among trends in \So between rivers by directly estimating the 
+correlation for each posterior iteration and then calculating the distribution of correlation values:
+
+\begin{equation}
+cor(Z_{1,r=1},Z_{1,r=2}) = \frac{cov(Z_{1,r=1}, Z_{1,r=2})}{\sqrt{var(Z_{1,r=1})}*\sqrt{var(Z_{1,r=2})}}\label{eq:corz1}
+\end{equation}
+
+where $Z_{1,r=1}$ and $Z_{1,r=2}$ are the $Z_1$ estimates, across years, for a
+single posterior iteration for rivers 1 and 2. 
+Given that the time series do not cover the same years, and that some rivers
+have missing years in the middle of the time series, only years with overlapping
+$Z_1$ estimates were used for the pairwise correlation estimates.
+We also estimated the correlation between the model parameters for each population using
+a similar approach as in equation~\ref{eq:corz1}. However we error-corrected the correlation estimates
+by including the error in the variance and covariance estimates.
+To get the error corrected correlations we need the error covariance ($cov_{error}$) and error
+variances ($var_{error}$). Estimates of these can be obtained by estimating the covariance of
+two parameters across all values in the posterior for a given year (rather
+than across years for each posterior iteration), which results in one
+covariance estimate for each year in each river, which then are averaged
+across years for each river. Likewise, the error variance of a parameter is estimated  
+across all values in the posterior for a given year and then are averaged
+across years for each river:
+
+\begin{equation}
+cor_{corrected}(P_{g,r},-Z_{1,r}) = \frac{cov(P_{g,r}, -Z_{1,r}) - cov_{error}(P_{g,r}, -Z_{1,r})}
+{\sqrt{var(P_{g,r})- var_{error}(P_{g,r})}*\sqrt{var(-Z_{1,r}) - var_{error}(-Z_{1,r})}}\label{eq:corparam}
+\end{equation}
+
+Lastly, to determine what parameters best explains the number of returns in
+the model (including the model uncertainty), we estimated a pseudo-$R^2$ by
+calculating the squared correlation between the estimated parameters and the
+estimated 1SW and 2SW returns ($R_{1,r,t}$ and $R_{2,r,t+1}$ in
+equation~\ref{eq:4}).
+
+% \begin{equation}
+% cor_{error}(P_{g,r},-Z_{1,r}) = \frac{cov_{error}(P_{g,r},-Z_{1,r})}{\sqrt{var_{error}(P_{g,r})}*\sqrt{var_{error}(-Z_{1,r})}}
+% \end{equation}
+
+% However correlations are not additive so to get the error corrected correlation you have to calculate
 
@@
         bars indicate the 90\% credible intervals. Year corresponds to the
-        year of 1SW return.} \label{fig:pr-faceted} \end{figure}
+        year of 1SW return.} \label{fig:pr-faceted} 
+\end{figure}
 
@@
 
-Most correlations between z-scored trends were not statistically significant, with only 
-three out of the 21 pair-wise comparisons having a p-value below 0.01 (Fig.~\ref{fig:s1-corr}a).
-When looking at the direction of the correlation, regardless whether they were
-significant or not, these spanned both positive and negative coefficients
-(Fig.~\ref{fig:s1-corr}b).
+Most correlations between trends were significant, with only five out of the
+21 pair-wise comparisons not having a p-value below 0.05
+(Fig.~\ref{fig:s1-corr}). When looking at the direction of the correlations,
+these spanned both positive and negative coefficients. All seven populations had both
+positive and negative correlations among them.
 
 \begin{figure}[htbp] \centering
-    \includegraphics[width=0.95\linewidth]{figures/corr-s1.png} \caption{
-        Correlation among Z-scores of median estimated trends in \So among
-        rivers. a) Pearson's correlation coefficients are shown in each square,
-        while colouring denotes significance of the correlation ($p \leq 0.01$), b)
-        colours denote direction and magnitude of correlation, while asterisks denote significance.}
+    \includegraphics[width=0.95\linewidth]{figures/corrplot-post-z1-corrected.png} \caption{
+        Correlations of estimated trends in $Z_1$ among rivers. 
+        Correlation coefficients are shown in each square, while colouring 
+        denotes significance of the correlation ($p \leq 0.05$) as well as the direction and magnitude.}
 \label{fig:s1-corr} 
+\end{figure}
+
+As is expected from Murphy's method, there is a strong negative correlation
+between survival in the second year at sea ($-Z_2$) and proportion returning
+as 1SW ($log(P_g)$, Fig.~\ref{fig:cor-params}). Some populations (Campbellton,
+LaHave, Nashwaak, and Saint-Jean) show a negative correlation between the
+estimated number of smolts and survival in the first year at sea, while the
+others fluctuate around zero. There is overall a positive correlation between
+survival in the first year (i.e. $-Z_1$) and proportion returning as 1SW,
+except for the Campbellton, where this correlation is negative.
+
+\begin{figure}[htbp] \centering
+    \includegraphics[width=0.95\linewidth]{figures/corrected-cor-ALL.png}
+    \caption{Estimated correlations among estimated model parameters.}
+    \label{fig:cor-params} 
+\end{figure}
+
+Overall $Z_1$ explained most of the variance in estimated 1SW returns they had
+the highest $R^2$ values (with the exception of the Saint-Jean,
+Fig.~\ref{fig:mu-rsq}).
+The low $R^2$ for the Saint-Jean is likely due to the negative correlation
+between survival and number of outmigrating smolts (Fig.~\ref{fig:cor-params}). 
+Populations with a lower proportion of 1SW in their returns (e.g. Trinit\'{e}
+and Saint-Jean) have a higher $R^2$ with estimated number of smolts than those
+that are 1SW-dominated, but overall, smolts estimates had low $R^2$ values in 
+all populations except the Saint-Jean (Fig.~\ref{fig:mu-rsq}).
+The proportion returning as 1SW explained a moderate amount of the variance in
+estimated 1SW returns in the Conne, LaHave, and Nashwaak, and much less in the
+other four populations.
+
+Values of $R^2$ were overall low for all parameters with regards to explaining
+the variance of estimated 2SW returns. Estimated pf $Z_2$ explained little
+variance in estimates for 2SW returns, with the Saint-Jean again being the
+exception.
+
+\begin{figure}[htbp] \centering
+    \includegraphics[width=0.95\linewidth]{figures/rsquared-mu-params.png}
+    \caption{Estimated $R^2$ values for the correlation between estimated
+        returns ($R_{1,r,t}$ and $R_{2,r,t+1}$) and the model parameters used
+        for their estimation (i.e. estimated smolts, mortality in the first
+        year at sea $Z_1$, mortality in the second year at sea $Z_2$, and
+        proportion returning as 1SW, \Pg). Note we did not estimate $R^2$
+        values between $R_{1,est}$ and $Z_2$ as the latter is not used to
+        estimate the former, and that only the non-1SW dominated rivers are
+        shown in the $R^2$ values of the model parameters and $R_{2,est}$.}
+    \label{fig:mu-rsq} 
 \end{figure}
@@
 relative to broader-scale, even ocean-wide, drivers.
-The synchrony reported for marine survival trends at broader spatial scales \citep{Olmos2019}
+The synchrony reported for marine survival trends at broader spatial scales \citep{Massiot-Granier2014,Olmos2019}
 might be attributable to the use of stock-recruitment relationships to estimate survival,
@@
 Atlantic salmon \citep{Jonsson1998,Gibson2006}, there could potentially be
-some density-dependent processes during parts of the post-smolt migration
-period, particularly for populations that are likely to be subjected to
+some density-dependent processes during parts of the marine phase, 
+particularly for populations that are likely to be subjected to
 declining per capita population growth rates ($r$) generated by Allee effects.
@@
 
+Thirdly, the accuracy and precision of the model's output could be different
+for different populations. For example, we know that survival in the first
+year at sea is an important driver of our estimates of 1SW returns, except for
+the Saint-Jean (Fig.~\ref{fig:mu-rsq}). Thus, estimates of \So for this
+population, which has the lowest proportion of 1SW returns out of all seven
+populations examined, are likely to be less accurate than those for the other
+populations.
+Furthermore, estimates of \St and \Pg are negatively correlated and highly uncertain,
+and this must be considered when making any inferences based on these parameters.
+
 % 5. Future directions
@@
 But the fact that it has remained at roughly similar levels as it was
-previously \emph{despite} reduced commercial fishing mortalities, suggests
+since the 1980s \emph{despite} reduced commercial fishing mortalities, suggests
 that there may well be an interaction between small population size (small
@@
 
-% We would like to thank Geir Bolstad, G\'{e}rald Chaput, Brian Dempson, and
-% Martha Robertson for their useful discussions on estimating marine survival
-% in Atlantic salmon, and Amanda Kissel for her helpful comments on the
-% manuscript. Brian Dempson and Geoff Venoit for providing the data Conne
-% River data.
-We would like to thank G\'{e}rald Chaput for his useful discussions on
-estimating marine survival in Atlantic salmon, Carmen David for her comments
+We would like to thank G\'{e}rald Chaput for his useful discussions and
+comments on the manuscript, Carmen David for her comments
 on the manuscript, and Sean Anderson for his help with implementing the

### 2020-04-25 analysis/05-corrected-covariances.R
Update correlation and covariance tests

@@ -1,3 +1,4 @@
+# Corrected covariances between mu1 and estimated parameters 
 
 mu1var <- tibble(river_name = rep(unique(river_names), each = 7),
                  correlation = rep(c("Var(logsmolts)/Var(mu)", 
@@ -201,15 +202,288 @@ mu1correctedcov %>%
 ggsave("figures/corrected_cor_mu1.png", height = 4.5, width = 8)
 
 
-mu1var %>%
-  mutate(river_name = fct_relevel(river_name, propgrilse$river_name),
-         correlation = fct_relevel(correlation, "mu sum", after = Inf)) %>%
-  ggplot(aes(correlation, median, color = river_name)) + 
-  geom_hline(yintercept = 0, col = "gray80") +
-  geom_point(position = position_dodge(-0.5)) + 
-  geom_errorbar(aes(ymax = upperCI, ymin = lowerCI), width = 0.0, position = position_dodge(-0.5)) +
-  ylab("Contribution to Variance") + xlab("Parameter") + ggtitle("Correlation with mu_1") +
-  coord_flip() + theme_cowplot()
-ggsave("figures/correlatedvariancemu1.png", height = 4.5, width = 8)
+# mu1var %>%
+#   mutate(river_name = fct_relevel(river_name, propgrilse$river_name),
+#          correlation = fct_relevel(correlation, "mu sum", after = Inf)) %>%
+#   ggplot(aes(correlation, median, color = river_name)) + 
+#   geom_hline(yintercept = 0, col = "gray80") +
+#   geom_point(position = position_dodge(-0.5)) + 
+#   geom_errorbar(aes(ymax = upperCI, ymin = lowerCI), width = 0.0, position = position_dodge(-0.5)) +
+#   ylab("Contribution to Variance") + xlab("Parameter") + ggtitle("Correlation with mu_1") +
+#   coord_flip() + theme_cowplot()
+# ggsave("figures/correlatedvariancemu1.png", height = 4.5, width = 8)
 
+
+
+mu2var <- tibble(river_name = rep(unique(river_names), each = 11),
+                 correlation = rep(c("Var(logsmolts)/Var(mu)", 
+                                     "Var(Z1)/Var(mu)", 
+                                     "Var(Z2)/Var(mu)", 
+                                     "Var(logPr)/Var(mu)",
[…]

### 2020-04-25 analysis/05-correlation-tests.R

@@ -142,6 +142,78 @@ for (i in unique(river_names)) {
   mu2out[mu2out$river_name == i & mu2out$correlation == "mu2 sum", 5] <- quantile(rsqsum2, 0.95)  
 }
 
+### PLOTTING PSEUDO R^2s
+
+print(mu1out, n = 28)
+print(mu2out, n = 28)
+
+mu1out %>%
+  mutate(param = word(correlation, 2)) %>%
+  mutate(river_name = fct_relevel(river_name, propgrilse$river_name)) %>%
+  filter(correlation != "mu sum") %>%
+  ggplot(aes(param, median, color = river_name)) + 
+  geom_point(position = position_dodge(-0.4)) + 
+  geom_errorbar(aes(ymax = upperCI, ymin = lowerCI), width = 0.0, position = position_dodge(-0.4)) +
+  ylab("R squared estimates") + xlab("Parameter") + ggtitle("Correlation with R_1") +
+  coord_flip() + theme_cowplot()
+
+
+mu2out %>%
+  mutate(param = word(correlation, 2)) %>%
+  mutate(river_name = fct_relevel(river_name, propgrilse$river_name)) %>%
+  filter(correlation != "mu sum") %>%
+  filter(!river_name %in% c("Campbellton River", "Conne River", "Western Arm Brook")) %>%
+  ggplot(aes(param, median, color = river_name)) + 
+  geom_point(position = position_dodge(-0.4)) + 
+  geom_errorbar(aes(ymax = upperCI, ymin = lowerCI), width = 0.0, position = position_dodge(-0.4)) +
+  ylab("R squared estimates") + xlab("Parameter") + ggtitle("Correlation with R_2") +
+  coord_flip() + theme_cowplot()
+
+param_names <- c("mu" = "italic(R)['1,est']", "mu2" = "italic(R)['2,est']")
+
+bind_rows(mu1out, mu2out) %>%
+  mutate(ret = word(correlation, 1), param = word(correlation, 2)) %>%
+  mutate(river_name = fct_relevel(river_name, propgrilse$river_name)) %>%
+  mutate(param = forcats::fct_relevel(param, rev(c("logsmolts", "Z1", "Z2", "Pr", "sum")))) %>%
+  # Need to replace with NA rather than filter out so the rivers align in both facets
+  mutate(median = ifelse((river_name %in% c("Campbellton River", "Conne River", "Western Arm Brook") & ret == "mu2"),
+NA, median)) %>%
[…]

### 2020-04-25 analysis/05-river-correlations.R

@@ -4,8 +4,22 @@ library(corrplot)
 
 fithierncp <- readRDS("data/stanfit_hier_ncp.rds") # hierarchical
 alldatastan <- readRDS("data/alldatastan.rds")
+
+estreturns <- tibble(river_name = rep(alldatastan$river_name, times = alldatastan$nyears),
+                     year = alldatastan$years,
+                     #logsmolts = alldatastan$logsmolts, 
+                     log1SW = alldatastan$loggrilse,
+                     log2SW = alldatastan$logSW2)
+
+propgrilse <- estreturns %>%
+  mutate(propgrilse = exp(log1SW)/(exp(log1SW)+exp(log2SW))) %>%
+  group_by(river_name) %>%
+  summarise(propgrilse = mean(propgrilse)) %>%
+  arrange(desc(propgrilse))
+
 river_names <- rep(alldatastan$river_name, times = alldatastan$nyears)
-unique_rivers <- unique(river_names) %>% sort
+#unique_rivers <- unique(river_names) %>% sort
+unique_rivers <- propgrilse$river_name
 
 z1mat <- as.matrix(fithierncp, pars = "Z1") 
 s1mat <- as.matrix(fithierncp, pars = "S1") 
@@ -22,7 +36,7 @@ riverpos <- tibble(year = alldatastan$years, river_name = river_names) %>%
 # matchingyears <- intersect(filter(riverpos, river_name == "Saint-Jean River") %>% pull(year), 
                            # filter(riverpos, river_name == "Western Arm Brook") %>% pull(year))
 
-riverpairs <- combn(unique_rivers %>% sort, 2, simplify = FALSE)
+riverpairs <- combn(unique_rivers, 2, simplify = FALSE)
 
 river_cor <- function(river1, river2) {
 # river1 <- "Saint-Jean River"
@@ -152,7 +166,7 @@ corrplot(medcor, method = "color", col = col(200),
          mar = c(0, 0, 1.8, 0),
          title = expression(paste("Corrected correlation of posterior estimates of ",italic(Z[1])))
          )
-dev.off()
+  dev.off()
 
 png("figures/corrplot-post-z1-uncorrected.png", width = 4, height = 3, units = "in", res = 300, pointsize = 7)
 corrplot(medcor_un, method = "color", col = col(200),  

### 2020-05-22 analysis/04-extract-posteriors.R
Update figures for draft

@@ -141,6 +141,24 @@ logitprcvquant <- logitprmupost %>%
   group_by(river_name) %>%
   arrange(param, as.numeric(pos))
 
+# Calculating CV in Pr
+# prcvquant <- prmupost %>%
+#   rename(Prmu = value) %>%
+#   mutate(Prsigma = prsigmapost$value,
+#          logitPrCV = logitPrsigma/abs(logitPrmu)) %>%
+#   select(pos, river_name, value = logitPrCV) %>%
+#   mutate(param = "logitPr_CV") %>%
+#   group_by(param, pos, river_name) %>%
+#   summarise(median = median(value),
+#             q05 = quantile(value, 0.05),
+#             q25 = quantile(value, 0.25),
+#             q75 = quantile(value, 0.75),
+#             q95 = quantile(value, 0.95)) %>%
+#   group_by(river_name) %>%
+#   arrange(param, as.numeric(pos))
+
+
+
 
 prmuquant <- bind_rows(logitprmupost, logitprsigmapost, prmupost) %>%
   group_by(param, pos, river_name) %>%

### 2020-05-22 analysis/05-correlation-tests.R

@@ -27,6 +27,7 @@ propgrilse <- estreturns %>%
   group_by(river_name) %>%
   summarise(propgrilse = mean(propgrilse)) %>%
   arrange(desc(propgrilse))
+saveRDS(propgrilse, file = "data/propgrilse.rds")
 
 s1post <- readRDS(file = "data/s1-posteriors.rds")
 s2post <- readRDS(file = "data/s2-posteriors.rds")
@@ -74,6 +75,11 @@ mu2out <- tibble(river_name = rep(unique(river_names), each = 5),
                  lowerCI = NaN,
                  upperCI = NaN)
 
+muallout <- tibble(river_name = rep(unique(river_names), each = 5),
+                 correlation = rep(c("muall Z1", "muall Z2", "muall Pr", "muall logsmolts", "muall sum"), 7),
+                 median = NaN,
+                 lowerCI = NaN,
+                 upperCI = NaN)
 
 # Estimating correlation squared among parameters 
 for (i in unique(river_names)) {
@@ -106,14 +112,13 @@ for (i in unique(river_names)) {
   mu1out[mu1out$river_name == i & mu1out$correlation == "mu sum", 5] <- quantile(rsqsum, 0.95)  
 }
 
-
 for (i in unique(river_names)) {
   z1sub <- filter(z1tib, river_name == i) %>% select(-river_name)
   z2sub <- filter(z2tib, river_name == i) %>% select(-river_name)
   prsub <- filter(prtib, river_name == i) %>% select(-river_name) 
   logsmoltssub <- filter(logsmoltstib, river_name == i) %>% select(-river_name) 
   mu2sub <- filter(mu2tib, river_name == i) %>% select(-river_name)
- # print(head(mu1sub))
+  # print(head(mu1sub))
   mu2z1 <- mapply(cor, mu2sub, z1sub)^2 
   mu2z2 <- mapply(cor, mu2sub, z2sub)^2 
   mu2pr <- mapply(cor, mu2sub, log(1 - prsub))^2
@@ -142,10 +147,58 @@ for (i in unique(river_names)) {
   mu2out[mu2out$river_name == i & mu2out$correlation == "mu2 sum", 5] <- quantile(rsqsum2, 0.95)  
 }
 
+
+
+# estumating correlation using all returns (R1 + R2)
+
+for (i in unique(river_names)) {
+  z1sub <- filter(z1tib, river_name == i) %>% select(-river_name)
+  z2sub <- filter(z2tib, river_name == i) %>% select(-river_name)
[…]

### 2020-05-22 analysis/05-figures-for-trends-paper.R

@@ -2,6 +2,7 @@
 
 library(tidyverse)
 library(cowplot)
+library(xtable)
 
 s1post <- readRDS(file = "data/s1-posteriors.rds")
 s1quant <- readRDS(file = "data/s1quant.rds")
@@ -144,3 +145,117 @@ s1quant %>%
 s1quant %>%
   filter(median >= 0.17)
 
+
+# Plotting individual population level posteriors of pr based on logitpr mu and logitpr sigma
+
+fithierncp <- readRDS("data/stanfit_hier_ncp.rds") # hierarchical
+propgrilse <- readRDS("data/propgrilse.rds")
+alldatastan <- readRDS("data/alldatastan.rds")
+
+logitpr_mu <- as.matrix(fithierncp, pars = "logitPr_mu") %>% 
+  t %>% as_tibble %>% mutate(river_name = alldatastan$river_name) %>%
+  pivot_longer(V1:V7500, names_to = "iteration") %>%
+  select(river_name, everything()) %>%
+  group_by(river_name) %>%
+  rename(`Population` = river_name, logitPr_mu = value)
+
+logitpr_mu
+
+logitpr_mu_quant <- logitpr_mu %>%
+  rename(value = logitPr_mu) %>%
+  ungroup() %>%
+  mutate(Population = fct_relevel(Population, propgrilse$river_name)) %>%
+  group_by(Population) %>%
+  summarise(Median = median(value), 
+            `2.5%` = quantile(value, probs = 0.025), 
+            `25%` = quantile(value, probs = 0.25), 
+            `75%` = quantile(value, probs = 0.75), 
+            `97.5%` = quantile(value, probs = 0.975)
+            # n = n()
+  )
+
+
+logitpr_sigma <- as.matrix(fithierncp, pars = "logitPr_sigma") %>% 
+  t %>% as_tibble %>% mutate(river_name = alldatastan$river_name) %>%
+  pivot_longer(V1:V7500, names_to = "iteration") %>%
+  select(river_name, everything()) %>%
+  group_by(river_name) %>%
+  rename(`Population` = river_name, logitPr_sigma = value)
+
+logitpr_sigma
+
+logitpr_sigma_quant <- logitpr_sigma %>%
+  rename(value = logitPr_sigma) %>%
+  ungroup() %>%
+  mutate(Population = fct_relevel(Population, propgrilse$river_name)) %>%
+  group_by(Population) %>%
+  summarise(Median = median(value), 
[…]

### 2020-05-22 ms/suppmat.tex
Add Supplementary Materials

--- 2020-03-28 ms/ms.tex
@@
 \usepackage[top=0.85in,left=1.0in,right=1.0in,footskip=0.75in]{geometry}
-%\usepackage[parfill]{parskip}
+\usepackage[parfill]{parskip}
 \usepackage{setspace}
@@
 \usepackage[hidelinks]{hyperref}
-%\onehalfspacing
-\doublespacing
+\onehalfspacing
 
-\usepackage[round,sectionbib]{natbib}
+\usepackage[round]{natbib}
 \setcitestyle{authoryear}
 \bibpunct{(}{)}{;}{a}{}{;}
-\bibliographystyle{fishfishnourl}
-
-%\usepackage{textcomp}
-%\usepackage{libertine}
-%\usepackage{inconsolata} % sans serif typewriter
-
-\usepackage{mathtools} % for dcases
-\usepackage{xcolor} % for textcolor
-\usepackage{makecell} % for \makecell and within cell line breaks (\thead)
-
-\usepackage[T1]{fontenc}
-
-%\makeatletter\let\expandableinput\@@input\makeatother % expandable input for \input inside tables
 
@@
 \usepackage[libertine,bigdelims,vvarbb]{newtxmath} % bb from STIX
-\usepackage[vvarbb]{newtxmath} % bb from STIX, removed bigdelims for ScholarOne rendering
 \usepackage[cal=boondoxo]{mathalfa} % mathcal
-%\useosf % osf for text, not math
-%\usepackage[supstfm=libertinesups,%
-%  supscaled=1.2,%
-%  raised=-.13em]{superiors}
+\useosf % osf for text, not math
+\usepackage[supstfm=libertinesups,%
+  supscaled=1.2,%
+  raised=-.13em]{superiors}
+
+\usepackage[utf8]{inputenc}
 
@@
 \usepackage{amssymb} % for \checkmark 
-\usepackage{amsmath} % for \checkmark 
 \usepackage{rotating}
-\usepackage[nolists,tablesfirst]{endfloat} % for putting figs and tables at end of document
-\DeclareDelayedFloatFlavor{sidewaystable}{table}
+%\usepackage[nolists,tablesfirst]{endfloat}
+%\DeclareDelayedFloatFlavor{sidewaystable}{table}
 \usepackage{makecell} % for \makecell in tables
-\usepackage{doi}
 
-%\usepackage{xr} % to obtain label references from supp materials file
-%\externaldocument[S-]{suppmat}
+% omit bibliography (\nobibliography{...})
+\usepackage{bibentry}
 
@@
 \newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-%\usepackage[detect-all]{siunitx} % for SI units
+\usepackage[detect-all]{siunitx} % for SI units
+
+\makeatletter\let\expandableinput\@@input\makeatother % expandable input for \input inside tables
+\setlength{\LTcapwidth}{\textwidth}
 
 % custom hyphenation:
-\hyphenation{inverse}
-\hyphenation{At-lantic}
 \hyphenation{elasmo-branchs}
@@
 \newcommand{\Pg}{$P_r$\xspace}
-\newcommand{\prmu}{$\mu_r$\xspace}
-\newcommand{\prsig}{$\sigma_r$\xspace}
 \newcommand{\Linf}{$L_{\infty}$}
@@
 \newcommand{\rsq}{$R^2$\xspace}
-\newcommand{\mytilde}{\raise.17ex\hbox{$\scriptstyle\mathtt{\sim}$}}
 % Select what to do with command \comment:  
-%\newcommand{\comment}[1]{}  % comment not shown
-\newcommand{\comment}[1]{\par {\bfseries \color{blue} #1 \par}} % comment shown
+\newcommand{\comment}[1]{}  % comment not shown
+%\newcommand{\comment}[1]{\par {\bfseries \color{blue} #1 \par}} % comment shown
 %% END MACROS SECTION
+
+%\bibliographystyle{model2-names}
+\bibliographystyle{fishfishnourl}
+
+%%% MY MACROS
+
+% Make supplementary tables and figures start with S1 and T2
+% use \beginsupplement to reset count
+\newcommand{\beginsupplement}{%
+        \setcounter{table}{0}
+        \renewcommand{\thetable}{S\arabic{table}}%
+        \setcounter{figure}{0}
+        \renewcommand{\thefigure}{S\arabic{figure}}%
+     }
 
@@
 
-\section*{Trends in marine survival of Atlantic salmon in eastern Canada}
+\noindent
+{\Large Supplementary Materials}
+\newline
 
-\textbf{Sebasti\'{a}n A. Pardo\textsuperscript{1*}, 
-        Geir H. Bolstad\textsuperscript{2}, 
-        J. Brian Dempson\textsuperscript{3}, 
-        Julien April\textsuperscript{4}, 
-        Ross A. Jones\textsuperscript{5}, 
-        Martha J. Robertson\textsuperscript{3}, 
-        Dustin Raab\textsuperscript{6}, 
-Jeffrey A. Hutchings\textsuperscript{1}}
+\noindent
+{\Large for}
+\newline
 
-\noindent\small{\textsuperscript{1} Department of Biology, Dalhousie University, Halifax, NS, Canada\\}
+\noindent
+{\large Trends in marine survival of Atlantic salmon populations in eastern Canada}
+\newline
+
+\noindent
+Sebasti\'{a}n A. Pardo\textsuperscript{1*}, Geir H. Bolstad\textsuperscript{2}, J. Brian Dempson\textsuperscript{3}, 
+        Julien April\textsuperscript{4}, Ross A. Jones\textsuperscript{5}, %Martha J. Robertson\textsuperscript{3}, 
+        Dustin Raab\textsuperscript{6}, Jeffrey A. Hutchings\textsuperscript{1}
+\newline
+
+\noindent
+\small{\textsuperscript{1} Department of Biology, Dalhousie University, Halifax, NS, Canada\\}
 \small{\textsuperscript{2} Norwegian Institute for Nature Research (NINA), Trondheim, Norway\\}
@@
 \small{\textsuperscript{*} Corresponding author: spardo@dal.ca}
+\beginsupplement
 
-\section*{Abstract}
+This Supplementary Material includes additional tables to accompany the publication.
 
-Declines in wild Atlantic salmon (\emph{Salmo salar}) abundance throughout the north
-Atlantic are primarily attributed to declining survival at sea. This
-hypothesis has proven challenging to test on a river-by-river basis because of
-the need to model data on both migrating smolts and returning adults and to
-simultaneously estimate multiple parameters, especially for salmon spending
-more than one winter at sea (1SW) before spawning. We fit a hierarchical Bayesian
-maturity schedule model to data (19-42 years) for seven populations in  
-Newfoundland and Labrador (NL), Qu\'{e}bec (QC), New Brunswick (NB), and Nova
-Scotia (NS), Canada. We estimate survival in the first (\So) and second year at sea (\St),
-and the proportion returning as 1SW adults (\Pg). Trends were not consistent
-among rivers. Since 1990, \So increased at Western Arm Brook (NL)
-and Rivi\`{e}re Saint-Jean (QC), but declined at Conne River (NL) and Rivi\'{e}re
-de la Trinit\'{e} (QC). Since the mid-1990s, \So increased at Campbellton River (NL),
-declined at LaHave River (NS), and fluctuated at Nashwaak
-River (NB). Estimates of \St were highly uncertain, particularly for
-1SW-dominated populations; \Pg was generally stable. These results challenge the
-narrative that marine survival has changed in a temporally consistent manner
-among spatially disparate populations. Our findings suggest that, at the
-population level, changes in abundance are attributable to temporal shifts in
-multiple components of individual fitness, including at-sea survival. If
-salmon populations do not respond in a consistent, uniform manner to changing
-environmental conditions throughout their range, future research initiatives
-should explore why.
+\clearpage
 
-% The marine phase of anadromous Atlantic salmon (\emph{Salmo salar}) is the
-% least known yet one of the most crucial with regards to population
-% persistence. Declines in many Atlantic salmon populations in eastern
-% Canada have often been attributed to changes in conditions at sea, negatively
-% affecting their survival. However, marine survival estimates are difficult to
-% obtain given that many individuals spend multiple winters in the ocean before
-% returning to freshwater to spawn, necessitating the estimation of multiple
-% parameters. To do so, we fit a hierarchical Bayesian maturity schedule model
-% to smolt and adult abundance time series for seven populations located in
-% Newfoundland and Labrador (NL), Qu\'{e}bec (QC), New Brunswick (NB), and Nova
-% Scotia (NS). The datasets ranged between 19 and 42 years in length. We
-% estimated three components of marine survival: survival in the first year at
-% sea (\So), survival in the second year at sea (\St), and proportion returning
-% as one sea-winter adults (\Pg). Controlling for time frame, trends in
-% estimates of \So were not consistent among rivers.
-% In the four populations for which data extended to 1990, marine survival
-% during the first year at sea predominantly increased over time in Western Arm
-% Brook (NL) and Rivi\`{e}re Saint-Jean (QC) but largely declined in Conne River
-% (NL) and Rivi\`{e}re Trinit\'{e} (QC). In the three other populations, \So
-% exhibited an increase in Campbellton River (NL), a decline in LaHave River
-% (NS), and fluctuating stability in Nashwaak River (NB) since approximately the
-% mid-1990s. 
-% Estimates of \St were highly uncertain, particularly for 1SW-dominated rivers
-% (Conne, Campbellton and WAB, where the abundance of 2SW returns is very low)
-% and thus the posterior distributions matched our choice of prior and trends
-% could not be assessed. 
-% \Pg was temporally stable within rivers, except for LaHave and Nashwaak Rivers. 
-% Our findings challenge the assumption that marine survival has changed in a consistent
-% manner among populations across a broad geographic scale, and suggest that
-% trends can be river-specific. 
-% The well established correlations between climate variables and abundance are
-% not being mediated solely by marine survival, and there can potentially be
-% some important indirect effects on fecundity.
-% Further work exploring potential correlates of marine survival and the
-% potential non-lethal effects of climate effects is warranted.
+\subsection*{Raw smolt data used in the model}
 
-Keywords: salmonid, survival at sea, natural mortality, marine mortality
+\input{smolts_tab.tex}
 
-%Running Head:
+\input{smoltscv_tab.tex}
 
-\section*{Introduction} % (4-5 paragraphs)
+\subsection*{Annual smolt abundances estimates derived from the Bayesian hierarchical model}
 
-%1. Declining salmon numbers
+%\begin{table}[ht]
+%\centering
+%\begin{longtable}{lrrrrrr}
+%\caption{Posterior annual smolt estimates for the seven populations examined. These
+%                    estimates differ from the smolt data used as they include both observation error and shrinkage from
+%                    the hierarchical model specification.} 
+%\hline
+\input{tableS1_smolts}
+%\end{longtable}
+%\end{table}
+\clearpage
 
-Reductions in fishing mortality, albeit necessary, are not always sufficient
-to facilitate population recovery. Experience with numerous commercially
-exploited marine fisheries since the early 1990s has shown that not all
-populations respond as favourably as anticipated to major reductions in
-exploitation \citep{Hutchings2017}. Gradual efforts to close commercial
-Atlantic salmon (\emph{Salmo salar}) fisheries in eastern Canada culminated in
-full moratoria in all regions, beginning in the Maritime provinces (1984) and
-following in Newfoundland (1992), Labrador (1998), and Qu\'{e}bec (2000). Since
-these closures, many populations have not increased as 
-expected \citep{Dempson2004, ICES2019}; some 
-have been assessed as threatened or endangered by the 
-Committee on the Status of Endangered Wildlife in Canada \citep[][]{Cosewic2010}, 
-Canada's science advisory body (to the national government) on
-species risk of extinction.
-While it is not fully understood what is driving population declines, the potential
-drivers of these are many \citep[see ][for a detailed discussion of possible
-causes]{Cairns2001}, including but not limited to: fishing mortality \citep{Dempson2004}, 
-damming of waterways and changes in the freshwater habitat \citep{Dunfield1985,Clarke2014a}, 
-acidification \citep[particularly in Nova Scotia's Southern Uplands, see][]{Gibson2010}, 
-predation by seals and birds \citep{Cairns2000}, negative
-effects of interbreeding or interactions with escaped farmed salmon
-\citep{Keyser2018}, and climate-driven changes in survival and productivity \citep{Mills2013}.
+\subsection*{Population-level estimates of proportion returning after one winter at sea}
 
-% 2.
-Over the past three decades, a narrative has emerged that marine survival of
-Atlantic salmon has declined throughout the North Atlantic \citep{ICES2019}.
-%\citep{Hansen1998,OMaoileidigh2003,Chaput2012a}.
-Based on multiple lines of evidence that climate conditions can directly and
-indirectly influence the abundance and productivity of Atlantic salmon
-populations \citep{Mills2013,Almodovar2019}, it has been presumed that oceanic climate effects are
-driving population dynamics primarily through changes in marine survival.
-Thus, an implicit assumption is that populations sharing a particular oceanic
-route during seaward migration will also show similar trends in marine
-survival. 
-Put another way, given that salmon from different rivers 
-are hypothesized to share marine habitat during some of their time at sea, it
-has been presumed that populations share similar temporal trends in
-at-sea mortality \citep{Friedland1993, Friedland1998, Russell2012}.  
-A recent study by \citet{Olmos2019} suggested that trends in post-smolt
-survival, when estimated at the stock level, are synchronously declining
-for all Atlantic salmon in eastern North America.
+\input{tableS2_smu}
 
-% This perception is widespread both in the scientific
-% literature as well as federal reports. The latest status report on Atlantic
-% salmon by COSEWIC states that ``While the mechanism(s) of marine mortality is
-% uncertain, what is clear is that the recent period of poor sea survival is
-% occurring in parallel with many widespread changes in the North Atlantic
-% ecosystem.'' \citep{Cosewic2010}. Consequently, there have several lines of
-% inquiry as to what the causes behind these declines might be attributable to
-% \citep{Friedland1993, Friedland1998}.
+\hspace{8em}
 
-In contrast to the narrative of widespread, demographically similar increases
-in at-sea mortality, the conservation status of Canadian salmon populations differs
-considerably. Populations in the southern part of their range are more
-likely to be assessed as being of conservation concern than those in more
-northerly regions \citep{Cosewic2010}. This latitudinal disparity in status
-suggests that if marine survival has been, or is, a key factor responsible for
-most population declines, these changes are not uniformly distributed across
-all populations.
+\input{tableS3_lsmu}
 
-% Nonetheless, most of the populations declines assessed by COSEWIC have
-% occurred in populations in the southern extent of its distribution:
-% populations in the Bay of Fundy, Anticosti Island, and the Atlantic coast of
-% Nova Scotia being assessed as Endangered, populations in the south coast of
-% Newfoundland assessed as Threatened, And the populations in the New Brunswick
-% and Qu\`{e}bec coasts of the Gulf of St. Lawrence assessed as Special Concern
-% \citep{Cosewic2010}. On the other hand, northermost populations have shown
-% stable, or even increasing population trends, suggesting that, if marine
-% survival were to be a factor in many of these declines, these changes are not
-% uniformly distributed across all populations.
+\input{tableS4_lssig}
 
-% While the purported decline of marine survival is mentioned widely in the 
-% scientific literature, only relatively few studies have quantified it in detail.
-% The basis for this premise is a limited number of time series that exhibit
-% temporal declines in a proxy of marine survival (i.e. return rates or
-% post-smolt survival).
+% \begin{figure}[htbp] \centering
+%     \includegraphics[width=0.9\linewidth]{simul_timeseries_declining.pdf} \caption{Simulated
+%         time series of returning adult salmon abundance in the six scenarios. The lines denote the
+%     simulated abundance estimates without observation error while the points are the same estimates
+% including observation error.}
+% \label{fig:simul-ts-dec} 
+% \end{figure}
 
-% 3. Nonetheless, this premise has never been examined in detail
-Given the logistical challenges associated with estimating at-sea survival, it
-is not surprising that the number of studies that have estimated temporal
-trends has been limited. An additional limitation has been the derivation of
-proxies (e.g., return rates), rather than direct model-based estimates, of
-marine survival.
-\citet{Chaput2012a}, for example, examined the return rate of smolts to adult salmon 
-as a metric of marine survival, finding that most Canadian populations 
-had experienced declining return rates. 
-However, examination of trends in return rates alone can mask changes in
-differential survival during different years at sea, as well as changes in the
-proportion of adults returning after one or two years at sea.
-Recently \citet{Olmos2019} suggested that trends in post-smolt
-survival, when estimated at the stock unit level, are synchronously declining
-for all Atlantic salmon in Eastern North America and Canada, a conclusion ultimately 
-grounded on the veracity of highly variable stock-recruitment relationships.
+% \begin{figure}[htbp] \centering
+%     \includegraphics[width=0.95\linewidth]{sir_out_18_declining.pdf} \caption{Yearly
+%         estimated \So, \St, and \Pg values in the six scenarios. True values
+%         are denoted by blue circles, black circles show median estimates,
+%         error bars indicate the 25\% and 75\% quantiles.}
+% \label{fig:sir-out-dec} 
+% \end{figure}
 
-%4. 
-In the present study, we compile data on the number of migrating smolts and number of returning adults 
-for seven wild Canadian populations of Atlantic salmon to model trends in marine survival.
-While some studies have previously used maturity-schedule models to estimate marine
-survival for a limited number of salmon populations \citep{Chaput2003b}, none
-have incorporated data extending over multiple decades, nor have they examined
-trends among more than two or three populations. 
-Here, we develop a hierarchical Bayesian model that uses Murphy's maturity
-schedule method, in conjunction with informative priors, to estimate yearly
-marine survival in salmon. In addition to accounting for observation error in
-smolt and return estimates, we estimate the proportion of salmon returning
-after one winter hierarchically.
+% \begin{figure}[htbp] \centering
+%     \includegraphics[width=0.7\linewidth]{simul_plots_6_declining.pdf}
+%     \caption{Comparison of estimated and true \So values in the six 
+%         scenarios. True \So values are deterministic, black circles show
+%         median \So estimates, error bars indicate the  25\% and 75\%
+%         quantiles, while the blue line denotes a linear model fit of the
+%         medians. The one-to-one relationship is shown by the gray dashed
+%         line.} 
+% \label{fig:simul-six-plots-dec} 
+% \end{figure}
 
-\section*{Methods}
-
-\subsection*{Data}
-
-We obtained time series data of outmigrating smolt and returning adult
-abundances for seven Atlantic salmon populations in eastern Canada, encompassing a
-wide range of the species' distribution (Fig.~\ref{fig:map}). 
-Populations included the LaHave River in Nova Scotia's (NS) Southern Uplands, 
-Nashwaak River, New Brunswick (NB), Rivi\`{e}re de la Trinit\'{e} (Trinit\'{e}) and
-Saint-Jean rivers in Qu\'{e}bec (QC), and  Western Arm Brook (also referred to as  WAB), 
-Campbellton, and Conne River, Newfoundland (NL). 
-Data were collected in NS, NB, and Newfoundland
-and Labrador (NL) by Fisheries and Oceans Canada (DFO) and in QC
-by the Minist\`{e}re des For\^{e}ts, de la Faune et des Parcs, Qu\'{e}bec.
-
-\begin{figure}[htbp] \centering
-    \includegraphics[width=0.85\linewidth]{figures/rivers-map2.png}
-    \caption{Locations of the seven rivers in eastern Canada with time series abundance data of outmigrating smolts and 
-    returning adults.} \label{fig:map} 
-\end{figure}
-
-\subsubsection*{Smolt and adult return abundance data}
-
-Smolt and adult return abundance estimates originate from a variety of
-sources. Smolt estimates from the Trinit\'{e}, Saint-Jean, LaHave, Nashwaak,
-and Conne populations were obtained using a mark-recapture approach, while
-estimates from the WAB and Campbellton populations were obtained by direct
-counts using fish counting fences.
-For further details on the data collection methodologies refer to 
-\citet{Dempson1991, Venoitt2018} for NL populations, 
-\citet{April2018}  for QC populations,
-\citet{Jones2014} for NB population,
-and \citet{Gibson2009} for NS population.
-
-Annual return data are often recorded in terms of two size groups: small ($< 63$ cm
-FL) and large ($\geq 63$ cm FL) salmon, as these closely represent different
-life-history strategies (i.e. 1SW and 2SW), but can be confounded with repeat
-spawners of different sizes. To correct for this in returns 
-reported as small and large salmon, rather than 1SW and 2SW fish, we estimated the abundance
-of 1SW and 2SW returns using yearly scale age data of a subsample of returns:
-
-\begin{equation}
-    p_{r,t,a} = \frac{\sum_{s}{(\frac{n_{r,t,s,a}}{n_{r,t,s}} * N_{r,y,s})}}{\sum_{s}{N_{r,t,s}}}
-\end{equation}
-
-where $p_{r,t,a}$ is the proportion of annual returns in river $r$, year $t$,
-and of spawning history $a$ (either 1SW or 2SW returns); $n_{r,t,s,a}$ is the
-number of samples in river $r$, year $t$, of spawning history $a$, and of size
-group $s$; $n_{r,t,s}$ is the total number of samples in river $r$, year $t$,
-and of size group $s$; and $N_{r,t,s}$ is the returns of salmon
-in river $r$, year $t$, and of size group $s$.
-
-\subsection*{Bayesian model}
-
-We developed a hierarchical Bayesian model that uses Murphy's maturity
-schedule method, in conjunction with informative priors, to estimate annual
-marine survival in seven populations of Atlantic salmon. We account for
-observation error in smolt and return estimates, as well as estimating the
-proportion returning after one winter (i.e. \Pg) hierarchically.
-There is an identifiability problem in the maturity schedule equations where
-the parameter estimates cannot be optimally solved \citep{Chaput2003a}.
-However, this issue can be mathematically overcome, at least partially, by
-using informative priors for all three marine survival parameters in a
-Bayesian framework.
-This method requires abundance estimates of smolts as well as abundance estimates
-of returning one-sea-winter (1SW) and two-sea-winter adults. With these data,
-it estimates three parameters: survival in the first year at sea (\So), survival
-in the second year at sea (\St), and the proportion of fish returning after one
-year at sea (\Pg).
-
-Our model does not include repeat spawners and assumes that no fish spend
-three or more winters at sea before returning to spawn for the first time.
-The model also assumes that mortality in the second winter at sea (\St)
-is additional to mortality in the first winter at sea in the previous year, 
-and therefore does not account for differences in environmental conditions experienced
-between 1SW and 2SW fish of the same smolt cohort during their overlapping first year at sea.
-In other words, our model assumes that the decision of returning occurs just before
-actually being counted as returns and that \St is any additional mortality in
-the subsequent year.
-
-%\comment{Other assumptions...}
-
-Observed smolt estimates were modelled hierarchically and included
-observation error:
-
-\begin{equation}
-log(smolts_{obs,t,r}) = log(smolts_{true,t,r}) + \epsilon_{t,r}
-\end{equation}
-
-where $smolts_{true,t,r}$ are the true smolt abundances for year $t$ and river
-$r$, and $\epsilon_{t,r}$ is the error term, which is empirically derived from
-by calculating the yearly coefficient of variation in the empirically derived
-smolt estimates (see Table S1 in the Supplementary material). 
-Where available, we used population-specific measurement error estimates for smolt abundances; if not 
-available, we set measurement error at 5\%. 
-The log-transformed true smolt abundances are
-normally distributed around a population-level mean and standard deviation:
-
-\begin{equation}
-log(smolts_{true,t,r}) \sim Normal(\mu_{smolts,r}, \sigma_{smolts,r})
-\end{equation}
-
-where $\mu_{smolts,r}$ and $\sigma_{smolts,r}$ are parameters estimated by the
-model for each river.
-
-Once we have yearly estimates of smolt, 1SW, and 2SW abundances, we estimate
-marine survival parameters using Murphy's maturity schedule method
-\citep{Murphy1952, Ricker1975}:
-
-\begin{align}
-    R_{1,t} &= smolts_{true,t-1} * S_{1,t} * Pr_t \label{eq:1}, \\
-    R_{2,t+1} &= smolts_{true,t-1} * S_{1,t} * (1 - Pr_t) * S_{2,t+1} \label{eq:2}
-\end{align}
-
-where $R_{1,t}$ and $R_{2,t+1}$ are the estimated abundances of 1SW and 2SW
-salmon returning in years $t$ and $t+1$, respectively, $smolts_{true,t-1}$ is the
-estimated number of outmigrating smolts in year $t-1$, $S_{1,t}$ is the proportion of
-salmon surviving in their first year ($t$) at sea, $Pr_t$ is the proportion of
-salmon that return to spawn at year $t$, $S_{2,t+1}$ is the survival in their
-second year at sea of the same cohort of salmon who did not return to spawn at
-year $t$.
-
-However, to allow for normally and log-normally-distributed parameters we
-log-transform equations~\ref{eq:1} and~\ref{eq:2} to obtain
-
-\begin{align}
-    log(R_{1,r,t}) &= log(smolts_{r,t-1}) + log(Pr_{r,t}) - Z_{1,r,t} \label{eq:3}, \\
-    log(R_{2,r,t+1}) &= log(smolts_{r,t-1}) - Z_{1,r,t} + log(1 - Pr_{r,t})  - Z_{2,r,t+1} \label{eq:4} 
-\end{align}
-
-where $Z_{1,r,t}$ and $Z_{2,r,t+1}$ are the instantaneous mortality rates.
-Process error was included as the standard deviation of the log-transformed
-return estimates from equation~\ref{eq:4}:
-
-\begin{align}
-log(R_{obs,1,t,r}) &\sim Normal(log(R_{1,t,r}), \epsilon_{1,r}), \\
-log(R_{obs,2,t,r}) &\sim Normal(log(R_{2,t,r}), \epsilon_{2,r}) \label{eq:5} 
-\end{align}
-
-where $R_{obs,1,t,r}$ and $R_{obs,2,t,r}$ are the observed return estimates
-for year $t$ and river $r$ of 1SW and 2SW fish, respectively, $\epsilon_{1,r}$
-and $\epsilon_{2,r}$ are the process error terms. 
-These error terms are normally distributed with a standard deviation
-of 0.01, which are almost equivalent to using a coefficient of variation in
-return estimates 1\%, given that return estimates in equations~\ref{eq:3}
-and~\ref{eq:4} are log-transformed:
-
-\begin{align}
-\epsilon_{1,r} &\sim Normal(0, 0.01) \\
-\epsilon_{2,r} &\sim Normal(0, 0.01).
-\end{align}
-
-Furthermore, we use instantaneous mortality rates in the model instead of survival probabilities
-as the model is more efficient in its parameter search in log-space, and instantaneous rates
-are easy to interpret. Instantaneous rates are easily converted to survival probabilities by
-
-\begin{align}
- S_{1} &= e^{-Z_1}, \\
- S_{2} &= e^{-Z_2}. 
-\end{align}
-
-We estimate population-level mean \Pg values around which the yearly \Pg
-values are normally distributed. We specify different informative hyperpriors
-for \prmu and \prsig based on whether the population is 1SW-dominated or not:
-
-\begin{align}
-    logit(P_{r,t}) &\sim Normal(\mu_r, \sigma_r) \\
-    \mu_r &\sim 
-    \begin{cases}
-       Normal(2.3, 0.4),  &\text{for 1SW-dominated populations} \\
-       Normal(0, 2.8), &\text{for non-1SW-dominated populations} \\
-   \end{cases} \\
-    \sigma_r &\sim halfNormal(0, 1).
-\end{align}
-
-The priors for $Z_1$ and $Z_2$ are specified as log-normal distributions:
-
-\begin{align}
-Z_1 &\sim logNormal(1, 0.22),   \\ 
-Z_2 &\sim logNormal(0.2, 0.3).
-\end{align}
-
-The model was written in Stan \citep{Carpenter2017} and run in R version 3.6.1
-\citep{RCoreTeam2019} using the \texttt{rstan} package version 2.19.2
-\citep{StanDevelopmentTeam2019}.
-The model was run with three chains and 3,000 iterations, with the first 1,500
-discarded as a burn-in. The models were considered to have converged when the
-$\hat R$ of all parameters were lower than 1.03 and the effective sample size 
-were higher than 500.
-
-\subsection*{Correlations among trends in survival}
-
-%\comment{How do were quantitatively compare trends among rivers? DFA is an
-%option, but so is a t-test of averages before and after a certain year}
-
-We looked at the correlation of trends in \So by calculating the Pearson's
-correlation coefficient for the Z-scores of these trends. Given that the time
-series do not cover the same years, and that some rivers have missing years in
-the middle of the time series, pair-wise Pearson's tests were done using only
-the years where there is data for both rivers.
-
-\section*{Results}
-
-%\subsubsection*{Model convergence}
-
-\subsubsection*{Trends in marine survival parameters}
-
-Trends in estimates of \So were highly variable within and among rivers
-(Fig.~\ref{fig:s1-dual}). The highest median posterior estimates of \So
-were for the Nashwaak River in 2006 and 2008, with values of 0.18 and 0.21,
-respectively. The lowest median \So estimate was in the Trinit\'{e} in 2001,
-with an estimate of \So of 0.007, while the Conne, LaHave, and Trinit\'{e} had
-years where estimates of \So varied between 0.01 and 0.02 (Fig.~\ref{fig:s1-faceted}).
-
-Trends among populations also varied: Campbellton,
-Saint-Jean, and Western Arm Brook populations showed increases in \So
-over time, Trinit\'{e}, Conne, and LaHave populations showed decreases,
-while at Nashwaak there was an increase in median \So during the early
-2000s but a decrease in the 2010s. Yearly estimates of \So had very little
-variability for one sea-winter dominated populations (Conne, Campbellton, and
-WAB), but were more variable (i.e. wider credible intervals) in the other
-populations.
-
-\begin{figure}[htbp] \centering
-    \includegraphics[width=0.95\linewidth]{figures/s1-trends-dual.png}
-    \caption{Trends in survival in the first year at sea (\So). a) Posterior
-        estimates of \So for the seven rivers examined, error bars indicate
-        the 90\% credible intervals, and dashed lines denote years of commercial
-        fishing moratoria for each province. b) Z-scores of median posterior
-        estimates.} \label{fig:s1-dual} \end{figure}
-
-\begin{figure}[htbp] \centering
-    \includegraphics[width=0.95\linewidth]{figures/s1-trends-faceted.png}
-    \caption{Posterior estimates of \So for the seven rivers examined, error
-        bars indicate the 90\% credible intervals.} \label{fig:s1-faceted}
-\end{figure}
-
-Estimates of \St were highly uncertain in all rivers, and trends 
-were not apparent in most rivers given the large range of the credible
-intervals in the yearly estimates (Fig.~\ref{fig:s2-faceted}). The estimates
-of \St for the Saint-Jean and Trinit\'{e} were considerably less uncertain
-than for the other populations, but showed no apparent trends through time.
-
-\begin{figure}[htbp] \centering
-    \includegraphics[width=0.95\linewidth]{figures/s2-trends-faceted.png}
-    \caption{Posterior estimates of \St for the seven rivers examined, error
-        bars indicate the 90\% credible intervals.} \label{fig:s2-faceted}
-\end{figure}
-
-Estimates of \Pg were mostly stable across time, except for the LaHave and
-Nashwaak Rivers; The estimates of \Pg were slightly lower in the last four
-years than in the previous ones, while in the Nashwaak the posterior estimates
-of \Pg in 2012 were much lower than in all other years
-(Fig.~\ref{fig:s2-faceted}). Uncertainty in yearly estimates was highest in
-the LaHave, Nashwaak, and Trinit\'{e}, and lowest in the 1SW-dominated
-populations.
-
-\begin{figure}[htbp] \centering
-    \includegraphics[width=0.95\linewidth]{figures/pr-trends-faceted.png}
-    \caption{Posterior estimates of \Pg for the seven rivers examined, error
-        bars indicate the 90\% credible intervals.} \label{fig:pr-faceted}
-\end{figure}
-
-Population-level estimates of \prmu and \prsig varied considerably among rivers. For all
-three 1SW-dominated rivers (Campbellton, Conne, and WAB), estimates of \prmu
-were very close to 1.0 and had little variability in \prsig (Fig~\ref{fig:prmu-post}).
-Estimates of \prmu were the lowest for the two QC rivers, particularly the Saint-Jean (median \prmu = 0.11).
-Estimates of \prmu for the Nashwaak and the LaHave Rivers were close to 0.5, with these two rivers having 
-the highest estimated values of \prsig, particularly the Nashwaak (Fig~\ref{fig:prmu-post}).
-
-\begin{figure}[htbp] \centering
-    \includegraphics[width=1.0\linewidth]{figures/pr-mu-posteriors.png}
-    \caption{Posterior estimates of the population-level parameters $Pr_{\mu}$
-       $logit(Pr_{\mu})$, and $logit(Pr_{\sigma})$. Dots denote median estimates, while the thick and thin error bars indicate
-       the 50\% and 90\% credible intervals, respectively.} 
-   \label{fig:prmu-post} 
-\end{figure}
-
-\subsubsection*{Correlations}
-
-Most correlations between z-scored trends were not statistically significant, with only 
-three out of the 21 pair-wise comparisons having a p-value below 0.01 (Fig.~\ref{fig:s1-corr}a).
-When looking at the direction of the correlation, regardless whether they were
-significant or not, these spanned both positive and negative coefficients
-(Fig.~\ref{fig:s1-corr}b).
-
-\begin{figure}[htbp] \centering
-    \includegraphics[width=0.95\linewidth]{figures/corr-s1.png} \caption{
-        Correlation among Z-scores of median estimated trends in \So among
-        rivers. a) Pearson's correlation coefficients are shown in each square,
-        while colouring denotes significance of the correlation ($p \leq 0.01$), b)
-        colours denote direction and magnitude of correlation, while asterisks denote significance.}
-\label{fig:s1-corr} 
-\end{figure}
-
-\section*{Discussion}
-
-% 1. Main findings
-% 2. How they compare to other publications
-% 3. Potential reasons
-% 4. Caveats
-% 5. Future directions
-
-% 1. Main findings
-Our results challenge the narrative that marine survival, specifically survival in the first year at sea, is declining
-uniformly throughout the range of Atlantic salmon in the northwest Atlantic.
-Temporal trends are not consistent among all populations. 
-Over the time periods for which data were available, some rivers show positive trends in survival in the
-first winter at sea (\So) while other exhibit highly variable yet stable trends, and some show
-declines. We could not assess trends in the second winter at sea (\St) or
-proportion returning as grilse (\Pg), as these parameter estimates were highly
-uncertain and were strongly influenced by the priors.
-Perhaps there is a need to rethink our understanding of Atlantic salmon
-population dynamics in light of the possibilities that (1) any real and persistent decline in marine survival was
-experienced by some but not necessarily all populations, (2) reductions in survival might
-have occurred over a relatively brief period of time and have not persisted, and (3) marine survival has
-been relatively stable, or increasing in some populations for one or more decades.
-
-% 2. How they compare to other publications
-Our results are contrary to those of \citet{Olmos2019}, who detected mostly positive
-correlations in post-smolt survival among spatially broad stock units. These differences
-could be due to a number of reasons: different model specifications and
-structure, different methods for estimating covariance, difference in the
-spatial scales of data sources (i.e. river vs province scales), and perhaps most importantly the use stock-recruitment relationships
-rather than empirical smolt count estimates to estimate marine survival.
-As trends in marine survival during the first winter at sea are highly independent
-among rivers on relatively small spatial scale, trends from broader
-geographical areas (i.e. province, state, or country-wide estimates) may not
-be representative.
-Interestingly, our estimates of \So and \St are very similar to those produced
-by \citet{Chaput2003b}, and our trends are almost identical for the
-overlapping time period that marine survival was estimated for in their study
-(1984-1998). 
-While \citet{Chaput2003b} separated abundance data for males and females
-and assumed their survival rates were the same (to be able to reach an
-analytical solution), our study reached almost the same results (albeit with
-slightly higher uncertainty), using a Bayesian approach with informative
-priors. These overlapping trends obtained with two different methods 
-suggest that our method is effective at estimating marine survival.
-
-Trends in marine survival
-among populations were compared by \citet{Chaput2012a} using adult return rates.
-He found that for 4 of 6 populations examined, return rates in the 1990s 
-were lower than those during the 1970s.
-\citet{Gibson2016} calculated higher return rates of 2SW for the Nashwaak in
-the 1970s than in the period since, with return rates of 1SW in the 1970s
-being comparable to those in the late 2000s.
-\citet{Friedland1993} compared return rates for a number of rivers in eastern
-North America between 1973 and 1988, and suggested there are similar trends among these. 
-However, the similarity in these trends was driven primarily by two years, 1977 and 1978, which
-show concurrent low and high relative return rates across rivers,
-respectively. Other years are much more variable relative to each other.
-\citeauthor{Friedland1993}'s \citeyear{Friedland1993} time series ends in  
-1988; thus there are only a few years for which to assess overlap with the
-time series in our study.
-While declines in return rates since the 1970s seem to be consistent among populations, 
-we were unable to assess if marine survival estimates were also higher in the 1970s as we lacked 
-data from this decade.
-In any event, we caution that the pooling of adult return rates \citep{Chaput2012a, Friedland1993,Gibson2016} 
-can mask inter-annual variability in marine survival,
-and hence might not produce an accurate depiction of marine survival trends.
-While river-specific return rate estimates are available for the populations
-used in this study \citep{ICES2019}, these estimates are only an approximation
-of marine survival with varying degrees of similarity depending on each
-population's life histories. The estimates of return rates of one sea-winter
-should approximate marine survival in the first year at sea for one
-sea-winter-dominated populations. However, as the proportion of fish that
-return as one sea-winter decreases, marine survival in the first year at sea
-can increasingly diverge from return rates of fish, and trends in return rate
-would be particularly misleading if the proportion returning as one sea-winter
-varies in time. 
-Furthermore, return rates of two sea-winter fish also include survival in the
-first winter at sea and the proportion returning as two sea-winter, thus
-providing a very coarse estimate of marine survival in the second year at sea. 
-In fact, changes in return rates of two sea-winter fish are a result of a
-combination of changes in survival in the first year at sea in the year prior,
-survival in the second year at sea, and in the proportion returning as two
-sea-winter, and an increase in any of those parameters while the others remain
-the same will result in an increase in return rates. Our model attempts to
-improve some of the shortcomings of using return rates as a proxy of marine
-survival by directly estimating marine survival in the first and second years
-at sea, as well as the proportion returning as 1SW.
-
-\citet{Dempson2003} described a general declining trend in marine survival for
-Newfoundland rivers (except WAB); we drew the same conclusion for 
-Conne River but not Campbellton River or WAB. It is not possible to draw broader conclusions
-with data from only three Newfoundland rivers, but it seems that among index rivers,
-those in Newfoundland are among those with the highest marine survival rates.
-
-% 3. Potential reasons
-There are a variety of potential explanations for the lack of synchronous
-trends in estimates of \So. 
-Marine survival in the first winter at sea could be highly variable between
-populations because of the predominance of spatially local environmental drivers of survival (e.g., temperature, predation, prey abundance) 
-relative to broader-scale, even ocean-wide, drivers.
-The synchrony reported for marine survival trends at broader spatial scales \citep{Olmos2019}
-might be attributable to the use of stock-recruitment relationships to estimate survival,
-relationships that may have been confounded by changes in recruitment dynamics.
-There is some evidence of a correlation between return rate and growth (as
-indicated by inter-circuli spacing on scales), where years of poor growth
-tended to also be years of poor survival \citep{Friedland1993}, supporting the
-idea that environmental variability can affect marine survival.
-Furthermore, among European salmon, there is evidence of a positive correlation
-between spring temperature in the Norwegian and North Seas and population abundance, suggesting warmer
-conditions favour post-smolts \citep{Friedland1998}, based on mapping the
-extent of area of suitable temperature (7-13 \textdegree C).
-
-Nonetheless, the causal mechanisms for why warming should affect post-smolt
-survival almost certainly differs depending on the difference between
-temperature experienced by the post-smolts and their respective
-population-specific thermal optima. 
-This difference could explain why populations in eastern North America are
-declining in the southern part of their range but potentially increasing
-further north, and also why some studies find positive correlations between
-temperature and abundance \citep{Friedland1998, Friedland1998b, Jonsson2004}
-while others find negative ones \citep{Friedland1993, Todd2008}.
-Putative associations between temperature and direct estimates of marine
-survival warrants further study at the population level.
-
-While there is little evidence that marine survival is density-dependent in
-Atlantic salmon \citep{Jonsson1998,Gibson2006}, there could potentially be
-some density-dependent processes during parts of the post-smolt migration
-period, particularly for populations that are likely to be subjected to
-declining per capita population growth rates ($r$) generated by Allee effects.
-Exploring relationships between survival and population size could potentially
-shed light about the processes that have caused many of the population
-declines that have been documented.
-
-Oceanic conditions have been correlated with abundance trends and growth
-\citep{Todd2008}, however, the mechanism by which such bottom-up effects,
-mediated by changes in food availability,
-affect population dynamics beyond marine survival needs to
-be thoroughly reassessed. If marine survival on its own cannot fully explain
-trends in abundance, then there are potential carry-on effects of oceanic
-conditions that manifest with regards to freshwater production. 
-For example, adults
-that return to spawn after spending suboptimal conditions at sea might be less
-likely to make it to their spawning grounds, successfully secure a mate,
-produce fewer eggs, or produce eggs with lower per capita fitness than those
-produced by adults which grew in optimal oceanic conditions.
-As larger females tend to be more productive, in terms of fecundity and total
-reproductive energy, than the same weight's worth of smaller females
-\citep{Barneche2018}, a small decrease in body condition resulting from bottom-up
-impacts on food availability and quality could potentially have
-disproportionate effects on fecundity and fitness of the offspring.
-
-% Furthermore, freshwater carryover effects acting on smolts, such as pH effects
-% on prey-avoidance behaviour, could potentially mediate post-smolt survival
-% \citep{Halfyard2012}.
-
-Egg-to-smolt survival in Atlantic salmon is highly variable \citep{Klemetsen2003,Chaput2015}.
-While this variability is likely attributable to changes in freshwater
-conditions (e.g. discharge, temperature, water quality) and uncertainty in
-spawner and smolt estimates, 
-changes in the oceanic conditions that spawners experience could also be
-contributing to this variability.
-Obviously, there would be a time lag (perhaps as much as a generation) in how such effects
-might be manifest at the adult stage.
-However, given that most correlations are between relatively monotonic declines
-in abundance coupled also monotonic increases in climatic indices
-over decadal time scales \citep[e.g.,][]{Friedland1998, Todd2008,
-    Beaugrand2012}, it would be expected that these correlations would be
-maintained even if salmon abundances were lagged by a generation length.
-
-% 4. Caveats
-As with all novel modelling approaches, there are caveats to acknowledge.
-The seven populations explored in the present study might not be representative
-of regional trends in marine survival. However, there are no other
-long-term time series of smolts and adult returns to draw inferences from.
-While there are analytical issues associated with the estimation of \So, \St, and \Pg,
-the assumption that \St is additive to \So could produce unrealistic results.
-We know there is a period of a few months where 1SW
-returns are subject to a different environment than those salmon that will
-return as 2SW the next year. 
-While this is not ideal,
-overcoming this assumption would require an additional parameter to be
-estimated, or an additional assumption as to what proportion of \So is not
-additive to \St (as the returning 1SW adults do not experience the same
-environment when they return to their natal streams as those fish who stayed
-at sea for an additional winter before returning to spawn).
-
-Secondly, the assumed hierarchical structure might not be the most appropriate
-for modelling smolt abundances. This approach results in shrinkage to the
-mean, which means that the variability of annual smolt estimates is less than
-it would be if not modelled hierarchically. 
-That said, this assumption is likely to result in more conservative trends in marine survival, as the
-variation is smolt estimates is reduced.
-It is important to note that the hierarchical structure in the estimation of
-\Pg seems like a reasonable assumption given that the probability of returning
-as grilse has a genetic component associated to it \citep{Aykanat2019} and is
-not expected to vary much, within a population, among years.
-
-% 5. Future directions
-Perhaps a reframing of the issue of marine survival is key to furthering our
-understanding of Atlantic salmon population dynamics. Marine survival may not
-have declined consistently, and over the same time periods, across all
-populations. 
-But the fact that it has remained at roughly similar levels as it was
-previously \emph{despite} reduced commercial fishing mortalities, suggests
-that there may well be an interaction between small population size (small
-relative to unfished population size or carrying capacity), recovery
-potential, and environmental stochasticity that has not been fully explored in
-Atlantic salmon. 
-All else being equal, relatively small populations are more vulnerable to
-demographic, environmental, and genetic stochasticity than large populations
-\citep{Lande1993, Hutchings2015}. Interactions between population size and the
-demographic consequences of environmental stochasticity appear to have
-affected recovery in many marine fishes that have exhibited impaired recovery
-since mitigation of the threat posed by fishing mortality
-\citep{Hutchings2017, Hutchings2020}. The possibility that similar
-interactions may be impairing the recovery of wild Atlantic salmon merits
-study.
-
-\section*{Acknowledgements}
-
-% We would like to thank Geir Bolstad, G\'{e}rald Chaput, Brian Dempson, and
-% Martha Robertson for their useful discussions on estimating marine survival
-% in Atlantic salmon, and Amanda Kissel for her helpful comments on the
-% manuscript. Brian Dempson and Geoff Venoit for providing the data Conne
-% River data.
-We would like to thank G\'{e}rald Chaput for his useful discussions on
-estimating marine survival in Atlantic salmon, Carmen David for her comments
-on the manuscript, and Sean Anderson for his help with implementing the
-non-centered parameterization of the model. This research was supported by the
-Atlantic Salmon Conservation Foundation and the Atlantic Salmon Research Joint
-Venture.
-
-\section*{Conflicts of Interest}
-
-The authors declare no conflicts of interest.
-
-\bibliography{subset}
-
-%\input{ms.bbl}
+%\clearpage
+%\subsection*{References}
+\nobibliography{subset}
 

### 2020-05-22 analysis/99-suppmat.R

@@ -1,0 +1,174 @@
+library(tidyverse)
+library(rstan)
+library(xtable)
+
+fithierncp <- readRDS("data/stanfit_hier_ncp.rds") # hierarchical
+
+alldatastan <- readRDS("data/alldatastan.rds")
+river_names <- rep(alldatastan$river_name, times = alldatastan$nyears)
+
+logsmoltsmat <- as.matrix(fithierncp, pars = "logsmolts_true")
+logsmoltstib <- logsmoltsmat %>% t %>% as_tibble %>% mutate(river_name = river_names, year = alldatastan$years) %>%
+  select(river_name, year, everything())
+
+estsmolts <- pivot_longer(logsmoltstib, V1:V7500, names_to = "iteration") %>%
+  mutate(year = year - 1) %>%
+  group_by(river_name, year) %>%
+  rename(`Population` = river_name, Year = year) %>%
+  summarise(Median = exp(median(value)), 
+            `2.5%` = exp(quantile(value, probs = 0.025)), 
+            `25%` = exp(quantile(value, probs = 0.25)), 
+            `75%` = exp(quantile(value, probs = 0.75)), 
+            `97.5%` = exp(quantile(value, probs = 0.975))
+           # n = n()
+            )
+
+allyears <- estsmolts %>%
+  group_by(Population) %>%
+  expand(Year = full_seq(Year, 1))
+
+estsmolts_mis <- right_join(estsmolts, allyears, by = c("Population", "Year")) %>%
+  mutate(Year = as.character(Year))
+
+print.xtable(xtable(estsmolts_mis, digits = c(0,0,0,1,1,1,1,1),
+                    display= c("s", "s", "s", rep("f", 5)),
+                    caption = "Posterior annual smolt estimates for the seven populations examined. These
+                    estimates differ from the smolt data used as they include both observation error and shrinkage from
+                    the hierarchical model specification."),
+            file = "ms/tableS1_smolts.tex",
+            # only.contents = TRUE,
+            NA.string = "-",
+            tabular.environment = "longtable", floating = FALSE,
+            size = "footnotesize",
+            format.args = list(big.mark = ","),
+            caption.placement = "top",
+       include.rownames = FALSE)
+
+
+logsmoltsmat <- as.matrix(fithierncp, pars = "logsmolts_mean")
+
+mu_smolts <- as.matrix(fithierncp, pars = "logsmolts_mean") %>% 
+  t %>% as_tibble %>% mutate(river_name = alldatastan$river_name) %>%
[…]

### 2020-05-22 ms/ms.tex
Edits before sending out draft ms, add logit pr and sigma tables

@@
 
+\usepackage[utf8]{inputenc} % to properly print special characters directly
+
 \usepackage{tabularx}
@@
 \newcommand{\St}{$S_{2}$\xspace}
-\newcommand{\Pg}{$P_r$\xspace}
-\newcommand{\prmu}{$\mu_r$\xspace}
-\newcommand{\prsig}{$\sigma_r$\xspace}
+\newcommand{\Pg}{$P_g$\xspace}
+\newcommand{\prmu}{$\mu_{Pg,r}$\xspace}
+\newcommand{\prsig}{$\sigma_{Pg,r}$\xspace}
 \newcommand{\Linf}{$L_{\infty}$}
@@
 
-\section*{Trends in marine survival of Atlantic salmon in eastern Canada}
+\section*{Trends in marine survival of Atlantic salmon populations \\ in eastern Canada}
 
@@
         Ross A. Jones\textsuperscript{5}, 
-        Martha J. Robertson\textsuperscript{3}, 
+        %Martha J. Robertson\textsuperscript{3}, 
         Dustin Raab\textsuperscript{6}, 
@@
 Declines in wild Atlantic salmon (\emph{Salmo salar}) abundance throughout the north
-Atlantic are primarily attributed to declining survival at sea. This
-hypothesis has proven challenging to test on a river-by-river basis because of
+Atlantic are primarily attributed to declining survival at sea. 
+Whether these declines in marine survival are highly variable among populations
+has not been straightforward to estimate because of
 the need to model data on both migrating smolts and returning adults and to
 simultaneously estimate multiple parameters, especially for salmon spending
-more than one winter at sea (1SW) before spawning. We fit a hierarchical Bayesian
-maturity schedule model to data (19-42 years) for seven populations in  
-Newfoundland and Labrador (NL), Qu\'{e}bec (QC), New Brunswick (NB), and Nova
-Scotia (NS), Canada. We estimate survival in the first (\So) and second year at sea (\St),
-and the proportion returning as 1SW adults (\Pg). Trends were not consistent
-among rivers. Since 1990, \So increased at Western Arm Brook (NL)
-and Rivi\`{e}re Saint-Jean (QC), but declined at Conne River (NL) and Rivi\'{e}re
-de la Trinit\'{e} (QC). Since the mid-1990s, \So increased at Campbellton River (NL),
-declined at LaHave River (NS), and fluctuated at Nashwaak
-River (NB). Estimates of \St were highly uncertain, particularly for
-1SW-dominated populations; \Pg was generally stable. These results challenge the
-narrative that marine survival has changed in a temporally consistent manner
-among spatially disparate populations. Our findings suggest that, at the
-population level, changes in abundance are attributable to temporal shifts in
-multiple components of individual fitness, including at-sea survival. If
-salmon populations do not respond in a consistent, uniform manner to changing
-environmental conditions throughout their range, future research initiatives
-should explore why.
+more than one winter at sea before spawning. We fit a hierarchical Bayesian
+maturity schedule model to data spanning between 19 and 42 years for seven populations in eastern 
+Canada. We estimate numbers of out-migrating smolts, 
+survival in the first and second year at sea, and the proportion returning after one year. 
+Trends in survival at sea were not consistent among populations, and we observe both
+positive, negative and no correlations in the trends. This suggest that large
+scale patterns of changes in marine survival are not necessarily representative for
+individual populations. 
+Variation in return abundances was mostly explained by marine survival in the
+first winter at sea in all but one population.
+However, variation in the other components were not negligible and
+their relative importance differed among populations. Variation in number of
+out-migrating smolts had a moderate contribution on returns, particularly
+after two winter at sea, and probability of returning after one year was
+important in some populations. 
+% Trends were not consistent among rivers. 
+% Since 1990, \So increased at Western Arm Brook (NL)
+% and Rivi\`{e}re Saint-Jean (QC), but declined at Conne River (NL) and Rivi\'{e}re
+% de la Trinit\'{e} (QC). Since the mid-1990s, \So increased at Campbellton River (NL),
+% declined at LaHave River (NS), and fluctuated at Nashwaak
+% River (NB). Estimates of \St were highly uncertain, particularly for
+% 1SW-dominated populations; \Pg was generally stable. 
+% These results challenge the narrative that marine survival has changed in a
+% temporally consistent manner among spatially disparate populations. 
+If salmon populations do not respond in a consistent, uniform manner to
+changing environmental (including oceanic) conditions throughout their range,
+future research initiatives should explore why.
 
@@
 species risk of extinction.
-While it is not fully understood what is driving population declines, the potential
+While the mechanisms behind population declines are not fully understood, the potential
 drivers of these are many \citep[see ][for a detailed discussion of possible
@@
 % 2.
-Over the past three decades, a narrative has emerged that marine survival of
-Atlantic salmon has declined throughout the North Atlantic \citep{ICES2019}.
+% It is widely perceived that,
+Several studies suggest that, over broad spatial scales,
+marine survival of Atlantic salmon has declined throughout the North Atlantic,
+particularly since the 1980's \citep{Massiot-Granier2014,ICES2019,Olmos2019}.
 %\citep{Hansen1998,OMaoileidigh2003,Chaput2012a}.
-Based on multiple lines of evidence that climate conditions can directly and
+These declines are thought to be driven by oceanic climate effects 
+based on multiple lines of evidence suggesting that climate conditions can directly and
 indirectly influence the abundance and productivity of Atlantic salmon
-populations \citep{Mills2013,Almodovar2019}, it has been presumed that oceanic climate effects are
-driving population dynamics primarily through changes in marine survival.
+populations \citep{Mills2013,Almodovar2019,Olmos2020}.
 Thus, an implicit assumption is that populations sharing a particular oceanic
-route during seaward migration will also show similar trends in marine
+route during seaward migration are likely to show similar trends in marine
 survival. 
@@
 at-sea mortality \citep{Friedland1993, Friedland1998, Russell2012}.  
-A recent study by \citet{Olmos2019} suggested that trends in post-smolt
-survival, when estimated at the stock level, are synchronously declining
-for all Atlantic salmon in eastern North America.
+% A recent study by \citet{Olmos2019} suggested that trends in post-smolt
+% survival, when estimated at the stock level, are synchronously declining
+% for all Atlantic salmon in eastern North America.
+
+% Recently \citet{Olmos2019} suggested that trends in post-smolt
+% survival, when estimated at the stock unit level, are synchronously declining
+% for all Atlantic salmon in Eastern North America and Canada, a conclusion ultimately 
+% grounded on the veracity of highly variable stock-recruitment relationships.
 
@@
 
-In contrast to the narrative of widespread, demographically similar increases
-in at-sea mortality, the conservation status of Canadian salmon populations differs
-considerably. Populations in the southern part of their range are more
+Despite the overall decreases in marine survival,
+the conservation status of Canadian salmon populations differ
+considerably. In Canada, populations in the southern part of their range are more
 likely to be assessed as being of conservation concern than those in more
@@
 % 3. Nonetheless, this premise has never been examined in detail
-Given the logistical challenges associated with estimating at-sea survival, it
-is not surprising that the number of studies that have estimated temporal
-trends has been limited. An additional limitation has been the derivation of
+
+Given the logistical challenges associated with estimating at-sea survival for
+individual populations, it is not surprising that the number of studies that
+have estimated temporal trends has been limited. 
+The evidence for widespread declines in marine survival mostly comes from studies
+over broad spatial scales that do not include empirical smolt abundance data, and
+rely on highly variable stock-recruitment relationships.
+An additional limitation has been the derivation of
 proxies (e.g., return rates), rather than direct model-based estimates, of
@@
 had experienced declining return rates. 
-However, return rates cannot directly be interpreted as marine survival rates and 
+However, with the exception of one-sea-winter (1SW) dominated populations 
+(such as most populations in Newfoundland) where return rates 
+closely approximate marine survival, return rates cannot directly be interpreted as marine survival rates and 
 examination of trends in return rates alone can mask changes in
@@
 proportion of adults returning after one or two years at sea \citep{Hubley2011}.
-Recently \citet{Olmos2019} suggested that trends in post-smolt
-survival, when estimated at the stock unit level, are synchronously declining
-for all Atlantic salmon in Eastern North America and Canada, a conclusion ultimately 
-grounded on the veracity of highly variable stock-recruitment relationships.
 
@@
 In the present study, we compile data on the number of migrating smolts and number of returning adults 
-for seven wild Canadian populations of Atlantic salmon to model trends in marine survival.
+for seven wild Canadian populations of Atlantic salmon to model population-level trends in marine survival
+and assess their among-population variation.
 While some studies have previously used maturity-schedule models to estimate marine
@@
 abundances for seven Atlantic salmon populations in eastern Canada, encompassing a
-wide range of the species' distribution (Fig.~\ref{fig:map}). 
+wide range of the species' distribution. 
 Populations included the LaHave River in Nova Scotia's (NS) Southern Uplands, 
-Nashwaak River, New Brunswick (NB), Rivi\`{e}re de la Trinit\'{e} (Trinit\'{e}) and
-Saint-Jean rivers in Qu\'{e}bec (QC), and  Western Arm Brook (also referred to as  WAB), 
-Campbellton, and Conne River, Newfoundland (NL). 
+Nashwaak River in New Brunswick (NB), Rivi\`{e}re de la Trinit\'{e} (Trinit\'{e}) and
+Rivi\`{e}re Saint-Jean in Qu\'{e}bec (QC), and Western Arm Brook (also referred to as WAB), 
+Campbellton, and Conne River, in Newfoundland and Labrador (NL) (Fig.~\ref{fig:map}). 
 Data were collected in NS, NB, and Newfoundland
@@
 FL) and large ($\geq 63$ cm FL) salmon, as these closely represent different
-life-history strategies (i.e. 1SW and 2SW), but can be confounded with repeat
+life-history strategies (i.e. one- and two-sea-winter), but can be confounded with repeat
 spawners of different sizes. To correct for this in returns 
@@
 This method requires abundance estimates of smolts as well as abundance estimates
-of returning one-sea-winter (1SW) and two-sea-winter adults. With these data,
+of returning one-sea-winter (1SW) and two-sea-winter (2SW) adults. With these data,
 it estimates three parameters: survival in the first year at sea (\So), survival
@@
 In other words, our model assumes that the decision of returning occurs just before
-actually being counted as returns and that \St is any additional mortality in
+actually being counted as returns and that \St is additional mortality in
 the subsequent year.
@@
 \begin{equation}
-log(smolts_{obs,t,r}) = log(smolts_{true,t,r}) + \epsilon_{t,r}
+log(smolts_{obs,t,r}) = log(smolts_{true,t,r}) + \epsilon_{t,r},
 \end{equation}
@@
 where $smolts_{true,t,r}$ are the true smolt abundances for year $t$ and river
-$r$, and $\epsilon_{t,r}$ is the error term, which is empirically derived from
-by calculating the yearly coefficient of variation in the empirically derived
-smolt estimates (see Table S1 in the Supplementary material). 
+$r$, and $\epsilon_{t,r}$ is the error term, which is calculated from the yearly 
+coefficient of variation in the empirically derived
+smolt estimates (see Tables S1 and S2 in the Supplementary material). 
 Where available, we used population-specific measurement error estimates for smolt abundances; if not 
-available, we set measurement error at 5\%. 
+available, we set measurement error between 5 and 10\%. 
 The log-transformed true smolt abundances are
@@
 
-where $\mu_{smolts,r}$ and $\sigma_{smolts,r}$ are parameters estimated by the
-model for each river.
+where $\mu_{smolts,r}$ and $\sigma_{smolts,r}$ are the mean and standard
+deviation of the of the hierarchical population-level log-smolt abundances
+estimated by the model for each population $r$.
 
@@
 \begin{align}
-    R_{1,t} &= smolts_{true,t-1} * S_{1,t} * Pr_t \label{eq:1}, \\
-    R_{2,t+1} &= smolts_{true,t-1} * S_{1,t} * (1 - Pr_t) * S_{2,t+1} \label{eq:2}
+    R_{1,t} &= smolts_{true,t-1} * S_{1,t} * P_{g,t} \label{eq:1}, \\
+    R_{2,t+1} &= smolts_{true,t-1} * S_{1,t} * (1 - P_{g,t}) * S_{2,t+1}, \label{eq:2}
 \end{align}
@@
 estimated number of outmigrating smolts in year $t-1$, $S_{1,t}$ is the proportion of
-salmon surviving in their first year ($t$) at sea, $Pr_t$ is the proportion of
+salmon surviving in their first year ($t$) at sea, $P_{g,t}$ is the proportion of
 salmon that return to spawn at year $t$, $S_{2,t+1}$ is the survival in their
@@
 These error terms are normally distributed with a standard deviation
-of 0.01, which are almost equivalent to using a coefficient of variation in
-return estimates 1\%, given that return estimates in equations~\ref{eq:3}
-and~\ref{eq:4} are log-transformed:
+of 0.01:
 
@@
 \end{align}
+
+These error terms are almost equivalent to using a coefficient of variation in
+return estimates 1\%, given that return estimates in equations~\ref{eq:3}
+and~\ref{eq:4} are log-transformed. As
+there is no associated uncertainty with most return estimates, we chose this value arbitrarily
+based on the expectation that uncertainty in returns estimates is likely lower than in
+smolt estimates.
 
@@
 These priors, when converted to yearly survival, are roughly constrained between
-0 and 0.2 for \So and between 0.2 and 0.5 for \St (Fig.~\ref{fig:priors}a).
+0 and 0.2 for \So and between 0.2 and 0.5 for \St (Fig.~\ref{fig:priors}a). They are
+informative only to the extent that they limit the estimates of marine survival
+to what what we considered to be biologically realistic.
 
@@
 \begin{align}
-    logit(P_{r,t}) &\sim Normal(\mu_r, \sigma_r) \\
-    \mu_r &\sim 
+    logit(P_{g,t}) &\sim Normal(\mu_{Pg,r}, \sigma_{Pg,r}) \\
+    \mu_{Pg,r} &\sim 
     \begin{cases}
@@
    \end{cases} \\
-    \sigma_r &\sim halfNormal(0, 1).
+    \sigma_{Pg,r} &\sim halfNormal(0, 1).
 \end{align}
 
-These priors, when converted to proportion returning, are narrowly constrained
+The three Newfoundland populations (WAB, Campbellton and Conne) are considered to be 
+1SW-dominated, while the other four have the more generic priors for $logit(P_{g,t})$.
+These priors, when converted back to a proportion, are narrowly constrained
 for 1SW-dominated populations, but relatively wide for all other populations
@@
         (\St) returns, and b) probability of returning as 1SW for
-        1SW-dominated rivers (proportion 1SW > 0.9) and other rivers.}
+        1SW-dominated populations (proportion 1SW > 0.9) and other populations.}
     \label{fig:priors} 
@@
 
-%\comment{How do were quantitatively compare trends among rivers? DFA is an
-%option, but so is a t-test of averages before and after a certain year}
-
-We looked at the correlation among trends in \So between rivers by directly estimating the 
-correlation for each posterior iteration and then calculating the distribution of correlation values:
+We looked at the correlation among trends in \So between populations by
+directly estimating the error-corrected correlation for each posterior
+iteration and then calculating the distribution of correlation values:
 
 \begin{equation}
-cor(Z_{1,r=1},Z_{1,r=2}) = \frac{cov(Z_{1,r=1}, Z_{1,r=2})}{\sqrt{var(Z_{1,r=1})}*\sqrt{var(Z_{1,r=2})}}\label{eq:corz1}
+cor_{corrected}(Z_{1,r=1},Z_{1,r=2}) = \frac{cov(Z_{1,r=1}, Z_{1,r=2}) - cov_{error}(Z_{1,r=1}, Z_{1,r=2})}
+    {\sqrt{var(Z_{1,r=1}) - var_{error}(Z_{1,r=1})}*\sqrt{var(Z_{1,r=2}) - var_{error}(Z_{1,r=2})}}, \label{eq:corz1}
 \end{equation}
 
+% \begin{equation}
+% cor(Z_{1,r=1},Z_{1,r=2}) = \frac{cov(Z_{1,r=1}, Z_{1,r=2})}{\sqrt{var(Z_{1,r=1})}*\sqrt{var(Z_{1,r=2})}}\label{eq:corz1-uncorrected}
+% \end{equation}
+
 where $Z_{1,r=1}$ and $Z_{1,r=2}$ are the $Z_1$ estimates, across years, for a
-single posterior iteration for rivers 1 and 2. 
-Given that the time series do not cover the same years, and that some rivers
-have missing years in the middle of the time series, only years with overlapping
-$Z_1$ estimates were used for the pairwise correlation estimates.
-We also estimated the correlation between the model parameters for each population using
-a similar approach as in equation~\ref{eq:corz1}. However we error-corrected the correlation estimates
-by including the error in the variance and covariance estimates.
-To get the error corrected correlations we need the error covariance ($cov_{error}$) and error
-variances ($var_{error}$). Estimates of these can be obtained by estimating the covariance of
+single posterior iteration for populations 1 and 2,
+$cov(Z_{1,r=1}, Z_{1,r=2})$ is the covariance between these overlapping $Z_1$ estimates, 
+$cov_{error}(Z_{1,r=1}, Z_{1,r=2})$ is the error covariance term,   
+$var(Z_{1,r=1})$ and $var(Z_{1,r=2})$ are the variances, while
+$var_{error}(Z_{1,r=1})$ and $var_{error}(Z_{1,r=2})$ and are error variance estimates.
+Estimates of the error covariance ($cov_{error}$) and error
+variances ($var_{error}$) are obtained by estimating the covariance of
 two parameters across all values in the posterior for a given year (rather
@@
 across all values in the posterior for a given year and then are averaged
-across years for each river:
+across years for each river.
+Given that the time series do not cover the same years, and that some populations
+have missing years in the middle of the time series, only years with overlapping
+$Z_1$ estimates were used for the pairwise correlation estimates, and thus
+each pair-wise correlation is specific for those years and does not include sample uncertainty due to sample size.
+
+We also estimated the correlation between the model parameters for each population using
+the same approach as in equation~\ref{eq:corz1},
+by including the error in the variance and covariance estimates:
 
@@
 cor_{corrected}(P_{g,r},-Z_{1,r}) = \frac{cov(P_{g,r}, -Z_{1,r}) - cov_{error}(P_{g,r}, -Z_{1,r})}
-{\sqrt{var(P_{g,r})- var_{error}(P_{g,r})}*\sqrt{var(-Z_{1,r}) - var_{error}(-Z_{1,r})}}\label{eq:corparam}
+{\sqrt{var(P_{g,r})- var_{error}(P_{g,r})}*\sqrt{var(-Z_{1,r}) - var_{error}(-Z_{1,r})}}\label{eq:corparam}.
 \end{equation}
@@
 
-% However correlations are not additive so to get the error corrected correlation you have to calculate
-
 \section*{Results}
@@
 
-Trends in estimates of \So were highly variable within and among rivers
+Trends in estimates of marine survival in the first winter at sea (\So) were 
+highly variable within and among populations
 (Fig.~\ref{fig:s1-dual}). The highest median posterior estimates of \So
@@
     \caption{Trends in survival in the first year at sea (\So). a) Posterior
-        estimates of \So for the seven rivers examined, error bars indicate
+        estimates of \So for the seven populations examined, error bars indicate
         the 90\% credible intervals, and dashed lines denote years of commercial
@@
     \includegraphics[width=0.95\linewidth]{figures/s1-trends-faceted.png}
-    \caption{Posterior estimates of \So for the seven rivers examined, error
+    \caption{Posterior estimates of \So for the seven populations examined, error
         bars indicate the 90\% credible intervals.} \label{fig:s1-faceted}
@@
 
-Estimates of \St were highly uncertain in all rivers, and trends 
-were not apparent in most rivers given the large range of the credible
+Estimates of \St were highly uncertain in all populations, and trends 
+were not apparent in most populations given the large range of the credible
 intervals in the yearly estimates (Fig.~\ref{fig:s2-faceted}). The estimates
@@
     \includegraphics[width=0.95\linewidth]{figures/s2-trends-faceted.png}
-    \caption{Posterior estimates of \St for the seven rivers examined, error
+    \caption{Posterior estimates of \St for the seven populations examined, error
         bars indicate the 90\% credible intervals.} \label{fig:s2-faceted}
@@
 Estimates of \Pg were mostly stable across time, except for the LaHave and
-Nashwaak Rivers; The estimates of \Pg were slightly lower in the last four
+Nashwaak populations; The estimates of \Pg were slightly lower in the last four
 years than in the previous ones, while in the Nashwaak the posterior estimates
 of \Pg in 2012 were much lower than in all other years
-(Fig.~\ref{fig:s2-faceted}). Uncertainty in yearly estimates was highest in
+(Fig.~\ref{fig:s2-faceted}). Variation in yearly estimates was highest in
 the LaHave, Nashwaak, and Trinit\'{e}, and lowest in the 1SW-dominated
@@
     \includegraphics[width=0.95\linewidth]{figures/pr-trends-faceted.png}
-    \caption{Posterior estimates of \Pg for the seven rivers examined, error
+    \caption{Posterior estimates of \Pg for the seven populations examined, error
         bars indicate the 90\% credible intervals. Year corresponds to the
@@
 
-Population-level estimates of \prmu and \prsig varied considerably among rivers. For all
-three 1SW-dominated rivers (Campbellton, Conne, and WAB), estimates of \prmu
-were very close to 1.0 and had little variability in \prsig (Fig~\ref{fig:prmu-post}).
-Estimates of \prmu were the lowest for the two QC rivers, particularly the Saint-Jean (median \prmu = 0.11).
-Estimates of \prmu for the Nashwaak and the LaHave Rivers were close to 0.5, with these two rivers having 
-the highest estimated values of \prsig, particularly the Nashwaak (Fig~\ref{fig:prmu-post}).
+Population-level estimates of \prmu and \prsig varied considerably among populations. For all
+three 1SW-dominated populations (Campbellton, Conne, and WAB), estimates of \prmu
+were very close to 1.0 and had little variability in \prsig (Fig~\ref{fig:prmu-post}; Table~\ref{tab:prmu}).
+Estimates of \prmu were the lowest for the two QC populations, particularly the Saint-Jean (median \prmu = 0.11).
+Estimates of \prmu for the Nashwaak and the LaHave populations were close to 0.5, with these two populations having 
+the highest estimated values of \prsig, particularly the Nashwaak (Fig~\ref{fig:prmu-post}; Table~\ref{tab:prsigma}).
+
+\input{table-logitprmu.tex}
+
+\input{table-logitprsigma.tex}
+
+% \begin{figure}[htbp] \centering
+%     \includegraphics[width=1.0\linewidth]{figures/pr-mu-posteriors.png}
+%     \caption{Posterior estimates of the population-level parameters $Pr_{\mu}$
+%        $logit(Pr_{\mu})$, and $logit(Pr_{\sigma})$. Dots denote median estimates, while the thick and thin error bars indicate
+%        the 50\% and 90\% credible intervals, respectively.} 
+%    \label{fig:prmu-post} 
+% \end{figure}
 
 \begin{figure}[htbp] \centering
-    \includegraphics[width=1.0\linewidth]{figures/pr-mu-posteriors.png}
-    \caption{Posterior estimates of the population-level parameters $Pr_{\mu}$
-       $logit(Pr_{\mu})$, and $logit(Pr_{\sigma})$. Dots denote median estimates, while the thick and thin error bars indicate
-       the 50\% and 90\% credible intervals, respectively.} 
+    \includegraphics[width=1.0\linewidth]{figures/logitpr-mu-sigma-post.png}
+    \caption{Posterior probability distributions resulting from the population-level estimates of the proportion returning after one winter at sea based on
+        the parameters \prmu and \prsig. Gray lines denote population-specific median estimates of \prmu and \prsig, while
+       the colored lines represent a sample of 20 posterior iteration draws for each population.} 
    \label{fig:prmu-post} 
@@
 positive and negative correlations among them.
+The corrected pair-wise correlations in $Z_1$ were almost identical to the
+uncorrected ones as the error variance and covariance terms were very small.
+The correlations between populations in the region were positive between Nashwaak and LaHave (Scotia-Fundy region), non-significant between the Trinit\'{e} and Saint-Jean (QC), and positive
+between WAB and Campbellton but negative between Conne and both WAB and Campbellton (NL).
 
@@
     \includegraphics[width=0.95\linewidth]{figures/corrplot-post-z1-corrected.png} \caption{
-        Correlations of estimated trends in $Z_1$ among rivers. 
+        Correlations of estimated trends in $Z_1$ among populations. 
         Correlation coefficients are shown in each square, while colouring 
@@
 As is expected from Murphy's method, there is a strong negative correlation
-between survival in the second year at sea ($-Z_2$) and proportion returning
+between survival in the second year at sea (i.e. $-Z_2$) and proportion returning
 as 1SW ($log(P_g)$, Fig.~\ref{fig:cor-params}). Some populations (Campbellton,
@@
 
-Overall $Z_1$ explained most of the variance in estimated 1SW returns they had
-the highest $R^2$ values (with the exception of the Saint-Jean,
-Fig.~\ref{fig:mu-rsq}).
+Overall, variation in return abundances was mostly explained by marine survival in the
+first winter at sea, with $R^2$ estimates between 0.5 and 0.75 for all populations
+with the exception of the Saint-Jean (Fig.~\ref{fig:mu-rsq}).
 The low $R^2$ for the Saint-Jean is likely due to the negative correlation
@@
 Values of $R^2$ were overall low for all parameters with regards to explaining
-the variance of estimated 2SW returns. Estimated pf $Z_2$ explained little
-variance in estimates for 2SW returns, with the Saint-Jean again being the
+the variance of estimated 2SW returns. Estimates of survival in the second winter at sea ($Z_2$)
+explained little variance in estimates for 2SW returns, with the Saint-Jean again being the
 exception.
@@
         values between $R_{1,est}$ and $Z_2$ as the latter is not used to
-        estimate the former, and that only the non-1SW dominated rivers are
+        estimate the former, and that only the non-1SW dominated populations are
         shown in the $R^2$ values of the model parameters and $R_{2,est}$.}
@@
 % 1. Main findings
-Our results challenge the narrative that marine survival, specifically survival in the first year at sea, is declining
-uniformly throughout the range of Atlantic salmon in the northwest Atlantic.
-Temporal trends are not consistent among all populations. 
-Over the time periods for which data were available, some rivers show positive trends in survival in the
-first winter at sea (\So) while other exhibit highly variable yet stable trends, and some show
-declines. We could not assess trends in the second winter at sea (\St) or
-proportion returning as grilse (\Pg), as these parameter estimates were highly
-uncertain and were strongly influenced by the priors.
-Perhaps there is a need to rethink our understanding of Atlantic salmon
-population dynamics in light of the possibilities that (1) any real and persistent decline in marine survival was
-experienced by some but not necessarily all populations, (2) reductions in survival might
-have occurred over a relatively brief period of time and have not persisted, and (3) marine survival has
-been relatively stable, or increasing in some populations for one or more decades.
+
+% Our results challenge the narrative that marine survival, specifically
+% survival in the first year at sea, is declining uniformly throughout the
+% range of Atlantic salmon in the northwest Atlantic.
+Our results indicate that trends in marine survival among Atlantic salmon
+populations in eastern Canada are highly variable, both temporally and
+spatially. Over the time periods for which data were available, some
+populations show positive trends in survival in the first winter at sea (\So)
+while other exhibit highly variable yet stable trends, and some show declines.
+We could not assess trends in the second winter at sea (\St) or proportion
+returning as grilse (\Pg), as these parameter estimates were highly uncertain
+and were strongly influenced by the priors. Perhaps there is a need to rethink
+our understanding of Atlantic salmon population dynamics in light of the
+possibilities that (1) any real and persistent declines in marine survival
+have been experienced by some but not necessarily all populations, (2)
+reductions in survival might have occurred over a relatively brief period of
+time and have not persisted, (3) marine survival has been relatively stable,
+or increasing in some populations for one or more decades, (4) variation in
+survival at the population level can be large even within a region.
 
 % 2. How they compare to other publications
-Our results are contrary to those of \citet{Olmos2019}, who detected mostly positive
-correlations in post-smolt survival among spatially broad stock units. These differences
-could be due to a number of reasons: different model specifications and
-structure, different methods for estimating covariance, difference in the
-spatial scales of data sources (i.e. river vs province scales), and perhaps most importantly the use stock-recruitment relationships
-rather than empirical smolt count estimates to estimate marine survival.
-As trends in marine survival during the first winter at sea are highly independent
-among rivers on relatively small spatial scale, trends from broader
-geographical areas (i.e. province, state, or country-wide estimates) may not
-be representative.
-Interestingly, our estimates of \So and \St are very similar to those produced
-by \citet{Chaput2003b}, and our trends are almost identical for the
-overlapping time period that marine survival was estimated for in their study
-(1984-1998). 
+% Population vs regional trends
+
+Interestingly, our estimates of both \So and \St for the Trinit\'{e}
+population are very similar to those produced by \citet{Chaput2003b} using a
+two sex model, and our trends are almost identical for the overlapping time
+period that marine survival was estimated for in their study (1984-1998). 
 While \citet{Chaput2003b} separated abundance data for males and females
-and assumed their survival rates were the same (to be able to reach an
-analytical solution), our study reached almost the same results (albeit with
+based on sex ratio information and assumed their survival rates were the 
+same (to be able to reach an analytical solution), our study reached almost the same results (albeit with
 slightly higher uncertainty), using a Bayesian approach with informative
 priors. These overlapping trends obtained with two different methods 
-suggest that our method is effective at estimating marine survival.
+suggest that our method is indeed effective at estimating marine survival.
 
@@
 being comparable to those in the late 2000s.
-\citet{Friedland1993} compared return rates for a number of rivers in eastern
+\citet{Friedland1993} compared return rates for a number of populations in eastern
 North America between 1973 and 1988, and suggested there are similar trends among these. 
 However, the similarity in these trends was driven primarily by two years, 1977 and 1978, which
-show concurrent low and high relative return rates across rivers,
+show concurrent low and high relative return rates across populations,
 respectively. Other years are much more variable relative to each other.
@@
 data from this decade.
-In any event, we caution that the pooling of adult return rates \citep{Chaput2012a, Friedland1993,Gibson2016} 
-can mask inter-annual variability in marine survival,
-and hence might not produce an accurate depiction of marine survival trends.
+
+In any event, we caution that the pooling of adult return rates
+\citep{Chaput2012a, Friedland1993,Gibson2016} can mask inter-annual
+variability in marine survival \citep{Hubley2011}, and hence might not produce
+an accurate depiction of marine survival trends.
 While river-specific return rate estimates are available for the populations
@@
 population's life histories. The estimates of return rates of one sea-winter
-should approximate marine survival in the first year at sea for one
-sea-winter-dominated populations. However, as the proportion of fish that
+approximate marine survival in the first year at sea for one
+sea-winter-dominated populations, and we see this by comparing the three NL populations
+with the return rates presented in \citet{ICES2019}. However, as the proportion of fish that
 return as one sea-winter decreases, marine survival in the first year at sea
@@
 providing a very coarse estimate of marine survival in the second year at sea. 
-In fact, changes in return rates of two sea-winter fish are a result of a
+As changes in return rates of two sea-winter fish are a result of a
 combination of changes in survival in the first year at sea in the year prior,
 survival in the second year at sea, and in the proportion returning as two
-sea-winter, and an increase in any of those parameters while the others remain
+sea-winter, an increase in any of those parameters while the others remain
 the same will result in an increase in return rates. Our model attempts to
@@
 \citet{Dempson2003} described a general declining trend in marine survival for
-Newfoundland rivers (except WAB); we drew the same conclusion for 
+Newfoundland populations (except WAB); we drew the same conclusion for 
 Conne River but not Campbellton River or WAB. It is not possible to draw broader conclusions
-with data from only three Newfoundland rivers, but it seems that among index rivers,
+with data from only three Newfoundland populations, but it seems that among index rivers,
 those in Newfoundland are among those with the highest marine survival rates.
 
+At first glance, the large variation in marine survival trends among the seven
+populations examined in our study seems to be at odds with the synchronous
+trends of declining marine survival detected at broad regional scales
+\citep{Olmos2019}. However, these two observations are not necessarily mutually
+exclusive: our study is only a small subsample of the populations in the region,
+and highly variable trends in marine survival among populations can still add
+up to overall negative trends at a regional scale.
+Aside from the overarching difference in the spatial scale of data sources
+(i.e. river vs province scales), the divergence in local versus regional
+trends could be further confounded by differences in model structure, such as
+the methodology for estimating covariance, and use stock-recruitment
+relationships rather than empirical smolt count estimates to estimate marine
+survival.
+% As trends in marine survival during the first winter at sea are highly
+% independent among rivers on relatively small spatial scale, 
+Our results suggest that trends from broader geographical areas (i.e.
+province, state, or country-wide estimates) may not be representative of
+individual populations, and concomitantly, trends of individual populations
+may also not be representative of region-wide trends.
+
 % 3. Potential reasons
-There are a variety of potential explanations for the lack of synchronous
-trends in estimates of \So. 
-Marine survival in the first winter at sea could be highly variable between
-populations because of the predominance of spatially local environmental drivers of survival (e.g., temperature, predation, prey abundance) 
+There are a variety of potential explanations for the highly variable trends
+in estimates marine survival in the first winter at sea.
+Environmental drivers of survival (e.g., temperature, predation, prey abundance) 
+could be local environmental drivers
 relative to broader-scale, even ocean-wide, drivers.
-The synchrony reported for marine survival trends at broader spatial scales \citep{Massiot-Granier2014,Olmos2019}
-might be attributable to the use of stock-recruitment relationships to estimate survival,
-relationships that may have been confounded by changes in recruitment dynamics.
+A recent study showed that correlations between oceanic temperature, primary
+productivity, and post smolt survival of Atlantic salmon in the North Atlantic
+are better explained by broad-scale environmental trends than regional ones
+\citep{Olmos2020}.
+However these local drivers are potentially much more difficult to quantify than 
+broader ones.
+% The synchrony reported for marine survival trends at broader spatial 
+% scales \citep{Massiot-Granier2014,Olmos2019} might be attributable to the 
+% use of stock-recruitment relationships to estimate survival,
+% relationships that may have been confounded by changes in recruitment dynamics.
 There is some evidence of a correlation between return rate and growth (as
@@
 tended to also be years of poor survival \citep{Friedland1993}, supporting the
-idea that environmental variability can affect marine survival.
+idea that marine survival is mediated by environmentally-driven changes in growth.
 Furthermore, among European salmon, there is evidence of a positive correlation
-between spring temperature in the Norwegian and North Seas and population abundance, suggesting warmer
+between spring temperature in the Norwegian and North Seas and population abundance, 
+suggesting warmer
 conditions favour post-smolts \citep{Friedland1998}, based on mapping the
@@
 temperature experienced by the post-smolts and their respective
-population-specific thermal optima. 
+population-specific thermal optimum. 
 This difference could explain why populations in eastern North America are
@@
 while others find negative ones \citep{Friedland1993, Todd2008}.
+\citet{Olmos2020} found a positive relationship between temperature and marine survival 
+in northern regions and a negative one in southern regions, providing 
+evidence of differing mechanisms across latitudes.
 Putative associations between temperature and direct estimates of marine
@@
 
-Oceanic conditions have been correlated with abundance trends and growth
-\citep{Todd2008}, however, the mechanism by which such bottom-up effects,
+Oceanic conditions have been correlated with abundance trends, growth, and marine survival
+in Atlantic salmon, which are thought
+to be mediated by bottom-up effects driving ocean productivity and food availability
+ \citep{Todd2008, Renkawitz2015,Olmos2020}. 
+However, the mechanism by which such bottom-up effects,
 mediated by changes in food availability,
 affect population dynamics beyond marine survival needs to
-be thoroughly reassessed. If marine survival on its own cannot fully explain
-trends in abundance, then there are potential carry-on effects of oceanic
-conditions that manifest with regards to freshwater production. 
+considered further. 
+There are potential carry-on effects of oceanic
+conditions that manifest with regards to freshwater production for individuals that survive
+the marine phase of their life cycle. 
 For example, adults
-that return to spawn after spending suboptimal conditions at sea might be less
-likely to make it to their spawning grounds, successfully secure a mate,
-produce fewer eggs, or produce eggs with lower per capita fitness than those
+that return to their natal streams after spending suboptimal conditions at sea might be less
+likely to make it to the spawning grounds or secure a mate, and might also
+produce fewer eggs or eggs with a lower energetic content than those
 produced by adults which grew in optimal oceanic conditions.
@@
 
-Egg-to-smolt survival in Atlantic salmon is highly variable \citep{Klemetsen2003,Chaput2015}.
+Egg-to-smolt survival in Atlantic salmon is highly variable
+\citep{Klemetsen2003,Chaput2015}.
 While this variability is likely attributable to changes in freshwater
@@
 contributing to this variability.
+
 Obviously, there would be a time lag (perhaps as much as a generation) in how such effects
@@
 While there are analytical issues associated with the estimation of \So, \St, and \Pg,
-the assumption that \St is additive to \So could produce unrealistic results.
+the assumption that survival in the second year at sea is additive to survival in the first 
+year at sea could produce unrealistic results.
 We know there is a period of a few months where 1SW 
@@
 it would be if not modelled hierarchically. 
-That said, this assumption is likely to result in more conservative trends in marine survival, as the
-variation is smolt estimates is reduced.
+That said, this assumption is likely to result in more conservative trends in
+marine survival, as the variation is smolt estimates is reduced.
+This shrinkage of annual smolt abundances is also reducing the $R^2$ values 
+between smolts and estimates returns, albeit only marginally.
 It is important to note that the hierarchical structure in the estimation of
@@
 for different populations. For example, we know that survival in the first
-year at sea is an important driver of our estimates of 1SW returns, except for
+year at sea is the variable that explains most of the variance in estimates of 1SW returns, except for
 the Saint-Jean (Fig.~\ref{fig:mu-rsq}). Thus, estimates of \So for this
@@
 
+Lastly, our model does not differentiate mortality resulting from natural
+causes (e.g. predation) or from fishing; we know the latter has decreased
+due to the sequential moratoria enacted across eastern Canada, and this
+confounds the estimated trends in marine survival because we know marine survival
+will be reduced in years where commercial fishing occurred (i.e. pre-2000s).
+Thus, any increasing trends in marine survival are confounded by the additional 
+fishing mortality before the turn of the century.
+However, for most populations, present-day marine survival is fluctuating around roughly 
+similar levels as it was since the 1980s, even after the fishing moratoria 
+were introduced as a response to low population sizes, thus
+suggesting a potential weakening of compensatory population dynamics.
+
 % 5. Future directions
@@
 have declined consistently, and over the same time periods, across all
-populations. 
-But the fact that it has remained at roughly similar levels as it was
-since the 1980s \emph{despite} reduced commercial fishing mortalities, suggests
+populations, even if the overall trend has been one of decline. 
+But the fact that, for most of the populations examined, it has remained at roughly 
+similar levels as it was since the 1980s \emph{despite} reduced commercial 
+fishing mortalities, suggests
 that there may well be an interaction between small population size (small
@@
 
-We would like to thank G\'{e}rald Chaput for his useful discussions and
-comments on the manuscript, Carmen David for her comments
-on the manuscript, and Sean Anderson for his help with implementing the
+We would like to thank Martha Robertson, G\'{e}rald Chaput, and Carmen David their useful 
+comments on the manuscript, and Sean Anderson for his help with implementing the
 non-centered parameterization of the model. This research was supported by the

### 2020-05-22 analysis/99-data-exports.R
Add data export code

@@ -1,0 +1,65 @@
+# Data export to share with Maria Tirronen for change point analysis
+library(tidyverse)
+library(R.matlab)
+alldata <- readRDS("data/alldata.rds")
+alldatastan <- readRDS("data/alldatastan.rds")
+
+
+str(alldata)
+
+alldata$logRS <- NULL
+alldata$logis_mu <- matrix(alldatastan$logis_mu, ncol = 1)
+alldata$logis_sigma <- matrix(alldatastan$logis_sigma, ncol = 1)
+
+names(alldata)[names(alldata) == "loggrilse"] <- "logSW1"
+
+saveRDS(alldata, "data/alldata_export.rds")
+
+writeMat("data/alldata.mat", alldata = alldata)
+writeMat("data/alldatasep.mat", 
+         years = alldata$years,
+         logsmolts = alldata$logsmolts,
+         logsmolts_cv = alldata$logsmolts_cv,
+         logSW1 = alldata$logSW1,
+         logSW2 = alldata$logSW2,
+         returns_cv = alldata$returns_cv,
+         nyears = alldata$nyears,
+         river_name = alldata$river_name,
+         logis_mu = alldata$logis_mu,
+         logis_sigma = alldata$logis_sigma
+         )
+
+# In the end we didn't use MATLAB export and went with a simple csv table instead:
+
+alldatastan <- readRDS("data/alldatastan.rds")
+
+returnsdata <- alldatastan[names(alldatastan) %in% c("years", "logsmolts", "logsmolts_cv", "loggrilse", "logSW2")] %>% do.call(cbind, .) %>% as_tibble
+
+otherdata <- tibble(nyears = alldatastan$nyears,
+       river_name = alldatastan$river_name,
+       returns_cv = alldatastan$returns_cv,
+       logis_mu = alldatastan$logis_mu,
+       logis_sigma = alldatastan$logis_sigma,
+       ny = nyears) %>% uncount(ny)
+
+alldata_table <- bind_cols(returnsdata, otherdata) %>%
+  select(river_name, year = years, logsmolts, logsmolts_cv, 
+         logSW1 = loggrilse, logSW2, returns_cv, everything())
+
+write.csv(alldata_table, file = "data/alldata_table.csv")
+
+
+# Save csvs for sharing data with colleagues
+
+# S1 median estimates for regime shift analysis (Anna)
+s1quant <- readRDS("data/s1quant.rds")
+
+s1medians <- s1quant %>%
[…]

### 2020-06-07 ms/ms.tex
Incorporate Geir's and Brian's suggestions

@@
 Despite the overall decreases in marine survival,
-the conservation status of Canadian salmon populations differ
+the conservation status of Canadian salmon populations differs
 considerably. In Canada, populations in the southern part of their range are more
@@
 For further details on the data collection methodologies refer to 
-\citet{Dempson1991, Venoitt2018} for NL populations, 
+\citet{Dempson1991}, \citet{Schwarz1994}, and \citet{Venoitt2018} for NL populations, 
 \citet{April2018}  for QC populations,
@@
 deterministically as shown above. However, sampling a parameter from a
-hypergeometric distribution is not possible in programming the model was
-written in (i.e. Stan).
+hypergeometric distribution is not possible in the programming language the model was
+written (i.e. Stan).
 
@@
 coefficient of variation in the empirically derived
-smolt estimates (see Tables S1 and S2 in the Supplementary material). 
+smolt estimates (see Table S1 in the Supplementary material). 
 Where available, we used population-specific measurement error estimates for smolt abundances; if not 
-available, we set measurement error between 5 and 10\%. 
+available, we set measurement error between 5 and 10\% (see Table S2 in the Supplementary material). 
 The log-transformed true smolt abundances are
@@
 These error terms are almost equivalent to using a coefficient of variation in
-return estimates 1\%, given that return estimates in equations~\ref{eq:3}
+return estimates of 1\%, given that return estimates in equations~\ref{eq:3}
 and~\ref{eq:4} are log-transformed. As
@@
 informative only to the extent that they limit the estimates of marine survival
-to what what we considered to be biologically realistic.
+to what we considered to be biologically realistic.
 
@@
 
-The three Newfoundland populations (WAB, Campbellton and Conne) are considered to be 
+The three Newfoundland populations (WAB, Campbellton and Conne) are 
 1SW-dominated, while the other four have the more generic priors for $logit(P_{g,t})$.
@@
 
-Trends in estimates of marine survival in the first winter at sea (\So) were 
+Estimates of marine survival probabilities in the first winter at sea (\So) were 
 highly variable within and among populations
@@
 with an estimate of \So of 0.007, while the Conne, LaHave, and Trinit\'{e} had
-years where estimates of \So varied between 0.01 and 0.02 (Fig.~\ref{fig:s1-faceted}).
-
-Trends among populations also varied: Campbellton,
+years where the lowest estimates of \So varied between 0.01 and 0.02 (Fig.~\ref{fig:s1-faceted}).
+
+Long-term trends among populations also varied: Campbellton,
 Saint-Jean, and Western Arm Brook populations showed increases in \So
@@
 while at Nashwaak there was an increase in median \So during the early
-2000s but a decrease in the 2010s. Yearly estimates of \So had very little
-variability for one sea-winter dominated populations (Conne, Campbellton, and
-WAB), but were more variable (i.e. wider credible intervals) in the other
+2000s but a decrease in the 2010s. Annual estimates of \So had little
+uncertainty for one sea-winter dominated populations (Conne, Campbellton, and
+WAB), but uncertainty was higher (i.e. wider credible intervals) in the other
 populations.
 
+% % Dual plot of trends in S1
+% \begin{figure}[htbp] \centering
+%     \includegraphics[width=0.95\linewidth]{figures/s1-trends-dual.png}
+%     \caption{Trends in survival in the first year at sea (\So). a) Posterior
+%         estimates of \So for the seven populations examined, error bars indicate
+%         the 90\% credible intervals, and dashed lines denote years of commercial
+%         fishing moratoria for each province. b) Z-scores of median posterior
+%         estimates.} \label{fig:s1-dual} 
+% \end{figure}
+
+% Single plot of trends in S1
 \begin{figure}[htbp] \centering
-    \includegraphics[width=0.95\linewidth]{figures/s1-trends-dual.png}
-    \caption{Trends in survival in the first year at sea (\So). a) Posterior
-        estimates of \So for the seven populations examined, error bars indicate
-        the 90\% credible intervals, and dashed lines denote years of commercial
-        fishing moratoria for each province. b) Z-scores of median posterior
-        estimates.} \label{fig:s1-dual} \end{figure}
+    \includegraphics[width=0.95\linewidth]{figures/s1-trends-all.png}
+    \caption{Median posterior estimates of survival in the first year at sea (\So) 
+        for the seven populations examined.} \label{fig:s1-dual} 
+\end{figure}
 
@@
     \includegraphics[width=0.95\linewidth]{figures/s1-trends-faceted.png}
-    \caption{Posterior estimates of \So for the seven populations examined, error
-        bars indicate the 90\% credible intervals.} \label{fig:s1-faceted}
+    \caption{Posterior estimates of survival probabilities in the first year at sea (\So). Dots indicate median estimates and error
+        bars indicate the 90\% credible intervals. Dashes vertical lines reflect the year the commercial fishing moratoria were enacted
+    in each province.} \label{fig:s1-faceted}
 \end{figure}
 
-Estimates of \St were highly uncertain in all populations, and trends 
+Estimates of survival in the second year at sea (\St) were highly uncertain in all populations, and trends 
 were not apparent in most populations given the large range of the credible
@@
     \includegraphics[width=0.95\linewidth]{figures/s2-trends-faceted.png}
-    \caption{Posterior estimates of \St for the seven populations examined, error
-        bars indicate the 90\% credible intervals.} \label{fig:s2-faceted}
+    \caption{Posterior estimates of survival in the second yyear at sea (\St).
+ Dots indicate median estimates and error bars indicate the 90\% credible
+ intervals. Dashes vertical lines reflect the year the commercial fishing
+ moratoria were enacted in each province.} \label{fig:s2-faceted}
 \end{figure}
@@
     \includegraphics[width=0.95\linewidth]{figures/pr-trends-faceted.png}
-    \caption{Posterior estimates of \Pg for the seven populations examined, error
-        bars indicate the 90\% credible intervals. Year corresponds to the
-        year of 1SW return.} \label{fig:pr-faceted} 
+    \caption{Posterior estimates of proportion returning after one winter at sea (\Pg).
+ Dots indicate median estimates and error
+        bars indicate the 90\% credible intervals. Dashes vertical lines reflect the year the commercial fishing moratoria were enacted
+    in each province.} \label{fig:pr-faceted} 
 \end{figure}
 
-Population-level estimates of \prmu and \prsig varied considerably among populations. For all
+Estimates of the population-level mean (\prmu) and standard deviation (\prsig) of proportion returning after one year at sea varied considerably among populations. 
+For all
 three 1SW-dominated populations (Campbellton, Conne, and WAB), estimates of \prmu
-were very close to 1.0 and had little variability in \prsig (Fig~\ref{fig:prmu-post}; Table~\ref{tab:prmu}).
+were very close to 1.0 and had little variation in \prsig (Fig~\ref{fig:prmu-post}; Table S7 in the Supplementary Materials).
 Estimates of \prmu were the lowest for the two QC populations, particularly the Saint-Jean (median \prmu = 0.11).
 Estimates of \prmu for the Nashwaak and the LaHave populations were close to 0.5, with these two populations having 
-the highest estimated values of \prsig, particularly the Nashwaak (Fig~\ref{fig:prmu-post}; Table~\ref{tab:prsigma}).
-
-\input{table-logitprmu.tex}
-
-\input{table-logitprsigma.tex}
+the highest estimated values of \prsig, particularly the Nashwaak (Fig~\ref{fig:prmu-post};  Table S8 in the Supplementary Materials).
+
+% \input{table-logitprmu.tex}
+
+% \input{table-logitprsigma.tex}
 
@@
 survival in the first year (i.e. $-Z_1$) and proportion returning as 1SW,
-except for the Campbellton, where this correlation is negative.
+except for the Campbellton population, where this correlation is negative.
 
@@
     \includegraphics[width=0.95\linewidth]{figures/corrected-cor-ALL.png}
-    \caption{Estimated correlations among estimated model parameters.}
+    \caption{Correlations among estimated model parameters. \emph{log(smolts)} = estimated smolt abundances in log-space, $log(P_g)$ = logit-transformed probability of
+        returning after one year at sea, $-Z_1$ = marine survival in the first year at sea, and $-Z_2$ = marine survival in the second year at sea.
+        Note the correlations derived from negative instantaneous total mortalities $-Z_1$ and $-Z_2$ are equivalent to those of marine survival $S_1$ and $S_2$.}
     \label{fig:cor-params} 
@@
 The low $R^2$ for the Saint-Jean is likely due to the negative correlation
-between survival and number of outmigrating smolts (Fig.~\ref{fig:cor-params}). 
+between survival (i.e. $-Z_1$) and number of outmigrating smolts (see Fig.~\ref{fig:cor-params}). 
+There are also moderately negative correlations between survival and smolts in
+the Campbellton, LaHave, and Nashwaak populations.
+These negative correlations could be indicative of density-dependent
+mortality at sea, however this is generally thought to be is unlikely.
 Populations with a lower proportion of 1SW in their returns (e.g. Trinit\'{e}
@@
 other four populations.
-
 Values of $R^2$ were overall low for all parameters with regards to explaining
-the variance of estimated 2SW returns. Estimates of survival in the second winter at sea ($Z_2$)
+the variance of estimated 2SW returns. Estimates of mortality in the second winter at sea ($Z_2$)
 explained little variance in estimates for 2SW returns, with the Saint-Jean again being the
@@
 
-% Our results challenge the narrative that marine survival, specifically
-% survival in the first year at sea, is declining uniformly throughout the
-% range of Atlantic salmon in the northwest Atlantic.
-Our results indicate that trends in marine survival among Atlantic salmon
+Our results show that trends in marine survival among monitored Atlantic salmon
 populations in eastern Canada are highly variable, both temporally and
@@
 populations show positive trends in survival in the first winter at sea (\So)
-while other exhibit highly variable yet stable trends, and some show declines.
-We could not assess trends in the second winter at sea (\St) or proportion
-returning as grilse (\Pg), as these parameter estimates were highly uncertain
-and were strongly influenced by the priors. Perhaps there is a need to rethink
-our understanding of Atlantic salmon population dynamics in light of the
-possibilities that (1) any real and persistent declines in marine survival
-have been experienced by some but not necessarily all populations, (2)
-reductions in survival might have occurred over a relatively brief period of
-time and have not persisted, (3) marine survival has been relatively stable,
-or increasing in some populations for one or more decades, (4) variation in
-survival at the population level can be large even within a region.
+while other exhibits highly variable yet stable trends, and some show declines.
+Variation in survival at the population level can be large even within
+a region, and there are no consistent declines in marine survival in any
+population after the onset of fishing moratoria, except perhaps in the LaHave
+River where the four last years in the time-series (2014-2017) had survival of
+less than 2.5\%.
+However, while correlations in marine survival trends between rivers were highly 
+variable (Fig.~\ref{fig:s1-corr}),
+there are some years where the populations appear to behave in concert:
+years 1997, 2001, and 2007 have consistently low survival across all populations 
+(i.e. <5\%, Fig.~\ref{fig:s1-dual}) and little variation across populations.
+Our results also show that variation of survival in first year at sea is the most
+important factor determining variation in number of returns, and suggests that 
+changes in abundance are primarily driven by changes in marine survival. 
+Nonetheless, variation in number of out-migrating smolts, and for some populations
+variation in probability of returning after one winter at sea, also explain
+variation in returns and thus contribute to the observed variation in return abundances.
+Furthermore, the negative correlation between marine survival and smolt
+abundances in some populations is suggestive of density-dependent processes
+occurring at sea.
+We could not assess trends for survival in the second winter at sea (\St) or
+proportion returning as grilse (\Pg), as these parameter estimates were highly
+uncertain and were strongly influenced by the priors.
+
+% Perhaps there is a need to rethink
+% our understanding of Atlantic salmon population dynamics in light of the
+% possibilities that (1) any real and persistent declines in marine survival
+% have been experienced by some but not necessarily all populations, (2)
+% reductions in survival might have occurred over a relatively brief period of
+% time and have not persisted, (3) marine survival has been relatively stable,
+% or increasing in some populations for one or more decades, (4) variation in
+% survival at the population level can be large even within a region.
 
@@
 % Population vs regional trends 
-
-Interestingly, our estimates of both \So and \St for the Trinit\'{e}
-population are very similar to those produced by \citet{Chaput2003b} using a
-two sex model, and our trends are almost identical for the overlapping time
-period that marine survival was estimated for in their study (1984-1998). 
-While \citet{Chaput2003b} separated abundance data for males and females
-based on sex ratio information and assumed their survival rates were the 
-same (to be able to reach an analytical solution), our study reached almost the same results (albeit with
-slightly higher uncertainty), using a Bayesian approach with informative
-priors. These overlapping trends obtained with two different methods 
-suggest that our method is indeed effective at estimating marine survival.
-
+At first glance, the large variation in marine survival trends among the seven
+populations examined in our study seems to be at odds with the synchronous
+trends of declining marine survival detected at broad regional scales
+\citep{Olmos2019}. However, these two observations are not necessarily mutually
+exclusive: our study is only a small subsample of the populations in the region,
+and, with one exception, over a much shorter period of time,
+and highly variable trends in marine survival among populations can still add
+up to overall negative trends at a regional scale.
+Aside from the overarching difference in the spatial scale of data sources
+(i.e. river vs province scales), the divergence in local versus regional
+trends could be further confounded by differences in model structure, such as
+the methodology for estimating covariance, and use stock-recruitment
+relationships rather than empirical smolt count estimates to estimate marine
+survival.
+% As trends in marine survival during the first winter at sea are highly
+% independent among rivers on relatively small spatial scale, 
+Our results suggest that trends from broader geographical areas (i.e.
+province, state, or country-wide estimates) may not be representative of
+individual populations, and concomitantly, trends of individual populations
+may also not be representative of region-wide trends.
+
+% Trends in survival at the population level
 Trends in marine survival
@@
 time series in our study.
-While declines in return rates since the 1970s seem to be consistent among populations, 
-we were unable to assess if marine survival estimates were also higher in the 1970s as we lacked 
-data from this decade.
-
+While declines in return rates since the 1970s seem to be consistent among
+populations, we were unable to assess if marine survival estimates were also
+higher in the 1970s as we lacked data from this decade.
+\citet{Dempson2003} described a general declining trend in marine survival for
+Newfoundland populations (except WAB); we drew the same conclusion for Conne
+River but not Campbellton River or WAB. It is not possible to draw broader
+conclusions with data from only three Newfoundland populations, but it seems
+that among index rivers, those in northern Newfoundland are among those with
+the highest marine survival rates.
+
+% Return rates vs marine survival estimates
 In any event, we caution that the pooling of adult return rates
 \citep{Chaput2012a, Friedland1993,Gibson2016} can mask inter-annual
-variability in marine survival \citep{Hubley2011}, and hence might not produce
+variation in marine survival \citep{Hubley2011}, and hence might not produce
 an accurate depiction of marine survival trends.
@@
 of marine survival with varying degrees of similarity depending on each
-population's life histories. The estimates of return rates of one sea-winter
+population's life histories. The estimates of return rates of one sea-winter salmon
 approximate marine survival in the first year at sea for one
-sea-winter-dominated populations, and we see this by comparing the three NL populations
+sea-winter-dominated populations, particularly for years in which 
+directed marine fisheries are largely non-existent,
+and we see this by comparing the three NL populations
 with the return rates presented in \citet{ICES2019}. However, as the proportion of fish that
 return as one sea-winter decreases, marine survival in the first year at sea
-can increasingly diverge from return rates of fish, and trends in return rate
+can increasingly diverge from return rates. Thus, trends in return rate
 would be particularly misleading if the proportion returning as one sea-winter
-varies in time. 
+varies in time;
+proportion returning as one sea-winter was in important factor driving number of 
+1SW returns in three of the populations examined (Fig.~\ref{fig:mu-rsq}).
 Furthermore, return rates of two sea-winter fish also include survival in the
@@
 
-\citet{Dempson2003} described a general declining trend in marine survival for
-Newfoundland populations (except WAB); we drew the same conclusion for 
-Conne River but not Campbellton River or WAB. It is not possible to draw broader conclusions
-with data from only three Newfoundland populations, but it seems that among index rivers,
-those in Newfoundland are among those with the highest marine survival rates.
-
-At first glance, the large variation in marine survival trends among the seven
-populations examined in our study seems to be at odds with the synchronous
-trends of declining marine survival detected at broad regional scales
-\citep{Olmos2019}. However, these two observations are not necessarily mutually
-exclusive: our study is only a small subsample of the populations in the region,
-and highly variable trends in marine survival among populations can still add
-up to overall negative trends at a regional scale.
-Aside from the overarching difference in the spatial scale of data sources
-(i.e. river vs province scales), the divergence in local versus regional
-trends could be further confounded by differences in model structure, such as
-the methodology for estimating covariance, and use stock-recruitment
-relationships rather than empirical smolt count estimates to estimate marine
-survival.
-% As trends in marine survival during the first winter at sea are highly
-% independent among rivers on relatively small spatial scale, 
-Our results suggest that trends from broader geographical areas (i.e.
-province, state, or country-wide estimates) may not be representative of
-individual populations, and concomitantly, trends of individual populations
-may also not be representative of region-wide trends.
-
-% 3. Potential reasons
+Interestingly, our estimates of survival in both first and second years at sea
+for the Trinit\'{e} population are very similar to those produced by
+\citet{Chaput2003b} using a two sex model, and our trends are almost identical
+for the overlapping time period that marine survival was estimated for in
+their study (1984-1998). While \citet{Chaput2003b} separated abundance data
+for males and females based on sex ratio information and assumed their
+survival rates were the same (to be able to reach an analytical solution), our
+study reached almost the same results (albeit with slightly higher
+uncertainty), using a Bayesian approach with informative priors. These
+overlapping trends obtained with two different methods suggest that our method
+is indeed effective at estimating marine survival.
+
+% 3. Potential drivers of patterns 
 There are a variety of potential explanations for the highly variable trends
 in estimates marine survival in the first winter at sea.
-Environmental drivers of survival (e.g., temperature, predation, prey abundance) 
-could be local environmental drivers
+Environmental drivers of survival (e.g., temperature, predation, prey abundance, 
+interactions with farmed salmon) could be highly localized 
 relative to broader-scale, even ocean-wide, drivers.
@@
 survival warrants further study at the population level.
-
-While there is little evidence that marine survival is density-dependent in
-Atlantic salmon \citep{Jonsson1998,Gibson2006}, there could potentially be
-some density-dependent processes during parts of the marine phase, 
-particularly for populations that are likely to be subjected to
-declining per capita population growth rates ($r$) generated by Allee effects.
-Exploring relationships between survival and population size could potentially
-shed light about the processes that have caused many of the population
-declines that have been documented.
 
@@
 
-Egg-to-smolt survival in Atlantic salmon is highly variable
-\citep{Klemetsen2003,Chaput2015}.
-While this variability is likely attributable to changes in freshwater
+Egg-to-smolt survival in Atlantic salmon is highly variable, and perhaps more so
+than marine survival \citep{Klemetsen2003,Chaput2015}.
+While this variation is likely attributable to changes in freshwater
 conditions (e.g. discharge, temperature, water quality) and uncertainty in
@@
 changes in the oceanic conditions that spawners experience could also be
-contributing to this variability.
-
+contributing to this variation.
+Nonetheless, our results suggests that while variation in number of smolts was 
+important for number of returns, it was considerable less important than
+survival the first year at sea, which is indicative that return abundances 
+are most strongly influenced by the marine phase of their life cycle.
 Obviously, there would be a time lag (perhaps as much as a generation) in how such effects
@@
 maintained even if salmon abundances were lagged by a generation length.
+
+While there is little evidence that marine survival is density-dependent in
+Atlantic salmon \citep{Jonsson1998,Gibson2006}, there could potentially be
+some density-dependent processes during parts of the marine phase, 
+particularly for populations that are likely to be subjected to
+declining per capita population growth rates ($r$) generated by Allee effects.
+While the correlations between return estimates and survival were negative for 
+some populations (Fig.~\ref{fig:cor-params})
+indicate that there may be some compensatory density dependent effects in the 
+marine phase, our analysis provide no evidence of Allee effects.
+Exploring relationships between survival and population size could potentially
+shed light about the processes that have caused many of the population
+declines that have been documented.
 
@@
 for modelling smolt abundances. This approach results in shrinkage to the
-mean, which means that the variability of annual smolt estimates is less than
+mean, which means that the variation in annual smolt estimates is less than
 it would be if not modelled hierarchically. 
@@
 Thirdly, the accuracy and precision of the model's output could be different
-for different populations. For example, we know that survival in the first
+for different populations. Survival in the first
 year at sea is the variable that explains most of the variance in estimates of 1SW returns, except for
-the Saint-Jean (Fig.~\ref{fig:mu-rsq}). Thus, estimates of \So for this
+the Saint-Jean (Fig.~\ref{fig:mu-rsq}).
+
+The model does not provide reliable parameter estimates for the Saint-Jean
+population (as indicated by the low $R^2$ values in Fig.~\ref{fig:mu-rsq}),
+which is the population with the highest proportion of 2SW returns.
+
+Thus, estimates of \So for this
 population, which has the lowest proportion of 1SW returns out of all seven
@@
 fishing mortality before the turn of the century.
-However, for most populations, present-day marine survival is fluctuating around roughly 
-similar levels as it was since the 1980s, even after the fishing moratoria 
-were introduced as a response to low population sizes, thus
+However, for most of our studied populations, 
+present-day marine survival fluctuates to values
+as high as those recorded in the past, thus
 suggesting a potential weakening of compensatory population dynamics.
@@
 populations, even if the overall trend has been one of decline. 
-But the fact that, for most of the populations examined, it has remained at roughly 
-similar levels as it was since the 1980s \emph{despite} reduced commercial 
-fishing mortalities, suggests
+But the fact that, for most of the populations examined, 
+present-day survival fluctuating around similar levels  
+similar levels as in the past, 
+\emph{despite} reduced commercial fishing mortalities, suggests
 that there may well be an interaction between small population size (small
@@
 
-We would like to thank Martha Robertson, G\'{e}rald Chaput, and Carmen David their useful 
-comments on the manuscript, and Sean Anderson for his help with implementing the
-non-centered parameterization of the model. This research was supported by the
-Atlantic Salmon Conservation Foundation and the Atlantic Salmon Research Joint
-Venture.
+We would like to thank Martha Robertson, G\'{e}rald Chaput, and Carmen David
+their useful comments on the manuscript, and Sean Anderson for his help with
+implementing the non-centered parameterization of the model. Hydro-Québec
+contributed to data acquisition from Rivière de la Trinité. This research was
+supported by the Atlantic Salmon Conservation Foundation and the Atlantic
+Salmon Research Joint Venture.
 

### 2020-06-07 analysis/05-figures-for-trends-paper.R

@@ -22,16 +22,24 @@ addnas <- tibble(pos = NA,
 
 s1quant <- bind_rows(s1quant, addnas)
 
+morat2 <- tibble(year = c(1984, 1984, 1998, 1998, 1998, 2000, 2000),
+                 region = c("NB","NS", "NL", "NL", "NL", "QC", "QC"),
+                 river_name = c("Nashwaak River", "LaHave River", 
+                                "Western Arm Brook", "Campbellton River", "Conne River", 
+                                "Trinité River", "Saint-Jean River"))
+
 s1trends <- ggplot(s1quant, aes(year, median)) + 
   geom_line(alpha = 0.5) + geom_point(size = 1, alpha = 0.7) + 
   ylab(expression(paste(S[1]~"estimates", sep = ""))) +
   geom_errorbar(aes(ymin = q05, ymax = q95),width= 0.1) +
+  geom_vline(data = morat2, aes(xintercept = year), alpha = 0.4, linetype = 2) +
   facet_wrap(~river_name, ncol = 2, scales = "free_y") + theme_bw() +
   xlab("Year") +
   labs(color = "River") +
   theme(legend.position = "none", 
         legend.text =  element_text(size = 12))
 s1trends
+
 ggsave("figures/s1-trends-faceted.png", height = 6, width = 5)
 
 s1trends + facet_wrap(~river_name, ncol = 3, scales = "free_y") 
@@ -39,12 +47,13 @@ ggsave("figures/s1-trends-presentation.png", height = 4, width = 8)
 
 
 
-
 morat <- tibble(year = c(1984, 1998, 2000),
                 region = c("NS+\nNB", "NL", "QC"),
                 river_name = c("Nashwaak River", "Western Arm Brook", "Trinité River"))
 
-s1trends <- ggplot(s1quant, aes(year, median, color = river_name)) + 
+
+
+s1trendsall <- ggplot(s1quant, aes(year, median, color = river_name)) + 
   geom_line(alpha = 0.5) + geom_point(size = 2, alpha = 0.8) +
   ylab(expression(paste(S[1]~"posterior estimates", sep = ""))) +
   geom_errorbar(aes(ymin = q05, ymax = q95), alpha = 0.5, width = 0.1) +
@@ -52,8 +61,24 @@ s1trends <- ggplot(s1quant, aes(year, median, color = river_name)) +
[…]

### 2020-06-07 analysis/07-prior-figs.R

@@ -24,7 +24,7 @@ s2df <- data.frame(Survival = S1,
                    Probability = Z2/2,
                    Parameter = "S2")
 
-df1 <- bind_rows(s1df, s2df) %>% as_tibble()
+df1 <- bind_rows(s1df, s2df) %>% as_tibble() 
 
 # Pr priors
 
@@ -55,6 +55,7 @@ prpriors
 ggsave("~/Dropbox/2020-ASCF-webinar/pr-priors.png", height = 4, width = 6)
 
 s1priors <- ggplot(df1, aes(Survival, Probability, color = Parameter)) + geom_line(size = 1.5) + 
+  scale_colour_discrete(labels = expression(italic(S[1]), italic(S[2]))) +
   theme_cowplot()
 s1priors
 ggsave("~/Dropbox/2020-ASCF-webinar/surv-priors.png", height = 4, width = 6)

### 2020-08-18 ms/ms.tex
Add changes to new epsilon methodology and Brian and Geir comments

@@
 the need to model data on both migrating smolts and returning adults and to
-simultaneously estimate multiple parameters, especially for salmon spending
-more than one winter at sea before spawning. We fit a hierarchical Bayesian
+simultaneously estimate multiple parameters. We fit a hierarchical Bayesian
 maturity schedule model to data spanning between 19 and 42 years for seven populations in eastern 
@@
 their relative importance differed among populations. Variation in number of
-out-migrating smolts had a moderate contribution on returns, particularly
-after two winter at sea, and probability of returning after one year was
+out-migrating smolts had a moderate contribution on returns and probability of returning after one year was
 important in some populations. 
@@
 are hypothesized to share marine habitat during some of their time at sea, it
-has been presumed that populations share similar temporal trends in
+has been presumed that populations would be more likely to experience similar temporal trends in
 at-sea mortality \citep{Friedland1993, Friedland1998, Russell2012}.  
@@
 Despite the overall decreases in marine survival,
-the conservation status of Canadian salmon populations differs
+the conservation status of Canadian Atlantic salmon populations differs
 considerably. In Canada, populations in the southern part of their range are more
-likely to be assessed as being of conservation concern than those in more
-northerly regions \citep{Cosewic2010}. This latitudinal disparity in status
+likely to be assessed as being of conservation concern (i.e. Threatened or Endangered) 
+than those in more
+northerly regions \citep{Cosewic2010}. This latitudinal disparity 
 suggests that if marine survival has been, or is, a key factor responsible for
@@
 abundances for seven Atlantic salmon populations in eastern Canada, encompassing a
-wide range of the species' distribution. 
+wide range of the species' westerly distribution. 
 Populations included the LaHave River in Nova Scotia's (NS) Southern Uplands, 
@@
 in river $r$, year $t$, and of size group $s$.
-While, ideally, we would account the uncertainty in the conversion between
-size group and returning age, this is being ignored by doing this conversion
-deterministically as shown above. However, sampling a parameter from a
-hypergeometric distribution is not possible in the programming language the model was
-written (i.e. Stan).
+Ideally, we would account the uncertainty in the conversion between
+size group and returning age in the model, but given limitations of the Bayesian 
+sampler we used, we incorporated this uncertainty indirectly as an estimate
+of the variance of annual log-smolt abundance.
 
@@
 \begin{align}
-log(R_{obs,1,t,r}) &\sim Normal(log(R_{1,t,r}), \epsilon_{1,r}), \\
-log(R_{obs,2,t,r}) &\sim Normal(log(R_{2,t,r}), \epsilon_{2,r}) \label{eq:5} 
+log(R_{obs,1,t,r}) &\sim Normal(log(R_{1,t,r}), \epsilon_{1,t,r}), \\
+log(R_{obs,2,t,r}) &\sim Normal(log(R_{2,t,r}), \epsilon_{2,t,r}) \label{eq:5} 
 \end{align}
@@
 where $R_{obs,1,t,r}$ and $R_{obs,2,t,r}$ are the observed return estimates
-for year $t$ and river $r$ of 1SW and 2SW fish, respectively, $\epsilon_{1,r}$
-and $\epsilon_{2,r}$ are the process error terms. 
-These error terms are normally distributed with a standard deviation
-of 0.01:
-
-\begin{align}
-\epsilon_{1,r} &\sim Normal(0, 0.01) \\
-\epsilon_{2,r} &\sim Normal(0, 0.01).
-\end{align}
-
-These error terms are almost equivalent to using a coefficient of variation in
-return estimates of 1\%, given that return estimates in equations~\ref{eq:3}
-and~\ref{eq:4} are log-transformed. As
-there is no associated uncertainty with most return estimates, we chose this value arbitrarily
-based on the expectation that uncertainty in returns estimates is likely lower than in
-smolt estimates.
+for year $t$ and river $r$ of 1SW and 2SW fish, respectively, $\epsilon_{1,t,r}$
+and $\epsilon_{2,t,r}$ are the process error terms. 
+These error terms are estimated indirectly by approximating the abundance of 1SW and 2SW returns
+from hypergeometric distributions based on the scale sample data. 
+We were able to obtain scale data for all but two of the populations (LaHave and Nashwaak).
+See the Supplementary Materials for a description of the methodology and the resulting estimates.
 
@@
 while at Nashwaak there was an increase in median \So during the early
-2000s but a decrease in the 2010s. Annual estimates of \So had little
+2000s but a decrease since 2010. Annual estimates of \So had little
 uncertainty for one sea-winter dominated populations (Conne, Campbellton, and
-WAB), but uncertainty was higher (i.e. wider credible intervals) in the other
+WAB), with the exception of the 2004-2011 periord in the Campbellton where
+there were very few scale samples, and hence high process error; but in general, 
+uncertainty was higher (i.e. wider credible intervals) in the other
 populations.
@@
 of \St for the Saint-Jean and Trinit\'{e} were considerably less uncertain
-than for the other populations, but showed no apparent trends through time.
+than for the other populations, with the latter showing an increase with a peak in the mid-1990s followed 
+by a decrease.
 
@@
 Estimates of \Pg were mostly stable across time, except for the LaHave and
-Nashwaak populations; The estimates of \Pg were slightly lower in the last four
-years than in the previous ones, while in the Nashwaak the posterior estimates
-of \Pg in 2012 were much lower than in all other years
-(Fig.~\ref{fig:s2-faceted}). Variation in yearly estimates was highest in
-the LaHave, Nashwaak, and Trinit\'{e}, and lowest in the 1SW-dominated
+Nashwaak populations; the estimates of \Pg were slightly lower in the last
+four years than in the previous ones, while in the Nashwaak the posterior
+estimates of \Pg in 2012 were much lower than in all other years
+(Fig.~\ref{fig:s2-faceted}). Variation in yearly estimates was highest in the
+LaHave, Nashwaak, and Trinit\'{e}, and lowest in the 1SW-dominated
 populations.
@@
 
-Most correlations between trends were significant, with only five out of the
-21 pair-wise comparisons not having a p-value below 0.05
+Roughly half (11) out of the 21 pair-wise correlations were significantt (p-value below 0.05)
 (Fig.~\ref{fig:s1-corr}). When looking at the direction of the correlations,
@@
 uncorrected ones as the error variance and covariance terms were very small.
-The correlations between populations in the region were positive between Nashwaak and LaHave (Scotia-Fundy region), non-significant between the Trinit\'{e} and Saint-Jean (QC), and positive
-between WAB and Campbellton but negative between Conne and both WAB and Campbellton (NL).
+The correlations between populations in the region were positive between Nashwaak and LaHave (Scotia-Fundy region), non-significant between the Trinit\'{e} and Saint-Jean (QC), 
+non-significant between WAB and Campbellton but negative between Conne and WAB (NL).
+The Campbellton had mostly non-significant pairwise correlations with other populations, even though there are both positive and negative correlation coefficients, except with the Saint-Jean
+where the correlation between $Z_1$ trends is positive and significant.
+Note that the uncertainty in these correlations reflect the uncertainty in the
+measurements, but not in the process (in this respect they are equivalent to
+sample variance and not population variance).
 
@@
         returning after one year at sea, $-Z_1$ = marine survival in the first year at sea, and $-Z_2$ = marine survival in the second year at sea.
-        Note the correlations derived from negative instantaneous total mortalities $-Z_1$ and $-Z_2$ are equivalent to those of marine survival $S_1$ and $S_2$.}
+        Note the correlations derived from negative instantaneous total mortalities $-Z_1$ and $-Z_2$ are representative to those of marine survival $S_1$ and $S_2$.}
     \label{fig:cor-params} 
@@
 These negative correlations could be indicative of density-dependent
-mortality at sea, however this is generally thought to be is unlikely.
+mortality at sea, however this is generally thought to be unlikely.
 Populations with a lower proportion of 1SW in their returns (e.g. Trinit\'{e}
@@
 while other exhibits highly variable yet stable trends, and some show declines.
-Variation in survival at the population level can be large even within
-a region, and there are no consistent declines in marine survival in any
-population after the onset of fishing moratoria, except perhaps in the LaHave
-River where the four last years in the time-series (2014-2017) had survival of
-less than 2.5\%.
-However, while correlations in marine survival trends between rivers were highly 
+Variation in survival at the population level can be large even within a
+region. With one possible exception (LaHave River), consistent declines in
+marine survival within populations are not evident since the onset of fishing
+moratoria.
+
+While correlations in marine survival trends between rivers are highly 
 variable (Fig.~\ref{fig:s1-corr}),
-there are some years where the populations appear to behave in concert:
-years 1997, 2001, and 2007 have consistently low survival across all populations 
+there are some years in which the populations appear to behave in concert:
+years 1997, 2007, and to some extent 2001 have consistently low survival across all populations 
 (i.e. <5\%, Fig.~\ref{fig:s1-dual}) and little variation across populations.
-Our results also show that variation of survival in first year at sea is the most
-important factor determining variation in number of returns, and suggests that 
-changes in abundance are primarily driven by changes in marine survival. 
-Nonetheless, variation in number of out-migrating smolts, and for some populations
+Our results also show that variation of survival in the first year at sea is the most
+important factor determining variation in number of returns, suggesting that changes in abundance are primarily driven by changes in marine survival. 
+Nonetheless, variation in number of out-migrating smolts and, for some populations,
 variation in probability of returning after one winter at sea, also explain
-variation in returns and thus contribute to the observed variation in return abundances.
+variation in returns, thus contributing to the observed variation in return abundances.
 Furthermore, the negative correlation between marine survival and smolt
-abundances in some populations is suggestive of density-dependent processes
-occurring at sea.
+abundances in some populations is suggestive of density-dependent processes.
 We could not assess trends for survival in the second winter at sea (\St) or
@@
 At first glance, the large variation in marine survival trends among the seven
-populations examined in our study seems to be at odds with the synchronous
+populations examined seems to be at odds with the synchronous
 trends of declining marine survival detected at broad regional scales
 \citep{Olmos2019}. However, these two observations are not necessarily mutually
-exclusive: our study is only a small subsample of the populations in the region,
-and, with one exception, over a much shorter period of time,
-and highly variable trends in marine survival among populations can still add
+exclusive: our study represents a subsample of the populations in the region,
+and, with one exception, encompasses a shorter period of time; highly variable 
+trends in marine survival among populations can still add
 up to overall negative trends at a regional scale.
@@
 trends could be further confounded by differences in model structure, such as
-the methodology for estimating covariance, and use stock-recruitment
-relationships rather than empirical smolt count estimates to estimate marine
-survival.
+the methodology for estimating covariance, and use of stock-recruitment
+relationships \citep{Olmos2019} rather than empirical smolt count estimates 
+(present study) to estimate marine survival.
 % As trends in marine survival during the first winter at sea are highly
@@
 Our results suggest that trends from broader geographical areas (i.e.
-province, state, or country-wide estimates) may not be representative of
+province, state, or country-wide estimates) might not be representative of
 individual populations, and concomitantly, trends of individual populations
-may also not be representative of region-wide trends.
+might also not be representative of region-wide trends.
 
@@
 were lower than those during the 1970s.
-\citet{Gibson2016} calculated higher return rates of 2SW for the Nashwaak in
+\citet{Gibson2016} calculated higher return rates of 2SW for Nashwaak River salmon in
 the 1970s than in the period since, with return rates of 1SW in the 1970s
@@
 respectively. Other years are much more variable relative to each other.
-\citeauthor{Friedland1993}'s \citeyear{Friedland1993} time series ends in  
-1988; thus there are only a few years for which to assess overlap with the
+The time series in \citet{Friedland1993} end in  
+1988; thus there are only a few years that overlap the
 time series in our study.
@@
 populations, we were unable to assess if marine survival estimates were also
-higher in the 1970s as we lacked data from this decade.
+higher in the 1970s because smolt count data from this decade are not available.
 \citet{Dempson2003} described a general declining trend in marine survival for
-Newfoundland populations (except WAB); we drew the same conclusion for Conne
+Newfoundland populations (except WAB). We drew the same conclusion for Conne
 River but not Campbellton River or WAB. It is not possible to draw broader
@@
 of marine survival with varying degrees of similarity depending on each
-population's life histories. The estimates of return rates of one sea-winter salmon
+population's life history.
+
+The estimates of return rates of one sea-winter salmon
 approximate marine survival in the first year at sea for one
@@
 directed marine fisheries are largely non-existent,
-and we see this by comparing the three NL populations
+and we see this by comparing $S_1$ in the three NL populations
 with the return rates presented in \citet{ICES2019}. However, as the proportion of fish that
@@
 varies in time;
-proportion returning as one sea-winter was in important factor driving number of 
+proportion returning as one sea-winter was an important factor driving number of 
 1SW returns in three of the populations examined (Fig.~\ref{fig:mu-rsq}).
@@
 providing a very coarse estimate of marine survival in the second year at sea. 
-As changes in return rates of two sea-winter fish are a result of a
+Given that changes in return rates of two sea-winter fish are a result of a
 combination of changes in survival in the first year at sea in the year prior,
@@
 for the Trinit\'{e} population are very similar to those produced by
-\citet{Chaput2003b} using a two sex model, and our trends are almost identical
+\citet{Chaput2003b}, who applied a two-sex model, and our trends are almost identical
 for the overlapping time period that marine survival was estimated for in
@@
 A recent study showed that correlations between oceanic temperature, primary
-productivity, and post smolt survival of Atlantic salmon in the North Atlantic
+productivity, and post-smolt survival in the North Atlantic
 are better explained by broad-scale environmental trends than regional ones
@@
 suggesting warmer
-conditions favour post-smolts \citep{Friedland1998}, based on mapping the
+conditions favor post-smolts \citep{Friedland1998}, based on mapping the
 extent of area of suitable temperature (7-13 \textdegree C).
@@
 This difference could explain why populations in eastern North America are
-declining in the southern part of their range but potentially increasing
+declining in the southern part of their range but potentially increasing, or
+remaining stable,
 further north, and also why some studies find positive correlations between
 temperature and abundance \citep{Friedland1998, Friedland1998b, Jonsson2004}
-while others find negative ones \citep{Friedland1993, Todd2008}.
-\citet{Olmos2020} found a positive relationship between temperature and marine survival 
+while others find negative associations \citep{Friedland1993, Todd2008}.
+\citet{Olmos2020} documented a positive relationship between temperature and marine survival 
 in northern regions and a negative one in southern regions, providing 
@@
 
-Oceanic conditions have been correlated with abundance trends, growth, and marine survival
-in Atlantic salmon, which are thought
-to be mediated by bottom-up effects driving ocean productivity and food availability
+Oceanic conditions have been correlated with abundance trends, growth, and
+marine survival in Atlantic salmon, which are thought to be mediated by
+bottom-up effects driving ocean productivity and food availability
 \citep{Todd2008, Renkawitz2015,Olmos2020}. 
-However, the mechanism by which such bottom-up effects,
-mediated by changes in food availability,
-affect population dynamics beyond marine survival needs to
-considered further. 
-There are potential carry-on effects of oceanic
-conditions that manifest with regards to freshwater production for individuals that survive
-the marine phase of their life cycle. 
-For example, adults
-that return to their natal streams after spending suboptimal conditions at sea might be less
-likely to make it to the spawning grounds or secure a mate, and might also
-produce fewer eggs or eggs with a lower energetic content than those
-produced by adults which grew in optimal oceanic conditions.
+However, the mechanism by which such bottom-up effects, mediated by changes in
+food availability, affect population dynamics beyond marine survival needs to
+be considered further. 
+There are potential carry-on effects of oceanic conditions that manifest with
+regards to freshwater production for individuals that survive the marine phase
+of their life cycle. 
+For example, adults that return to their natal streams after spending
+suboptimal conditions at sea might be less likely to make it to the spawning
+grounds or secure a mate, and might also produce fewer eggs or eggs with a
+lower energetic content than those produced by adults which grew in optimal
+oceanic conditions.
 As larger females tend to be more productive, in terms of fecundity and total
@@
 Nonetheless, our results suggests that while variation in number of smolts was 
-important for number of returns, it was considerable less important than
-survival the first year at sea, which is indicative that return abundances 
+important for number of returns, it was considerably less important than
+survival during the first year at sea, which suggests that return abundances 
 are most strongly influenced by the marine phase of their life cycle.
@@
 some density-dependent processes during parts of the marine phase, 
-particularly for populations that are likely to be subjected to
-declining per capita population growth rates ($r$) generated by Allee effects.
-While the correlations between return estimates and survival were negative for 
-some populations (Fig.~\ref{fig:cor-params})
-indicate that there may be some compensatory density dependent effects in the 
-marine phase, our analysis provide no evidence of Allee effects.
+particularly for populations that are 
+near historically low levels of abundance. One potential mechanism might be a
+``predator pit''. When prey populations are very small, predator-induced
+mortality can also be low because the prey are simply not abundant enough to
+be generally consumed by an optimal forager (for example, search costs may be
+too high relative to the fitness benefits of consuming the prey). But as prey
+abundance increases from very low levels, predator-induced mortality might
+also increase as preference for the prey increases. 
+Correlations between return estimates and survival were negative for 
+some populations (Fig.~\ref{fig:cor-params}),
+indicating that there may be some compensatory density dependent effects during the 
+marine phase.
 Exploring relationships between survival and population size could potentially
@@
 As with all novel modelling approaches, there are caveats to acknowledge.
-The seven populations explored in the present study might not be representative
-of regional trends in marine survival. However, there are no other
-long-term time series of smolts and adult returns to draw inferences from.
+The seven populations examined in the present study represent a small subset of the total number of salmon rivers in eastern Canada and hence might not be representative
+of the overall regional trends in marine survival. However, there are no other
+long-term time series of smolts and adult returns from which to draw inferences from.
 While there are analytical issues associated with the estimation of \So, \St, and \Pg,
-the assumption that survival in the second year at sea is additive to survival in the first 
+the assumption that survival in the second year at sea is multiplicative to survival in the first 
 year at sea could produce unrealistic results.
@@
 estimated, or an additional assumption as to what proportion of \So is not
-additive to \St (as the returning 1SW adults do not experience the same
+multiplicative to \St (as the returning 1SW adults do not experience the same
 environment when they return to their natal streams as those fish who stayed
@@
 \Pg seems like a reasonable assumption given that the probability of returning
-as grilse has a genetic component associated to it \citep{Aykanat2019} and is
-not expected to vary much, within a population, among years.
+as grilse has a genetic component associated to it \citep{Barson2015} and its inter-annual
+variation in a given population is likely partly constrained by this.
 
 Thirdly, the accuracy and precision of the model's output could be different
-for different populations. Survival in the first
-year at sea is the variable that explains most of the variance in estimates of 1SW returns, except for
-the Saint-Jean (Fig.~\ref{fig:mu-rsq}).
-
-The model does not provide reliable parameter estimates for the Saint-Jean
+for different populations. Survival in the first year at sea is the variable
+that explains most of the variance in estimates of 1SW returns, except for the
+Saint-Jean; 
+the model does not provide reliable parameter estimates for this population
 population (as indicated by the low $R^2$ values in Fig.~\ref{fig:mu-rsq}),
 which is the population with the highest proportion of 2SW returns.
-
-Thus, estimates of \So for this
-population, which has the lowest proportion of 1SW returns out of all seven
-populations examined, are likely to be less accurate than those for the other
-populations.
-Furthermore, estimates of \St and \Pg are negatively correlated and highly uncertain,
-and this must be considered when making any inferences based on these parameters.
+Furthermore, estimates of \St and \Pg are negatively correlated and highly
+uncertain, and this must be considered when making any inferences based on
+these parameters for any of the populations examined.
 
@@
 causes (e.g. predation) or from fishing; we know the latter has decreased
-due to the sequential moratoria enacted across eastern Canada, and this
+dramatically due to the sequential moratoria enacted across eastern Canada, and this
 confounds the estimated trends in marine survival because we know marine survival
@@
 fishing mortality before the turn of the century.
-However, for most of our studied populations, 
-present-day marine survival fluctuates to values
-as high as those recorded in the past, thus
-suggesting a potential weakening of compensatory population dynamics.
 
@@
 But the fact that, for most of the populations examined, 
-present-day survival fluctuating around similar levels  
-similar levels as in the past, 
+present-day survival fluctuates around levels similar to those that have occurred in the past, 
 \emph{despite} reduced commercial fishing mortalities, suggests

### 2020-08-18 ms/suppmat.tex
Incorporate Geir's and Brian's suggestions
Start new pages after each table
Add changes to new epsilon methodology and Brian and Geir comments

@@
 
+\subsection*{Estimation of uncertainty in return estimates based on scale data}
+
+We incorporated the uncertainty in the conversion from size group (small and large returns) to age class (1SW and 2SW returns) by 
+approximating the variance of log return abundance of each maiden age class 1SW and 2SW by assuming that the probability 
+of the true number of returns of a given age, river, year, and size group ($N_{r,t,s,a}$) can be derived from a hypergeometric distribution
+
+\begin{equation}
+p[N_{r,t,s,a}|0:N_{r,t,s}] \sim Hypergeometric(n_{r,t,s}, 0:N_{r,t,s}, N_{r,t,s}:0, n_{r,t,s,a}),
+\end{equation}
+
+where $n_{r,t,s,a}$ is the number of scale samples in river r, year t, of returning age a, and of size group s;
+$n_{r,t,s}$ is the total number of scale samples in river r, year t, and of size group s; and
+$N_{r,t,s}$ is the total number of returns in river, year t, and of size group s.
+
+We therefore obtain a vector of the probabilities of the returns of a given river r, year t, size group s and age a (i.e. $N_{r,t,s,a}$)
+across the whole range of potential values, i.e. from zero to $N_{r,t,s}$, which provides the weighting probabilities to generate 10,000 random draws of $N_{r,t,s,a}$. 
+We then add the draws of the same age from both size groups:
+to obtain the estimates of number of returns of a given age a (i.e. 1SW or 2SW) for a given river and year ($N_{r,t,a}$):
+
+\begin{equation}
+N_{r,t,a} = N_{r,t,a,s=small}+ N_{r,t,a,s=large}.
+\end{equation}
+
+Lastly, as the coefficient of variation are comparable to standard deviations of less than 0.3 in the log-scale, we obtain
+the estimate of $\epsilon_{r,t,a}$ as
+
+\begin{equation}
+\epsilon_{r,t,a} \approx \sigma_{log(N_{r,t,a})}.
+\end{equation}
+
+The $\epsilon_{r,t,a}$ parameter is used as the variance term in equations~8 and 9 of the manuscript.
+The data used is in Table~\ref{tab:scales} while the estimates of $\epsilon_{r,t,a}$ are provided in Tables~\ref{tab:epsilon1} and~\ref{tab:epsilon2}.
+
+There are a number of years with very few scales samples, particularly for the
+Campbellton population, and one year for this population (2011) with no scales
+samples. Under these conditions, the model provides high estimates of
+$\epsilon$, but which are somewhat constrained by the log-transformation of
+abundance draws of a uniform distribution. Thus, the value of $\epsilon$ for
+2011, the year with no scale samples, is of 0.711 (Table~\ref{tab:epsilon1}),
+which is likely an underestimate of true uncertainty given the absence of
+data. Nonetheless, the model results in small $\epsilon$ values for years
+where there are ten or more scale samples. 
+For years where there are no scale nor return samples to draw inferences from (e.g. a year for which there are no small 2SW scales, nor large scale samples),
+sometimes the error for 2SW returns cannot be estimated; in these cases we set the value of $\epsilon_{2}$ to 0.999.
+
+\begin{figure}[htbp] \centering
+    \includegraphics[width=0.95\linewidth]{figures/camp-scale-epsilon-example.png} \caption{Example
+        of the estimation of $\epsilon_{1}$ for two different years of data for the Campbellton population: year 2011 where
+        there were no scale samples, and year 2014 with a large number of scales samples. a) Distribution of draws of potential
+    1SW abundances, adding the estimates from both small and large returns, for the year 2011. Note that the distribution spans the whole range of
+returns (up to 5443). b) The same distribution but log-transformed, which results in a right-skewed distribution. The blue line represents the normal
+distribution fitted to the log-transformed data, from which $\epsilon_{1}$ is equal to its standard deviation. b) Distribution of draws of potential
+    1SW abundances, adding the estimates from both small and large returns, for the year 2014, which is much tightly constrained. d) The distribution
+of log-transformed abundance draws for year 2014. In this case, the normal distribution fitted is much tighter and more closely resembles the underlying histogram,
+and results in a much lower estimate of $\epsilon_{1}$.}
+    \label{fig:priors} 
+\end{figure}
+
+\clearpage
+
+\input{scales_tab.tex}
+
+\clearpage
+
+\input{epsilon1_tab.tex}
+
+\clearpage
+
+\input{epsilon2_tab.tex}
+
+\clearpage
+
 \subsection*{Raw smolt data used in the model}
@@
 
+\clearpage
+
+\subsection*{Error in smolt abundances used in the model}
+
 \input{smoltscv_tab.tex}
+
+\clearpage
 
@@
 
-\subsection*{Population-level estimates of proportion returning after one winter at sea}
+\subsection*{Population-level estimates of smolt abundance}
 
@@
 
-\hspace{8em}
+%\hspace{8em}
+
+\clearpage
 
@@
 \input{tableS4_lssig}
+
+\clearpage
+
+\subsection*{Population-level estimates of proportion returning after one winter at sea}
+
+\input{table-logitprmu.tex}
+
+\hspace{8em}
+
+\input{table-logitprsigma.tex}
 

### 2020-08-18 analysis/00-campbellton-wrangling.R
Code update to include new epsilon estimates in model

@@ -1,5 +1,6 @@
 ### WAB data
 
+source("analysis/99-hypergeometric-bootstrap-function.R")
 library(tidyverse)
 
 campsmolts <- read_csv("raw-data/NL-smolt-estimates-Kelly-et-al-2017-ResDoc.csv",
@@ -12,11 +13,120 @@ campsmolts <- read_csv("raw-data/NL-smolt-estimates-Kelly-et-al-2017-ResDoc.csv"
 # Loading estimated returns of 1SW, 2Sw, and RS
 camp <- readRDS(file = "~/Dropbox/salmon-lh-variation/output/NL/campbellton-estimated-returns.rds")
 
+# Loading raw scale age data
+campraw <- readRDS("~/Dropbox/salmon-lh-variation/output/NL/campraw.rds") %>%
+  select(year, sl, M1, M2p, RStotal)
+
+campwide <- pivot_wider(campraw, names_from = sl, values_from = c(M1, M2p, RStotal)) 
+campwide[is.na(campwide)] <- 0 # replacing NAs with zeros
+
+campjoin <- left_join(camp, campwide, by = "year")
+
+campscales <- campjoin %>%
+  mutate_all(replace_na, 0) %>%
+  rename(n_small = small, 
+         n_large = large, 
+         scales_small_1SW = M1_small, 
+         scales_small_2SW = M2p_small,
+         scales_small_other = RStotal_small,
+         scales_large_1SW = M1_large, 
+         scales_large_2SW = M2p_large, 
+         scales_large_other = RStotal_large)
+
+
+campsds <-  pmap(campscales, hyper_boot) %>% bind_rows %>%
+  select(year, starts_with("sd"))
+
+# replacing NaNs (from years without samples) with mean value
+campsds[is.nan(campsds$sdlog2SW),"sdlog2SW"] <- 0.99 #  mean(wabsds$sdlog2SW, na.rm = TRUE)
+
+
+campepsilon <- select(campscales, year, n_small, n_large, totalreturns, scales_small_1SW:scales_large_other) %>%
+  left_join(campsds, by = "year") %>%
+  mutate(river_name = "Campbellton River") %>%
+  select(river_name, everything())
+
+saveRDS(campepsilon, file = "data/campepsilon.rds")
+
+campscales2011 <- filter(campscales, year == 2011)
+campscales2014 <- filter(campscales, year == 2014)
+
+
+debug(hyper_boot)
+campsds2011 <- hyper_boot(year = campscales2011$year,
+         campscales2011$n_small,
[…]

### 2020-08-18 analysis/00-conne-wrangling.R

@@ -1,23 +1,65 @@
 # Conne data
 
+source("analysis/99-hypergeometric-bootstrap-function.R")
 library(tidyverse)
+
 
 # Loading estimated returns of 1SW, 2Sw, and RS
 conne <- readRDS(file = "~/Dropbox/salmon-lh-variation/output/NL/conne-estimated-returns.rds")
 
+# Loading raw scale age data
+conneraw <- readRDS("~/Dropbox/salmon-lh-variation/output/NL/conneraw.rds") %>%
+  select(year, sl, M1, M2p, RStotal)
+
 # Loading small, large, and smolt data
 # Smolt years missing: up to 1988, 2016 and 2017, which
 # is the duration of the dataset, so no need to remove years in the middle
-connewithsmolts <- readRDS(file = "~/Dropbox/salmon-inverse-matrix/data/connereturns.rds")
+connewithsmolts <- readRDS(file = "~/Dropbox/salmon-inverse-matrix/olddata/connereturns.rds")
 
 # Adding missing years based on Dempson et al 2003 (Ch 1)
 connewithsmolts[connewithsmolts$year == 1987, "smolt"] <- 74585
 connewithsmolts[connewithsmolts$year == 1988, "smolt"] <- 65692
 
 
+connewide <- pivot_wider(conneraw, names_from = sl, values_from = c(M1, M2p, RStotal)) 
+connewide[is.na(connewide)] <- 0 # replacing NAs with zeros
+
+connejoin <- left_join(connewithsmolts, connewide, by = "year") %>%
+   filter(year >= 1986 & year < 2017) %>% 
+  select(-smolt) 
+
+connescales <- connejoin %>%
+  rename(n_small = small, 
+         n_large = large, 
+         scales_small_1SW = M1_small, 
+         scales_small_2SW = M2p_small,
+         scales_small_other = RStotal_small,
+         scales_large_1SW = M1_large, 
+         scales_large_2SW = M2p_large, 
+         scales_large_other = RStotal_large)
+
+
+connesds <-  pmap(connescales, hyper_boot) %>% bind_rows %>%
+  select(year, starts_with("sd"))
+
+# replacing NaNs (from years without samples) with mean value
+connesds[is.nan(connesds$sdlog2SW),"sdlog2SW"] <- 0.99 #  mean(wabsds$sdlog2SW, na.rm = TRUE)
+
+conneepsilon <- connescales %>%
+  mutate(totalreturns = n_small + n_large) %>%
[…]

### 2020-08-18 analysis/00-lahave-wrangling.R

@@ -19,6 +19,10 @@ lhbiochar <- lahave_bio_adult %>%
   filter(!is.na(POST)) %>% # removing those with no sea ages
   mutate(LS = ifelse(RS == TRUE, "RS", paste0(POST, "SW"))) 
 
+# There are more large salmon scale samples than wild large retruns this year, so need to subset scales
+lh1995 <- lahave_bio_adult %>%
+  filter(RCYEAR == 1995)
+
 # Test that only Xs are present
 lhbiochar %>% 
   mutate(RS2 = paste0(SP1, SP2, SP3, SP4, SP5)) %>%
@@ -49,11 +53,9 @@ lhbiowide <- lhbiochar %>%
   mutate(propRS = ifelse(is.na(propRS), 0, propRS)) %>%
   filter(year >= 1995)
 
-lhbiowide
-
 # Reading return counts of WILD FISH ONLY!!!
 lhreturns <- read_excel("raw-data/lahave/Morgan Falls adult returns.xlsx", 
-           skip = 2, na = c("", "--"), n_max = 50) %>%
+                        skip = 2, na = c("", "--"), n_max = 50) %>%
   select(year = `...1`, small = `Wild Small`, large = `Wild Large`)
 
 lhreturnslong <- lhreturns %>%
@@ -74,7 +76,35 @@ lhreturns_corrected <- left_join(lhreturnslong, lhbiowide, by = c("year", "smla"
             estRS = sum(estRS),
             totalsmall = first(returns),
             totallarge = last(returns))
-  
+
+
+
+# Group numbers of scales for CV calculation
+lhwide <- lhbiowide %>%
+  select(year, smla, `1SW`:RS) %>% 
+  pivot_wider(names_from = smla, values_from = c(`1SW`:RS))
+
+lhwide[is.na(lhwide)] <- 0 # replacing NAs with zeros
+
+lhjoin <- left_join(lhreturns_corrected, lhwide, by = "year")
+
+lhscales <- lhjoin %>%
+  mutate_all(replace_na, 0) %>%
+  mutate(n_small = totalsmall, 
+         n_large = totallarge,
+         scales_small_1SW = `1SW_small`,
+         scales_small_2SW = `2SW_small`,
+         scales_small_other = `3SW_small` + `4SW_small` + `5SW_small` + RS_small,
+         scales_large_1SW = `1SW_large`, 
+         scales_large_2SW = `2SW_large`, 
[…]

### 2020-08-18 analysis/00-nashwaak-wrangling.R

@@ -3,7 +3,7 @@
 library(tidyverse)
 
 # Load Naskwaak data from salmon-inverse-matrix repo
-load("~/Dropbox/salmon-inverse-matrix/data/nashwaak.rda")
+load("~/Dropbox/salmon-inverse-matrix/olddata/nashwaak.rda")
 
 # We use the return data frame that includes wild salmon only, as it
 nashwaak_wild_returns
@@ -53,6 +53,8 @@ nashwaakdata <- list(years = nashwaakreturns$year,
   #N =  length(connereturns$small[-1]),
   #Pr = rep(0.8, nrow(connereturns)))
   returns_cv = 0.01, # ASSUMPTION, CV of return abundance counts
+  logSW1_cv = rep(0.01, length(nashwaakreturns$year)), # NO DATA SO USING VALUES BASED ON MEAN TRINITE ESTIMATES
+  logSW2_cv =  rep(0.06, length(nashwaakreturns$year)),
   nyears = length(nashwaakreturns$year),
   river_name = "Nashwaak River",
   allreturns = nashwaakdat  %>% filter(year >= 1980) %>% mutate(river_name = "Nashwaak River")

### 2020-08-18 analysis/00-qc-smolt-data.R

@@ -1,0 +1,121 @@
+# wrangling smolt data for Saint-Jean and Trinité rivers
+
+library(tidyverse)
+library(readxl)
+
+source("analysis/99-hypergeometric-bootstrap-function.R")
+
+# smolts data
+# qcsmolts <- read_excel("raw-data/Smolt_St-Jean-Trinite.xlsx")
+# names(qcsmolts) <- c("river", "year", "smolts")
+# 
+# table(qcsmolts$river)
+# 
+# ggplot(qcsmolts, aes(year, smolts, color = river)) + geom_point() + geom_line() + ylim(0,160000)
+# 
+# saveRDS(qcsmolts, file = "data/qc-smolts.rds")
+
+# returns data
+qcdata <- read_excel("raw-data/données_multifrayeur_Trinité et St-Jean_Quebec.xlsx")
+
+tail(qcdata, n = 7)
+qcdata <- qcdata[1:(nrow(qcdata) - 6), ] # removing last 6 rows
+tail(qcdata, n = 7)
+
+qcadults <- qcdata %>% select(Rivière, 
+                     Année, 
+                     `Montaison Mad`, 
+                     `Montaison Réd`, 
+                     `Montaison totale`,
+                     `Age Nb échan. Diber`,
+                     `Age Nb échan. Triber`,
+                     `Age Nb échan. Fraie antérieure`,
+                     `Age Nb en montaison Diber`, 
+                     `Age Nb en montaison Triber`, 
+                     `Age Nb en montaison Fraie antérieure`, 
+                     `Reproducteur Mad Nb`, 
+                     `Reproducteur Mad % Femelle`, 
+                     `Reproducteur Mad Femelle Estimation`, 
+                    `Reproducteur Réd Nb`, 
+                    `Reproducteur Réd % Femelle`,	
+                    `Reproducteur Réd Femelle Estimation`) %>%
+          rename(river = Rivière, 
+                 year = Année, 
+                 returns_grilse = `Montaison Mad`, 
+                 returns_msw = `Montaison Réd`, 
+                 returns_total = `Montaison totale`,
+                 scales_2sw = `Age Nb échan. Diber`,
+                 scales_3sw = `Age Nb échan. Triber`,
+                 scales_rs = `Age Nb échan. Fraie antérieure`,
+                 returns_2sw_est = `Age Nb en montaison Diber`, 
[…]

### 2020-08-18 analysis/00-saintjean-wrangling.R

@@ -29,18 +29,26 @@ sjsmolts <- read_excel("raw-data/Estimation dévalaison St-Jean-Trinité 1984-20
 
 excel_sheets("raw-data/Estimation dévalaison St-Jean-Trinité 1984-2018.xlsx")
 
+sjsds <- readRDS("data/sj-scales-sd.rds") %>%
+  select(-river_name) %>% as_tibble()
 
-saintjeanreturns <- saintjean %>%
+saintjeanreturns <- left_join(saintjean, sjsds, by = "year") %>%
   mutate(smoltlag = lag(smolts),
          #smolt025lag = lag(smolt_025),
          #smolt975lag = lag(smolt_975),
          est_1SW = returns_grilse,
          est_2SWlead = lead(returns_2sw_est),
          est_RSlead = lead(returns_rs_est),
+         sdlog2SWlead = lead(sdlog2SW)
          #logsmoltsdlag = sqrt(300) * (log(smolt975lag) - log(smolt025lag))/3.92
   ) %>% 
   filter(year %in% 1990:2015) %>% # updated from 2014 to 2015
   filter(year != 1998) # removing smotl year 1997 as its missing smolt data
+
+# replacing NA and NaN values in sds with averages
+saintjeanreturns[is.na(saintjeanreturns$sdlog2SWlead),"sdlog2SWlead"] <- mean(saintjeanreturns$sdlog2SWlead, na.rm = TRUE)
+
+select(saintjeanreturns, year, contains("sd"))
 
 # Removing zeros in 2SW-RS estimates as they cannot be log transformed
 filter(saintjeanreturns, est_RSlead != 0) %>% pull(est_RSlead) %>% min(.)/2 # picking min(RS)/2
@@ -64,6 +72,8 @@ saintjeandata <- list(years = saintjeanreturns$year,
                     #N =  length(saintjeanreturns$small[-1]),
                     #Pr = rep(0.8, nrow(connereturns)))
                     returns_cv = 0.01, # ASSUMPTION, CV of return abundance counts
+                    logSW1_cv = saintjeanreturns$sdlog1SW,
+                    logSW2_cv = saintjeanreturns$sdlog2SWlead,
                     nyears = length(saintjeanreturns$year),
                     river_name = "Saint-Jean River",
                     allreturns = saintjean %>% mutate(river_name = "Saint-Jean River")

### 2020-08-18 analysis/00-trinite-wrangling.R

@@ -17,14 +17,18 @@ trinite %>%
   geom_line() +
   scale_y_continuous("return number", sec.axis = sec_axis(~ . * 50, name = "smolt number"))
 
+trisds <- readRDS("data/trinite-scales-sd.rds") %>%
+  select(-river_name) %>% as_tibble()
 
 trinitereturns <- trinite %>%
+  left_join(trisds, by = "year") %>%
   mutate(smoltlag = lag(smolts),
          #smolt025lag = lag(smolt_025),
          #smolt975lag = lag(smolt_975),
          est_1SW = returns_grilse,
          est_2SWlead = lead(returns_2sw_est),
          est_RSlead = lead(returns_rs_est),
+         sdlog2SWlead = lead(sdlog2SW)
          #logsmoltsdlag = sqrt(300) * (log(smolt975lag) - log(smolt025lag))/3.92
         ) %>% 
   filter(year %in% 1985:2016) %>% # lead 2017 is missing 2SW and RS
@@ -55,6 +59,8 @@ trinitedata <- list(years = trinitereturns$year,
                   #N =  length(connereturns$small[-1]),
                   #Pr = rep(0.8, nrow(connereturns)))
                   returns_cv = 0.01, # ASSUMPTION, CV of return abundance counts
+                  logSW1_cv = trinitereturns$sdlog1SW,
+                  logSW2_cv = trinitereturns$sdlog2SWlead,
                   nyears = length(trinitereturns$year),
                   river_name = "Trinité River",
                   allreturns = trinite %>% mutate(river_name = "Trinité River")

### 2020-08-18 analysis/00-wab-wrangling.R

@@ -1,5 +1,5 @@
 ### WAB data
-
+source("analysis/99-hypergeometric-bootstrap-function.R")
 library(tidyverse)
 
 wabsmolts <- read_csv("raw-data/NL-smolt-estimates-Kelly-et-al-2017-ResDoc.csv",
@@ -11,8 +11,50 @@ wabsmolts <- read_csv("raw-data/NL-smolt-estimates-Kelly-et-al-2017-ResDoc.csv",
 # Loading estimated returns of 1SW, 2Sw, and RS
 wab <- readRDS(file = "~/Dropbox/salmon-lh-variation/output/NL/wab-estimated-returns.rds")
 
+# Loading raw scale age data
+wabraw <- readRDS("~/Dropbox/salmon-lh-variation/output/NL/wabraw.rds") %>%
+  select(year, sl, M1, M2p, RStotal)
+
+
+wabwide <- pivot_wider(wabraw, names_from = sl, values_from = c(M1, M2p, RStotal)) 
+wabwide[is.na(wabwide)] <- 0 # replacing NAs with zeros
+
+# Editing this year as there were no large fish reported in the returns, yet there was
+# a single fish who was scale sampled and measured 64.8 cm and was still labelled as a grilse
+wabwide[wabwide$year == 1990, "M1_small"] <- 47 # Originally 46
+wabwide[wabwide$year == 1990, "M1_large"] <- 0 # Originally 1
+wabwide[wabwide$year == 1995, "RStotal_large"] <- 30 # originally 31, one removed to match total large return count
+wabjoin <- left_join(wab, wabwide, by = "year")
+
+wabscales <- wabjoin %>%
+  mutate_all(replace_na, 0) %>%
+  rename(n_small = small, 
+         n_large = large, 
+         scales_small_1SW = M1_small, 
+         scales_small_2SW = M2p_small,
+         scales_small_other = RStotal_small,
+         scales_large_1SW = M1_large, 
+         scales_large_2SW = M2p_large, 
+         scales_large_other = RStotal_large)
+
+
+wabsds <-  pmap(wabscales, hyper_boot, debug = FALSE) %>% bind_rows %>%
+  select(year, starts_with("sd"))
+
+# replacing NaNs (from years without samples) with mean value
+wabsds[is.nan(wabsds$sdlog2SW),"sdlog2SW"] <- 0.99 #  mean(wabsds$sdlog2SW, na.rm = TRUE)
+
+
[…]

### 2020-08-18 analysis/03-stan-run-models.R

@@ -14,18 +14,38 @@ alldatastan <- readRDS("data/alldatastan.rds")
 
 # Loading and running hierarchical model with non-centered parameterization
 modhierncp <- stan_model("analysis/survival_hier_ncp.stan")
-fithierncp <- sampling(object = modhierncp, data = alldatastan, iter = 5000, chains = 3,
+fithierncp <- sampling(object = modhierncp, data = alldatastan, iter = 500, chains = 3,
                        control = list(adapt_delta = 0.95, max_treedepth = 15))
 saveRDS(fithierncp, "data/stanfit_hier_ncp.rds")
 # fithierncp <- readRDS("data/stanfit_hier_ncp.rds")
+
+
+# Loading and running hierarchical model with non-centered parameterization AND new annual cv values for 1SW & 2SW separately
+modhierncp_acv <- stan_model("analysis/survival_hier_ncp_annualcvs.stan")
+fithierncp_acv <- sampling(object = modhierncp_acv, data = alldatastan, iter = 5000, chains = 3,
+                       control = list(adapt_delta = 0.95, max_treedepth = 15))
+print(fithierncp_acv)
+print(fithierncp)
+
+# fithierncpold <- fithierncp
+# fithierncp <- fithierncp_acv
+
+saveRDS(fithierncp_acv, "data/stanfit_hier_ncp_acv_full.rds")
+saveRDS(fithierncpold, "data/stanfit_hier_ncp_500.rds")
+
+saveRDS(fithierncp, "data/stanfit_hier_ncp.rds")
+# fithierncp <- readRDS("data/stanfit_hier_ncp.rds")
+
+
+
 
 # Investigating model object
 print(fithierncp, pars = "S1")
 print(fithierncp, pars = paste0("S1[",101:186,"]"))
 
-mcmc_intervals(fithierncp, pars = paste0("Pr_mu[",1:7,"]")) + 
+mcmc_intervals(fithierncp_acv, pars = paste0("Pr_mu[",1:7,"]")) + 
   scale_y_discrete(labels = alldatastan$river_name)
-traceplot(fithierncp, pars = paste0("S1[",170:186,"]"), inc_warmup = TRUE)
+traceplot(fithierncp_acv, pars = paste0("S1[",170:186,"]"), inc_warmup = TRUE)
 
 
 # Fitting non-hierarchical model (still with ncp)

### 2020-08-18 analysis/05-corrected-covariances.R

@@ -423,7 +423,7 @@ for (i in unique(river_names)) {
 }
 
 #saveRDS(mu2correctedcov, file = "data/corrected-correlations.rds")
-mu2correctedcov <- readRDS("data/corrected-correlations.rds")
+#mu2correctedcov <- readRDS("data/corrected-correlations.rds")
 library(scales)
 
 mu2correctedcov

### 2020-08-18 analysis/05-river-correlations.R

@@ -97,22 +97,27 @@ vec2mat <- function(M) {
 # corr1 <- river_cor("Saint-Jean River", "Western Arm Brook")
 
 rivercors <- lapply(riverpairs, function(x) river_cor(x[1], x[2]))
+lapply(rivercors, function(x) anyNA(x$cor))
 
-lapply(rivercors, function(x) quantile(x$cor, c(0.025, 0.5, 0.975)))
-medcor <- lapply(rivercors, function(x) quantile(x$cor, 0.5)) %>% unlist %>% vec2mat()
-corlow <- lapply(rivercors, function(x) quantile(x$cor, 0.025)) %>% unlist() %>% vec2mat()
-corhi <- lapply(rivercors, function(x) quantile(x$cor, 0.975)) %>% unlist() %>% vec2mat()
-pvals <- lapply(rivercors, function(x) quantile(x$cor, c(0.025, 0.5, 0.975))) %>%
+which(is.na(rivercors[[3]]$cor))
+which(is.na(rivercors[[4]]$cor))
+which(is.na(rivercors[[6]]$cor))
+
+lapply(rivercors, function(x) quantile(x$cor, c(0.025, 0.5, 0.975), na.rm = TRUE))
+medcor <- lapply(rivercors, function(x) quantile(x$cor, 0.5, na.rm = TRUE)) %>% unlist %>% vec2mat()
+corlow <- lapply(rivercors, function(x) quantile(x$cor, 0.025, na.rm = TRUE)) %>% unlist() %>% vec2mat()
+corhi <- lapply(rivercors, function(x) quantile(x$cor, 0.975, na.rm = TRUE)) %>% unlist() %>% vec2mat()
+pvals <- lapply(rivercors, function(x) quantile(x$cor, c(0.025, 0.5, 0.975), na.rm = TRUE)) %>%
   lapply(function(x) sign(x[1]*x[3])) %>% unlist %>% unname() %>% {ifelse(. == 1, 0.001, 0.5)} %>%
   vec2mat()
 
 ci.up <- 0.975
 ci.low <- 0.025
 
-medcor_un <- lapply(rivercors, function(x) quantile(x$uncor, 0.5)) %>% unlist %>% vec2mat()
-corlow_un <- lapply(rivercors, function(x) quantile(x$uncor, 0.025)) %>% unlist() %>% vec2mat()
[…]

### 2020-08-18 analysis/99-hypergeometric-bootstrap-function.R

@@ -1,0 +1,81 @@
+### Function to estimate coefficient of variation for 1SW and 2SW based on 
+### the total number of estimated small and large returns as well as a
+### subset of the scale data to assign small and large as  1SW and 2SW returns.
+### This function uses a hypergeometric distribution 
+
+hyper_boot <- function(year = NA, n_small, n_large, scales_small_1SW, scales_small_2SW,
+                       scales_small_other,
+                       scales_large_1SW, scales_large_2SW, 
+                       scales_large_other,
+                       iter = 10000, 
+                       debug = FALSE,
+                       list = FALSE,
+                       ...) {
+  
+  #if (!is.na(year) & year == 1995) browser()
+  
+  scales_small_total <- scales_small_1SW + scales_small_2SW + scales_small_other # total small scales samples
+  scales_large_total <- scales_large_1SW + scales_large_2SW + scales_large_other  # total large scales samples
+  
+  if (scales_small_total > n_small) stop(paste("There are more small scale samples than returns for year", year))
+  if (scales_large_total > n_large) stop(paste("There are more large scale samples than returns for year", year))
+  
+  # vectors of all potential abundances for small and large returns
+  if (n_small > 0) {vec_nsmall <- 0:n_small} else {vec_nsmall <- 0}
+  if (n_large > 0) {vec_nlarge <- 0:n_large} else {vec_nlarge <- 0}
+  
+  
+  prob_sm1 <- dhyper(scales_small_1SW, vec_nsmall, n_small - vec_nsmall, scales_small_total) # prob of small 1SW
+  prob_sm2 <- dhyper(scales_small_2SW, vec_nsmall, n_small - vec_nsmall, scales_small_total) # prob of small 2SW
+  prob_la1 <- dhyper(scales_large_1SW, vec_nlarge, n_large - vec_nlarge, scales_large_total) # prob of large 1SW
+  prob_la2 <- dhyper(scales_large_2SW, vec_nlarge, n_large - vec_nlarge, scales_large_total) # prob of large 2SW
+
+  prob_la1[is.nan(prob_la1)] <- 1
+  prob_la2[is.nan(prob_la2)] <- 1
+  
+  # Sampling vectors of abundances based on probabilities of hypergeometric distribution
+  hyp_draws_sm1 <- base::sample(vec_nsmall, iter, replace = TRUE, prob = prob_sm1)
[…]

### 2020-08-18 analysis/99-suppmat.R

@@ -23,6 +23,9 @@ estsmolts <- pivot_longer(logsmoltstib, V1:V7500, names_to = "iteration") %>%
            # n = n()
             )
 
+# Writing csv of (unlogged) smolts_true for use as input in CPM
+write_csv(estsmolts, path = "data/smolts-true-posteriors.csv")
+
 allyears <- estsmolts %>%
   group_by(Population) %>%
   expand(Year = full_seq(Year, 1))
@@ -132,7 +135,8 @@ logsmolts_cv <- tibble(river_name = rep(alldatastan$river_name, times = alldatas
 
 (smolts_tab <- pivot_wider(logsmolts, names_from = river_name, values_from = logsmolts) %>%
   arrange(year) %>%
-  rename(`WAB` = "Western Arm Brook") %>%
+    mutate(year = year - 1) %>%
+    rename(`WAB` = "Western Arm Brook") %>%
   rename_all(. %>% gsub(" River", "", .)) %>%
   rename(Year = year) %>%
   mutate(Year = as.character(Year)))
@@ -152,6 +156,7 @@ print.xtable(xtable(smolts_tab, digits = c(0,0,rep(0,7)),
 
 (smoltscv_tab <- pivot_wider(logsmolts_cv, names_from = river_name, values_from = logsmolts_cv) %>%
     arrange(year) %>%
+    mutate(year = year - 1) %>%
     rename(`WAB` = "Western Arm Brook") %>%
     rename_all(. %>% gsub(" River", "", .)) %>%
     mutate_at(.vars = vars(-year), .funs = list(~ . * 100)) %>%
@@ -170,5 +175,112 @@ print.xtable(xtable(smoltscv_tab, digits = c(0,0,rep(1,7)),
              include.rownames = FALSE)
 
 
+##### EPSILON ESTIMATES TABLE
+
+conneepsilon <- readRDS("data/conneepsilon.rds") %>% filter(year %in% na.omit(alldatastan$yearmat[3,]))
+wabepsilon <- readRDS("data/wabepsilon.rds")  %>% filter(year %in% na.omit(alldatastan$yearmat[5,]))
+sjepsilon <- readRDS("data/qcepsilon.rds") %>% 
+  filter(river_name == "Saint-Jean River" & year %in% na.omit(alldatastan$yearmat[1,]))
+triepsilon <- readRDS("data/qcepsilon.rds") %>% 
[…]

### 2020-10-07 ms/ms.tex
Add Geir and Brian edits Aug 17 draft

@@
 Declines in wild Atlantic salmon (\emph{Salmo salar}) abundance throughout the north
-Atlantic are primarily attributed to declining survival at sea. 
-Whether these declines in marine survival are highly variable among populations
+Atlantic are primarily attributed to decreases in survival at sea. 
+Whether these changes in marine survival are highly variable among populations
 has not been straightforward to estimate because of
@@
 trends among more than two or three populations. 
-Here, we develop a hierarchical Bayesian model that uses Murphy's maturity
+Accordingly, we develop a hierarchical Bayesian model that uses Murphy's maturity
 schedule method \citep{Murphy1952}, in conjunction with informative priors, to estimate yearly
@@
 
-We obtained time series data of outmigrating smolt and returning adult
+We obtained time series data of out-migrating smolt and returning adult
 abundances for seven Atlantic salmon populations in eastern Canada, encompassing a
@@
     \includegraphics[width=0.85\linewidth]{figures/rivers-map2.png}
-    \caption{Locations of the seven rivers in eastern Canada with time series abundance data of outmigrating smolts and 
-    returning adults.} \label{fig:map} 
+    \caption{Locations of the seven rivers in eastern Canada with time series abundance data of out-migrating smolts and 
+    returning adult Atlantic salmon.} \label{fig:map} 
 \end{figure}
@@
 salmon returning in years $t$ and $t+1$, respectively, $smolts_{true,t-1}$ is the
-estimated number of outmigrating smolts in year $t-1$, $S_{1,t}$ is the proportion of
+estimated number of out-migrating smolts in year $t-1$, $S_{1,t}$ is the proportion of
 salmon surviving in their first year ($t$) at sea, $P_{g,t}$ is the proportion of
@@
 The low $R^2$ for the Saint-Jean is likely due to the negative correlation
-between survival (i.e. $-Z_1$) and number of outmigrating smolts (see Fig.~\ref{fig:cor-params}). 
+between survival (i.e. $-Z_1$) and number of out-migrating smolts (see Fig.~\ref{fig:cor-params}). 
 There are also moderately negative correlations between survival and smolts in

### 2021-04-04 ms/suppmat.tex
Add changes up to ICES JMS submission

@@
 We incorporated the uncertainty in the conversion from size group (small and large returns) to age class (1SW and 2SW returns) by 
-approximating the variance of log return abundance of each maiden age class 1SW and 2SW by assuming that the probability 
-of the true number of returns of a given age, river, year, and size group ($N_{r,t,s,a}$) can be derived from a hypergeometric distribution
+approximating the variance of log return abundances of each maiden age class (i.e. 1SW and 2SW) assuming that the probability 
+of the true number of returns of a given age, river, year, and size group ($N_{r,t,s,a}$) can be derived by bootstrapping a distribution
+of number of returns from the hypergeometric distribution
+
+% \begin{equation}
+% p[N_{r,t,s,a}|0:N_{r,t,s}] \sim Hypergeometric(n_{r,t,s}, 0:N_{r,t,s}, N_{r,t,s}:0, n_{r,t,s,a}),
+% \end{equation}
 
 \begin{equation}
-p[N_{r,t,s,a}|0:N_{r,t,s}] \sim Hypergeometric(n_{r,t,s}, 0:N_{r,t,s}, N_{r,t,s}:0, n_{r,t,s,a}),
+P(N_{obs,r,t,s,a}|n_{r,t,s,a}) \sim Hypergeometric(N_{r,t,s,a}, N_{r,t,s} - N_{r,t,s,a}, n_{r,t,s}),
 \end{equation}
 
-where $n_{r,t,s,a}$ is the number of scale samples in river r, year t, of returning age a, and of size group s;
-$n_{r,t,s}$ is the total number of scale samples in river r, year t, and of size group s; and
-$N_{r,t,s}$ is the total number of returns in river, year t, and of size group s.
+where $P(N_{obs,r,t,s,a}|n_{r,t,s,a})$ is the probability of having a specific number of returns $N_{r,t,s,a}$ given 
+$n_{r,t,s,a}$, which is the number of scale samples in river r, year t, of returning age a, and of size group s;
+$N_{r,t,s}$ is the total number of returns in river r, year t, and of size group s;
+$N_{r,t,s} - N_{r,t,s,a}$ is the total numbers of returns in river r, year t, of size group s that are \textbf{not} of returning age a;
+and $n_{r,t,s}$ is the total number of scale samples in river r, year t, and of size group s.
+We iterated through all possible values of $N_{r,t,s,a}$ (i.e. from zero to $N_{r,t,s}$), to obtain a probavility for each value, which
+we then used as weighting probabilities to sample 10,000 random draws of $N_{r,t,s,a}$, which we denote as $\hat{N}_{r,t,s,a}$. 
+We then add the draws of the same age from both size groups to obtain a vector of estimates of number of returns of a given age a (i.e. 1SW or 2SW) for a given river and year ($\hat{N}_{r,t,a}$):
 
-We therefore obtain a vector of the probabilities of the returns of a given river r, year t, size group s and age a (i.e. $N_{r,t,s,a}$)
-across the whole range of potential values, i.e. from zero to $N_{r,t,s}$, which provides the weighting probabilities to generate 10,000 random draws of $N_{r,t,s,a}$. 
-We then add the draws of the same age from both size groups:
-to obtain the estimates of number of returns of a given age a (i.e. 1SW or 2SW) for a given river and year ($N_{r,t,a}$):
+% k is the number of successes ($n_{r,t,s,a}$) (successes)
+% N is the total number of individuals of size s ($N_{r,t,s}$) (total count) 
+% K is the number of age a individual of size s ($N_{r,t,s,a}$) (total successes)
+% n is the total number of scale samples ($n_{r,t,s}$) (trials)
+
+% where $n_{r,t,s,a}$ is the number of scale samples in river r, year t, of returning age a, and of size group s;
+% $n_{r,t,s}$ is the total number of scale samples in river r, year t, and of size group s; and
+% $N_{r,t,s}$ is the total number of returns in river, year t, and of size group s.
+
+% We therefore obtain a vector of the probabilities of the returns of a given river r, year t, size group s and age a (i.e. $N_{r,t,s,a}$)
+% across the whole range of potential values, i.e. from zero to $N_{r,t,s}$, which provides the weighting probabilities to generate 10,000 random draws of $N_{r,t,s,a}$.
 
 \begin{equation}
-N_{r,t,a} = N_{r,t,a,s=small}+ N_{r,t,a,s=large}.
+    \hat{N}_{r,t,a} = \hat{N}_{r,t,a,s=small }+ \hat{N}_{r,t,a,s=large}.
 \end{equation}
 
-Lastly, as the coefficient of variation are comparable to standard deviations of less than 0.3 in the log-scale, we obtain
-the estimate of $\epsilon_{r,t,a}$ as
+Lastly, we obtain the estimates of $\epsilon_{r,t,a}$ by calculating the standard deviation of log-transformed $\hat{N}_{r,t,a}$:
 
 \begin{equation}
-\epsilon_{r,t,a} \approx \sigma_{log(N_{r,t,a})}.
+\epsilon_{r,t,a} = \sigma_{log(\hat{N}_{r,t,a})}.
 \end{equation}

### 2021-04-04 analysis/00-campbellton-wrangling.R

@@ -2,6 +2,7 @@
 
 source("analysis/99-hypergeometric-bootstrap-function.R")
 library(tidyverse)
+library(zoo)
 
 campsmolts <- read_csv("raw-data/NL-smolt-estimates-Kelly-et-al-2017-ResDoc.csv",
          na = c("-", "NA")) %>%
@@ -15,10 +16,32 @@ camp <- readRDS(file = "~/Dropbox/salmon-lh-variation/output/NL/campbellton-esti
 
 # Loading raw scale age data
 campraw <- readRDS("~/Dropbox/salmon-lh-variation/output/NL/campraw.rds") %>%
-  select(year, sl, M1, M2p, RStotal)
+  select(year, sl, M1, M2p, RStotal) 
+campraw[is.na(campraw)] <- 0 
 
-campwide <- pivot_wider(campraw, names_from = sl, values_from = c(M1, M2p, RStotal)) 
-campwide[is.na(campwide)] <- 0 # replacing NAs with zeros
+
+
+
+campwide <- pivot_wider(campraw, names_from = sl, values_from = c(M1, M2p, RStotal)) %>%
+  ungroup() %>%
+  complete(year = full_seq(year, 1))
+campwide[is.na(campwide)] <- 0 # replacing NAs with zeros 
+
+campwide %>% 
+  mutate(M1_small_mean = rollapply(M1_small, 3, mean, partial = TRUE) %>% round(),
+         M1_large_mean = rollapply(M1_large, 3, mean, partial = TRUE) %>% round(),
+         M2p_small_mean = rollapply(M2p_small, 3, mean, partial = TRUE) %>% round(),
+         M2p_large_mean = rollapply(M2p_large, 3, mean, partial = TRUE) %>% round(),
+         RStotal_small_mean = rollapply(RStotal_small, 3, mean, partial = TRUE) %>% round(),
+         RStotal_large_mean = rollapply(RStotal_large, 3, mean, partial = TRUE) %>% round()) %>%
+  mutate(M1_large = if_else(M1_small <= 3, M1_large_mean, M1_large),
+         M2p_small = if_else(M1_small <= 3, M2p_small_mean, M2p_small),
+         M2p_large = if_else(M1_small <= 5, M2p_large_mean,  M2p_large),
+         RStotal_small = if_else(M1_small <= 3, RStotal_small_mean, RStotal_small),
+         RStotal_large = if_else(M1_small <= 3, RStotal_large_mean, RStotal_large),
[…]

### 2021-04-04 analysis/05-figures-for-trends-paper.R

@@ -22,7 +22,7 @@ addnas <- tibble(pos = NA,
 
 s1quant <- bind_rows(s1quant, addnas)
 
-morat2 <- tibble(year = c(1984, 1984, 1998, 1998, 1998, 2000, 2000),
+morat2 <- tibble(year = c(1984, 1984, 1992, 1992, 1992, 2000, 2000),
                  region = c("NB","NS", "NL", "NL", "NL", "QC", "QC"),
                  river_name = c("Nashwaak River", "LaHave River", 
                                 "Western Arm Brook", "Campbellton River", "Conne River", 
@@ -47,7 +47,7 @@ ggsave("figures/s1-trends-presentation.png", height = 4, width = 8)
 
 
 
-morat <- tibble(year = c(1984, 1998, 2000),
+morat <- tibble(year = c(1984, 1992, 2000),
                 region = c("NS+\nNB", "NL", "QC"),
                 river_name = c("Nashwaak River", "Western Arm Brook", "Trinité River"))
 

### 2021-04-04 analysis/99-data-exports.R

@@ -31,20 +31,19 @@ writeMat("data/alldatasep.mat",
 
 # In the end we didn't use MATLAB export and went with a simple csv table instead:
 
-alldatastan <- readRDS("data/alldatastan.rds")
-
-returnsdata <- alldatastan[names(alldatastan) %in% c("years", "logsmolts", "logsmolts_cv", "loggrilse", "logSW2")] %>% do.call(cbind, .) %>% as_tibble
+returnsdata <- alldatastan[names(alldatastan) %in% 
+                             c("years", "logsmolts", "logsmolts_cv", "loggrilse", "logSW2",
+                               "logSW1_cv", "logSW2_cv")] %>% do.call(cbind, .) %>% as_tibble
 
 otherdata <- tibble(nyears = alldatastan$nyears,
        river_name = alldatastan$river_name,
-       returns_cv = alldatastan$returns_cv,
        logis_mu = alldatastan$logis_mu,
        logis_sigma = alldatastan$logis_sigma,
        ny = nyears) %>% uncount(ny)
 
 alldata_table <- bind_cols(returnsdata, otherdata) %>%
   select(river_name, year = years, logsmolts, logsmolts_cv, 
-         logSW1 = loggrilse, logSW2, returns_cv, everything())
+         logSW1 = loggrilse, logSW2, logSW1_cv, logSW2_cv, everything())
 
 write.csv(alldata_table, file = "data/alldata_table.csv")
 

### 2021-04-04 analysis/99-suppmat.R

@@ -164,7 +164,7 @@ print.xtable(xtable(smolts_tab, digits = c(0,0,rep(0,7)),
     mutate(Year = as.character(Year)))
 
 print.xtable(xtable(smoltscv_tab, digits = c(0,0,rep(1,7)),
-                    caption = "Annual coefficient of variation (CV, \\%) estimates for the seven populations
+                    caption = "Annual coefficient of variation (CV, \\%) in smolt estimates for the seven populations
                     examined. Where possible, we estimated CV directly
                     from uncertainty in smolts estimates."),
              file = "ms/smoltscv_tab.tex",
@@ -215,10 +215,9 @@ epsilons2 <- select(epsilons, river_name, year, sdlog2SW)
     mutate(Year = as.character(Year)))
 
 print.xtable(xtable(epsilon1_tab, digits = c(0,0,rep(3,7)),
-                    caption = "Annual estimates of error in log-transformed abundance of 1SW returns 
-                    ($\\epsilon_{1}$) for the seven populations
-                    examined. For all populations except Nashwaak and LaHave, we estimated the CV 
-                    from the uncertainty in the conversion from size group to age using aged scale data.",
+                    caption = "Annual error estimates derived from the standard deviation of log-transformed abundance 
+                    in 1SW returns ($\\epsilon_{1}$), derived from size group to age conversion using aged scale data 
+                    (except for Nashwaak and LaHave populations).",
                     label = "tab:epsilon1"),
              file = "ms/epsilon1_tab.tex",
              NA.string = "-",
@@ -245,10 +244,9 @@ print.xtable(xtable(epsilon1_tab, digits = c(0,0,rep(3,7)),
 
 
 print.xtable(xtable(epsilon2_tab, digits = c(0,0,rep(3,7)),
-                    caption = "Annual estimates of error in log-transformed abundance of 2SW returns 
-                    ($\\epsilon_{2}$) for the seven populations
-                    examined. For all populations except Nashwaak and LaHave, we estimated the CV 
-                    from the uncertainty in the conversion from size group to age using aged scale data.",
+                    caption = "Annual error estimates derived from the standard deviation of log-transformed abundance 
+                    in 2SW returns ($\\epsilon_{2}$), derived from size group to age conversion using aged scale data 
+                    (except for Nashwaak and LaHave populations).",
                     label = "tab:epsilon2"),
              file = "ms/epsilon2_tab.tex",
              NA.string = "-",

### 2021-04-23 ms/suppmat.tex
Edits resulting from comments to reviewers ICES JMS

@@
         Julien April\textsuperscript{4}, Ross A. Jones\textsuperscript{5}, %Martha J. Robertson\textsuperscript{3},
-        Dustin Raab\textsuperscript{6}, Jeffrey A. Hutchings\textsuperscript{1}
+        Dustin Raab\textsuperscript{6}, Jeffrey A. Hutchings\textsuperscript{1,7,8}
 \newline
@@
 \small{\textsuperscript{6} Fisheries and Oceans Canada, Dartmouth, NS, Canada\\}
+\small{\textsuperscript{7} Institute of Marine Research, Flødevigen Marine Research Station, N-4817 His, Norway\\}
+\small{\textsuperscript{8} Centre for Coastal Research, University of Agder, N-4604 Kristiansand, Norway\\}
 \small{\textsuperscript{*} Corresponding author: spardo@dal.ca}
@@
 and $n_{r,t,s}$ is the total number of scale samples in river r, year t, and of size group s.
-We iterated through all possible values of $N_{r,t,s,a}$ (i.e. from zero to $N_{r,t,s}$), to obtain a probavility for each value, which
+We iterated through all possible values of $N_{r,t,s,a}$ (i.e. from zero to $N_{r,t,s}$), to obtain a probability for each value, which
 we then used as weighting probabilities to sample 10,000 random draws of $N_{r,t,s,a}$, which we denote as $\hat{N}_{r,t,s,a}$.
@@
 abundance draws of a uniform distribution. Thus, the value of $\epsilon$ for
-2011, the year with no scale samples, is of 0.711 (Table~\ref{tab:epsilon1}),
+2011, the year with no scale samples, is of 0.719 (Table~\ref{tab:epsilon1}),
 which is likely an underestimate of true uncertainty given the absence of
@@
 returns (up to 5443). b) The same distribution but log-transformed, which results in a right-skewed distribution. The blue line represents the normal
-distribution fitted to the log-transformed data, from which $\epsilon_{1}$ is equal to its standard deviation. b) Distribution of draws of potential
-    1SW abundances, adding the estimates from both small and large returns, for the year 2014, which is much tightly constrained. d) The distribution
+distribution fitted to the log-transformed data, from which $\epsilon_{1}$ is equal to its standard deviation. c) Distribution of draws of potential
+    1SW abundances, adding the estimates from both small and large returns, for the year 2014, which is much more tightly constrained. d) The distribution
 of log-transformed abundance draws for year 2014. In this case, the normal distribution fitted is much tighter and more closely resembles the underlying histogram,
@@
 
-\subsection*{Error in smolt abundances used in the model}
+\subsection*{Uncertainty in smolt abundances used in the model}
 

### 2021-04-23 analysis/05-figures-for-trends-paper.R

@@ -281,8 +281,8 @@ pop_logitpr <- median_logitpr %>%
 ggplot(pop_logitpr, aes(x, y, color = Population)) + geom_line(aes(group = Population), size = 2)
 
 p1 + geom_line(inherit.aes = FALSE, data = pop_logitpr, aes(x, y, group = Population), 
-               size = 1.3, color = "gray40", alpha = 0.5) + 
+               size = 1.3, color = "gray20", alpha = 0.5) + 
   coord_cartesian(ylim = c(0, 5)) + guides(color = guide_legend(override.aes = list(size = 1.3, alpha = 1))) +
   theme_cowplot() +
-  xlab("Proportion returning after first winter at sea") + ylab("Probability")
+  xlab("Proportion returning after first winter at sea") + ylab("Density")
 ggsave("figures/logitpr-mu-sigma-post.png", width = 7, height = 4)

### 2021-04-23 analysis/99-data-exports.R

@@ -1,6 +1,6 @@
 # Data export to share with Maria Tirronen for change point analysis
 library(tidyverse)
-library(R.matlab)
+# library(R.matlab)
 alldata <- readRDS("data/alldata.rds")
 alldatastan <- readRDS("data/alldatastan.rds")
 
@@ -15,19 +15,19 @@ names(alldata)[names(alldata) == "loggrilse"] <- "logSW1"
 
 saveRDS(alldata, "data/alldata_export.rds")
 
-writeMat("data/alldata.mat", alldata = alldata)
-writeMat("data/alldatasep.mat", 
-         years = alldata$years,
-         logsmolts = alldata$logsmolts,
-         logsmolts_cv = alldata$logsmolts_cv,
-         logSW1 = alldata$logSW1,
-         logSW2 = alldata$logSW2,
-         returns_cv = alldata$returns_cv,
-         nyears = alldata$nyears,
-         river_name = alldata$river_name,
-         logis_mu = alldata$logis_mu,
-         logis_sigma = alldata$logis_sigma
-         )
+# writeMat("data/alldata.mat", alldata = alldata)
+# writeMat("data/alldatasep.mat", 
+#          years = alldata$years,
+#          logsmolts = alldata$logsmolts,
+#          logsmolts_cv = alldata$logsmolts_cv,
+#          logSW1 = alldata$logSW1,
+#          logSW2 = alldata$logSW2,
+#          returns_cv = alldata$returns_cv,
+#          nyears = alldata$nyears,
+#          river_name = alldata$river_name,
+#          logis_mu = alldata$logis_mu,
+#          logis_sigma = alldata$logis_sigma
+#          )
 
 # In the end we didn't use MATLAB export and went with a simple csv table instead:
 

### 2021-04-23 analysis/99-suppmat.R

@@ -24,7 +24,7 @@ estsmolts <- pivot_longer(logsmoltstib, V1:V7500, names_to = "iteration") %>%
             )
 
 # Writing csv of (unlogged) smolts_true for use as input in CPM
-write_csv(estsmolts, path = "data/smolts-true-posteriors.csv")
+write_csv(estsmolts, file = "data/smolts-true-posteriors.csv")
 
 allyears <- estsmolts %>%
   group_by(Population) %>%
@@ -143,7 +143,7 @@ logsmolts_cv <- tibble(river_name = rep(alldatastan$river_name, times = alldatas
 
 print.xtable(xtable(smolts_tab, digits = c(0,0,rep(0,7)),
                     display= c("s", "s", rep("f",7)),
-                    caption = "Empirical annual smolt estimates for the seven populations examined."),
+                    caption = "Empirically estimated annual smolt abundances for the seven populations examined. Smolt abundance estimates from the Trinité, Saint-Jean, LaHave, Nashwaak, and Conne populations were obtained using a mark-recapture approach, while estimates from the WAB and Campbellton populations were obtained by direct counts using fish counting fences."),
              file = "ms/smolts_tab.tex",
              size = "footnotesize",
              format.args = list(big.mark = ","),
@@ -187,7 +187,7 @@ campepsilon <- readRDS("data/campepsilon.rds") %>% filter(year %in% na.omit(alld
 
 allscales <- bind_rows(campepsilon, wabepsilon, conneepsilon, sjepsilon, triepsilon)
 
-epsilonslh <- tibble(river_name = "LaHafootnoteve River",  
+epsilonslh <- tibble(river_name = "LaHave River",  
                      year = na.omit(alldatastan$yearmat[7,]),
                      sdlog1SW = 0.01, sdlog2SW = 0.06)
 

### 2021-05-24 analysis/05-corrected-covariances.R
Update indexing of tibbles to work with R 4.0

@@ -107,28 +107,28 @@ for (i in unique(river_names)) {
   
   # storing values, na.rm = TRUE to ignore NaNs created when square-rooting a few negative values
   mu1correctedcov[mu1correctedcov$river_name == i & 
-                    mu1correctedcov$correlation == "Corrected cov(logsmolts,-Z1)", 3:5] <- quantile(corrected_corlogsmoltsz1, 
-                                                                                                    c(0.5, 0.05, 0.95), na.rm = TRUE)
+                    mu1correctedcov$correlation == "Corrected cov(logsmolts,-Z1)", 3:5] <- as.list(quantile(corrected_corlogsmoltsz1, 
+                                                                                                    c(0.5, 0.05, 0.95), na.rm = TRUE))
   mu1correctedcov[mu1correctedcov$river_name == i & 
-                    mu1correctedcov$correlation == "Corrected cov(logsmolts,logPr)", 3:5] <- quantile(corrected_corlogsmoltslogpr, 
-                                                                                                      c(0.5, 0.05, 0.95), na.rm = TRUE)
+                    mu1correctedcov$correlation == "Corrected cov(logsmolts,logPr)", 3:5] <- as.list(quantile(corrected_corlogsmoltslogpr, 
+                                                                                                      c(0.5, 0.05, 0.95), na.rm = TRUE))
   mu1correctedcov[mu1correctedcov$river_name == i & 
-                    mu1correctedcov$correlation == "Corrected cov(-Z1,logPr)", 3:5] <- quantile(corrected_corz1logpr, 
-                                                                                                c(0.5, 0.05, 0.95), na.rm = TRUE)
+                    mu1correctedcov$correlation == "Corrected cov(-Z1,logPr)", 3:5] <- as.list(quantile(corrected_corz1logpr, 
+                                                                                                c(0.5, 0.05, 0.95), na.rm = TRUE))
   
    
     # storing values, na.rm = TRUE to ignore NaNs created when square-rooting a few negative values
   mu1correctedrsq[mu1correctedrsq$river_name == i & 
-                    mu1correctedrsq$correlation == "R2(mu1,logsmolts)", 3:5] <- quantile(corrected_r2_mu1logsmolts,
[…]

### 2021-05-24 ms/suppmat.tex
Add prior sensitivity tests

@@
 \newcommand{\St}{$S_{2}$\xspace}
-\newcommand{\Pg}{$P_r$\xspace}
+\newcommand{\Pg}{$P_g$\xspace}
+\newcommand{\prmu}{$\mu_{Pg,r}$\xspace}
 \newcommand{\Linf}{$L_{\infty}$}
@@
 
-% \begin{figure}[htbp] \centering
-%     \includegraphics[width=0.9\linewidth]{simul_timeseries_declining.pdf} \caption{Simulated
-%         time series of returning adult salmon abundance in the six scenarios. The lines denote the
-%     simulated abundance estimates without observation error while the points are the same estimates
-% including observation error.}
-% \label{fig:simul-ts-dec}
-% \end{figure}
+\clearpage
 
-% \begin{figure}[htbp] \centering
-%     \includegraphics[width=0.95\linewidth]{sir_out_18_declining.pdf} \caption{Yearly
-%         estimated \So, \St, and \Pg values in the six scenarios. True values
-%         are denoted by blue circles, black circles show median estimates,
-%         error bars indicate the 25\% and 75\% quantiles.}
-% \label{fig:sir-out-dec}
-% \end{figure}
+\subsection*{Posterior estimates with weaker priors}
 
-% \begin{figure}[htbp] \centering
-%     \includegraphics[width=0.7\linewidth]{simul_plots_6_declining.pdf}
-%     \caption{Comparison of estimated and true \So values in the six
-%         scenarios. True \So values are deterministic, black circles show
-%         median \So estimates, error bars indicate the  25\% and 75\%
-%         quantiles, while the blue line denotes a linear model fit of the
-%         medians. The one-to-one relationship is shown by the gray dashed
-%         line.}
-% \label{fig:simul-six-plots-dec}
-% \end{figure}
+To assess the sensitivity of our model parameter estimates to the choice of
+priors, we assessed posterior estimate under two weaker prior scenarios:
+one with wider priors for \So, \St (Fig.~\ref{fig:zweak}), and a second one with wide priors
+of \prmu for all populations regardless of whether they were considered
+1SW-dominated or not (Fig.~\ref{fig:prweak}).
+
+\begin{figure}[htbp] \centering
+    \includegraphics[width=0.55\linewidth]{figures/z-weak-prior.png}
+    \caption{Priors used for marine survival of one sea-winter (\So) and two
+        sea-winter (\St) returns. Solid lines denote strong priors used in the
+        study, while dotted lines denote weaker priors used to assess
+        sensitivity to prior choice.} \label{fig:zweak} \end{figure}
+
+\begin{figure}[htbp] \centering
+    \includegraphics[width=0.55\linewidth]{figures/pr-weak-prior.png}
+    \caption{Prior of the probability of returning as 1SW for all populations
+        under the weak \prmu prior scenario. Note this prior is identical to
+        that used in the study for non 1SW-dominated populations.}
+    \label{fig:prweak} \end{figure}
+
+The model runs with weaker priors did not differ mayorly from those with
+stronger priors, particularly for posterior estimates of \So, which show the
+same overall trends for all populations and some yearly \So estimates with
+high uncertainty (i.e., large error bars) show greater uncertainty uncertainty
+under the weak \So and \St scenario (Fig.~\ref{fig:s1-trends-priors}).
+Not unexpectedly, the weak \Pg scenario had no visually discernible effect on
+posterior estimates of \So.
+
+\begin{figure}[htbp] \centering
+    \includegraphics[width=0.95\linewidth]{figures/s1-trends-priors.png}
+    \caption{Median posterior estimates of survival in the first year at sea
+        (\So) for the seven populations examined under three scenarios with
+        different priors. The strong prior scenario is the one presented in
+        our study, the weak \Pg priors uses the same weak prior for all
+        populations as shown by Fig.~\ref{fig:prweak}, while the weak \So \&
+        \St scenario used wider survival priors as shown in Fig.~\ref{fig:zweak}.} 
+    \label{fig:s1-trends-priors} \end{figure}
+
+Having weaker \So and \St priors did considerable widen the posterior
+estimates of \St for all populations (Fig.~\ref{fig:s2-trends-priors}). The
+Saint-Jean River, which had the least uncertain \St estimates among all
+populations, had considerably larger uncertainties in \St while still
+maintaining a very similar temporal trend in \St. The Trinité River also
+maintained a similar temporal trend in \St, but the weaker priors resulted in
+annual estimates of \St showing a larger spread, particularly in 1994 and 1995
+which were the years withe highest estimated \St.
+
+\begin{figure}[htbp] \centering
+    \includegraphics[width=0.95\linewidth]{figures/s2-trends-priors.png}
+    \caption{Median posterior estimates of survival in the second year at sea
+        (\St) for the seven populations examined under three scenarios with
+        different priors. The strong prior scenario is the one presented in
+        our study, the weak \Pg priors uses the same weak prior for all
+        populations as shown by Fig.~\ref{fig:prweak}, while the weak \So \&
+        \St scenario used wider survival priors as shown in Fig.~\ref{fig:zweak}.} 
+    \label{fig:s2-trends-priors} \end{figure}
+
+Estimates of \Pg under the weak \Pg scenario look almost identical to those
+under the strong priors scenario, however the weak \So and \St scenario
+resulted in more uncertain \Pg estimates for non 1SW-dominated populations
+(Fig.~\ref{fig:pr-trends-priors}). For these populations, the temporal trends
+observed in the strong prior scenario were maintained under the weak \So and
+\St scenario, albeit weakened by the larger uncertainties in annual \Pg
+estimates.
+
+Overall, weakening the \prmu priors did not have an observable effect on model
+parameter estimates, while weakening \So and \St did result in more uncertain
+estimates of \St and \Pg, with estimates of \So being mostly unchanged.
+
+\begin{figure}[htbp] \centering
+    \includegraphics[width=0.95\linewidth]{figures/pr-trends-priors.png}
+    \caption{Median posterior estimates of proportion returning after one winter at sea 
+        (\Pg) for the seven populations examined under three scenarios with
+        different priors. The strong prior scenario is the one presented in
+        our study, the weak \Pg priors uses the same weak prior for all
+        populations as shown by Fig.~\ref{fig:prweak}, while the weak \So \&
+        \St scenario used wider survival priors as shown in Fig.~\ref{fig:zweak}.} 
+    \label{fig:pr-trends-priors} \end{figure}
+
+We calculated the correlations of estimated trends in $Z_1$ among populations
+under the weak \So and \St scenario. While the direction of the correlation
+stayed the same between all populations when compared with the strong prior
+scenario, there were differences in the correlation coefficients (Fig.~\ref{fig:s1-corr-weakz}).
+Among these, the correlation coefficients differed significantly for three pair-wise
+comparisons: the correlations between Campbellton and Nashwaak as well
+Campbellton and Trinité changed from non-significant to positively
+significant, while the correlation between LaHave and Saint-Jean changed from negatively significant to
+non-significant even though the coefficient value decreased further into the
+negative.
+Twelve out of the 21 pair-wise correlations were significant (p-value below 0.05) yet under the  
+weak \So and \St scenario 9 out of the twelve significant correlations were positive.
+
+\begin{figure}[htbp] \centering
+    \includegraphics[width=0.95\linewidth]{figures/corrplot-post-z1-corrected-weakz.png} \caption{
+        Correlations of estimated trends in $Z_1$ among populations with weak \So and
+\St priors. Correlation coefficients are shown in each square, while colouring
+        denotes significance of the correlation ($p \leq 0.05$) as well as the direction and magnitude.}
+\label{fig:s1-corr-weakz}
+\end{figure}
 
@@
 %\subsection*{References}
+
 \nobibliography{subset}

### 2021-05-24 analysis/04-run-prior-sensitivity-models.R

@@ -1,0 +1,96 @@
+# Running Stan models with differing prior assumptions to assess sensitivity
+# to prior choice
+#  Sensitivity tests: 
+# (1) use a relatively weak informative prior for Z (equations 12 and 13), e.g., using sigma = 0.5 or 0.8; 
+# (2) increase/decrease the median values for Z by 10% or 20%; 
+# (3) you may also do the similar tests for parameter Pg,r,t in equations 14-16. 
+  
+library(tidyverse)
+library(rstan)
+library(bayesplot)
+
+options(mc.cores = parallel::detectCores())
+rstan_options(auto_write = TRUE)
+
+# Loading data
+alldatastan <- readRDS("data/alldatastan.rds")
+
+# (1) use a relatively weak informative prior for Z (equations 12 and 13), e.g., using sigma = 0.5 
+# (2) increase/decrease the median values for Z by 10% or 20%; 
+# lower values of mu (towards -Inf result in higher S values skewed towards 1)
+# values closer to zero center S around 0.5. 
+# Basically, this is weakening the prior towards a median of 0.5 with wider tails
+alldatastan_zmu_h <- alldatastan
+alldatastan_zmu_h$z1_mu <- 0.5
+alldatastan_zmu_h$z1_sd <- 0.5
+alldatastan_zmu_h$z2_mu <- 0.1
+alldatastan_zmu_h$z2_sd <- 0.5
+
+x <- seq (0.001, 0.999, by =0.001)
+hist(exp(-rlnorm(5000, alldatastan_zmu_h$z1_mu, alldatastan_zmu_h$z1_sd)))
+hist(exp(-rlnorm(5000, alldatastan_zmu_h$z2_mu, alldatastan_zmu_h$z2_sd)))
+
+# In the original stan model:
+# Z1 ~ lognormal(1, 0.22);    
+# Z2 ~ lognormal(0.2, 0.3);
+
+plot(x, dlnorm(-log(x), 1, 0.22))
+plot(x, dlnorm(-log(x), alldatastan_zmu_h$z1_mu, alldatastan_zmu_h$z1_sd))
+plot(x, dlnorm(-log(x), 0.2, 0.3))
+plot(x, dlnorm(-log(x), alldatastan_zmu_h$z2_mu, alldatastan_zmu_h$z2_sd))
+
+
[…]

### 2021-05-24 analysis/06-prior-sensitivity-tests.R

@@ -1,0 +1,370 @@
+# Sensitivity analyses of models with differing priors
+library(tidyverse)
+library(rstan)
+
+alldatastan <- readRDS("data/alldatastan.rds")
+
+# Loading stanfit objects 
+fithierncp <- readRDS("data/stanfit_hier_ncp_acv_full.rds")
+fithierncp_prweak <- readRDS("data/stanfit_hier_ncp_acv_prweak.rds")
+fithierncp_zsdmu <- readRDS("data/stanfit_hier_ncp_acv_zsdmu.rds")
+
+
+# Indexing positions of years for each river
+indexdf <- tibble(pos = as.character(1:length(alldatastan$years)),
+                  river_name = rep(alldatastan$river_name, 
+                                   times = alldatastan$nyears), 
+                  year = alldatastan$years)
+
+# For indexing population level posteriors
+riverindexdf <- tibble(pos = as.character(1:length(alldatastan$river_name)),
+                       river_name = alldatastan$river_name)
+
+# Extracting posterior distributions
+extract_post <- function(stanfit, pars = "S1", index = indexdf) {
+  as.matrix(stanfit, pars = pars) %>%
+    as_tibble %>%
+    gather(parameter, value) %>%
+    mutate(pos = str_extract(parameter, "(?<=\\[).+?(?=\\])"),
+           param = str_extract(parameter, ".+?(?=\\[)")) %>%
+    left_join(index, by = "pos")
+}
+
+addnas <- tibble(pos = NA,
+                 river_name = c("Saint-Jean River", "Trinité River", rep("LaHave River", 2)),
+                 year = c(1998, 2007, 2012:2013),
+                 median = c(NA),
+                 q05 = c(NA),
+                 q25 = c(NA),
+                 q75 = c(NA),
+                 q95 = c(NA),
+                 zscore = NA
+)
+
+morat2 <- tibble(year = c(1984, 1984, 1992, 1992, 1992, 2000, 2000),
+                 region = c("NB","NS", "NL", "NL", "NL", "QC", "QC"),
+                 river_name = c("Nashwaak River", "LaHave River", 
+                                "Western Arm Brook", "Campbellton River", "Conne River", 
[…]

### 2021-05-24 ms/ms.tex
Update manuscript to include Geir comments and prior sensitivity text

@@
 \bibpunct{(}{)}{;}{a}{}{;}
-\bibliographystyle{fishfishnourl}
+\bibliographystyle{icesjms}
 
@@
         Dustin Raab\textsuperscript{6},
-Jeffrey A. Hutchings\textsuperscript{1}}
+Jeffrey A. Hutchings\textsuperscript{1,7,8}}
 
@@
 Atlantic are primarily attributed to decreases in survival at sea.
-Whether these changes in marine survival are highly variable among populations
-has not been straightforward to estimate because of
-the need to model data on both migrating smolts and returning adults and to
-simultaneously estimate multiple parameters. We fit a hierarchical Bayesian
-maturity schedule model to data spanning between 19 and 42 years for seven populations in eastern 
-Canada. We estimate numbers of out-migrating smolts, 
+However, comparing trends in marine survival among populations
+is challenging as data on both migrating smolts and returning adults is sparse and
+models are difficult to parameterize due to their varied life histories. We fit a hierarchical Bayesian
+maturity schedule model to data from seven populations in eastern
+Canada to estimate numbers of out-migrating smolts,
 survival in the first and second year at sea, and the proportion returning after one year.
-Trends in survival at sea were not consistent among populations, and we observe both
-positive, negative and no correlations in the trends. This suggest that large
+Trends in survival at sea were not consistent among populations; we observe
+positive, negative and no correlations in these, suggesting that large
 scale patterns of changes in marine survival are not necessarily representative for
@@
 However, variation in the other components were not negligible and
-their relative importance differed among populations. Variation in number of
-out-migrating smolts had a moderate contribution on returns and probability of returning after one year was
-important in some populations. 
-% Trends were not consistent among rivers. 
-% Since 1990, \So increased at Western Arm Brook (NL)
-% and Rivi\`{e}re Saint-Jean (QC), but declined at Conne River (NL) and Rivi\'{e}re
-% de la Trinit\'{e} (QC). Since the mid-1990s, \So increased at Campbellton River (NL),
-% declined at LaHave River (NS), and fluctuated at Nashwaak
-% River (NB). Estimates of \St were highly uncertain, particularly for
-% 1SW-dominated populations; \Pg was generally stable. 
-% These results challenge the narrative that marine survival has changed in a
-% temporally consistent manner among spatially disparate populations. 
-If salmon populations do not respond in a consistent, uniform manner to
-changing environmental (including oceanic) conditions throughout their range,
+their relative importance differed among populations.
+If salmon populations do not respond in a uniform manner to
+changing environmental conditions throughout their range,
 future research initiatives should explore why.
+
+% Declines in wild Atlantic salmon (\emph{Salmo salar}) abundance throughout the north
+% Atlantic are primarily attributed to decreases in survival at sea.
+% Whether these changes in marine survival are highly variable among populations
+% has not been straightforward to estimate because of
+% the need to model data on both migrating smolts and returning adults and to
+% simultaneously estimate multiple parameters. We fit a hierarchical Bayesian
+% maturity schedule model to data spanning between 19 and 42 years for seven populations in eastern
+% Canada. We estimate numbers of out-migrating smolts,
+% survival in the first and second year at sea, and the proportion returning after one year.
+% Trends in survival at sea were not consistent among populations, and we observe both
+% positive, negative and no correlations in the trends. This suggest that large
+% scale patterns of changes in marine survival are not necessarily representative for
+% individual populations.
+% Variation in return abundances was mostly explained by marine survival in the
+% first winter at sea in all but one population.
+% However, variation in the other components were not negligible and
+% their relative importance differed among populations. Variation in number of
+% out-migrating smolts had a moderate contribution on returns and probability of returning after one year was
+% important in some populations.
+% % Trends were not consistent among rivers.
+% % Since 1990, \So increased at Western Arm Brook (NL)
+% % and Rivi\`{e}re Saint-Jean (QC), but declined at Conne River (NL) and Rivi\'{e}re
+% % de la Trinit\'{e} (QC). Since the mid-1990s, \So increased at Campbellton River (NL),
+% % declined at LaHave River (NS), and fluctuated at Nashwaak
+% % River (NB). Estimates of \St were highly uncertain, particularly for
+% % 1SW-dominated populations; \Pg was generally stable.
+% % These results challenge the narrative that marine survival has changed in a
+% % temporally consistent manner among spatially disparate populations.
+% If salmon populations do not respond in a consistent, uniform manner to
+% changing environmental (including oceanic) conditions throughout their range,
+% future research initiatives should explore why.
 
@@
 marine survival.
-\citet{Chaput2012a}, for example, examined the return rate of smolts to adult salmon 
+For example, \citet{Chaput2012a} examined return rates (i.e., ratio of adult salmon to outgoing smolts)
 as a metric of marine survival, finding that most Canadian populations
@@
 \begin{equation}
-    p_{r,t,a} = \frac{\sum_{s}{(\frac{n_{r,t,s,a}}{n_{r,t,s}} * N_{r,y,s})}}{\sum_{s}{N_{r,t,s}}}
+    p_{r,t,a} = \frac{\sum_{s}{(\frac{n_{r,t,s,a}}{n_{r,t,s}} * N_{r,t,s})}}{\sum_{s}{N_{r,t,s}}}
 \end{equation}
@@
 in river $r$, year $t$, and of size group $s$.
-Ideally, we would account the uncertainty in the conversion between
-size group and returning age in the model, but given limitations of the Bayesian 
-sampler we used, we incorporated this uncertainty indirectly as an estimate
-of the variance of annual log-smolt abundance.
+For years where scale data of a given size group is lacking,
+we averaged the proportions of annual returns
+for the closest ten years for which there was data.
+We incorporated the uncertainty in the conversion between
+size group and returning age in the model indirectly as an estimate
+of the variance of annual log-1SW and 2SW return abundance (see Supplementary Materials).
 
@@
 the parameter estimates cannot be optimally solved \citep{Chaput2003a}.
-However, this issue can be mathematically overcome, at least partially, by
+However, this issue can be mathematically overcome by
 using informative priors for the two marine survival parameters and the maturation
@@
 three or more winters at sea before returning to spawn for the first time.
+Assuming that there are no age 3+ maiden spawners is likely to result in
+slightly lower S1 and S2 estimates as these older maiden spawners that would
+be returning in later years are assumed to have died. Nonetheless, in the
+populations examined these fish comprise a very small fraction of maiden spawners,
+and we do not expect these to substantially affect our results.
 The model also assumes that mortality in the second winter at sea (\St) for 2SW returns
@@
 \begin{equation}
-log(smolts_{obs,t,r}) = log(smolts_{true,t,r}) + \epsilon_{t,r},
+log(smolts_{obs,r,t}) = log(smolts_{true,r,t}) + \epsilon_{r,t},
 \end{equation}
 
-where $smolts_{true,t,r}$ are the true smolt abundances for year $t$ and river
-$r$, and $\epsilon_{t,r}$ is the error term, which is calculated from the yearly 
+where $smolts_{true,r,t}$ are the true smolt abundances for river
+$r$ and year $t$, and $\epsilon_{r,t}$ is the error term, which is calculated from the yearly
 coefficient of variation in the empirically derived
-smolt estimates (see Table S1 in the Supplementary material). 
+smolt estimates (see Tables S4-6 in the Supplementary Material).
 Where available, we used population-specific measurement error estimates for smolt abundances; if not
-available, we set measurement error between 5 and 10\% (see Table S2 in the Supplementary material). 
+available, we set the measurement error at 5 or 10\%, depending on whether the smolt
+abundance estimates originated from direct counts or mark-recapture studies, respectively
+(see Table S5 in the Supplementary Material).
 The log-transformed true smolt abundances are
@@
 \begin{equation}
-log(smolts_{true,t,r}) \sim Normal(\mu_{smolts,r}, \sigma_{smolts,r})
+log(smolts_{true,r,t}) \sim Normal(\mu_{smolts,r}, \sigma_{smolts,r})
 \end{equation}
@@
 \begin{align}
-    R_{1,t} &= smolts_{true,t-1} * S_{1,t} * P_{g,t} \label{eq:1}, \\
-    R_{2,t+1} &= smolts_{true,t-1} * S_{1,t} * (1 - P_{g,t}) * S_{2,t+1}, \label{eq:2}
+    R_{r,1,t} &= smolts_{true,r,t-1} * S_{1,r,t} * P_{g,r,t} \label{eq:1}, \\
+    R_{r,2,t+1} &= smolts_{true,r,t-1} * S_{1,r,t} * (1 - P_{g,r,t}) * S_{2,r,t+1}, \label{eq:2}
 \end{align}
 
-where $R_{1,t}$ and $R_{2,t+1}$ are the estimated abundances of 1SW and 2SW
-salmon returning in years $t$ and $t+1$, respectively, $smolts_{true,t-1}$ is the
-estimated number of out-migrating smolts in year $t-1$, $S_{1,t}$ is the proportion of
-salmon surviving in their first year ($t$) at sea, $P_{g,t}$ is the proportion of
-salmon that return to spawn at year $t$, $S_{2,t+1}$ is the survival in their
+where $R_{1,r,t}$ and $R_{2,r,t+1}$ are the estimated abundances of 1SW and 2SW
+salmon returning for river $r$ in years $t$ and $t+1$, respectively, $smolts_{true,r,t-1}$ is the
+estimated number of out-migrating smolts for river $r$ in year $t-1$, $S_{1,r,t}$ is the proportion of
+salmon surviving in their first year ($t$) at sea for river $r$, $P_{g,t}$ is the proportion of
+salmon that return to spawn at year $t$ for river $r$, and $S_{2,r,t+1}$ is the survival in their
 second year at sea of the same cohort of salmon who did not return to spawn at
-year $t$.
-
-However, to allow for normally and log-normally-distributed parameters we
-log-transform equations~\ref{eq:1} and~\ref{eq:2} to obtain
+year $t$ for river $r$.
+
+We log-transform equations~\ref{eq:1} and~\ref{eq:2} so the model
+is linear on the log-scale:
 
 \begin{align}
-    log(R_{1,r,t}) &= log(smolts_{r,t-1}) + log(Pr_{r,t}) - Z_{1,r,t} \label{eq:3}, \\
-    log(R_{2,r,t+1}) &= log(smolts_{r,t-1}) - Z_{1,r,t} + log(1 - Pr_{r,t})  - Z_{2,r,t+1} \label{eq:4} 
+    log(R_{1,r,t}) &= log(smolts_{r,t-1}) + log(P_{g,r,t}) - Z_{1,r,t} \label{eq:3}, \\
+    log(R_{2,r,t+1}) &= log(smolts_{r,t-1}) - Z_{1,r,t} + log(1 - P_{g,r,t})  - Z_{2,r,t+1} \label{eq:4}
 \end{align}
@@
 \begin{align}
-log(R_{obs,1,t,r}) &\sim Normal(log(R_{1,t,r}), \epsilon_{1,t,r}), \\
-log(R_{obs,2,t,r}) &\sim Normal(log(R_{2,t,r}), \epsilon_{2,t,r}) \label{eq:5} 
+log(R_{obs,1,r,t}) &\sim Normal(log(R_{1,r,t}), \epsilon_{1,r,t}), \\
+log(R_{obs,2,r,t}) &\sim Normal(log(R_{2,r,t}), \epsilon_{2,r,t}) \label{eq:5}
 \end{align}
 
-where $R_{obs,1,t,r}$ and $R_{obs,2,t,r}$ are the observed return estimates
-for year $t$ and river $r$ of 1SW and 2SW fish, respectively, $\epsilon_{1,t,r}$
-and $\epsilon_{2,t,r}$ are the process error terms. 
-These error terms are estimated indirectly by approximating the abundance of 1SW and 2SW returns
+where $R_{obs,1,r,t}$ and $R_{obs,2,r,t}$ are the observed return estimates
+for year $t$ and river $r$ of 1SW and 2SW fish, respectively, $\epsilon_{1,r,t}$
+and $\epsilon_{2,r,t}$ are the process error terms.
+These error terms are estimated by approximating the abundance of 1SW and 2SW returns
 from hypergeometric distributions based on the scale sample data.
-We were able to obtain scale data for all but two of the populations (LaHave and Nashwaak).
+We assume that the error in the total number of returns is zero;
+the error in the total number of returns is likely minor
+compared to the error in the number of outmigrating smolts and the error due to
+a small scale sample size. Hence, ignoring the error in the returns is
+unlikely to change our results significantly.
+We were able to obtain scale data for all but two of the populations (LaHave and Nashwaak), and we
+bootstrapped a distribution of abundances of annual 1SW and 2SW returns for each populations from hypergeometric distributions,
+from which we estimated the variance in log-space.
 See the Supplementary Materials for a description of the methodology and the resulting estimates.
@@
 \begin{align}
- S_{1} &= e^{-Z_1}, \\
- S_{2} &= e^{-Z_2}. 
+ S_{1,r,t} &= e^{-Z_{1,r,t}}, \\
+ S_{2,r,t} &= e^{-Z_{2,r,t}}.
 \end{align}
@@
 \begin{align}
-Z_1 &\sim logNormal(1, 0.22),   \\ 
-Z_2 &\sim logNormal(0.2, 0.3).
+Z_{1,r,t} &\sim logNormal(1, 0.22),   \\
+Z_{2,r,t} &\sim logNormal(0.2, 0.3).
 \end{align}
@@
 informative only to the extent that they limit the estimates of marine survival
-to what we considered to be biologically realistic.
-
-We estimate population-level mean logit \Pg values around which the yearly \Pg
+to what we considered to be biologically realistic based on previous studies on
+return rates of 1SW-dominated populations in NL, as well as the couple studies
+that have explored survival in the second winter at sea of Atlantic salmon in
+eastern Canada.
+Although two published values are higher than 0.8 \citep{Chaput2003b,Hubley2011}, most estimates of S2 encompass a range between 0.1 and 0.6; this is the empirical basis for our prior.
+While there is circularity in setting priors based on data for the same rivers,
+and is far from ideal, these are the only data we have available to set priors
+and the models used previously are considerably different to the one in the present study.
+
+We estimate population-level mean probit \Pg values around which the yearly \Pg
 values are normally distributed. We specify different informative hyperpriors
@@
 \begin{align}
-    logit(P_{g,t}) &\sim Normal(\mu_{Pg,r}, \sigma_{Pg,r}) \\
-    \mu_{Pg,r} &\sim 
+    probit(P_{g,t}) &\sim Normal(\mu_{Pg,r}, \sigma_{Pg,r}) \\
+    \mu_{P_{g,r}} &\sim
     \begin{cases}
@@
 The three Newfoundland populations (WAB, Campbellton and Conne) are
-1SW-dominated, while the other four have the more generic priors for $logit(P_{g,t})$.
+1SW-dominated, while the other four have the more generic priors for $probit(P_{g,t})$.
 These priors, when converted back to a proportion, are narrowly constrained
@@
 \end{figure}
-
-The model was written in Stan \citep{Carpenter2017} and run in R version 3.6.1
-\citep{RCoreTeam2019} using the \texttt{rstan} package version 2.19.2
-\citep{StanDevelopmentTeam2019}.
-The model was run with three chains and 3,000 iterations, with the first 1,500
-discarded as a burn-in. The models were considered to have converged when the
-$\hat R$ of all parameters were lower than 1.03 and the effective sample size 
-were higher than 500.
 
@@
 Lastly, to determine what parameters best explains the number of returns in
-the model (including the model uncertainty), we estimated a pseudo-$R^2$ by
+the model (including the model uncertainty), we estimated the variance in returns explained by each parameter by
 calculating the squared correlation between the estimated parameters and the
 estimated 1SW and 2SW returns ($R_{1,r,t}$ and $R_{2,r,t+1}$ in
-equation~\ref{eq:4}).
+equations~\ref{eq:3} and~\ref{eq:4}).
+
+The model was written in Stan \citep{Carpenter2017} and all analyses were run in R version 3.6.1
+\citep{RCoreTeam2019} using the \texttt{rstan} package version 2.19.2
+\citep{StanDevelopmentTeam2019}.
+The model was run with three chains and 3,000 iterations, with the first 1,500
+discarded as a burn-in. The models were considered to have converged when the
+$\hat R$ of all parameters were lower than 1.03 and the effective sample size
+were higher than 500. The data and code are available at \url{https://github.com/sebpardo/salmon-marine-survival}.
 
@@
 
-Long-term trends among populations also varied: Campbellton,
+Visual inspection of long-term trends among populations also varied: Campbellton,
 Saint-Jean, and Western Arm Brook populations showed increases in \So
@@
 Estimates of \Pg were mostly stable across time, except for the LaHave and
-Nashwaak populations; the estimates of \Pg were slightly lower in the last
+Nashwaak populations; for the La Have the estimates of \Pg were slightly lower in the last
 four years than in the previous ones, while in the Nashwaak the posterior
 estimates of \Pg in 2012 were much lower than in all other years
-(Fig.~\ref{fig:s2-faceted}). Variation in yearly estimates was highest in the
+(Fig.~\ref{fig:pr-faceted}). Variation in yearly estimates was highest in the
 LaHave, Nashwaak, and Trinit\'{e}, and lowest in the 1SW-dominated
@@
 
-Estimates of the population-level mean (\prmu) and standard deviation (\prsig) of proportion returning after one year at sea varied considerably among populations. 
+Estimates of the population-level mean (\prmu) and standard deviation (\prsig)
+of proportion returning after one year at sea varied considerably among populations.
 For all
 three 1SW-dominated populations (Campbellton, Conne, and WAB), estimates of \prmu
-were very close to 1.0 and had little variation in \prsig (Fig~\ref{fig:prmu-post}; Table S7 in the Supplementary Materials).
+were very close to 1.0 and had little variation in \prsig (Fig~\ref{fig:prmu-post}; Table S10 in the Supplementary Materials).
 Estimates of \prmu were the lowest for the two QC populations, particularly the Saint-Jean (median \prmu = 0.11).
 Estimates of \prmu for the Nashwaak and the LaHave populations were close to 0.5, with these two populations having
-the highest estimated values of \prsig, particularly the Nashwaak (Fig~\ref{fig:prmu-post};  Table S8 in the Supplementary Materials).
+the highest estimated values of \prsig, particularly the Nashwaak (Fig~\ref{fig:prmu-post};  Table S11 in the Supplementary Materials).
+
+Uniquely for the Saint-Jean population, yearly variation in \St is being
+estimated by the model but there is almost no variation in \Pg estimates
+(Fig.~\ref{fig:pr-faceted}) and a very low estimate of \prsig (Table S11
+in the Supplementary Materials). In other words, the variation in the relative
+proportion of 1SW and 2SW returns for Saint-Jean is being manifested as
+variation in \So and \St but not in \Pg rather than being manifested in \So
+and \Pg but not \St, which is the case for the other six populations. This
+difference is likely a result of the identifiability issue with the maturity
+schedule method, which for the Saint-Jean population data, combined with the priors
+used, results in the model estimating variation in \Pg rather than \St.
+
+\ref{fig:pr-faceted}
 
@@
     \caption{Posterior probability distributions resulting from the population-level estimates of the proportion returning after one winter at sea based on
-        the parameters \prmu and \prsig. Gray lines denote population-specific median estimates of \prmu and \prsig, while
+        the parameters \prmu and \prsig. Gray lines denote population-specific median estimates of $P_{g,r}$, while
        the colored lines represent a sample of 20 posterior iteration draws for each population.}
@@
 
-Roughly half (11) out of the 21 pair-wise correlations were significantt (p-value below 0.05)
+Roughly half (11) out of the 21 pair-wise correlations were significant (p-value below 0.05)
 (Fig.~\ref{fig:s1-corr}). When looking at the direction of the correlations,
@@
 measurements, but not in the process (in this respect they are equivalent to
-sample variance and not population variance).
+population variance and not sample variance).
+Our choice of priors affected these correlations:
+using weaker priors for \So and \St resulted in slightly different coefficients and significance thresholds (see Fig. S7 of Supplementary materials).
 
@@
 
-As is expected from Murphy's method, there is a strong negative correlation
+As is expected from Murphy's method, there are strong negative, yet highly uncertain, correlations
 between survival in the second year at sea (i.e. $-Z_2$) and proportion returning
@@
     \includegraphics[width=0.95\linewidth]{figures/corrected-cor-ALL.png}
-    \caption{Correlations among estimated model parameters. \emph{log(smolts)} = estimated smolt abundances in log-space, $log(P_g)$ = logit-transformed probability of
+    \caption{Correlations among estimated model parameters. \emph{log(smolts)} = estimated smolt abundances in log-space, $probit(P_g)$ = probit-transformed probability of
         returning after one year at sea, $-Z_1$ = marine survival in the first year at sea, and $-Z_2$ = marine survival in the second year at sea.
@@
 populations show positive trends in survival in the first winter at sea (\So)
-while other exhibits highly variable yet stable trends, and some show declines.
+while other exhibits highly variable yet stationary trends, and some show declines.
 Variation in survival at the population level can be large even within a
@@
 
-While correlations in marine survival trends between rivers are highly 
+While correlations in marine survival trends among rivers are highly
 variable (Fig.~\ref{fig:s1-corr}),
@@
 Our results also show that variation of survival in the first year at sea is the most
-important factor determining variation in number of returns, suggesting that changes in abundance are primarily driven by changes in marine survival. 
+important factor determining variation in number of returns, suggesting that
+changes in abundance are primarily driven by changes in early marine survival.
 Nonetheless, variation in number of out-migrating smolts and, for some populations,
@@
 proportion returning as grilse (\Pg), as these parameter estimates were highly
-uncertain and were strongly influenced by the priors.
+uncertain and in most cases the posteriors encompassed the entire range of
+the prior (suggesting that \St were strongly influenced by the prior).
 
@@
 (i.e. river vs province scales), the divergence in local versus regional
-trends could be further confounded by differences in model structure, such as
-the methodology for estimating covariance, and use of stock-recruitment
+trends could be further confounded by differences in model structure (e.g.
+methods used for estimating covariance), use of stock-recruitment
 relationships \citep{Olmos2019} rather than empirical smolt count estimates
-(present study) to estimate marine survival.
+(present study) to estimate marine survival, and separation between fishing mortality
+and natural mortality, which was not done in our study.
 % As trends in marine survival during the first winter at sea are highly
@@
 Newfoundland populations (except WAB). We drew the same conclusion for Conne
-River but not Campbellton River or WAB. It is not possible to draw broader
+River, but not Campbellton River or WAB. It is not possible to draw broader
 conclusions with data from only three Newfoundland populations, but it seems
@@
 study reached almost the same results (albeit with slightly higher
-uncertainty), using a Bayesian approach with informative priors. These
-overlapping trends obtained with two different methods suggest that our method
-is indeed effective at estimating marine survival.
+uncertainty), using a Bayesian approach with informative priors. These overlapping
+trends obtained with two different methods suggest that our method is indeed
+a viable alternative approach for estimating marine survival.
 
@@
 relative to broader-scale, even ocean-wide, drivers.
-A recent study showed that correlations between oceanic temperature, primary
+A large fraction of total marine mortality occurs as post-smolts in the early
+part of their outmigration through estuarine and coastal areas \citep{Halfyard2012, Chaput2019}.
+These differences in migration routes and resulting early survival could be one
+of the main drivers of the marine survival variation observed in the present study.
+Concomitantly, correlations between oceanic temperature, primary
 productivity, and post-smolt survival in the North Atlantic
 are better explained by broad-scale environmental trends than regional ones
-\citep{Olmos2020}.
-However these local drivers are potentially much more difficult to quantify than 
-broader ones.
+\citep{Olmos2020}, suggesting that ocean-wide changes also impact overall marine survival.
+It is important to consider that local drivers are potentially much more difficult to quantify than
+broader ones, which could impact our ability to detect local effects on post-smolt survival.
 % The synchrony reported for marine survival trends at broader spatial
@@
 
-Egg-to-smolt survival in Atlantic salmon is highly variable, and perhaps more so
+Egg-to-smolt survival in Atlantic salmon is also highly variable, and perhaps more so
 than marine survival \citep{Klemetsen2003,Chaput2015}.
-While this variation is likely attributable to changes in freshwater
+While this variation is attributable to changes in freshwater
 conditions (e.g. discharge, temperature, water quality) and uncertainty in
@@
 changes in the oceanic conditions that spawners experience could also be
-contributing to this variation.
-Nonetheless, our results suggests that while variation in number of smolts was 
+contributing to this variation through, for example, decreases in body condition, fecundity and hence fitness.
+However, while variation in number of smolts was
 important for number of returns, it was considerably less important than
-survival during the first year at sea, which suggests that return abundances 
-are most strongly influenced by the marine phase of their life cycle.
-Obviously, there would be a time lag (perhaps as much as a generation) in how such effects
-might be manifest at the adult stage.
-However, given that most correlations are between relatively monotonic declines
-in abundance coupled also monotonic increases in climatic indices
-over decadal time scales \citep[e.g.,][]{Friedland1998, Todd2008,
-    Beaugrand2012}, it would be expected that these correlations would be
-maintained even if salmon abundances were lagged by a generation length.
-
-While there is little evidence that marine survival is density-dependent in
-Atlantic salmon \citep{Jonsson1998,Gibson2006}, there could potentially be
-some density-dependent processes during parts of the marine phase, 
+survival during the first year at sea, suggesting that return abundances
+are mostly influenced by the marine phase of their life cycle.
+While there would be a generational lag in how climate-driven freshwater effects
+might manifest at the subsequent adult stages,
+a correlation between freshwater effects and adult abundances might be expected
+as most climate-salmon abundance correlations are between relatively monotonic declines
+in abundance coupled with monotonic increases in climatic indices
+over decadal time scales \citep[e.g.,][]{Friedland1998, Todd2008, Beaugrand2012}.
+There is little evidence that marine survival is density-dependent in
+Atlantic salmon \citep{Jonsson1998,Gibson2006}, but these
+density-dependent processes could occurr during parts of the marine phase,
 particularly for populations that are
 near historically low levels of abundance. One potential mechanism might be a
-``predator pit''. When prey populations are very small, predator-induced
+``predator pit'': when prey populations are very small, predator-induced
 mortality can also be low because the prey are simply not abundant enough to
@@
 also increase as preference for the prey increases.
+Another potential mechanism is if smolt quantity and quality are negatively
+correlated; abundant smolt cohorts likely experience high competition in rivers and thus
+attain smaller size or reduced body condition than less abundant cohorts.
 Correlations between return estimates and survival were negative for
@@
 
-Secondly, the assumed hierarchical structure might not be the most appropriate
-for modelling smolt abundances. This approach results in shrinkage to the
-mean, which means that the variation in annual smolt estimates is less than
-it would be if not modelled hierarchically. 
-That said, this assumption is likely to result in more conservative trends in
-marine survival, as the variation is smolt estimates is reduced.
-This shrinkage of annual smolt abundances is also reducing the $R^2$ values 
-between smolts and estimates returns, albeit only marginally.
-It is important to note that the hierarchical structure in the estimation of
-\Pg seems like a reasonable assumption given that the probability of returning
-as grilse has a genetic component associated to it \citep{Barson2015} and its inter-annual
-variation in a given population is likely partly constrained by this.
-
-Thirdly, the accuracy and precision of the model's output could be different
+Secondly, our error estimates of 1SW and 2SW returns might not approximate the
+underlying uncertainty. Nonetheless, our approach likely overestimates
+uncertainty as we did not model error in annual 1SW and 2SW returns
+hierarchically within each population (see Supplementary Materials for
+details), which should produce conservative results.
+
+% Secondly, the assumed hierarchical structure might not be the most appropriate
+% for modelling smolt abundances. This approach results in shrinkage to the
+% mean, which means that the variation in annual smolt estimates is less than
+% it would be if not modelled hierarchically.
+% That said, this assumption is likely to result in more conservative trends in
+% marine survival, as the variation is smolt estimates is reduced.
+% This shrinkage of annual smolt abundances is also reducing the $R^2$ values
+% between smolts and estimates returns, albeit only marginally.
+% It is important to note that the hierarchical structure in the estimation of
+% \Pg seems like a reasonable assumption given that the probability of returning
+% as grilse has a genetic component associated to it \citep{Barson2015} and its inter-annual
+% variation in a given population is likely partly constrained by this.
+
+Thirdly, the accuracy and precision of the model's output is different
 for different populations. Survival in the first year at sea is the variable
 that explains most of the variance in estimates of 1SW returns, except for the
-Saint-Jean; 
-the model does not provide reliable parameter estimates for this population
-population (as indicated by the low $R^2$ values in Fig.~\ref{fig:mu-rsq}),
+Saint-Jean; it seems
+the model does not provide reliable parameter estimates for this population,
 which is the population with the highest proportion of 2SW returns.
-Furthermore, estimates of \St and \Pg are negatively correlated and highly
+Furthermore, estimates of \St and \Pg in all populations are negatively correlated and highly
 uncertain, and this must be considered when making any inferences based on
@@
 Lastly, our model does not differentiate mortality resulting from natural
-causes (e.g. predation) or from fishing; we know the latter has decreased
-dramatically due to the sequential moratoria enacted across eastern Canada, and this
-confounds the estimated trends in marine survival because we know marine survival
-will be reduced in years where commercial fishing occurred (i.e. pre-2000s).
-Thus, any increasing trends in marine survival are confounded by the additional 
+causes (e.g. predation) or from anthropogenic factors such as fishing or
+interactions with aquaculture. Fishing mortality has decreased
+significantly due to the sequential moratoria enacted across eastern Canada, and
+aquaculture, which has grown over recent decades, can negatively impact salmon survival through sea lice
+\citep{Shephard2020,Bohn2020} or genetic introgression \citep{Glover2017,Vollset2020}.
+Thus, estimated trends in marine survival are confounded because we know marine
+survival will be reduced in years where commercial fishing occurred (i.e.
+pre-2000s), while the trends in aquaculture-related mortality are unknown.
+In other words, any increasing trends in marine survival are confounded by the additional
 fishing mortality before the turn of the century.
@@
 Perhaps a reframing of the issue of marine survival is key to furthering our
-understanding of Atlantic salmon population dynamics. Marine survival may not
-have declined consistently, and over the same time periods, across all
+understanding of Atlantic salmon population dynamics. Marine survival has not
+declined consistently, and over the same time periods, across all
 populations, even if the overall trend has been one of decline.
@@
 Salmon Research Joint Venture.
+GHB was funded by the Norwegian Research Council (project 275862 and 280308).
 

### 2021-05-29 ms/ms.tex
Final edits before acceptance

@@
 we averaged the proportions of annual returns
-for the closest ten years for which there was data.
+for the closest ten years for which there were data.
 We incorporated the uncertainty in the conversion between
@@
 schedule method, in conjunction with informative priors, to estimate annual
-marine survival in seven populations of Atlantic salmon. We account for
+marine survival. We account for
 observation error in smolt and return estimates, as well as estimating the
@@
 coefficient of variation in the empirically derived
-smolt estimates (see Tables S4-6 in the Supplementary Material).
+smolt estimates (see Tables S4-6 in the Supplementary Materials).
 Where available, we used population-specific measurement error estimates for smolt abundances; if not
@@
 abundance estimates originated from direct counts or mark-recapture studies, respectively
-(see Table S5 in the Supplementary Material).
+(see Table S5 in the Supplementary Materials).
 The log-transformed true smolt abundances are
@@
 to what we considered to be biologically realistic based on previous studies on
-return rates of 1SW-dominated populations in NL, as well as the couple studies
+return rates of 1SW-dominated populations in NL, as well as the few studies
 that have explored survival in the second winter at sea of Atlantic salmon in
@@
 (Fig.~\ref{fig:priors}b).
+
+We explored the effect of using weaker priors for $Z_1$, $Z_2$ and \Pg, and found
+that the posteriors of $Z_1$ were not sensitive to the choice of prior, while the 
+ posteriors $Z_2$ and \Pg became more uncertain with weaker priors while maintaining
+ the same temporal trends observed when using stronger priors. See Figs S2-7 in the Supplementary Materials.
 
@@
 
-Lastly, to determine what parameters best explains the number of returns in
+Lastly, to determine which parameters best explain the number of returns in
 the model (including the model uncertainty), we estimated the variance in returns explained by each parameter by
@@
 non-significant between WAB and Campbellton but negative between Conne and WAB (NL).
-The Campbellton had mostly non-significant pairwise correlations with other populations, even though there are both positive and negative correlation coefficients, except with the Saint-Jean
+Campbellton had mostly non-significant pairwise correlations with other populations, even though there are both positive and negative correlation coefficients, except with the Saint-Jean
 where the correlation between $Z_1$ trends is positive and significant.
@@
 Our choice of priors affected these correlations:
-using weaker priors for \So and \St resulted in slightly different coefficients and significance thresholds (see Fig. S7 of Supplementary materials).
+using weaker priors for \So and \St resulted in slightly different coefficients and significance thresholds (see Fig. S7 of Supplementary Materials).
 
@@
     \includegraphics[width=0.95\linewidth]{figures/corrplot-post-z1-corrected.png} \caption{
-        Correlations of estimated trends in $Z_1$ among populations.
+        Correlations of estimated trends in instantaneous natural mortality in the first winter at sea, $Z_1$, among populations.
         Correlation coefficients are shown in each square, while colouring
@@
 populations show positive trends in survival in the first winter at sea (\So)
-while other exhibits highly variable yet stationary trends, and some show declines.
+while others exhibit highly variable yet stationary trends, and some show declines.
 Variation in survival at the population level can be large even within a
@@
 uncertain, and this must be considered when making any inferences based on
-these parameters for any of the populations examined.
+these parameters for any of the populations examined, at least for those population appreciable contributions of 2SW salmon.
 
@@
 
+\section*{Data Availability Statement}
+
+The data and code underlying this article are available in GitHub, at \url{https://github.com/sebpardo/salmon-marine-survival}.
+
 \section*{Acknowledgements}
@@
 Salmon Research Joint Venture.
-GHB was funded by the Norwegian Research Council (project 275862 and 280308).
+GHB was funded by the Norwegian Research Council (projects 275862 and 280308).
 

### 2021-05-29 ms/suppmat.tex
Final edits before acceptance

@@
 
-The model runs with weaker priors did not differ mayorly from those with
+The model runs with weaker priors did not differ majorly from those with
 stronger priors, particularly for posterior estimates of \So, which show the
@@
     \includegraphics[width=0.95\linewidth]{figures/corrplot-post-z1-corrected-weakz.png} \caption{
-        Correlations of estimated trends in $Z_1$ among populations with weak \So and
+        Correlations of estimated trends in in instantaneous natural mortality in the first winter at sea, $Z_1$, among populations with weak \So and
 \St priors. Correlation coefficients are shown in each square, while colouring
